### 2020-10-13 comments.tex
Initial Overleaf Import

\documentclass[12pt]{article}
\begin{document}
\section*{Section 2}
\begin{enumerate}
    \item Should we use ``stationary measure'' or ``stationary distribution'' for the SRB measure, instead of ``equilibrium distribution''? This is because physically, the dynamical systems we consider may not be in equilibrium: systems with an attractor, when ``energy'' is dissipated, may be in nonequilibrium steady states. A fluid dynamics/physics audience may assume that we are concerned only with systems that attain an equilibrium, if we call the SRB measure 
    an equilibrium measure.
    \item Another point about the SRB measure: the way they are introduced in section 2 (as the probability distribution of a long trajectory) is the best possible. When we talk about the Birkhoff ergodic theorem, it may not be accurate to introduce the SRB measure right afterwards because the existence of the SRB measure does not follow from Birkhoff's theorem. Secondly, without going into mathematical details, we can insert information that forces us to define SRB measure: that it is a special stationary measure which is physical. That is, we can say it is the particular measure, which is smooth along the attractor filaments, and this property is enough to guarantee that it is also the same as $\mu$ defined using the long trajectory. I have given this a shot by adding a few sentences but if this does not add value to the paper and/or is confusing, please remove them!
\end{enumerate}
\subsection*{Section 2.2: Lorenz map}
\begin{enumerate}
    \item To numerically find periodic orbits of the Lorenz system using the Lorenz map, we obtain the a value of $u_3$ for a point on a periodic orbit. However, there is numerical error in our reading of $u_3$ and hence in $u_1,$ $u_2$ as well. Does this make it so that the generated orbit is not periodic but a typical chaotic orbit? The effect of numerical error may be more pronounced for periodic orbits of longer period.  
    \item I am curious also about how to find local maxima in a program. Is the `scipy.find\_peaks` function a good way?
\end{enumerate} 
\end{document}

### 2020-10-13 main.tex
Initial Overleaf Import

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
\usepackage[utf8]{inputenc}
\usepackage{graphicx, float}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{xcolor}
\definecolor{darkgreen}{rgb}{0.0, 0.5, 0.0}
\newcommand{\nisha}[1]{{\color{darkgreen} #1}}
\begin{document}

\title{On the probability of finding a nonphysical solution through shadowing}
\date{June 2020}
\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Qiqi Wang}

%\authorrunning{Short form of author list} % if too long for running head

\institute{F. Author \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{fauthor@example.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           S. Author \at
              second address
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
Insert your abstract here. Include keywords, PACS and mathematical
subject classification numbers as needed.
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\maketitle
\section{Shadowing and some applications}

Shadowing refers to the relationship between a pair of solutions to slightly different governing equations. The difference between the governing equations can be due to parameter perturbation or numerical error. The solution to one governing equation is said to shadow a solution to a second, slightly different equation if
the first solution stays close to the second solution for some amount of time.

The existence of arbitrarily long shadowing solutions for sufficiently similar governing equations
has been proven for invertible hyperbolic maps by Anosov (1967) and Bowen (1970, 1978).
For non-hyperbolic systems, it is often possible to estimate the length of time for which shadowing
solutions exist. (Grebogi, Hammel, Yorke and Sauer (1990), Sauer, Grebogi and Yorke (1997)).

The existence of shadowing trajectories underlies many applications.
Numerical simulations of turbulent flows have been widely used to study its statistical behavior.
It is argued, based on shadowing, that such numerical simulations of chaotic dynamical systems can be
useful, despite the butterfly effect. Consider a numerical solution to a deterministic chaotic system with an initial condition $x$.
Because of numerical and modeling errors, there will typically be a small difference between the
true governing physics and the equations solved on a computer.  As the system is evolved forward
in time, this difference will be amplified exponentially due to the chaotic nature of the system.
The question therefore arises as to whether or not numerical solutions to chaotic systems, such
as turbulent flows, are related to the physics at all.  The existence of shadowing solutions
is used to argue for the usefulness of such numerical solutions.
When certain conditions for shadowing theorems are met, the numerical solution would be an
approximation to a ``true'' solution that satisfies the real governing physics.

Shadowing is also used in sensitivity analysis of chaotic dynamical systems.  In particular,
it is used in computing how long-time-averages in a chaotic system
respond to small perturbations in the governing equation. In this application, derivative
of the long-time-averages is computed using a solution to the perturbed governing equation
that shadows a solution to the unperturbed equation.  The Least Squares Shadowing \cite{qiqi-lss} method
uses this concept.

There is an implicit assumption in both of these applications.  The assumption is that
the shadowing trajectory is a ``physical'' trajectory, a trajectory on which
the long-time-average of a quantity is equal to the physical \nisha{can we remove the word ``physical'' before ensemble average?} ensemble average.
Not all solutions satisfying the physical governing equation are considered ``physical''
in this sense.  In high-Reynolds number fluid flows, for example, a steady-state, laminar flow solution may satisfy the Navier-Stokes equation. But such a solution would never be observed in reality because it is unstable, and any small perturbation
would trip it into turbulence. Unstable steady-state solutions are not the only
non-physical solution. Many chaotic dynamical systems have infinitely many periodic
solutions that are, similar to their steady counterpart, unstable. These trajectories can have a probability distribution that is remarkably different from that of a typically observed solution of the governing equation. It is also possible to effect significant change in the statistics of the true governing dynamics by introducing a minor parameter perturbation. Thus, although two trajectories, at slightly different parameters, may shadow each other, the long-time-averages of observables calculated using them, may not be close.    
As we will see
in the next section, even solutions that look ``chaotic'', i.e., unsteady and aperiodic, may not be physical.

\section{What are physical and nonphysical solutions?}

Intuitively, we call a solution to a governing equation ``physical'' if
it represents what one would observe in a physical experiment.  In particular, the statistics measured from a physical solution match the statistics observed
in an experiment.   Not all solutions are physical.  A laminar flow
solution, despite satisfying the governing equation, does not produce the
turbulent statistics one would observe in a high-Reynolds number experiment. Such solutions are thus called nonphysical.

In this section, we first describe mathematically what distinguishes a physical
and a nonphysical solution.  We also explain theoretically why it is unlikely
to observe a nonphysical solution in an experiment.  We then give a few
examples of these almost-never-observed nonphysical solutions.

\subsection{Physical solutions}
What we typically call ``physical'' solutions to a governing equation satisfy the following
two criteria: \nisha{Is the term ``physical'' to mean the following common in fluid mechanics? If not, can we emphasize that we are defining it here? e.g. }
\begin{enumerate}
    \item Long-time-averaged, ``statistical'' quantities converges as time goes to infinity.
    Consider $u(t)$ as the solution to a chaotic governing equation, then for \nisha{a continuous/regular observable?}
    a smooth quantity of interest $J(u(t))$,
    \begin{equation} \label{longtimeavg}
        \lim_{T\to\infty}\frac1T \int_0^T J(u(t))\,dt
    \end{equation}
    exists.
    \item For almost any small perturbation to the initial condition $u(0)\to u'(0)$, the perturbed
    solution $u'(t)$ should have the same statistics, i.e.,
    \begin{equation}
        \lim_{T\to\infty}\frac1T \int_0^T J(u(t))\,dt
      = \lim_{T\to\infty}\frac1T \int_0^T J(u'(t))\,dt
    \end{equation}
\end{enumerate}

\begin{figure}[H] \centering
\includegraphics[width=0.48\textwidth]{figure/lorenz_trajectory_1000.0.png}
\hspace{0.02\textwidth}
\includegraphics[width=0.48\textwidth]{figure/lorenz_ensemble_10.png}
\caption{
Left: An ensemble of initial conditions distributed uniformly in the box $u_1\in[0,1], u_2\in[0,1], u_3\in[28,39]$ 
after 10 time units of evolution, shown in in the $u_1-u_3$ plane.
Right: distribution of a trajectory of $1000$ time units in length.
In both plots, the color represent the number of samples in a
$2048\times2048$ uniform grid.  The trajectory is sampled every 0.001 time units.
}
\label{fig:lorenz_ergodicity1}
\end{figure}
\begin{figure}[H] \centering
\includegraphics[width=0.48\textwidth]{figure/lorenz_trajectory_10000.0.png}
\hspace{0.02\textwidth}
\includegraphics[width=0.48\textwidth]{figure/lorenz_ensemble_15.png}
\caption{
Left: distribution of the same ensemble as in Figure \ref{fig:lorenz_ergodicity1}
after another 5 time units of evolution.
Right: distribution of the same trajectory as in Figure \ref{fig:lorenz_ergodicity1}
evolved for $10,000$ time units.
}
\label{fig:lorenz_ergodicity2}
\end{figure}
\begin{figure}[H] \centering
\includegraphics[width=0.48\textwidth]{figure/lorenz_trajectory_1000000.0.png}
\hspace{0.02\textwidth}
\includegraphics[width=0.48\textwidth]{figure/lorenz_ensemble_50.png}
\caption{
Left: distribution of the same ensemble as in Figures \ref{fig:lorenz_ergodicity1}
and \ref{fig:lorenz_ergodicity2} after a total of 50 time units of evolution.
Right: distribution of the same trajectory as in Figures \ref{fig:lorenz_ergodicity1}
and \ref{fig:lorenz_ergodicity2} evolved for $100,000$ time units.
}
\label{fig:lorenz_ergodicity3}
\end{figure}

The left column of Figures \ref{fig:lorenz_ergodicity1}-\ref{fig:lorenz_ergodicity3}
illustrates a physical solution of the Lorenz system,
\begin{equation}
\label{eqn:lorenz}
\frac{du}{dt} =
    \frac{d}{dt}\begin{pmatrix}u_1\\u_2\\u_3\end{pmatrix} =
    \begin{pmatrix}
        \sigma (u_2-u_1) \\
        u_1(\rho -u_3) - u_2 \\
        u_1\,u_2 - \beta\,u_3
\end{pmatrix}
\end{equation}
where $\sigma=10,\rho=28$, and $\beta=8/3$.  The solution starts at the initial
condition $u(0)=(0.01, 0.01, 28)$.  After a long time evolution, we observe that
the solution has visited a large portion of the $u_1-u_3$ plane with a varying, but well-defined, frequency.  We can use a probability distribution, $\mu$, to quantify
how frequently a physical solution visits portions the phase space.
Specifically, for a subset $A$ of the phase space, $\mu(A)$ measures the
fraction of time a very long physical solution spends inside the subset $A$.
With this distribution defined, the infinite-time average of any quantity $J(u)$ can be represented as an average of $J$ over the entire phase space $U$, weighted
by this statistical distribution $\mu$.  Mathematically,
\begin{equation}
    \frac1T \int_0^T J(u(t))\,dt \xrightarrow{T\to\infty}
    \int_U J(u)\: d\mu(u)
\end{equation}

Remarkably, the distribution $\mu$ not only characterizes the history of
a long physical solution, but also describes the settled state of
an ensemble of solutions.
This is illustrated on the right column of Figures \ref{fig:lorenz_ergodicity1}--\ref{fig:lorenz_ergodicity3}.
We generate these plots by starting from an ensemble of about one billion initial
conditions, randomly and uniformly spaced in a small three-dimensional box.  All these billion solutions are evolved by solving Eq. \ref{eqn:lorenz} for 10, 15, and 50 time units, to obtain the plots. 
We observe that, as time evolves, the ensemble spreads over an increasingly
larger portion of the $u_1-u_3$ plane. After a long time, the ensemble settles into a set known as an \emph{attractor} that does not change with time, and contains the \emph{unstable manifold}, which appears as filaments forming the attractor. The distribution of the ensemble on the attractor becomes identical to the distribution of a single, very long physical
solution, which is also contained in the attractor (Figure \ref{fig:lorenz_ergodicity3}).

This remarkable agreement has been thoroughly studied under the subject
of ergodic theory. Under surprisingly weak conditions, a solution starting from
{\bf almost} any initial condition, chosen randomly from a set enclosing the attractor, is a physical solution \cite{young}.   
Meanwhile, an ensemble of trajectories starting from any distribution with a 
finite density also evolves towards the same final distribution, $\mu$. Due to expansion of a volume of solutions tangent to the attractor filaments, a finite density under long-time evolution becomes absolutely continuous on the unstable manifold, i.e., the likelihood of a trajectory visiting any set not intersecting the attractor filaments is zero. The 
stationary distribution achieved on long-time evolution of the ensemble that has the absolute continuity property is called the Sinai-Ruelle-Bowen
(SRB) measure. The absolute continuity property is sufficient to ensure that the SRB measure is the same $\mu.$ That is, the SRB measure is physically observed in the sense that  
\nisha{what do we mean by ``under sufficient regularity conditions'' here? Suggesting a modification to the previous sentence} physical solutions produce long-time-averages which are expectations with respect to the SRB measure.

Note that almost any, not any, initial condition leads to a physical solution. A set of special initial conditions
contained in a neighborhood of the attractor may exist starting from which physical solutions are not generated. These initial conditions do not produce the same statistics as the physical solutions. This special set of initial conditions is Lebesgue measure zero -- one
has zero chance of finding such an initial condition by randomly sampling.
Nonphysical solutions thus take an effort to find.  Nevertheless, they
turn out to be important when discussing shadowing, the topic of this paper.
We first introduce a type of obviously nonphysical solutions in the next
subsection, before discussing a less obvious type in Section \ref{sec:quasiphysical}.

\subsection{Nonphysical solutions Type I: Periodic Solutions}

A periodic solution is nonphysical because it 
does not visit as much
of the phase space as a physical solution does.
Figure \ref{fig:lorenz_periodic} shows
a few periodic solutions of the Lorenz equation. Comparing these solutions
to the physical solution visualized in Figures \ref{fig:lorenz_ergodicity1}--\ref{fig:lorenz_ergodicity3},
we see that the periodic solutions are significantly more limited in their
extent of exploration.

\begin{figure}[H]\centering
\includegraphics[width=0.48\textwidth]{figure/lorenz_periodic.png}
\caption{Periodic solutions of the Lorenz equation, overlaid on top of
its SRB distribution.  The solid line, dashed line, and dotted line
represent three distinct periodic solutions.}
\label{fig:lorenz_periodic}
\end{figure}

Because periodic solutions have a more limited extent in the phase space, their
statistics are different from physical solutions.
Here we illustrate the difference using the mean of two
quantities of interest
\begin{align}
    J_1(u) = u_3\;,\quad J_2(u) = e^{-\dfrac{u_3^2}2}.
\end{align}
Table \ref{tab:lorenz_periodic_stats} shows these quantities
of interest averaged over the three periodic solutions shown in Figure
\ref{fig:lorenz_periodic}, compared against those averaged over a
physical solution.  Here, Periodic \#1, \#2, and \#3 correspond to
the solid, dashed, and dotted lines, respectively.
We observe from the table that although the mean of $J_1$ over the periodic
solutions are different but comparable to the mean over physical solutions.  The mean of $J_2$ over the periodic solutions, on the other hand,
is orders of magnitude different from that of the physical solutions.
These differences disqualify the periodic solutions from being physical.
\begin{table}[H]
    \centering
    \begin{tabular}{c|c|c|c|c}
             & Periodic \#1 & Periodic \#2
             & Periodic \#3 & Physical solutions \\
    \hline
        $J_1$& 23.05 &
               23.19 &
               23.37 &
               23.67 \\
        $J_2$&$4\times 10^{-35}$&
              $7\times 10^{-28}$&
              $3\times 10^{-22}$&
              $1.58\times 10^{-05}$
    \end{tabular}
    \caption{Comparison of statistics computed from periodic solutions of the Lorenz equation with the statistics computed from physical solutions}
    \label{tab:lorenz_periodic_stats}
\end{table}

While periodic solutions are generally difficult to find,
the Lorenz equation has a special feature that makes the task significantly
easier.  In a typical solution to the Lorenz equation, the 
$u_3(t)$ component oscillates in a pattern that appears neither regular nor random.  Lorenz
observed that the height of one peak in the oscillation can predict the height of the next peak.  He quantified his observation by the Lorenz map,
as shown in Figure \ref{fig:lorenz_map}.

\begin{figure}[H]\centering
\includegraphics[width=0.48\textwidth]{figure/lorenz_zmax_10_28_2.6666666666666665.png}
\hspace{0.02\textwidth}
\includegraphics[width=0.48\textwidth]{figure/lorenz_zmax2_10_28_2.6666666666666665.png}
\caption{Left: the Lorenz map.  The x-axis is the $n$-th local maximum of $u_3(t)$ over a
long solution; the y-axis is the $(n+1)$-th local maximum of $u_3(t)$.  The intersection of this curve
with the diagonal dashed line indicates the initial condition for the solid line
in Figure \ref{fig:lorenz_periodic}.
Right: the Lorenz map iterated twice.  The x-axis is the $n$-th local maximum
of $u_3(t)$; the y-axis is the $(n+2)$-th maximum.  The intersection with the diagonal dashed
line indicates the initial conditions for both the solid and dashed lines in
Figure \ref{fig:lorenz_periodic}. \nisha{In Figure 4, the solid line appears only to have one local maximum, which is shown as the fixed point on the left plot of this figure. The two fixed points of the 2-time-iterate of the Lorenz map, shown on the right plot of this figure, both appear to correspond to the local maxima of the dashed curve in Figure 4.}}
\label{fig:lorenz_map}
\end{figure}

The Lorenz map provides us a tool to find as many periodic solutions as we want.
By intersecting the map with a diagonal line, we can find a local maximum of $u_3(t)$
for which the next maximum is almost the same value.  We then look up the values of
$u_1(t)$ and $u_2(t)$ when $u_3(t)$ achieved this maximum.  This gives us an initial
condition starting from which the solution is nearly periodic.   We can similarly
intersect the second iterate of the Lorenz map (right plot of Figure \ref{fig:lorenz_map},)
and the third iterate, etc, with a diagonal line to find increasingly
complex periodic solutions.

This Lorenz map is more than a tool to study the Lorenz equation.  It is a chaotic
dynamical system all by itself. Unlike the Lorenz equation, which is a continuous-time
dynamical system in three dimensions, the Lorenz map is a discrete-time dynamical
system in one dimension.  It exhibits the same sensitivity to initial condition
that characterizes the Lorenz equation.  One can readily observe in the right
plot of Figure \ref{fig:lorenz_map} that a small perturbation in the x-axis
can lead to a large change in the y-axis. This sensitivity grows exponentially
for the $n$-th iterate of this map. A solution to the Lorenz map can be obtained
by extracting the consecutive local maxima of a solution to the Lorenz equation.
If we extract from a physical solution to the Lorenz equation, we obtain a physical solution of the Lorenz map.  It will visit the
interval between 29 and 49 with varying, well-defined frequencies.
By contrast, if one extracts a solution to the Lorenz map from a periodic solution
to the Lorenz equation, the solution will visit only a discrete set of points.
It is thus a periodic, non-physical solution.  What we learned about
the Lorenz equation could have all been learned from the Lorenz map.

Having discussed periodic solutions in this section,
we move to a second type of nonphysical solutions.
This type is more difficult to find and study than the periodic solutions.
To make it easier, we switch our example to a one-dimensional, discrete time dynamical
systems like the Lorenz map.  Because the Lorenz map lacks a closed form,
we construct, in the next subsection, a one-dimensional discrete-time dynamical system
with a closed form, one that qualitatively resembles the Lorenz map.  This map
will help us, in Section \ref{sec:quasiphysical}, to study a more insidious type
of nonphysical solutions, a type that hides in the shadow.
\nisha{I wish all papers were this intriguing!}
\subsection{Tent map: periodic and physical solutions}
\label{sec:tent}
The tent map has been widely studied \cite{tent1}\cite{tent2}.  It is qualitatively similar
to the Lorenz map and has a simple, analytic form:
\begin{equation} \label{tentmap}
    \varphi(x) := \begin{cases}
    2x \quad & x < 1 \\
    4 - 2x \quad & 2 \ge x \ge 1
    \end{cases}
\end{equation}
Figure \ref{fig:tent_map} shows the tent map.
\begin{figure}[H]\centering
\includegraphics[width=0.48\textwidth]{figure/tent_map.png}
\caption{The tent map $\varphi$.}
\label{fig:tent_map}
\end{figure}

The tent map $\varphi$ is chaotic because a trajectory $x_0,x_1,\ldots$ satisfying
$x_{i+1} = \varphi(x_i)$ exhibits exponential sensitivity to initial condition. An infinitesimal perturbation of absolute value $\delta x$, applied to an initial condition $x_0$, generates a trajectory that is $2 \delta x$ away from $x_1,$ $4 \delta x$ from $x_2$, $8 \delta x$ from $x_3$ and so on. This exponential divergence of two trajectories that start infinitesimally apart is the butterfly effect that characterizes
chaotic dynamics.

It is easy to find periodic solutions for the tent map.
$\frac43$ maps to itself; $\frac45$ and $\frac85$ map to each other;
$\frac47, \frac87$, and $\frac{12}7$ map circularly.  In fact, any rational number evolves into periodic solutions that visit
only a finite set of rational numbers with the same denominator.
This is compatible
with ergodic theory because all rational numbers in $[0,2]$ comprise a subset of Lebesgue measure zero. That is, we would have zero likelihood of getting a
rational number if sampled the Lebesgue measure (the uniform distribution) on $[0,2]$.  Instead, with a 100\% probability, one would get an irrational number
that, when iterated under the tent map, leads to a physical trajectory that distributes uniformly on the interval $[0,2]$.

To understand why a physical solution visits the interval $[0,2]$ at a uniform
frequency, it is helpful to view the tent map from a different perspective.
For a randomly chosen $x_0\in[0,2]$, we can represent it in binary form:
\begin{equation}
    x_0 = \sum_{j=0}^{\infty} \frac{x_{0,j}}{2^j}
\end{equation}
where $x_{0,0}\in\{0,1\}$ is the integer component of $x_0$ and $x_{0,j}\in\{0,1\}$ is the $j$th bit after the binary point.
Let $x_{i+1} = \varphi(x_i)$, then it is straightforward to verify from the definition
of the map that its binary representation
\begin{equation}
    x_i = \sum_{j=0}^{\infty} \frac{x_{i,j}}{2^j}
\end{equation}
would satisfy
\begin{equation} \label{tent_xor}
    x_{i+1,j} = x_{i,0} \veebar x_{i,j+1}
\end{equation}
where $\veebar$ is the xor operator. To see why, note that 
multiplication by 2 is a left-shift operator in binary, and 
subtraction from 4 flips every bit after the binary point. If $x_0$ is chosen uniformly in $[0,2]$, then each bit $x_{0,j},j=0,1,\ldots$ of $x_0$
has equal probability of being 0 or 1, and each bit is independent of other bits.
It follows from Eq. \ref{tent_xor} that each bit $x_{i,j},j=0,1,\ldots$ of $x_i,
i=1,2,\ldots$ has equal probability of being 0 or 1,
and each bit is still independent of the other bits.
Thus, every $x_i$ in this physical solution is uniformly distributed in $[0,2]$.
As the map iterates starting from almost any $x_0$, a physical solution explores the entire interval $[0,2]$ uniformly.

\subsection{Nonphysical solutions Type II: Quasi-physical Solutions}
\label{sec:quasiphysical}
The simplicity of the tent map, as well as its binary decimal form (\ref{tent_xor}),
enables us to study non-physical solutions that are not periodic.
As in the last section, consider an
\begin{equation}
    x_0 = \sum_{j=0}^{\infty} \frac{x_{0,j}}{2^j}
\end{equation}
the digits of which are not independent of each other.  Instead,
suppose each digit is more likely to repeat the previous digit than to be different.
That is, $x_{0,j+1}=x_{0,j}$ with probability $p>\frac12$ for $j=0,1,2,\ldots$.
Then, by Equation (\ref{tent_xor}), the digits of every $x_i, i=1,2,\ldots$
have digits that repeats its previous digit with probability $p$.
Moreover, consider Equation (\ref{tent_xor}) for $j=0$ and any $i$:
\begin{equation}
    x_{i+1,0} = x_{i,0} \veebar x_{i,1}
\end{equation}
Because $x_{i,0}=x_{i,1}$ with probability $p>\frac12$, $x_{i+1,0}=0$ with probability 
$p>\frac12$.  The solution, starting from $x_1$, visits $[0,1]$ with
probability $p>\frac12$.  Instead of visiting $[0,1]$ and $[1,2]$ with
equal probability, as a physical solution does, this solution favors $[0,1]$.
Because it visits the interval $[0,2]$ with a provably different frequency from that
of a physical solution, this is a nonphysical solution.

Note that the solution we just constructed is both nonphysical and aperiodic.
The digits of $x_0$, though correlated with each other,
still can exhibit an infinite variety of patterns.  As the map iterates, these patterns
shift towards more significant digits, and the solution visits an infinite set of
points.  We call such aperiodic nonphysical solutions ``quasi-physical'' solutions.

\begin{figure}[H]\centering
\includegraphics[width=0.48\textwidth]{figure/tent_quasiphysical_hist_p_0.5001.png}
\hspace{0.02\textwidth}
\includegraphics[width=0.48\textwidth]{figure/tent_quasiphysical_hist_p_0.51.png}
\\ \vspace{0.05\textwidth}
\includegraphics[width=0.48\textwidth]{figure/tent_quasiphysical_hist_p_0.55.png}
\hspace{0.02\textwidth}
\includegraphics[width=0.48\textwidth]{figure/tent_quasiphysical_hist_p_0.9.png}
\caption{Empirical density function of long solutions (a billion steps)
starting from four point whose binary digits have probability $p$ of
repeating the previous digit.
The solution shown in the top-left plot starts from a point with $p=0.5001$;
top-right: $p=0.51$; bottom-left: $p=0.55$; bottom-right: $p=0.9$.
Note that the empirical densities at larger $p$s go well beyond 1.}
\label{fig:tent_quasiphysical}
\end{figure}

Figure \ref{fig:tent_quasiphysical} shows the empirical density functions of
four such quasi-physical solutions.  When $p=0.5001$,
the statistical distribution of the quasi-physical solution
is approximately uniform.  Recall that a physical solution explores $[0,2]$
uniformly.  In this case, the quasi-physical solution has very similar
statistical behavior as a physical solution.  When $p=0.51$, the empirical
distribution becomes ``hairy''.  An apparently fractal pattern emerges.
This fractal pattern further amplifies when $p=0.55$.  Meanwhile,
the density on the left side, $[0,1]$, becomes obviously higher than
the density on the right side, $[1,2]$.  This is consistent with our
theoretical analysis in the first paragraph of this subsection.
When $p=0.9$, the fractal pattern is so intensified that most
of the solution seem to concentrate in a collection of tiny intervals.
These plots exposes the diversity of quasi-physical solutions.
Their statistical distribution can be either
very similar or completely different from that of physical solutions.

So far we have constructed one class of quasi-physical solutions.
It is noteworthy that there are infinite variations to how we
construct quasi-physical solutions.  In the binary decimal
representation of the initial condition $x_0$, any statistical
deviation from equal probability of 0 or 1, or any statistical
dependence among digits would lead to quasi-physical solutions.
One could, for example, construct an $x_0$ whose digit is
more likely to be 1 than 0 only if it follows two consecutive 0's.
Such $x_0$ would lead to a nonphysical solution.  Its statistical
distribution would differ from any of the plots in Figure
\ref{fig:tent_quasiphysical}.  Nevertheless, as can be
observed in Figure \ref{fig:tent_alternative}, it would show
a remarkable resemblance in its ``hairiness''; some kind of fractal
pattern would emerge from the distribution.  A fractal density
appears to be a hallmark of quasi-physical solutions.
\begin{figure}[H]\centering
    \centering
    \includegraphics[width=0.48\textwidth]{figure/tent_quasiphysical_alternaive1_hist_p_0.6.png}
    \hspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tent_quasiphysical_alternaive2_hist_p_0.6.png}
    \caption{Empirical distribution function of other quasi-physical solutions.
    The solution shown in the left plot starts from an initial condition
    whose binary decimal digits are independent and have probability 0.6 of being 0.
    The solution shown in the right plot starts from an initial condition
    whose binary decimal digits have probability of 0.6 of being 1 only following two
    consecutive 1's; otherwise it is 0 or 1 with equal probability of 0.5.}
    \label{fig:tent_alternative}
\end{figure}

We have only seem quasi-physical solutions for the tent map.
It is difficult to analytically construct quasi-physical solutions
to the Lorenz equation and other, more complex, chaotic governing
equations.  Neverthless, the similarity between the tent map and the
Lorenz map, shown in Figure \ref{fig:lorenz_map}, suggests that
quasi-physical solutions may exist for the Lorenz map, and by
extension, the Lorenz equation.  It is then natural to conjecture
that such aperiodic, non-physical, quasi-physical solutions
exist in general for chaotic dynamical systems.

These quasi-physical solutions raise doubt to the usefulness of shadowing
in some applications.  For example, even if a numerical
solution is shadowed by a solution to the true governing physics,
is the shadowing solution physical or quasi-physical?  Such doubt may first
seem paranoid; since almost all solutions are physical, it appears
that one has to be extremely unlucky for the shadowing solution to
be nonphysical. It seems reasonable to argue that because the set of
all nonphysical solutions is measure-zero, the probability of finding
a nonphysical solution through shadowing is zero percent.  Such argument,
as we show in the next section, is wrong.  The probability of
finding a nonphysical solution through shadowing can be, instead of
zero percent, one hundred percent.

\section{Are shadowing solutions physical?}

\subsection{Example of a shadowing solution for the tent map}
To illustrate the concept of shadowing, consider the tent map, defined
by Equation (\ref{tentmap}), and a scaled version of the map, defined by
\begin{equation} \label{tent_scaled}
    \hat\varphi_s(x) := \begin{cases}
    2x \quad & x < 1+s \\
    4(1+s) - 2x \quad & x \ge 1+s
    \end{cases}
\end{equation}
where $s<<1$.  Note that tiny differences in $s$ can lead to drastic
different solutions, if the solutions start from the same initial condition.
As demonstrated in Figure \ref{fig:tent_ivp}, this sensitivity to small
perturbations reflects the chaotic nature of the governing equation.

\begin{figure}[H]   \centering
    \includegraphics[width=0.48\textwidth]{figure/scaled_tent_map.png}
    \hspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tent_ivp.png}
    \caption{Sensitivity to small perturbation in the governing
    equation.
    The y-axis shows the absolute value of the difference between
    two solutions, one satisfying Equation (\ref{tent_scaled}),
    one for $s=0$ and the other for $s=10^{-5}$.
    The x-axis shows the iteration number.  The initial condition
    is at $x_0=\pi/2$.
    }
    \label{fig:tent_ivp}
\end{figure}

We can avoid this sensitivity to the governing equation
by a coordinated perturbation to the initial condition.
Consider two solutions satisfying Equation (\ref{tent_scaled})
for different values of $s$.  Instead of starting from the same
initial condition, these two solutions start from $x_0 (1 + s)$ with
the same $x_0$ but their respective values of $s$.  It can be shown that
the solution of these two equations would be
$x_i (1+s), i=0,1,\ldots$ with their respective values of $s$,
where $x_{i+1}=\varphi(x_i)$ satisfies the original
tent map (\ref{tentmap}).  When the values of $s$ are similar between
the two solutions, this pair of solutions stay uniformly close to each
other forever.  Such pair of solutions, satisfying slightly different
governing equations, yet keep close to each other over a long time,
are known as shadowing solutions.

For every solution $\hat{x}_i,i=0,\ldots$ satisfying Equation
(\ref{tent_scaled}), there is a shadowing solution satisfying Equations
(\ref{tentmap}): $x_i:= \hat{x}_i/(1+s)$.  The map,
\begin{equation} \label{conjugate_scaled}
    \hat{h}_s(x):=x/(1+s)
\end{equation}
is known as the conjugate map, because it satisfies
\begin{equation} \label{conjugate_def}
    \varphi(\hat{h}_s(x)) = \hat{h}_s(\hat\varphi_s(x))\;,\quad \forall x
\end{equation}
or equivalently, $\varphi\circ \hat{h}_s \equiv \hat{h}_s\circ\hat\varphi_s$.
Such conjugate maps can generally help us construct shadowing solutions.
If a conjugate map $h$ exists for two maps $\varphi$ and $\psi$, i.e.,
$\varphi\circ h \equiv h\circ\psi$, then for every solution satisfying
$\psi$, $h$ maps it to a shadowing solution satisfying $\varphi$.

Is the shadowing solution a physical solution?  In this example,
the answer is almost surely positive.  We can demonstrate that
almost any solution of the scaled tent map (\ref{tent_scaled}) is uniformly
distributed in $[0, 2(1+s)]$ -- we can repeat our analysis
in Section \ref{sec:tent} but represent our initial condition as
$x_0= (1+s) \sum x_{0,j} / 2^j$.  Its shadowing solution,
satisfying the original tent map (\ref{tentmap}), can be obtained through the
conjugate map (\ref{conjugate_scaled}).  Thus the distribution
of the shadowing solution can be obtained by mapping a uniform
distribution in $[0,2(1+s)]$ through the conjugate map.
This leads to a uniformly distribution in $[0,2]$,
precisely the distribution of a physical solution of the tent map.

This simple example is useful in illustrating the concept of shadowing
and the utility of the conjugate map.  This example is nevertheless special
in that the shadowing solution is almost always physical.  Our next
example introduces a tilted version of the tent map.  Although
the tilting perturbation to the tent map seems as simple as the
scaling perturbation, the shadowing solution, as we will see,
is almost always a quasi-physical solution.

\subsection{An example of quasi-physical shadowing solution}

The tilted tent map is defined as
\begin{equation} \label{tent_tilted}
    \hat\varphi_s(x) := \begin{cases}
    \frac2{1+s} x \quad & x < 1+s \\
    \frac2{1-s}(2-s) \quad & x \ge 1+s
    \end{cases}
\end{equation}
When $s=0$, this map is identical to the tent map (\ref{tentmap}).
Let $\tilde{x}_{i+1}=\hat\varphi_s(\tilde{x}_i), i=0,1,\ldots$ be a solution.
As we explained in the previous subsection,
this shadowing solution can be found through a conjugate map $\tilde{h}_s$.
If $\varphi\circ \tilde{h}_s \equiv \tilde{h}_s\circ\tilde\varphi_s$,
then $x_i=\tilde{h}_s(\tilde{x}_i), i=0,1,\ldots$ is a shadowing solution
that satisfies $x_{i+1} = \varphi(x_i)$.

The conjugate map, although more complex than the one
in the last subsection, has the following closed form:
\begin{equation} \label{conjugate_tilted}
    \tilde{h}_s(x) = \sum_{j=0}^{\infty} \frac{\xi_{s,j}(x)}{2^j}\;,
\end{equation}
where
\begin{equation} \label{conjugate_tilted_helper}
    \xi_{s,j}(x) := \begin{cases}
    0 & j=0, x<1+s \\
    1 & j=0, x\ge 1+s \\
    \xi_{s,j-1}(x) & j>0, \tilde\varphi_s^j(x) < 1+s \\
    1 - \xi_{s,j-1}(x) & j>0, \tilde\varphi_s^j(x) \ge 1+s
    \end{cases}
\end{equation}
To see why this map satisfies the definition of the conjugate map --
$\tilde{h}_s(\tilde\varphi_s(x)) = \varphi(\tilde{h}_s(x))$ for all $x$ --
we need to analyze two cases: $x<1+s$ and $x\ge 1+s$.

When $x<1+s$, $\xi_{s,0}(x)=0$; thus
$\tilde{h}_s(x) = \sum_{j=1}^{\infty} \frac{\xi_{s,j}(x)}{2^j}$.
Inside this infinite series, $\xi_{s,1}(x)=\xi_{s,0}(x)=0$ if $\tilde\varphi_s(x)<1+s$,
or $\xi_{s,1}(x)=1-\xi_{s,0}(x)=1$ if $\tilde\varphi_s(x)\ge 1+s$.
Thus, $\xi_{s,1}(x) = \xi_{s,0}(\varphi_s(x))$ according to the definition of $\xi_{s,0}$.
Using this as the base case, one can inductively verify that
$\xi_{s,j+1}(x) = \xi_{s,j}(\tilde\varphi_s(x))$
for all $j>0$, using just the definition of $\xi_{s,j}$ and $\xi_{s,j+1}$.
Therefore,
$\tilde{h}_s(x) = \frac12 \sum_{j=0}^{\infty} \frac{\xi_{s,j}(\tilde\varphi_s(x))}{2^j}
=\frac12 \tilde{h}_s(\tilde\varphi_s(x))$.
On the other hand, because
$\tilde{h}_s(x) = \sum_{j=1}^{\infty} \frac{\xi_{s,j}(x)}{2^j}\le1$,
$\varphi(\tilde{h}_s(x)) = 2\tilde{h}_s(x)$ according
to the definition of $\varphi$.  Thus,
$\varphi(\tilde{h}_s(x)) = \tilde{h}_s(\tilde\varphi_s(x))$ holds when $x<1+s$.

When $x\ge1+s$, $\xi_{s,0}(x)=1$; thus
$\tilde{h}_s(x) = 1+ \sum_{j=1}^{\infty} \frac{\xi_{s,j}(x)}{2^j}
                = 2- \sum_{j=1}^{\infty} \frac{1-\xi_{s,j}(x)}{2^j}$.
Inside this infinite series, $1-\xi_{s,1}(x)=1-\xi_{s,0}(x)=0$ if $\tilde\varphi_s(x)<1+s$,
or $1-\xi_{s,1}(x)=\xi_{s,0}(x)=1$ if $\tilde\varphi_s(x)\ge 1+s$.
Thus, $1-\xi_{s,1}(x) = \xi_{s,0}(\varphi_s(x))$ according to the definition of $\xi_{s,0}$.
Using this as the base case, one can inductively verify that
$1-\xi_{s,j+1}(x) = \xi_{s,j}(\tilde\varphi_s(x))$
for all $j>0$, using just the definition of $\xi_{s,j}$ and $\xi_{s,j+1}$.
Therefore,
$\tilde{h}_s(x) = 2 - \frac12 \sum_{j=0}^{\infty} \frac{\xi_{s,j}(\tilde\varphi_s(x))}{2^j}
=2 - \frac12 \tilde{h}_s(\tilde\varphi_s(x))$.
On the other hand, because
$\tilde{h}_s(x) = 1+\sum_{j=1}^{\infty} \frac{\xi_{s,j}(x)}{2^j}\ge 1$,
$\varphi(\tilde{h}_s(x)) = 2(1-\tilde{h}_s(x))$ according
to the definition of $\varphi$.  Thus,
$\varphi(\tilde{h}_s(x)) = \tilde{h}_s(\tilde\varphi_s(x))$ also holds when $x\ge 1+s$.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.48\textwidth]{figure/tilted_tent_map.png}
    \hspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tilted_conjugate_map.png}
    \\ \vspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tilted_conjugate_density_0.1.png}
    \hspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tilted_conjugate_density_0.5.png}
    \caption{Conjugate map $\tilde{h}_s$ between the tent map (\ref{tentmap})
    and the tilted tent map (\ref{tent_tilted}), evaluated using
    Equation (\ref{conjugate_tilted}-\ref{conjugate_tilted_helper}).
    Starting from the diagonal line, $s=0, 0.1, 0.2, 0.3, 0.4,$ and 0.5.}
    \label{fig:tilted_tent_conjugate}
\end{figure}

Figure \ref{fig:tilted_tent_conjugate} shows how fractal the conjugate map is.
This fractal conjugate map transforms a uniform distribution in $[0,2]$ into
a fractal distribution similar to the ones plotted in Section \ref{sec:quasiphysical}.
In fact, we will show that the fractal distribution obtained by mapping a uniform
distribution under $\tilde{h}_s$ is exactly the family of distributions
shown in Figure \ref{fig:tent_quasiphysical}.  We will also show that
a physical solution of the tilted tent map (\ref{tent_tilted}) is uniformly
distributed in $[0,2]$.  Thus, for almost any physical solution of the
tilted tent map, a shadowing solution of the original tent map, obtained
through the conjugate map $\tilde{h}_s$, has a fractal distribution.
Such shadowing solution is therefore a quasi-physical solution of
the tent map.

We first show that a physical solution of the tilted tent map 
(\ref{tent_tilted}) is uniformly distributed in $[0,2]$, for
any $0\le s<1$.
According to ergodic theory, we only need to show that the uniform
distribution is stationary under the tilted tent map.
If $x$ is uniformly distributed in $[0,2]$ with a density of 0.5,
the left and right sides of the map, from $x=0$ to $x=1+s$ and
from $x=1+s$ to 2 respectively, both has constant derivatives.
The resulting density of $\tilde\varphi_s(x)$ is
therefore uniform in $[0,2]$.

We then show that the conjugate map $\tilde{h}_s$ maps a uniform
distribution to a fractal.  Here we use the closed form
(\ref{conjugate_tilted}-\ref{conjugate_tilted_helper}).
If $x$ is a random number drawn uniformly in $[0,2]$,
it has $\frac{1+s}2$ probability of being less than $1+s$.
Thus, $\xi_{s,0}(x)=0$ with probability $\frac{1+s}2$.
This means, according to Equation (\ref{conjugate_tilted}),
$\tilde{h}_s(x)<1$ with probability $\frac{1+s}2$ --
$\tilde{h}_s(x)$ is more likely to lie in the left half
of the domain $[0,2]$.
Furthermore, for $j\ge 0$, $\xi_{s,j+1}(x)=\xi_{s,j}(x)$
with probability $\frac{1+s}2$.  This probability has direct
implication, again by Equation (\ref{conjugate_tilted}),
to the binary decimal representation of $\tilde{h}_s(x)$ --
each binary digit repeat the previous one with probability
$\frac{1+s}2$.  Now we understand that for a uniformly random
$x$, $\tilde{h}_s(x)$ is exactly the kind of initial condition
we used to construct the quasi-physical solution
in Section \ref{sec:quasiphysical}, with $p=\frac{1+s}2$.

This analysis implies that the shadowing solution is
almost surely nonphysical.  Let $x_{s,i},i=0,1,\ldots$ be a solution
to the tilted tent map with an initial condition randomly chosen
in $[0,1]$.  Then with a hundred percent probability
it is a physical solution that uniformly visits the domain $[0,2]$.
Its shadowing solution $\tilde{h}_s(x_{s,i}),i=0,1,\ldots$, however,
is a quasi-physical solution, also with a hundred percent probability.
It explores $[0,2]$ at a nonuniform frequency with a fractal density function.

\subsection{Are general shadowing solutions physical?}

Here we completed our example.  For almost every solution to a perturbed governing
equation, its shadowing solution, satisfying the real governing equation, will
be nonphysical.  The shadowing theorems do guarantee the existence of shadowing
solutions.  But are we sure it is the kind of solution we want?

\begin{figure}[H]
    \centering
    \includegraphics[width=0.48\textwidth]{figure/tent_tilted_shadow_density_0.01.png}
    \hspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tent_tilted_shadow_density_0.05.png}
    \\ \vspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tent_tilted_shadow_density_0.4.png}
    \hspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tent_tilted_shadow_density_0.7.png}
    \caption{Caption}
    \label{fig:tent_tilted_shadow}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.48\textwidth]{figure/squashed_tent_map.png}
    \caption{Caption}
    \label{fig:tent_squashed}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.48\textwidth]{figure/tent_squashed_shadow_density_0.01.png}
    \hspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tent_squashed_physical_density_0.01.png}
    \\ \vspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tent_squashed_shadow_density_0.2.png}
    \hspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tent_squashed_physical_density_0.2.png}
    \caption{Caption}
    \label{fig:tent_squashed_shadow}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.48\textwidth]{figure/pinched_tent_map.png}
    \caption{Caption}
    \label{fig:tent_pinched}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.48\textwidth]{figure/tent_pinched_shadow_density_0.05.png}
    \hspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tent_pinched_physical_density_0.05.png}
    \\ \vspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tent_pinched_shadow_density_0.2.png}
    \hspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tent_pinched_physical_density_0.2.png}
    \caption{Caption}
    \label{fig:tent_pinched_shadow}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.48\textwidth]{figure/wave_tent_map.png}
    \caption{Caption}
    \label{fig:tent_wave}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.48\textwidth]{figure/tent_wave_shadow_density_0.05.png}
    \hspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tent_wave_physical_density_0.05.png}
    \\ \vspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tent_wave_shadow_density_0.2.png}
    \hspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tent_wave_physical_density_0.2.png}
    \caption{Caption}
    \label{fig:tent_wave_shadow}
\end{figure}

\subsection{What about the Lorenz equation?}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.32\textwidth]{figure/lorenz_zmax_sigma.png}
    \hspace{0.005\textwidth}
    \includegraphics[width=0.32\textwidth]{figure/lorenz_zmax_rho.png}
    \hspace{0.005\textwidth}
    \includegraphics[width=0.32\textwidth]{figure/lorenz_zmax_beta.png}
    \caption{The left, center, and right plots show the effect of the
    parameters $\sigma, \rho$, and $\beta$ on the Lorenz map, respectively.
    In the left plot, the blue, orange, and green lines represent
    $\sigma=10,11$, and $12$, respectively, while $\rho$ and $\beta$ stay
    at 28 and $8/3$.
    In the center plot, the blue, orange, and green lines represent
    $\rho=28,29$ and $30$, respectively, while $\sigma$ and $\beta$ stays at 10
    and $8/3$.
    In the right plot, the blue, orange, and green lines represent
    $\sigma=8/3,3$ and $10/3$ respectively, while $\sigma$ and $\rho$ stays at 10
    and $28$.}
    \label{fig:lorenz_params}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.32\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_15_28_2.67.json.png}
    \hspace{0.005\textwidth}
    \includegraphics[width=0.32\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_10_30_2.67.json.png}
    \hspace{0.005\textwidth}
    \includegraphics[width=0.32\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_10_28_3.33.json.png}
    \caption{
    }
    \label{fig:lorenz_params_scaled}
\end{figure}
----------------------------

\section{Discussions}
Shown through a rigorous counter-example, we conclude that shadowing can lead
to nonphysical solutions.  The conclusion has instant implication for
sensitivity analysis methods based on the concept shadowing.
If the goal is to compute derivative of ensemble averages, where the ensemble
is distributed according to the physical measure, i.e., the SRB measure,
then shadowing-based methods can give wrong results.  This has been observed
before and appropriately attributed as ``ergodicity breaking error''.
This paper merely gave a rigorous analytical example of it.

Even more troubling is the implication for any numerical simulation of chaotic
governing equations.  We now understand that a numerical solution, satisfying
a governing equation perturbed due to numerics, is not expected to shadow a
physical solution of the real governing physics.  What is, if any, the
relation of a numerical solution, such as DNS of turbulent flows, to
the true physics?
[Nisha: Use tent map as example for point 1. For $10^{-18}$ perturbation (trunction numerically), we get completely different statistics. How can we trust numerical solutions]

[Nisha: Eq. 18 and Eq. 19 - shadowing sensitivities are incorrect.]

\begin{figure}
    \centering
    \includegraphics[width=0.35\textwidth,height=2in]{figure/mapn0s0.png}
    \includegraphics[width=0.35\textwidth,height=2in]{figure/mapn5s0.1.png}
    \caption{Two chaotic maps (in red) and their associated stationary  distributions (in purple). The map on the right is only an \emph{infinitesimal} perturbation of that on the left; for more information see \cite{desmos}.}
    \label{fig:map}
\end{figure}
[Nisha: shadowing perturabtions predict tiny effects. Real effect is a lot more dramatic. Desmos plot.]
% give example of numerics that destroys physicalness
\bibliographystyle{spmpsci}
\bibliography{refs.bib}
\end{document}

### 2020-10-13 code/lorenz_density/histogram.h

@@ -1,0 +1,130 @@
+// density.h
+#pragma once
+#include<cinttypes>
+                                                                                 
+struct ValueWeight {
+    double x, y, weight;
+};
+
+__device__ __forceinline__                                                       
+void pIncrement(                                                                 
+        float * tmpCounts,                                                    
+        double x0, double y0,
+        double dx, double dy,
+        uint32_t nx, uint32_t ny,                          
+        double x, double y, double w
+) {                                                                              
+    if (x < x0 or y < y0)
+        return;
+
+    uint32_t i = uint32_t((x - x0) / dx);                                            
+    uint32_t j = uint32_t((y - y0) / dy);                                         
+                                                                             
+    if (i < nx and j < ny) {                                                 
+        uint32_t ind = i * ny + j;                                               
+        atomicAdd(tmpCounts + ind, w);                                           
+    }                                                                            
+}                                                                                
+                                                                                 
+template<bool accumulate, class Map, class ObjFunc, uint32_t mapDim>                                             
+__global__                                                                       
+void pRun(                                                                       
+        Map map, ObjFunc obj, uint32_t nIters,                                                
+        uint32_t nPoints, double (*points)[mapDim],                                   
+        float * tmpCounts,                                                    
+        double x0, double y0,
+        double dx, double dy,
+        uint32_t nx, uint32_t ny                           
+) {                                                                              
+    uint32_t ind = blockIdx.x * blockDim.x + threadIdx.x;                        
+                                                                                 
+    if (ind < nPoints) {                                                         
+        for (uint32_t iIter = 0; iIter < nIters; ++iIter) {                      
+            map.map(points[ind]);
+            ValueWeight valWgt = obj.eval(points[ind]);
+
+            if (accumulate) {                                                    
+                pIncrement(tmpCounts, x0, y0, dx, dy, nx, ny,
+                           valWgt.x, valWgt.y, valWgt.weight);                      
+            }                                                                    
+        }                                                                        
+    }                                                                            
+}                                                                                
+                                                                                 
+template<uint32_t mapDim>
+struct Counter {                                                                 
+  public:                                                                        
+    const uint32_t nx, ny;                                                       
+    const double x0, y0, dx, dy;                                                         
+    double* counts;                                                              
+                                                                                 
+  private:                                                                       
+    float* tmpCounts;                                                         
+    float* cpuCounts;                                                         
[…]

### 2020-10-13 code/lorenz_density/lorenz.h

@@ -1,0 +1,31 @@
+struct Lorenz {
+    const double sigma, rho, beta, dt;
+
+    void init(double xyz[3]) {
+        for (int i = 0; i < 3; ++i)
+            xyz[i] = (rand() / double(RAND_MAX) + rand()) / double(RAND_MAX) / 50;
+        xyz[2] += 28;
+    }
+
+    __device__ __forceinline__
+    void ddt(double ddtxyz[3], const double xyz[3]) {
+        ddtxyz[0] = sigma * (xyz[1] - xyz[0]);
+        ddtxyz[1] = xyz[0] * (rho - xyz[2]) - xyz[1];
+        ddtxyz[2] = xyz[0] * xyz[1] - beta * xyz[2];
+    }
+
+    __device__ __forceinline__
+    void map(double xyz[3]) {
+        double ddtxyz[3];
+        ddt(ddtxyz, xyz);
+        double xyzMid[3];
+
+        for (uint8_t i = 0; i < 3; ++i)
+            xyzMid[i] = xyz[i] + 0.5 * dt * ddtxyz[i];
+
+        ddt(ddtxyz, xyzMid);
+
+        for (uint8_t i = 0; i < 3; ++i)
+            xyz[i] += dt * ddtxyz[i];
+    }
+};

### 2020-10-13 code/lorenz_density/lorenz_density.cpp

@@ -1,0 +1,52 @@
+// density/bakers_density.cpp
+#include<cstdio>
+#include<cassert>
+#include<algorithm>
+
+#include"histogram.h"
+#include"lorenz.h"
+
+struct ReturnXZ {
+    __device__ __forceinline__
+    static ValueWeight eval(double xyz[3]) {
+        return ValueWeight{xyz[0], xyz[2], 1.0};
+    }
+};
+
+int main() {
+    uint32_t iDevice;
+    assert(fread(&iDevice, sizeof(uint32_t), 1, stdin) == 1);
+    cudaSetDevice(iDevice);
+    fprintf(stderr, "Set To Device %u\n", iDevice);
+
+    uint32_t randSeed;
+    assert(fread(&randSeed, sizeof(uint32_t), 1, stdin) == 1);
+    srand(randSeed);
+
+    const int nx = 2048, nz = 2048;
+    Counter<3> counter(nx, nz, -20., 0., 40./nx, 50./nz);
+
+    float parameters[5];
+    assert(fread(parameters, sizeof(float), 5, stdin) == 5);
+    Lorenz lorenz{parameters[0], parameters[1], parameters[2], parameters[3]};
+    ReturnXZ obj;
+    fprintf(stderr, "Ready with parameters (%f %f %f) with dt %f for %f steps\n",
+            parameters[0], parameters[1], parameters[2], parameters[3], parameters[4]);
+
+    uint32_t nIters;
+    assert(fread(&nIters, sizeof(uint32_t), 1, stdin) == 1);
+
+    for (uint32_t iIter = 0; iIter < nIters; iIter++) {
+        if (iIter % 128 == 0) {
+            fprintf(stderr, "%u/%u iterations\n", iIter, nIters);
+        }
+
+        counter.init(lorenz, 80000);
+        counter.run(lorenz, obj, static_cast<int>(parameters[4]), false);
+
+        counter.run(lorenz, obj, 1, true);
+    }
+
+    fwrite(counter.counts, sizeof(double), nx * nx, stdout);
+    return 0;
+}

### 2020-10-13 code/lorenz_density/run_binary.py

@@ -1,0 +1,33 @@
+# density/run_binary.py
+import os
+import sys
+import json
+import subprocess
+from numpy import *
+
+config = json.load(open(sys.argv[1]))
+config = config[sys.argv[2]]
+binary = config['binary']
+params = list(config['parameters'])
+
+procs = []
+nIter = 256
+gpus = subprocess.check_output(
+    ['nvidia-smi', '--query-gpu=name', '--format=csv,noheader'])
+numGpus = len(gpus.strip().splitlines())
+for i in range(numGpus):
+    p = subprocess.Popen(binary, stdin=subprocess.PIPE, stdout=subprocess.PIPE)
+    p.stdin.write(array([i, random.randint(1 << 30)], uint32).tobytes())
+    p.stdin.write(array(params, float32).tobytes())
+    p.stdin.write(array(nIter, uint32).tobytes())
+    p.stdin.flush()
+    procs.append(p)
+
+n = 2048
+out = [frombuffer(p.stdout.read(n * n * 8), double) for p in procs]
+histogram = sum(out, 0).reshape([n, n])
+myname = subprocess.check_output('hostname').decode().strip()
+params = ('_'.join(['{}'] * len(params))).format(*tuple(params))
+fname = '{}_{}_{}_{}_{}.npy'.format(os.path.basename(binary),
+        params, myname, nIter, random.rand())
+save(fname, histogram)

### 2020-10-13 code/lorenz_map/fit.py

@@ -1,0 +1,131 @@
+import os
+import sys
+import json
+from pylab import *
+from numpy import *
+from scipy.optimize import curve_fit
+
+z = load(sys.argv[1])[:,2]
+imax = z[1:].argmax()
+def fit0(dz, p):
+    return pow(dz*1000, p)
+
+zSep, zMax = z[imax:imax+2]
+zMin = z.min()
+print(zSep, zMax, file=sys.stderr)
+
+sel = (z[1:] > percentile(z, 90)).nonzero()[0]
+zSep, pwr = curve_fit(
+        lambda z, zSep, p : zMax - fit0(abs(z - zSep), p),
+        z[sel], z[sel+1], (zSep, 1./3))[0]
+print(zSep, zMax, pwr, file=sys.stderr)
+
+iL = (z[:-1] <= zSep).nonzero()[0]
+iR = (z[:-1] >= zSep).nonzero()[0]
+
+figure(figsize=(12,12))
+plot(z[iL], z[iL+1], '.', ms=1)
+plot(z[iR], z[iR+1], '.', ms=1)
+
+figure(figsize=(12,12))
+loglog(zSep - z[iL], zMax - z[iL+1], '.', ms=1)
+loglog(z[iR] - zSep, zMax - z[iR+1], '.', ms=1)
+
+figure(figsize=(12,12))
+plot(z[iL], zMax - z[iL+1] - fit0(zSep - z[iL], pwr), '.k', ms=1)
+plot(z[iR], zMax - z[iR+1] - fit0(z[iR] - zSep, pwr), '.k', ms=1)
+gca().set_xticks(linspace(30,45,4))
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(30)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(30)
+
+pL, pR = (0, 0.01091624, 0.44126003, -0.30532497), \
+         (0, 0.01260884, 0.4280573, -0.28548578)
+print(pL, pR, file=sys.stderr)
+
+selL = iL[(z[iL+1] < percentile(z, 20)).nonzero()[0]]
+selR = iR[(z[iR+1] < percentile(z, 20)).nonzero()[0]]
+pL = curve_fit(
+        lambda dz, *p : fit0(dz, pwr) + polyval(p, dz),
+        zSep - z[selL], zMax - z[selL + 1], pL)[0]
+pR = curve_fit(
[…]

### 2020-10-13 code/lorenz_map/interpolate.py

@@ -1,0 +1,47 @@
+import os
+import sys
+import json
+from pylab import *
+from numpy import *
+from scipy.optimize import curve_fit
+
+params0 = json.load(open(sys.argv[1]))
+params1 = json.load(open(sys.argv[2]))
+
+def fit0(dz, pwr):
+    return pow(dz*1000, pwr)
+
+def fit1(dz, pwr, p):
+    return fit0(dz, pwr) + polyval(p, dz)
+
+def fit2(dz, pwr, p, a, b):
+    return fit1(dz, pwr, p) + polyval(a, dz) / polyval(b, dz)
+
+dz = hstack([linspace(0, 0.1, 501), linspace(0.1, 1, 901)[1:], linspace(1, 10, 901)[1:]])
+eta = linspace(0, 1, 1001)
+z0, z1 = [], []
+for i in range(eta.size):
+    pwr = params0['pwr'] * (1 - eta[i]) + params1['pwr'] * eta[i]
+    zMax = params0['zMax'] * (1 - eta[i]) + params1['zMax'] * eta[i]
+    zSep = params0['zSep'] * (1 - eta[i]) + params1['zSep'] * eta[i]
+    pL = array(params0['pL']) * (1 - eta[i]) + array(params1['pL']) * eta[i]
+    pR = array(params0['pR']) * (1 - eta[i]) + array(params1['pR']) * eta[i]
+    aL = array(params0['aL']) * (1 - eta[i]) + array(params1['aL']) * eta[i]
+    aR = array(params0['aR']) * (1 - eta[i]) + array(params1['aR']) * eta[i]
+    bL = array(params0['bL']) * (1 - eta[i]) + array(params1['bL']) * eta[i]
+    bR = array(params0['bR']) * (1 - eta[i]) + array(params1['bR']) * eta[i]
+    z0.append(hstack([zSep - dz[::-1], zSep + dz[1:]]))
+    z1.append(hstack([zMax - fit2(dz[::-1], pwr, pL, aL, bL), zMax - fit2(dz[1:], pwr, pR, aR, bR)]))
+
+z0, z1 = array([z0, z1])
+figure(figsize=(24,24))
+contourf(z0, eta[:,newaxis] * ones_like(z0), z1, 1000)
+cbar = colorbar()
+cbar.ax.tick_params(labelsize=50)
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(50)
[…]

### 2020-10-13 code/lorenz_map/lorenz_map.py

@@ -1,0 +1,50 @@
+import os
+import sys
+import numba
+from numpy import *
+
+@numba.jit(nopython=True)
+def ddt(xyz):
+    x, y, z = xyz
+    dxdt = sigma * (y - x)
+    dydt = x * (rho - z) - y
+    dzdt = x * y - beta * z
+    return array([dxdt, dydt, dzdt])
+
+@numba.jit(nopython=True)
+def step(xyz0):
+    ddt0 = ddt(xyz0)
+    dt = 0.001
+    xyz1 = xyz0 + 0.5 * dt * ddt0
+    ddt1 = ddt(xyz1)
+    xyz1 = xyz0 + dt * ddt1
+    ddt1 = ddt(xyz1)
+    if ddt0[2] >= 0 and ddt1[2] < 0:
+        if xyz0[2] > xyz1[2]:
+            zmax = xyz0.copy()
+        else:
+            zmax = xyz1.copy()
+    else:
+        zmax = zeros(3)
+    xyz0[:] = xyz1
+    return zmax
+
+@numba.jit(nopython=True)
+def evolve(xyz, nsteps):
+    zmax = []
+    for i in range(10000):
+        step(xyz)
+    for i in range(nsteps):
+        z = step(xyz)
+        if z[2] > 0:
+            zmax.append(z)
+    return xyz, zmax
+
+sigma = 10
+rho = 28
+beta = 8./3
+nbatches = int(sys.argv[1])
+xyz = array([1, 1, 28.])
+for i in range(nbatches):
+    xyz, zmax = evolve(xyz, 1000000000)
+    save('lorenz_zmax_{}_{}_{}_{:05d}.npy'.format(sigma, rho, beta, i), zmax)

### 2020-10-13 code/lorenz_map/shadow.py

@@ -1,0 +1,111 @@
+import os
+import sys
+import json
+import numba
+from pylab import *
+from numpy import *
+from scipy.optimize import curve_fit
+
+@numba.jit(nopython=True)
+def polyval(p, x):
+    y, dy = 0, 0
+    for pi in p:
+        dy *= x
+        dy += y
+        y *= x
+        y += pi
+    return y, dy
+
+@numba.jit(nopython=True)
+def fit0(dz, pwr):
+    return pow(dz * 1000, pwr), 1000 * pwr * pow(dz * 1000, pwr - 1)
+
+@numba.jit(nopython=True)
+def fit1(dz, pwr, p):
+    f0, f0grad = fit0(dz, pwr)
+    pz, pzgrad = polyval(p, dz)
+    return f0 + pz, f0grad + pzgrad
+
+@numba.jit(nopython=True)
+def fit2(dz, pwr, p, a, b):
+    f1, f1grad = fit1(dz, pwr, p)
+    num, numgrad = polyval(a, dz)
+    den, dengrad = polyval(b, dz)
+    return f1 + num / den, f1grad + numgrad / den - num * dengrad / den / den
+
+@numba.jit(nopython=True)
+def find2(dzNext, dzMin, dzMax, pwr, p, a, b):
+    if dzNext <= dzMin or dzNext <= 0:
+        return 0
+    dz0 = 10.0
+    for idz0 in range(4):
+        dz0 /= 10
+        dz = dz0
+        #print()
+        for i in range(10):
+            if dz <= dzMax:
+                f2, f2grad = fit2(dz, pwr, p, a, b)
+            else:
+                f2Max, f2grad = fit2(dzMax, pwr, p, a, b)
+                f2 = f2Max + f2grad * (dz - dzMax)
+            #print(dzNext, f2, f2grad, dz, exp((dzNext - f2) / f2grad / dz))
+            if abs(dzNext - f2) < 1E-8 or abs(dzNext - f2) < 1E-2 and dz < 1E-8:
+                return dz
+            dz *= exp((dzNext - f2) / f2grad / dz)
+    print(dzNext, dzMin, dzMax, pwr, p, a, b)
+    return inf
+
+@numba.jit(nopython=True)
+def shadow(zTraj, zSep0, z0, zMin, zMax, zSep, pwr, pL, pR, aL, aR, bL, bR):
+    zShadow = zeros_like(zTraj)
+    z = z0
+    dzMinL = fit2(0, pwr, pL, aL, bL)[0]
[…]

### 2020-10-13 code/lorenz_periodic/lorenz_periodic.py

@@ -1,0 +1,49 @@
+import os
+import sys
+import numba
+from numpy import *
+
+@numba.jit(nopython=True)
+def ddt(xyz):
+    x, y, z = xyz
+    dxdt = sigma * (y - x)
+    dydt = x * (rho - z) - y
+    dzdt = x * y - beta * z
+    return array([dxdt, dydt, dzdt])
+
+@numba.jit(nopython=True)
+def step(xyz0):
+    ddt0 = ddt(xyz0)
+    dt = 0.001
+    xyz1 = xyz0 + 0.5 * dt * ddt0
+    ddt1 = ddt(xyz1)
+    xyz0 += dt * ddt1
+
+@numba.jit(nopython=True)
+def evolve(xyz0, steps):
+    history = []
+    xyz = xyz0
+    for i in range(steps):
+        history.append(xyz.copy())
+        step(xyz)
+    history.append(xyz)
+    return history
+
+sigma = 10
+rho = 28
+beta = 8./3
+
+xyz0, steps = array([-13.92609234,  -7.54780951,  39.82352798]), 1558
+xyz = array(evolve(xyz0, steps))
+save('lorenz_periodic_1.npy', xyz, '-')
+plot(xyz[:,0], xyz[:,2])
+
+xyz0, steps = array([12.99988069,   7.76732599,  37.74858325]), 3083
+xyz = array(evolve(xyz0, steps))
+save('lorenz_periodic_2.npy', xyz)
+plot(xyz[:,0], xyz[:,2], '--')
+
+xyz0, steps = array([-13.14298792,  -7.7604085 ,  38.04801711]), 4566
+xyz = array(evolve(xyz0, steps))
+save('lorenz_periodic_3.npy', xyz)
+plot(xyz[:,0], xyz[:,2], ':')

### 2020-10-13 code/lorenz_shadow/shadow_step1.py

@@ -1,0 +1,55 @@
+from pylab import *
+from numpy import *
+
+zmax0 = load('../../data/lorenz_zmax_10_28_2.6666666666666665.npy')
+#zmax1 = load('../../data/lorenz_zmax_10_30_2.6666666666666665.npy')
+#zmax1 = load('../../data/lorenz_zmax_12_28_2.6666666666666665.npy')
+#zmax1 = load('../../data/lorenz_zmax_10_28_3.3333333333333335.npy')
+
+zgrid = linspace(28, 53, 2401)
+
+#density0 = histogram(zmax0[:,2], zgrid[::2])[0] / zmax0.shape[0] * (zgrid.size //2) / (zgrid[-1] - zgrid[0])
+#fill_between(zgrid[1::2], density0)
+#
+#figure()
+#density1 = histogram(zmax1[:,2], zgrid[::2])[0] / zmax0.shape[0] * (zgrid.size //2) / (zgrid[-1] - zgrid[0])
+#fill_between(zgrid[1::2], density1)
+
+def shadow_density(zmax0, zmax1, fname):
+    isCross0 = (zmax0[:-1,0] > 0) ^ (zmax0[1:,0] > 0)
+    isCross1 = (zmax1[:-1,0] > 0) ^ (zmax1[1:,0] > 0)
+    
+    iL1 = (~isCross1).nonzero()[0][zmax1[1:,2][~isCross1].argsort()]
+    iR1 = (isCross1).nonzero()[0][zmax1[1:,2][isCross1].argsort()]
+    zL1 = zmax1[iL1+1,2]
+    zR1 = zmax1[iR1+1,2]
+    zL1[-1] = inf
+    zR1[-1] = inf
+    
+    z = zmax0[-1,2]
+    i1seq = zeros(zmax0.shape[0]-1, int)
+    for i0 in range(1, zmax0.shape[0]):
+        if isCross0[-i0]:
+            i1 = iR1[searchsorted(zR1, z)]
+        else:
+            i1 = iL1[searchsorted(zL1, z)]
+        z = zmax1[i1, 2]
+        i1seq[-i0] = i1
+
+    save(fname, zmax1[i1seq])
+    
+    figure()
+    zgrid = linspace(32, 50, 2401)
+    density1 = histogram(zmax1[i1seq,2], zgrid[::2])[0] / zmax0.shape[0] * (zgrid.size //2) / (zgrid[-1] - zgrid[0])
+    fill_between(zgrid[1::2], density1)
[…]

### 2020-10-13 code/lorenz_trajectory/lorenz_trajectory.py

@@ -1,0 +1,60 @@
+import os
+import sys
+import numba
+from numpy import *
+
+spec = [
+    ('x0', numba.float64),
+    ('y0', numba.float64),
+    ('dx', numba.float64),
+    ('dy', numba.float64),
+    ('nx', numba.int32),
+    ('ny', numba.int32),
+    ('count', numba.uint64[:,:]),
+]
+
+@numba.experimental.jitclass(spec)
+class Hist2d:
+    def __init__(self, x0, y0, dx, dy, nx, ny):
+        self.x0 = x0
+        self.y0 = y0
+        self.dx = dx
+        self.dy = dy
+        self.nx = nx
+        self.ny = ny
+        self.count = zeros((nx, ny), dtype=uint64)
+
+    def accumulate(self, x, y):
+        i, j = int((x - self.x0) / self.dx), int((y - self.y0) / self.dy)
+        if i >= 0 and j >= 0 and i < self.nx and j < self.ny:
+            self.count[i, j] += 1
+
+@numba.jit(nopython=True)
+def ddt(xyz):
+    sigma = 10
+    rho = 28
+    beta = 8./3
+    x, y, z = xyz
+    dxdt = sigma * (y - x)
+    dydt = x * (rho - z) - y
+    dzdt = x * y - beta * z
+    return array([dxdt, dydt, dzdt])
+
+@numba.jit(nopython=True)
+def step(xyz):
+    dt = 0.001
+    xyz1 = xyz + 0.5 * dt * ddt(xyz)
+    xyz += dt * ddt(xyz1)
+
+@numba.jit(nopython=True)
+def evolve(nsteps):
+    nx = nz = 2048
+    hist = Hist2d(-20, 0, 40./nx, 50./nz, nx, nz)
+    xyz = array([0.01, 0.01, 28.])
+    for i in range(nsteps):
+        step(xyz)
+        hist.accumulate(xyz[0], xyz[2])
+    return hist.count
+
+nsteps = int(sys.argv[1])
+save('lorenz_trajectory_{}.npy'.format(nsteps), evolve(nsteps))

### 2020-10-13 code/tent_quasiphysical/hist0.py

@@ -1,0 +1,68 @@
+import numba
+from pylab import *
+from numpy import *
+
+width = 2**13
+
+@numba.jit(nopython=True)
+def transform(x, p):
+    addition = (random.rand() > p) ^ (x % 2)
+    x = (x * 2) + addition
+    if x >= width:
+        x = (2 * width - x - 1) % width
+    return x
+
+@numba.jit(nopython=True)
+def accumulate(n, p):
+    xhist = zeros(width)
+    x = random.randint(0,width)
+    for i in range(n):
+        xhist[x] += 1
+        x = transform(x, p)
+    return xhist
+
+n = 10000000000
+
+figure(figsize=(16,18))
+density = array(accumulate(n, 0.5001), float) / n * width / 2
+x = linspace(0,2,width*2+1)[1:-1:2]
+fill_between(x, density)
+ylim([0,1])
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+savefig('tent_quasiphysical_hist_p_0.5001.png')
+
+figure(figsize=(16,18))
+density = array(accumulate(n, 0.51), float) / n * width / 2
+x = linspace(0,2,width*2+1)[1:-1:2]
+fill_between(x, density)
+ylim([0,1])
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+savefig('tent_quasiphysical_hist_p_0.51.png')
+
+figure(figsize=(16,18))
+density = array(accumulate(n, 0.55), float) / n * width / 2
+x = linspace(0,2,width*2+1)[1:-1:2]
+fill_between(x, density)
+ylim([0,1])
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+savefig('tent_quasiphysical_hist_p_0.55.png')
+
+figure(figsize=(16,18))
+density = array(accumulate(n, 0.9), float) / n * width / 2
+x = linspace(0,2,width*2+1)[1:-1:2]
+fill_between(x, density)
+ylim([0,1])
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
[…]

### 2020-10-13 code/tent_quasiphysical/hist1.py

@@ -1,0 +1,41 @@
+import numba
+from pylab import *
+from numpy import *
+
+width = 2**13
+
+@numba.jit(nopython=True)
+def transform(x, p, flip):
+    if flip:
+        addition = (random.rand() < p)
+    else:
+        addition = (random.rand() > p)
+    x = (x * 2) + addition
+    if x >= width:
+        x = (2 * width - x - 1) % width
+        return x, not flip
+    else:
+        return x, flip
+
+@numba.jit(nopython=True)
+def accumulate(n, p):
+    xhist = zeros(width)
+    x = random.randint(0,width)
+    flip = False
+    for i in range(n):
+        xhist[x] += 1
+        x, flip = transform(x, p, flip)
+    return xhist
+
+n = 100000000
+
+figure(figsize=(16,18))
+density = array(accumulate(n, 0.6), float) / n * width / 2
+x = linspace(0,2,width*2+1)[1:-1:2]
+fill_between(x, density)
+ylim([0,1])
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+savefig('tent_quasiphysical_alternaive1_hist_p_0.6.png')

### 2020-10-13 code/tent_quasiphysical/hist2.py

@@ -1,0 +1,43 @@
+import numba
+from pylab import *
+from numpy import *
+
+width = 2**13
+
+@numba.jit(nopython=True)
+def transform(x, p, flip):
+    if not flip and x & 0b11 == 0b11:
+        addition = (random.rand() < p)
+    elif flip and (~x) & 0b11 == 0b11:
+        addition = (random.rand() > p)
+    else:
+        addition = random.randint(0,2)
+    x = (x * 2) + addition
+    if x >= width:
+        x = (2 * width - x - 1) % width
+        return x, not flip
+    else:
+        return x, flip
+
+@numba.jit(nopython=True)
+def accumulate(n, p):
+    xhist = zeros(width)
+    x = random.randint(0,width)
+    flip = False
+    for i in range(n):
+        xhist[x] += 1
+        x, flip = transform(x, p, flip)
+    return xhist
+
+n = 10000000000
+
+figure(figsize=(16,18))
+density = array(accumulate(n, 0.6), float) / n * width / 2
+x = linspace(0,2,width*2+1)[1:-1:2]
+fill_between(x, density)
+ylim([0,1])
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+savefig('tent_quasiphysical_alternaive2_hist_p_0.6.png')

### 2020-10-13 code/tent_shadow/initial_value_problem.py

@@ -1,0 +1,27 @@
+from pylab import *
+from numpy import *
+
+x = pi/2 * ones(2)
+s = array([0, 1E-5])
+dx = [x[1] - x[0]]
+for i in range(25):
+    x = minimum(2 * x, 4 * (1+s) - 2 * x)
+    dx.append(x[1] - x[0])
+
+figure(figsize=(12,12))
+semilogy(absolute(dx), 'o')
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(30)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(30)
+savefig('tent_ivp.png')
+
+figure(figsize=(12,12))
+for s in [0, 0.1, 0.2, 0.3, 0.4, 0.5]:
+    plot([0, 1+s, 2+2*s], [0,2+2*s,0])
+axis('scaled')
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(30)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(30)
+savefig('scaled_tent_map.png')

### 2020-10-13 code/tent_shadow/tent_physical_pinched.py

@@ -1,0 +1,34 @@
+import numba
+from pylab import *
+from numpy import *
+
+@numba.jit(nopython=True)
+def maps(x, s):
+    if x > pi:
+        x = 2 * pi - x
+    return 4 * x / (s + 1 + sqrt((s + 1)**2 - 4 * s * x/pi))
+
+@numba.jit(nopython=True)
+def physical(n, s, width):
+    count = zeros(width, dtype=uint64)
+    x = random.rand() * 2 * pi
+    for i in range(10):
+        x = maps(x, s)
+    for i in range(n):
+        x = maps(x, s)
+        count[int(floor(x * width / (2 * pi)))] += 1
+    return count
+
+n = 10000000000
+for s in [0.01, 0.05, 0.2, 0.5]:
+    width = 2**13
+    density = physical(n, s, width) * width / 2 / n
+    x = linspace(0,2,width*2+1)[1:-1:2]
+    figure(figsize=(16,18))
+    fill_between(x, density)
+    ylim([0,1])
+    for tick in gca().xaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    for tick in gca().yaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    savefig('tent_pinched_physical_density_{}.png'.format(s))

### 2020-10-13 code/tent_shadow/tent_physical_squashed.py

@@ -1,0 +1,35 @@
+import numba
+from pylab import *
+from numpy import *
+
+@numba.jit(nopython=True)
+def maps(x, s):
+    if x < pi:
+        return x * (2-s)
+    else:
+        return (2*pi - x) * (2-s)
+
+@numba.jit(nopython=True)
+def physical(n, s, width):
+    count = zeros(width, dtype=uint64)
+    x = random.rand() * 2 * pi
+    for i in range(100):
+        x = maps(x, s)
+    for i in range(n):
+        x = maps(x, s)
+        count[int(floor(x * width / (2*pi)))] += 1
+    return count
+
+n = 10000000000
+for s in [0.01, 0.05, 0.2, 0.5]:
+    width = 2**13
+    density = physical(n, s, width) * width / 2 / n
+    x = linspace(0,2,width*2+1)[1:-1:2]
+    figure(figsize=(16,18))
+    fill_between(x, density)
+    ylim([0,1])
+    for tick in gca().xaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    for tick in gca().yaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    savefig('tent_squashed_physical_density_{}.png'.format(s))

### 2020-10-13 code/tent_shadow/tent_physical_wave.py

@@ -1,0 +1,36 @@
+import numba
+from pylab import *
+from numpy import *
+
+@numba.jit(nopython=True)
+def maps(x, s):
+    if x < pi:
+        return 4 * x / (1 + s + sqrt((1 + s)**2 - 4 * s * x/pi))
+    else:
+        x = 2*pi - x
+        return 4 * x / (1 - s + sqrt((1 - s)**2 + 4 * s * x/pi))
+
+@numba.jit(nopython=True)
+def physical(n, s, width):
+    count = zeros(width, dtype=uint64)
+    x = random.rand() * 2 * pi
+    for i in range(100):
+        x = maps(x, s)
+    for i in range(n):
+        x = maps(x, s)
+        count[int(floor(x * width / (2*pi)))] += 1
+    return count
+
+n = 10000000000
+for s in [0.01, 0.05, 0.2, 0.5]:
+    width = 2**13
+    density = physical(n, s, width) * width / 2 / n
+    x = linspace(0,2,width*2+1)[1:-1:2]
+    figure(figsize=(16,18))
+    fill_between(x, density)
+    ylim([0,1])
+    for tick in gca().xaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    for tick in gca().yaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    savefig('tent_wave_physical_density_{}.png'.format(s))

### 2020-10-13 code/tent_shadow/tent_shadow_pinched.py

@@ -1,0 +1,53 @@
+import numba
+from pylab import *
+from numpy import *
+
+@numba.jit(nopython=True)
+def find(x, y, s):
+    if x < 1:
+        return (y + y * (2-y) * s / 2) / 2
+    else:
+        return 2 - (y + y * (2-y) * s / 2) / 2
+
+@numba.jit(nopython=True)
+def shadow(n, s, width):
+    count = zeros(width, dtype=uint64)
+    x = random.rand() * 2
+    y = x
+    for i in range(100):
+        y = find(x, y, s)
+        x = x / 2 + random.randint(0,2)
+    for i in range(n):
+        y = find(x, y, s)
+        count[int(floor(y * width / 2))] += 1
+        x = x / 2 + random.randint(0,2)
+    return count
+
+n = 10000000000
+for s in [0.01, 0.05, 0.2, 0.9]:
+    width = 2**13
+    density = shadow(n, s, width) * width / 2 / n
+    x = linspace(0,2,width*2+1)[1:-1:2]
+
+    figure(1, figsize=(12,12))
+    y = linspace(0,2,1025)
+    plot(hstack([(y + y * (2-y) * s / 2) / 2,
+                 2 - (y + y * (2-y) * s / 2)[-1::-1] / 2]),
+         hstack([y, y[-1::-1]]))
+
+    figure(figsize=(16,18))
+    fill_between(x, density)
+    ylim([0,1])
+    for tick in gca().xaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    for tick in gca().yaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    savefig('tent_pinched_shadow_density_{}.png'.format(s))
+
+figure(1)
+axis('scaled')
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(30)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(30)
+savefig('pinched_tent_map.png')

### 2020-10-13 code/tent_shadow/tent_shadow_squashed.py

@@ -1,0 +1,50 @@
+import numba
+from pylab import *
+from numpy import *
+
+@numba.jit(nopython=True)
+def find(x, y, s):
+    if x < 1:
+        return y / (2-s)
+    else:
+        return 2 - y / (2-s)
+
+@numba.jit(nopython=True)
+def shadow(n, s, width):
+    count = zeros(width, dtype=uint64)
+    x = random.rand() * 2
+    y = x
+    for i in range(100):
+        y = find(x, y, s)
+        x = x / 2 + random.randint(0,2)
+    for i in range(n):
+        y = find(x, y, s)
+        count[int(floor(y * width / 2))] += 1
+        x = x / 2 + random.randint(0,2)
+    return count
+
+n = 10000000000
+for s in [0.01, 0.05, 0.2, 0.5]:
+    figure(1, figsize=(12,12))
+    y = linspace(0,2,1025)
+    plot([0, 1, 2], [0, 2-s, 0])
+
+    width = 2**13
+    density = shadow(n, s, width) * width / 2 / n
+    x = linspace(0,2,width*2+1)[1:-1:2]
+    figure(figsize=(16,18))
+    fill_between(x, density)
+    ylim([0,1])
+    for tick in gca().xaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    for tick in gca().yaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    savefig('tent_squashed_shadow_density_{}.png'.format(s))
+
+figure(1)
+axis('scaled')
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(30)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(30)
+savefig('squashed_tent_map.png')

### 2020-10-13 code/tent_shadow/tent_shadow_wave.py

@@ -1,0 +1,53 @@
+import numba
+from pylab import *
+from numpy import *
+
+@numba.jit(nopython=True)
+def find(x, y, s):
+    if x < 1:
+        return (y + y * (2-y) * s / 2) / 2
+    else:
+        return 2 - (y - y * (2-y) * s / 2) / 2
+
+@numba.jit(nopython=True)
+def shadow(n, s, width):
+    count = zeros(width, dtype=uint64)
+    x = random.rand() * 2
+    y = x
+    for i in range(100):
+        y = find(x, y, s)
+        x = x / 2 + random.randint(0,2)
+    for i in range(n):
+        y = find(x, y, s)
+        count[int(floor(y * width / 2))] += 1
+        x = x / 2 + random.randint(0,2)
+    return count
+
+n = 10000000000
+for s in [0.01, 0.05, 0.2, 0.9]:
+    width = 2**13
+    density = shadow(n, s, width) * width / 2 / n
+    x = linspace(0,2,width*2+1)[1:-1:2]
+
+    figure(1, figsize=(12,12))
+    y = linspace(0,2,1025)
+    plot(hstack([(y + y * (2-y) * s / 2) / 2,
+                 2 - (y - y * (2-y) * s / 2)[-1::-1] / 2]),
+         hstack([y, y[-1::-1]]))
+
+    figure(figsize=(16,18))
+    fill_between(x, density)
+    ylim([0,1])
+    for tick in gca().xaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    for tick in gca().yaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    savefig('tent_wave_shadow_density_{}.png'.format(s))
+
+figure(1)
+axis('scaled')
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(30)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(30)
+savefig('wave_tent_map.png')

### 2020-10-13 code/tent_shadow/tilted_conjugate_map.py

@@ -1,0 +1,61 @@
+from pylab import *
+from numpy import *
+
+
+for s in [0, 0.1, 0.2, 0.3, 0.4, 0.5]:
+    xAll, yAll = array([[0, 2], [0, 2]])
+    for i in range(22):
+        # xNew = xAll[1::2] / 2
+        yNew = (yAll[1:] + yAll[:-1]) / 2
+        # yNew = yAll[1::2] / 2 + minimum(yAll[1::2], 1 - yAll[1::2]) * s
+        weightLeft, weightRight = zeros([2, yAll.size - 1])
+        weightLeft[::2] = (1-s)
+        weightRight[::2] = (1+s)
+        weightLeft[1::2] = (1+s)
+        weightRight[1::2] = (1-s)
+        xNew = (xAll[:-1] * weightLeft + xAll[1:] * weightRight) / 2
+        xAll = hstack([xAll, xNew])
+        yAll = hstack([yAll, yNew])
+        iSort = xAll.argsort()
+        xAll = xAll[iSort]
+        yAll = yAll[iSort]
+    
+    figure(1, figsize=(12,12))
+    plot(xAll, yAll)
+
+    figure(2, figsize=(12,12))
+    plot([0, 1+s, 2], [0,2,0])
+
+    nSample = 2**20
+    x = linspace(0,2,nSample*2+1)[1:-1:2]
+    i = searchsorted(xAll, x)
+    assert i.min() > 0
+    r = (xAll[i] - x) / (xAll[i] - xAll[i-1])
+    y = yAll[i-1] * r + yAll[i] * (1 - r)
+
+    figure(figsize=(16,18))
+    width = 2**13
+    density = histogram(y, linspace(0,2,width+1))[0] / nSample * width / 2
+    fill_between(linspace(0,2,width*2+1)[1:-1:2], density)
+    ylim([0,1])
+    for tick in gca().xaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    for tick in gca().yaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    savefig('tilted_conjugate_density_{}.png'.format(s))
+
+figure(1)
+axis('scaled')
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(30)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(30)
+savefig('tilted_conjugate_map.png')
+
+figure(2)
+axis('scaled')
[…]

### 2020-10-13 data/collect.py

@@ -1,0 +1,16 @@
+# collect.py
+import os
+import argparse
+from numpy import *
+
+parser = argparse.ArgumentParser()
+parser.add_argument('-i', type=str, nargs='+')
+parser.add_argument('-o', type=str)
+args = parser.parse_args()
+
+a = 0
+for fname in args.i:
+    a = a + load(fname)
+for fname in args.i:
+    os.remove(fname)
+save(args.o, a)

### 2020-10-13 data/compute_lorenz_stats.py

@@ -1,0 +1,15 @@
+from numpy import *
+
+for i in range(3):
+    t = load('lorenz_periodic_{}.npy'.format(i+1))
+    print(t[:,2].mean(), exp(-t[:,2]**2/2).mean())
+
+a = load('lorenz_density_10.0_28.0_2.6666666666666665_0.01_5000.npy')
+
+za = linspace(0, 50, 2048*2+1)[1:-1:2]
+xa = linspace(-20, 20, 2048*2+1)[1:-1:2]
+z_mean, gaussian_mean = (a * za).sum() / a.sum(), (a * exp(-za**2/2)).sum() / a.sum()
+z2_mean, gaussian2_mean = (a * za**2).sum() / a.sum(), (a * exp(-za**2/2)**2).sum() / a.sum()
+z_std, gaussian_std = sqrt(z2_mean - z_mean**2), sqrt(gaussian2_mean - gaussian_mean**2) 
+print(z_mean, gaussian_mean)
+print(z_std / sqrt(a.sum()), gaussian_std / sqrt(a.sum()))

### 2020-10-13 data/plot_ensemble.py

@@ -1,0 +1,18 @@
+from pylab import *
+from matplotlib.colors import LogNorm
+from numpy import *
+
+for i in [10,15,50]:
+    a = load('lorenz_density_10.0_28.0_2.6666666666666665_0.01_{}.npy'.format(i*100))
+    a[a == 0] = 0.1
+    figure(figsize=(24,24))
+    imshow(a.T, extent=[-20,20,0,50],
+           norm=LogNorm(vmin=0.2, vmax=a.max()), cmap='inferno', origin='lower')
+    axis('scaled')
+    cbar = colorbar()
+    cbar.ax.tick_params(labelsize=50)
+    for tick in gca().xaxis.get_major_ticks():
+        tick.label.set_fontsize(50)
+    for tick in gca().yaxis.get_major_ticks():
+        tick.label.set_fontsize(50)
+    savefig('lorenz_ensemble_{}.png'.format(i))

### 2020-10-13 data/plot_lorenz_periodic.py

@@ -1,0 +1,20 @@
+from pylab import *
+from matplotlib.colors import LogNorm
+from numpy import *
+
+a = load('lorenz_density_10.0_28.0_2.6666666666666665_0.01_5000.npy')
+a[a == 0] = 0.1
+figure(figsize=(24,24))
+imshow(a.T, extent=[-20,20,0,50],
+       norm=LogNorm(vmin=0.2, vmax=a.max()), cmap='inferno', origin='lower')
+
+styles = ['-', '--', ':']
+for i in range(3):
+    t = load('lorenz_periodic_{}.npy'.format(i+1))
+    plot(t[:,0], t[:,2], styles[i] + 'b', lw=4)
+axis('scaled')
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(50)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(50)
+savefig('lorenz_periodic.png')

### 2020-10-13 data/plot_lorenz_zmax_map.py

@@ -1,0 +1,196 @@
+from pylab import *
+from numpy import *
+
+z = load('lorenz_zmax_10_28_2.6666666666666665.npy')[:,2]
+'''
+figure(figsize=(16,16))
+plot(z[:-1], z[1:], '.', ms=1)
+axis('scaled')
+axis([28,49,28,49])
+xticks(linspace(29, 49, 5))
+yticks(linspace(29, 49, 5))
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+plot([28,49], [28,49], '--k')
+savefig('lorenz_zmax_10_28_2.6666666666666665.png')
+
+figure(figsize=(16,16))
+plot(z[:-2], z[2:], '.', ms=1)
+axis('scaled')
+axis([28,49,28,49])
+xticks(linspace(29, 49, 5))
+yticks(linspace(29, 49, 5))
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+plot([28,49], [28,49], '--k')
+savefig('lorenz_zmax2_10_28_2.6666666666666665.png')
+
+figure(figsize=(16,16))
+plot(z[:-3], z[3:], '.', ms=1)
+axis('scaled')
+axis([28,49,28,49])
+xticks(linspace(29, 49, 5))
+yticks(linspace(29, 49, 5))
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+for tick in gca().yaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+plot([28,49], [28,49], '--k')
+savefig('lorenz_zmax3_10_28_2.6666666666666665.png')
+
+figure(figsize=(16,16))
+plot([0, 1, 2], [0, 2, 0])
+axis('scaled')
+axis([0,2,0,2])
+xticks(linspace(0, 2, 3))
+yticks(linspace(0, 2, 3))
+for tick in gca().xaxis.get_major_ticks():
+    tick.label.set_fontsize(40)
+for tick in gca().yaxis.get_major_ticks():
[…]

### 2020-10-13 data/plot_trajectory.py

@@ -1,0 +1,18 @@
+from pylab import *
+from matplotlib.colors import LogNorm
+from numpy import *
+
+for i in [6,7,9]:
+    a = array(load('lorenz_trajectory_{}.npy'.format(10**i)), float)
+    a[a == 0] = 0.1
+    figure(figsize=(24,24))
+    imshow(a.T, extent=[-20,20,0,50],
+           norm=LogNorm(vmin=0.2, vmax=a.max()), cmap='inferno', origin='lower')
+    axis('scaled')
+    cbar = colorbar()
+    cbar.ax.tick_params(labelsize=50)
+    for tick in gca().xaxis.get_major_ticks():
+        tick.label.set_fontsize(50)
+    for tick in gca().yaxis.get_major_ticks():
+        tick.label.set_fontsize(50)
+    savefig('lorenz_trajectory_{}.png'.format(10**i/1000))

### 2020-10-13 figure/autocrop.py

@@ -1,0 +1,32 @@
+#!/usr/bin/python3
+
+import os
+import sys
+import argparse
+from glob import glob
+from PIL import Image
+import numpy as np
+
+parser = argparse.ArgumentParser(description='Autocrop png images.')
+parser.add_argument('filenames', metavar='filename', type=str, nargs='+',
+                    help='png image filenames to be cropped')
+
+args = parser.parse_args()
+
+for arg in args.filenames:
+    for f in glob(arg):
+        image=Image.open(f)
+        image = image.convert('RGBA')
+
+        image_data = np.asarray(image)
+        not_white = (image_data.min(axis=2) < 255)
+        if image_data.shape[2] > 3:
+            not_white *= (image_data[:,:,3] > 0)
+        non_empty_columns = np.where(not_white.max(axis=0))[0]
+        non_empty_rows = np.where(not_white.max(axis=1))[0]
+        cropBox = (min(non_empty_rows), max(non_empty_rows), min(non_empty_columns), max(non_empty_columns))
+
+        image_data_new = image_data[cropBox[0]:cropBox[1]+1, cropBox[2]:cropBox[3]+1 , :]
+
+        cropped = Image.fromarray(image_data_new)
+        cropped.save(f)

### 2020-10-13 main.tex
make edits up to quasi-physical solutions.

@@
 \label{sec:quasiphysical}
-The simplicity of the tent map, as well as its binary decimal form (\ref{tent_xor}),
-enables us to study non-physical solutions that are not periodic.
+The simplicity of the tent map, as well as its binary form (\ref{tent_xor}),
+enables us to study nonphysical solutions that are not periodic.
 As in the last section, consider an
@@
 \end{equation}
-the digits of which are not independent of each other.  Instead,
-suppose each digit is more likely to repeat the previous digit than to be different.
+in which the bits $x_{0,j}$ are not independent of each other.  Instead,
+suppose each bit is more likely to be identical to the previous bit than to be different.
 That is, $x_{0,j+1}=x_{0,j}$ with probability $p>\frac12$ for $j=0,1,2,\ldots$.
-Then, by Equation (\ref{tent_xor}), the digits of every $x_i, i=1,2,\ldots$
-have digits that repeats its previous digit with probability $p$.
-Moreover, consider Equation (\ref{tent_xor}) for $j=0$ and any $i$:
+Then, by Eq. \ref{tent_xor}, the bits of every $x_i, i=1,2,\ldots$ follow the same pattern, namely, each bit repeats the previous bit
+with probability $p$. Moreover, consider Eq. \ref{tent_xor} 
+for $j=0$ and any $i$:
 \begin{equation}
@@
 Because $x_{i,0}=x_{i,1}$ with probability $p>\frac12$, $x_{i+1,0}=0$ with probability 
-$p>\frac12$.  The solution, starting from $x_1$, visits $[0,1]$ with
+$p>\frac12$.  Starting from $x_1$, this solution visits $[0,1]$ with
 probability $p>\frac12$.  Instead of visiting $[0,1]$ and $[1,2]$ with
-equal probability, as a physical solution does, this solution favors $[0,1]$.
-Because it visits the interval $[0,2]$ with a provably different frequency from that
+equal probability, as a physical solution does, this solution favors $[0,1]$. Since it provably visits the interval $[0,2]$ with a different frequency from that
 of a physical solution, this is a nonphysical solution.
 
-Note that the solution we just constructed is both nonphysical and aperiodic.
-The digits of $x_0$, though correlated with each other,
-still can exhibit an infinite variety of patterns.  As the map iterates, these patterns
+If $p = 1$, the bits of $x_0$ are either all zeros or all ones.
+So, all further iterates when $x_0$ is 0, are 0, and this is a trivial nonphysical solution of type 1.
+When $x_0$ consists of all ones, the solution eventually becomes 0, and hence is also a nonphysical solution, which is eventually constant (of period 1).
+Even when this solution does not
+attain zero in a large number of iterations, it does not 
+distribute uniformly on $[0,2]$ like a physical solution 
+because all $x_i$ starting from $i = 1$ are dyadic numbers; other numbers are not visited. Now when $p > \frac12$ but strictly less than 1, 
+the solutions we just constructed are both nonphysical and aperiodic.
+The bits of $x_0$, though correlated with each other,
+still can exhibit an infinite variety of patterns. 
+This implies that further iterates may not eventually converge to 
+any fixed point, nor show any repetitive patterns.  
+ As the map iterates, these patterns
 shift towards more significant digits, and the solution visits an infinite set of
-points.  We call such aperiodic nonphysical solutions ``quasi-physical'' solutions.
+points. 
+ Nevertheless, the solutions observed along a trajectory also 
+ do not conform to a uniform distribution on $[0,2]$ since they 
+ preferentially visit the first half of this interval.
+We call such aperiodic nonphysical solutions ``quasi-physical'' solutions.
 

### 2020-10-13 main.tex
edits up to fig 7

@@
 So, all further iterates when $x_0$ is 0, are 0, and this is a trivial nonphysical solution of type 1.
-When $x_0$ consists of all ones, the solution eventually becomes 0, and hence is also a nonphysical solution, which is eventually constant (of period 1).
-Even when this solution does not
-attain zero in a large number of iterations, it does not 
+When $x_0$ consists of all ones, the solution eventually becomes 0, and hence is also a nonphysical solution that is eventually periodic (of period 1).
+Even when this solution does not attain zero in a large number of iterations, it does not
 distribute uniformly on $[0,2]$ like a physical solution 
@@
 still can exhibit an infinite variety of patterns. 
-This implies that further iterates may not eventually converge to 
+This implies that the solution may not eventually converge to 
 any fixed point, nor show any repetitive patterns.  
@@
 \caption{Empirical density function of long solutions (a billion steps)
-starting from four point whose binary digits have probability $p$ of
+starting from four points whose binary digits have probability $p$ of
 repeating the previous digit.
@@
 top-right: $p=0.51$; bottom-left: $p=0.55$; bottom-right: $p=0.9$.
-Note that the empirical densities at larger $p$s go well beyond 1.}
+Note that the empirical densities at larger $p'$s are well above 1.}
 \label{fig:tent_quasiphysical}
@@
 the density on the right side, $[1,2]$.  This is consistent with our
-theoretical analysis in the first paragraph of this subsection.
+theoretical analysis at the beginning of this subsection.
 When $p=0.9$, the fractal pattern is so intensified that most
 of the solution seem to concentrate in a collection of tiny intervals.
-These plots exposes the diversity of quasi-physical solutions.
+These plots expose the diversity of quasi-physical solutions.
 Their statistical distribution can be either

### 2020-10-13 main.tex
Start editing section 3

@@
 distribute uniformly on $[0,2]$ like a physical solution 
-because all $x_i$ starting from $i = 1$ are dyadic numbers; other numbers are not visited. Now when $p > \frac12$ but strictly less than 1, 
+because all $x_i, i\geq 1$ are of the form $1/2^k$ for some $k \in \mathbb{N}$; other numbers are not visited. Now when $p > \frac12$ but strictly less than 1, 
 the solutions we just constructed are both nonphysical and aperiodic.
@@
 So far we have constructed one class of quasi-physical solutions.
-It is noteworthy that there are infinite variations to how we
-construct quasi-physical solutions.  In the binary decimal
-representation of the initial condition $x_0$, any statistical
+It is noteworthy that there are infinite ways to 
+construct quasi-physical solutions.  In the binary representation 
+of the initial condition $x_0$, any statistical
 deviation from equal probability of 0 or 1, or any statistical
 dependence among digits would lead to quasi-physical solutions.
-One could, for example, construct an $x_0$ whose digit is
+One could, for example, construct an $x_0$ in which a bit is
 more likely to be 1 than 0 only if it follows two consecutive 0's.
@@
 \ref{fig:tent_quasiphysical}.  Nevertheless, as can be
-observed in Figure \ref{fig:tent_alternative}, it would show
+observed in Figure \ref{fig:tent_alternative}, it shows
 a remarkable resemblance in its ``hairiness''; some kind of fractal
-pattern would emerge from the distribution.  A fractal density
+pattern emerges from the distribution.  A fractal distribution
 appears to be a hallmark of quasi-physical solutions.
@@
     The solution shown in the left plot starts from an initial condition
-    whose binary decimal digits are independent and have probability 0.6 of being 0.
+    whose bits are independent and have probability 0.6 of being 0.
     The solution shown in the right plot starts from an initial condition
-    whose binary decimal digits have probability of 0.6 of being 1 only following two
-    consecutive 1's; otherwise it is 0 or 1 with equal probability of 0.5.}
+    whose bits have a probability of 0.6 of being 1 only following two
+    consecutive 1's; otherwise a bit is 0 or 1 with equal probability.}
     \label{fig:tent_alternative}
@@
 
-We have only seem quasi-physical solutions for the tent map.
+We have only seen quasi-physical solutions for the tent map.
 It is difficult to analytically construct quasi-physical solutions
-to the Lorenz equation and other, more complex, chaotic governing
-equations.  Neverthless, the similarity between the tent map and the
-Lorenz map, shown in Figure \ref{fig:lorenz_map}, suggests that
+to the Lorenz equation and other, more complex, governing
+equations which typically produce chaotic solutions.  Neverthless, the similarity 
+between the tent map and the Lorenz map \cite{tent2}, shown in Figure \ref{fig:lorenz_map}, suggests that
 quasi-physical solutions may exist for the Lorenz map, and by
@@
 
-These quasi-physical solutions raise doubt to the usefulness of shadowing
+These quasi-physical solutions raise doubts over the usefulness of shadowing
 in some applications.  For example, even if a numerical
 solution is shadowed by a solution to the true governing physics,
-is the shadowing solution physical or quasi-physical?  Such doubt may first
-seem paranoid; since almost all solutions are physical, it appears
-that one has to be extremely unlucky for the shadowing solution to
-be nonphysical. It seems reasonable to argue that because the set of
-all nonphysical solutions is measure-zero, the probability of finding
-a nonphysical solution through shadowing is zero percent.  Such argument,
+is the shadowing solution physical or quasi-physical? At first glance, this may 
+seem to be a nonissue because almost all solutions are physical. 
+It seems reasonable to argue that because the set of
+all nonphysical solutions is Lebesgue measure-zero, the probability of finding
+a nonphysical solution through shadowing is zero percent.  Such an argument,
 as we show in the next section, is wrong.  The probability of
@@
 To illustrate the concept of shadowing, consider the tent map, defined
-by Equation (\ref{tentmap}), and a scaled version of the map, defined by
+in Eq. \ref{tentmap}, and a scaled version of the map, defined by
 \begin{equation} \label{tent_scaled}
@@
 \end{equation}
-where $s<<1$.  Note that tiny differences in $s$ can lead to drastic
-different solutions, if the solutions start from the same initial condition.
+where $s<<1$.  Note that a small change in $s$ can lead to drastic differences 
+between solutions starting from the same initial condition.
 As demonstrated in Figure \ref{fig:tent_ivp}, this sensitivity to small
@@
     \includegraphics[width=0.48\textwidth]{figure/tent_ivp.png}
-    \caption{Sensitivity to small perturbation in the governing
+    \caption{L: The scaled tent map $\hat{\varphi}_s$ at different values of $s$ between 0 and 1. R: sensitivity to small perturbation in the governing
     equation.
     The y-axis shows the absolute value of the difference between
-    two solutions, one satisfying Equation (\ref{tent_scaled}),
+    two solutions, one satisfying Eq. \ref{tent_scaled},
     one for $s=0$ and the other for $s=10^{-5}$.

### 2020-10-14 main.tex
Starting to edit from Eq. 15

@@
 \subsection{Example of a shadowing solution for the tent map}
+\label{sec:sub:shadow-example}
 To illustrate the concept of shadowing, consider the tent map, defined
@@
 
-We can avoid this sensitivity to the governing equation
-by a coordinated perturbation to the initial condition.
-Consider two solutions satisfying Equation (\ref{tent_scaled})
-for different values of $s$.  Instead of starting from the same
+We can avoid this sensitivity to the governing equation using a coordinated perturbation to the initial condition.
+Consider two solutions satisfying Eq. \ref{tent_scaled} at different values of $s$.  Instead of starting from the same
 initial condition, these two solutions start from $x_0 (1 + s)$ with
@@
 where $x_{i+1}=\varphi(x_i)$ satisfies the original
-tent map (\ref{tentmap}).  When the values of $s$ are similar between
-the two solutions, this pair of solutions stay uniformly close to each
-other forever.  Such pair of solutions, satisfying slightly different
-governing equations, yet keep close to each other over a long time,
-are known as shadowing solutions.
+tent map (Eq. \ref{tentmap}).  When the values of $s$ are similar between
+the two solutions, this pair of solutions stays uniformly close to each
+other forever.  Such a true solution of the governing equation
+that stays close to a given perturbed solution over a long time, is known as a \emph{shadowing} solution. In this example, the perturbed solution satisfies a 
+slightly different governing equation.
 
-For every solution $\hat{x}_i,i=0,\ldots$ satisfying Equation
-(\ref{tent_scaled}), there is a shadowing solution satisfying Equations
-(\ref{tentmap}): $x_i:= \hat{x}_i/(1+s)$.  The map,
+For every solution $\hat{x}_i,i=0,\ldots$ satisfying Eq. \ref{tent_scaled}, there is a shadowing solution satisfying Eq. \ref{tentmap}: $x_i:= \hat{x}_i/(1+s)$.  The map,
 \begin{equation} \label{conjugate_scaled}
-    \hat{h}_s(x):=x/(1+s)
+    h_s(x):=x/(1+s)
 \end{equation}
-is known as the conjugate map, because it satisfies
+is known as a \emph{conjugacy} between the two maps $\varphi$ and $\hat{\varphi_s}$ 
+because it satisfies
 \begin{equation} \label{conjugate_def}
-    \varphi(\hat{h}_s(x)) = \hat{h}_s(\hat\varphi_s(x))\;,\quad \forall x
+    \varphi(h_s(x)) = h_s(\hat\varphi_s(x))\;,\quad \forall\: x
 \end{equation}
-or equivalently, $\varphi\circ \hat{h}_s \equiv \hat{h}_s\circ\hat\varphi_s$.
-Such conjugate maps can generally help us construct shadowing solutions.
-If a conjugate map $h$ exists for two maps $\varphi$ and $\psi$, i.e.,
-$\varphi\circ h \equiv h\circ\psi$, then for every solution satisfying
-$\psi$, $h$ maps it to a shadowing solution satisfying $\varphi$.
+or equivalently, $\varphi\circ h_s \equiv h_s\circ\hat\varphi_s$.
+Such conjugacy maps can generally help us construct shadowing solutions.
+For every solution satisfying
+$\hat{\varphi}_s$, $h_s$ maps it to a shadowing solution satisfying $\varphi$.
 
@@
 the answer is almost surely positive.  We can demonstrate that
-almost any solution of the scaled tent map (\ref{tent_scaled}) is uniformly
+almost any solution of the scaled tent map (Eq. \ref{tent_scaled}) is uniformly
 distributed in $[0, 2(1+s)]$ -- we can repeat our analysis
 in Section \ref{sec:tent} but represent our initial condition as
-$x_0= (1+s) \sum x_{0,j} / 2^j$.  Its shadowing solution,
-satisfying the original tent map (\ref{tentmap}), can be obtained through the
-conjugate map (\ref{conjugate_scaled}).  Thus the distribution
+$x_0= (1+s) \sum_{j=0}^\infty x_{0,j} / 2^j$.  Its shadowing solution,
+which satisfies the original tent map (Eq. \ref{tentmap}), can be obtained through the
+conjugacy (Eq. \ref{conjugate_scaled}).  Thus, the distribution
 of the shadowing solution can be obtained by mapping a uniform
-distribution in $[0,2(1+s)]$ through the conjugate map.
-This leads to a uniformly distribution in $[0,2]$,
+distribution in $[0,2(1+s)]$ through the conjugacy map.
+This leads to a uniformly distribution in $[0,2]$, which is
 precisely the distribution of a physical solution of the tent map.
@@
 This simple example is useful in illustrating the concept of shadowing
-and the utility of the conjugate map.  This example is nevertheless special
+and the utility of the conjugacy map.  But it is rather special
 in that the shadowing solution is almost always physical.  Our next
@@
 \subsection{An example of quasi-physical shadowing solution}
-
+\label{sec:sub:quasi-physical-shadow}
 The tilted tent map is defined as
@@
     \frac2{1+s} x \quad & x < 1+s \\
-    \frac2{1-s}(2-s) \quad & x \ge 1+s
+    \frac2{1-s}(2-x) \quad & x \ge 1+s
     \end{cases}
 \end{equation}
-When $s=0$, this map is identical to the tent map (\ref{tentmap}).
+When $s=0$, this map is identical to the tent map (Eq. \ref{tentmap}).
 Let $\tilde{x}_{i+1}=\hat\varphi_s(\tilde{x}_i), i=0,1,\ldots$ be a solution.
@@
     \includegraphics[width=0.48\textwidth]{figure/tilted_conjugate_density_0.5.png}
-    \caption{Conjugate map $\tilde{h}_s$ between the tent map (\ref{tentmap})
+    \caption{Upper-left: the tilted tent map (\ref{tent_tilted}) for $s=0, 0.1, 0.2, 0.3, 0.4,$ and 0.5.
+    Upper-right: the conjugate map $\tilde{h}_s$ between the tent map (\ref{tentmap})
     and the tilted tent map (\ref{tent_tilted}), evaluated using
-    Equation (\ref{conjugate_tilted}-\ref{conjugate_tilted_helper}).
-    Starting from the diagonal line, $s=0, 0.1, 0.2, 0.3, 0.4,$ and 0.5.}
+    Equation (\ref{conjugate_tilted}-\ref{conjugate_tilted_helper}), for the same set of $s$ as the
+    upper-left plot. Lower-left: the density of a trajectory satisfying (\ref{tentmap})
+    that shadows a random trajectory satisfying (\ref{tent_tilted}) for $s=0.1$.
+    Lower-right: the density of a trajectory satisfying (\ref{tentmap})
+    that shadows a random trajectory satisfying (\ref{tent_tilted}) for $s=0.5$.
+    }
     \label{fig:tilted_tent_conjugate}
@@
 
-Here we completed our example.  For almost every solution to a perturbed governing
-equation, its shadowing solution, satisfying the real governing equation, will
-be nonphysical.  The shadowing theorems do guarantee the existence of shadowing
-solutions.  But are we sure it is the kind of solution we want?
+In the previous subsection, we examined a tilting perturbation to the tent map.
+We derived the fractal conjugate map, and showed that a typical shadowing trajectory
+is exactly the quasi-physical solution we analyzed in Section \ref{sec:quasiphysical}.
+This example contrast with Section \ref{sec:sub:shadow-example}, in which the
+shadowing solutions of a different perturbation, the scaling perturbation,
+are physical solutions.  Which situation is more typical?  When we make other
+types of perturbations, do we expect to observe physical or nonphysical
+shadowing solutions?
 
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.48\textwidth]{figure/tent_tilted_shadow_density_0.01.png}
-    \hspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tent_tilted_shadow_density_0.05.png}
-    \\ \vspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tent_tilted_shadow_density_0.4.png}
-    \hspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tent_tilted_shadow_density_0.7.png}
-    \caption{Caption}
-    \label{fig:tent_tilted_shadow}
-\end{figure}
+%%
+%%\begin{figure}[H]
+%%    \centering
+%%    \includegraphics[width=0.48\textwidth]{figure/tent_tilted_shadow_density_0.01.png}
+%%    \hspace{0.02\textwidth}
+%%    \includegraphics[width=0.48\textwidth]{figure/tent_tilted_shadow_density_0.05.png}
+%%    \\ \vspace{0.02\textwidth}
+%%    \includegraphics[width=0.48\textwidth]{figure/tent_tilted_shadow_density_0.4.png}
+%%    \hspace{0.02\textwidth}
+%%    \includegraphics[width=0.48\textwidth]{figure/tent_tilted_shadow_density_0.7.png}
+%%    \caption{Caption}
+%%    \label{fig:tent_tilted_shadow}
+%%\end{figure}
+%%
 
@@
     \includegraphics[width=0.48\textwidth]{figure/squashed_tent_map.png}
-    \caption{Caption}
+    \caption{The ``squashed'' tent map (\ref{eqn:tent_squashed}) for $s=0.01, 0.05, 0,2,$ and 0.5.}
     \label{fig:tent_squashed}
 \end{figure}
+
+This section attempts to answer this question by investigating several other
+perturbations, the first of which scales the height, but not the width,
+of the tent map.  The resulting ``squashed'' tent map, illustrated
+in Figure \ref{fig:tent_squashed}, has a closed form of
+\begin{equation} \label{eqn:tent_squashed}
+    \widehat{\varphi^{sq}}_s(x) := \begin{cases}
+    (2-s)x \quad & x < 1 \\
+    (2-s)(2 - x) \quad & x \ge 1
+    \end{cases}
+\end{equation}
 

### 2020-10-14 comments.tex
Start editing section 3
Add a comment on including perturbed solutions.

@@
 \end{enumerate} 
+\subsection*{Section 3.1}
+\begin{enumerate}
+    \item Reading the text around the definition of the shadowing solution, 
+    the reader may wonder when a true solution is called a shadowing solution.
+    What does it shadow? To clarify, we can add that a solution of $\tilde{\varphi}_s$
+    is a perturbed solution (a pseudoorbit) of $\varphi.$ The shadowing solution 
+    is a true trajectory of $\varphi$ that stays close, forever, to the perturbed solution.
+\end{enumerate}
+
 \end{document}

### 2020-10-14 main.tex
Add a comment on including perturbed solutions.

@@
 we need to analyze two cases: $x<1+s$ and $x\ge 1+s$.
-
+\begin{figure}[H]
+  \centering
+    \includegraphics[width=0.48\textwidth]{figure/tilted_tent_map.png}
+    \hspace{0.02\textwidth}
+    \includegraphics[width=0.48\textwidth]{figure/tilted_conjugate_map.png}
+    \\ \vspace{0.02\textwidth}
+     \caption{Upper-left: the tilted tent map (\ref{tent_tilted}) for $s=0, 0.1, 0.2, 0.3, 0.4,$ and 0.5.
+    Upper-right: the conjugate map $\tilde{h}_s$ between the tent map (\ref{tentmap})
+    and the tilted tent map (\ref{tent_tilted}), evaluated using
+    Equation (\ref{conjugate_tilted}-\ref{conjugate_tilted_helper}), for the same set of $s$ as the
+    upper-left plot.}
+     \label{fig:tilted_tent_conjugate}
+\end{figure}
 When $x<1+s$, $\xi_{s,0}(x)=0$; thus
@@
 \begin{figure}[H]
-    \centering
-    \includegraphics[width=0.48\textwidth]{figure/tilted_tent_map.png}
-    \hspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tilted_conjugate_map.png}
-    \\ \vspace{0.02\textwidth}
    \includegraphics[width=0.48\textwidth]{figure/tilted_conjugate_density_0.1.png}
@@
     \includegraphics[width=0.48\textwidth]{figure/tilted_conjugate_density_0.5.png}
-    \caption{Upper-left: the tilted tent map (\ref{tent_tilted}) for $s=0, 0.1, 0.2, 0.3, 0.4,$ and 0.5.
-    Upper-right: the conjugate map $\tilde{h}_s$ between the tent map (\ref{tentmap})
-    and the tilted tent map (\ref{tent_tilted}), evaluated using
-    Equation (\ref{conjugate_tilted}-\ref{conjugate_tilted_helper}), for the same set of $s$ as the
-    upper-left plot. Lower-left: the density of a trajectory satisfying (\ref{tentmap})
+   \caption{
+     Lower-left: the density of a trajectory satisfying (\ref{tentmap})
     that shadows a random trajectory satisfying (\ref{tent_tilted}) for $s=0.1$.
@@
     }
-    \label{fig:tilted_tent_conjugate}
+    \label{fig:tilted_tent_density}
 \end{figure}
 
-Figure \ref{fig:tilted_tent_conjugate} shows how fractal the conjugate map is.
+Figure \ref{fig:tilted_tent_density} shows how fractal the conjugate map is.
 This fractal conjugate map transforms a uniform distribution in $[0,2]$ into

### 2020-10-14 main.tex
fix typos upto defining tilted conjugacy map

@@
 \begin{equation} \label{conjugate_scaled}
-    h_s(x):=x/(1+s)
+    \hat{h}_s(x):=x/(1+s)
 \end{equation}
@@
 \begin{equation} \label{conjugate_def}
-    \varphi(h_s(x)) = h_s(\hat\varphi_s(x))\;,\quad \forall\: x
+    \varphi(\hat{h}_s(x)) = \hat{h}_s(\hat\varphi_s(x))\;,\quad \forall\: x
 \end{equation}
@@
 \begin{equation} \label{tent_tilted}
-    \hat\varphi_s(x) := \begin{cases}
+    \tilde\varphi_s(x) := \begin{cases}
     \frac2{1+s} x \quad & x < 1+s \\
@@
 When $s=0$, this map is identical to the tent map (Eq. \ref{tentmap}).
-Let $\tilde{x}_{i+1}=\hat\varphi_s(\tilde{x}_i), i=0,1,\ldots$ be a solution.
+Let $\tilde{x}_{i+1}=\tilde\varphi_s(\tilde{x}_i), i=0,1,\ldots$ be a solution of 
+$\tilde{\varphi}_s$ that is also a perturbed solution of $\varphi$.
 As we explained in the previous subsection,
-this shadowing solution can be found through a conjugate map $\tilde{h}_s$.
+the corresponding 
+shadowing solution, which solves $\varphi$, can be found through a conjugacy map $\tilde{h}_s$ that connects the solutions of $\varphi$ and $\tilde{\varphi}_s$.
 If $\varphi\circ \tilde{h}_s \equiv \tilde{h}_s\circ\tilde\varphi_s$,
-then $x_i=\tilde{h}_s(\tilde{x}_i), i=0,1,\ldots$ is a shadowing solution
-that satisfies $x_{i+1} = \varphi(x_i)$.
+then $x_i=\tilde{h}_s(\tilde{x}_i), i=0,1,\ldots$ is the shadowing solution 
+that satisfies $x_{i+1} = \varphi(x_i)$, corresponding to the perturbed solution 
+$\tilde{x}_i,i=0,1,\cdots$.
 
-The conjugate map, although more complex than the one
+The conjugacy map, although more complex than the one
 in the last subsection, has the following closed form:
@@
     \xi_{s,j-1}(x) & j>0, \tilde\varphi_s^j(x) < 1+s \\
-    1 - \xi_{s,j-1}(x) & j>0, \tilde\varphi_s^j(x) \ge 1+s
+    1 - \xi_{s,j-1}(x) & j>0, \tilde\varphi_s^j(x) \ge 1+s.
     \end{cases}
 \end{equation}
-To see why this map satisfies the definition of the conjugate map --
+In the above expression, $\tilde{\varphi}^j_s$ refers to the $j$-time 
+composition of $\tilde{\varphi}_s.$ That is, if $\tilde{x}_{j+1} = \tilde{\varphi}_s(\tilde{x}_j), j = 0,1,\cdots,$ is a solution, $\tilde{\varphi}_s^i(\tilde{x}_j) = \tilde{x}_{j+i},\quad i,j = 0,1,\cdots$.
+To see why $\tilde{h}_s$ constructed in Eq. \ref{conjugate_tilted} satisfies the definition of a conjugacy map --
 $\tilde{h}_s(\tilde\varphi_s(x)) = \varphi(\tilde{h}_s(x))$ for all $x$ --
@@
      \caption{Upper-left: the tilted tent map (\ref{tent_tilted}) for $s=0, 0.1, 0.2, 0.3, 0.4,$ and 0.5.
-    Upper-right: the conjugate map $\tilde{h}_s$ between the tent map (\ref{tentmap})
-    and the tilted tent map (\ref{tent_tilted}), evaluated using
-    Equation (\ref{conjugate_tilted}-\ref{conjugate_tilted_helper}), for the same set of $s$ as the
+    Upper-right: the conjugate map $\tilde{h}_s$ between the tent map (Eq. \ref{tentmap})
+    and the tilted tent map (Eq. \ref{tent_tilted}), evaluated using
+    Eq. \ref{conjugate_tilted}-\ref{conjugate_tilted_helper}, for the same set of $s$ as the
     upper-left plot.}

### 2020-10-14 code/tent_shadow/tent_conjugate_helper.py
start to plot conjugate helper equation

@@ -1,0 +1,38 @@
+import numba
+from pylab import *
+from numpy import *
+
+@numba.jit(nopython=True)
+def find(x, y, s):
+    if x < 1:
+        return y / 2 * (1 + s)
+    else:
+        return 2 - y / 2 * (1 - s)
+
+@numba.jit(nopython=True)
+def shadow(n, s, width):
+    count = zeros(width, dtype=uint64)
+    x = random.rand() * 2
+    y = x
+    for i in range(100):
+        y = find(x, y, s)
+        x = x / 2 + random.randint(0,2)
+    for i in range(n):
+        y = find(x, y, s)
+        count[int(floor(y * width / 2))] += 1
+        x = x / 2 + random.randint(0,2)
+    return count
+
+n = 10000000000
+for s in [0.01, 0.05, 0.4, 0.7]:
+    width = 2**13
+    density = shadow(n, s, width) * width / 2 / n
+    x = linspace(0,2,width*2+1)[1:-1:2]
+    figure(figsize=(16,18))
+    fill_between(x, density)
+    ylim([0,1])
+    for tick in gca().xaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    for tick in gca().yaxis.get_major_ticks():
+        tick.label.set_fontsize(40)
+    savefig('tent_tilted_shadow_density_{}.png'.format(s))

### 2020-10-14 code/tent_shadow/tent_conjugate_helper.py
Plot conjugacy helpers. The nth conjugacy helper is activated at an x if varphi^n(x) is in the later half.

@@ -3,36 +3,38 @@ from pylab import *
 from numpy import *
 
 @numba.jit(nopython=True)
-def find(x, y, s):
-    if x < 1:
-        return y / 2 * (1 + s)
+def next(x, s):
+    if x < 1+s:
+        return 2*x/(1+s)
     else:
-        return 2 - y / 2 * (1 - s)
+        return 4/(1-s) - 2*x/(1-s)
 
 @numba.jit(nopython=True)
-def shadow(n, s, width):
-    count = zeros(width, dtype=uint64)
-    x = random.rand() * 2
-    y = x
-    for i in range(100):
-        y = find(x, y, s)
-        x = x / 2 + random.randint(0,2)
-    for i in range(n):
-        y = find(x, y, s)
-        count[int(floor(y * width / 2))] += 1
-        x = x / 2 + random.randint(0,2)
-    return count
+def xi(x, s, j):
+    if j==0 and x<1+s:
+        return 0
+    elif j==0 and x>=(1+s):
+        return 1
+    xj = x
+    for k in range(j):
+        xj = next(xj,s)
+    if xj < 1+s:
+        return xi(x, s, j-1)
+    return 1 - xi(x,s,j-1)
 
-n = 10000000000
-for s in [0.01, 0.05, 0.4, 0.7]:
-    width = 2**13
-    density = shadow(n, s, width) * width / 2 / n
-    x = linspace(0,2,width*2+1)[1:-1:2]
-    figure(figsize=(16,18))
-    fill_between(x, density)
-    ylim([0,1])
-    for tick in gca().xaxis.get_major_ticks():
-        tick.label.set_fontsize(40)
-    for tick in gca().yaxis.get_major_ticks():
-        tick.label.set_fontsize(40)
-    savefig('tent_tilted_shadow_density_{}.png'.format(s))
+
+n = 500
+x = linspace(0.,2, n)
+for s in [0, 0.05, 0.4, 0.7]:
+    fig = figure(figsize=(16,18))
+    ax = fig.add_subplot(111)
+    xij = zeros_like(x)
+    for j in range(5,6):
+        for k,xk in enumerate(x):
+            xij[k] = xi(xk, s, j)
+        ax.plot(x, xij, label=r'$\xi_{{{},{}}}$'.format(s,j), lw=2.5)
+        ax.xaxis.set_tick_params(labelsize=40)
+        ax.yaxis.set_tick_params(labelsize=40)
[…]

### 2020-10-14 comments.tex
Writing comment on interpretation of shadowing coding

@@
 \end{enumerate}
-
+\subsection*{Section 3.2}
+\begin{enumerate}
+    \item Given two nearby maps, how can we construct a conjugacy between their trajectories?
+    I think there is a Newton's method for conjugacies, can we explain this with an example?
+    Since finding conjugacies is not directly relevant to analyzing the physicality of shadowing, it may suffice to point to references. 
+    \item Can we give intuition for $\tilde{h}_s?$ The conjugacy helpers $\xi_{s, j}$
+    are bits in the binary representation of $\tilde{h}_s.$ Does this observation provide 
+    any intuition for the shadowing map? Additionally, $\xi_{s, j}$ is a recursive function that codes the orbits of $\tilde{\varphi}_s.$ The coding is as follows. If an orbit 
+    visits the interval $[1+s, 2]$ an even number of times upto and including time $j$,
+    then, $\xi_{s,j} =0;$ otherwise, it is one.
+Can $\tilde{h}_s(x)$ ever be $x$ at some point?
+This means that all the bits ($\xi_{s,j}(x)$) must be the same as the binary representation of $x$ itself.
+Also, $\tilde{h}_s$
+\end{enumerate}
 \end{document}

### 2020-10-14 main.tex
Writing comment on interpretation of shadowing coding

@@
     \centering
-    \includegraphics[width=0.32\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_15_28_2.67.json.png}
+    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_15_28_2.67.json.png}
     \hspace{0.005\textwidth}
-    \includegraphics[width=0.32\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_10_30_2.67.json.png}
+    \includegraphics[width=0.48\textwidth]{figure/baseline_density_lorenz_zmax_15_28_2.67.json.png}
+    \\
+    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_10_30_2.67.json.png}
     \hspace{0.005\textwidth}
-    \includegraphics[width=0.32\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_10_28_3.33.json.png}
+    \includegraphics[width=0.48\textwidth]{figure/baseline_density_lorenz_zmax_10_30_2.67.json.png}
+    \\
+    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_10_28_3.33.json.png}
+    \hspace{0.005\textwidth}
+    \includegraphics[width=0.48\textwidth]{figure/baseline_density_lorenz_zmax_10_28_3.33.json.png}
     \caption{

### 2020-10-14 code/lorenz_map/lorenz_map.py

@@ -12,21 +12,33 @@ def ddt(xyz):
     return array([dxdt, dydt, dzdt])
 
 @numba.jit(nopython=True)
-def step(xyz0):
+def step(xyz0_zint_tint):
+    xyz0 = xyz0_zint_tint[:3]
     ddt0 = ddt(xyz0)
     dt = 0.001
     xyz1 = xyz0 + 0.5 * dt * ddt0
     ddt1 = ddt(xyz1)
     xyz1 = xyz0 + dt * ddt1
+    zint = xyz0_zint_tint[3] + (xyz0[2] + xyz1[2]) / 2 * dt
+    tint = xyz0_zint_tint[4] + dt
     ddt1 = ddt(xyz1)
+    zmax = zeros(5)
     if ddt0[2] >= 0 and ddt1[2] < 0:
         if xyz0[2] > xyz1[2]:
-            zmax = xyz0.copy()
+            zmax[:3] = xyz0
+            zmax[3] = xyz0_zint_tint[3]
+            zmax[4] = xyz0_zint_tint[4]
+            zint = (xyz0[2] + xyz1[2]) / 2 * dt
+            tint = dt
         else:
-            zmax = xyz1.copy()
-    else:
-        zmax = zeros(3)
-    xyz0[:] = xyz1
+            zmax[:3] = xyz1
+            zmax[3] = zint
+            zmax[4] = tint
+            zint = 0
+            tint = 0
+    xyz0_zint_tint[3] = zint
+    xyz0_zint_tint[4] = tint
+    xyz0_zint_tint[:3] = xyz1
     return zmax
 
 @numba.jit(nopython=True)

### 2020-10-14 code/lorenz_map/plot.py

@@ -1,0 +1,15 @@
+from pylab import *
+from numpy import *
+import sys
+
+density = load(sys.argv[1])
+print(density.shape)
+width = 2100
+assert density.size == width
+x = linspace(29,50,width*2+1)[1:-1:2]
+
+figure(figsize=(16,18))
+fill_between(x, density)
+xticks(fontsize=40)
+yticks(fontsize=40)
+savefig('{}.png'.format(sys.argv[1].rsplit('.', maxsplit=1)[0]))

### 2020-10-14 code/lorenz_map/shadow.py

@@ -98,7 +98,7 @@ if __name__ == '__main__':
     
     zs0 = (zMax + zSep) / 2
     for i in range(99999, -1, -1):
-        fname = 'shadow_baseline_{}.npy'.format(i)
+        fname = 'shadow_baseline_{}_{}.npy'.format(sys.argv[0].rsplit('.', r)[0], i)
         if os.path.exists(fname):
             print(i)
             z = load(fname)

### 2020-10-14 code/lorenz_map/shadow_baseline.py

@@ -1,0 +1,76 @@
+import os
+import sys
+import json
+import numba
+from pylab import *
+from numpy import *
+from scipy.optimize import curve_fit
+
+params0 = json.load(open(sys.argv[1]))
+
+@numba.jit(nopython=True)
+def polyval(p, x):
+    y = 0
+    for pi in p:
+        y *= x
+        y += pi
+    return y
+
+@numba.jit(nopython=True)
+def fit0(dz, pwr):
+    return pow(dz * 1000, pwr)
+
+@numba.jit(nopython=True)
+def fit1(dz, pwr, p):
+    f0 = fit0(dz, pwr)
+    pz = polyval(p, dz)
+    return f0 + pz
+
+@numba.jit(nopython=True)
+def fit2(dz, pwr, p, a, b):
+    f1 = fit1(dz, pwr, p)
+    num = polyval(a, dz)
+    den = polyval(b, dz)
+    return f1 + num / den
+
+@numba.jit(nopython=True)
+def trajectory(z, n, zMax, zSep, pwr, pL, pR, aL, aR, bL, bR):
+    for i in range(100):
+        if z > zSep:
+            z = zMax - fit2(z - zSep, pwr, pR, aR, bR)
+        else:
+            z = zMax - fit2(zSep - z, pwr, pL, aL, bL)
+    
+    zT = zeros(n)
+    for i in range(n):
+        if z > zSep:
+            z = zMax - fit2(z - zSep, pwr, pR, aR, bR)
+        else:
+            z = zMax - fit2(zSep - z, pwr, pL, aL, bL)
+        zT[i] = z
+    return zT
+
+pwr = params0['pwr']
+zMin = params0['zMin']
+zMax = params0['zMax']
+zSep = params0['zSep']
+pL = array(params0['pL'])
+pR = array(params0['pR'])
+aL = array(params0['aL'])
+aR = array(params0['aR'])
+bL = array(params0['bL'])
+bR = array(params0['bR'])
+
+width = 2100
+x = linspace(29,50,width*2+1)[1:-1:2]
+density = zeros_like(x)
+
+n = 10000000
+z = (zSep + zMax) / 2
+for i in range(200):
+    z = trajectory(z, n, zMax, zSep, pwr, pL, pR, aL, aR, bL, bR)
+    # save('shadow_baseline_{}_{}.npy'.format(sys.argv[1].rsplit('.', 1)[0], i), z)
[…]

### 2020-10-14 main.tex
Edits up to proof of conjugacy.

@@
     \\ \vspace{0.02\textwidth}
-     \caption{Upper-left: the tilted tent map (\ref{tent_tilted}) for $s=0, 0.1, 0.2, 0.3, 0.4,$ and 0.5.
-    Upper-right: the conjugate map $\tilde{h}_s$ between the tent map (Eq. \ref{tentmap})
+     \caption{L: the tilted tent map Eq. \ref{tent_tilted} for $s=0, 0.1, 0.2, 0.3, 0.4,$ and 0.5. R: the conjugacy $\tilde{h}_s$ between the tent map (Eq. \ref{tentmap})
     and the tilted tent map (Eq. \ref{tent_tilted}), evaluated using
     Eq. \ref{conjugate_tilted}-\ref{conjugate_tilted_helper}, for the same set of $s$ as the
-    upper-left plot.}
+    left plot.}
      \label{fig:tilted_tent_conjugate}
@@
 or $\xi_{s,1}(x)=1-\xi_{s,0}(x)=1$ if $\tilde\varphi_s(x)\ge 1+s$.
-Thus, $\xi_{s,1}(x) = \xi_{s,0}(\varphi_s(x))$ according to the definition of $\xi_{s,0}$.
+Thus, $\xi_{s,1}(x) = \xi_{s,0}(\tilde{\varphi}_s(x))$ according to the definition of $\xi_{s,0}$.
 Using this as the base case, one can inductively verify that
 $\xi_{s,j+1}(x) = \xi_{s,j}(\tilde\varphi_s(x))$
-for all $j>0$, using just the definition of $\xi_{s,j}$ and $\xi_{s,j+1}$.
+for all $j>0$, using just the definitions of $\xi_{s,j}$ and $\xi_{s,j+1}$.
 Therefore,
@@
 or $1-\xi_{s,1}(x)=\xi_{s,0}(x)=1$ if $\tilde\varphi_s(x)\ge 1+s$.
-Thus, $1-\xi_{s,1}(x) = \xi_{s,0}(\varphi_s(x))$ according to the definition of $\xi_{s,0}$.
+Thus, $1-\xi_{s,1}(x) = \xi_{s,0}(\tilde{\varphi}_s(x))$ according to the definition of $\xi_{s,0}$.
 Using this as the base case, one can inductively verify that
 $1-\xi_{s,j+1}(x) = \xi_{s,j}(\tilde\varphi_s(x))$
-for all $j>0$, using just the definition of $\xi_{s,j}$ and $\xi_{s,j+1}$.
+for all $j>0$, using just the definitions of $\xi_{s,j}$ and $\xi_{s,j+1}$.
 Therefore,
@@
 $\tilde{h}_s(x) = 1+\sum_{j=1}^{\infty} \frac{\xi_{s,j}(x)}{2^j}\ge 1$,
-$\varphi(\tilde{h}_s(x)) = 2(1-\tilde{h}_s(x))$ according
+$\varphi(\tilde{h}_s(x)) = 4 - 2 \tilde{h}_s(x)$ according
 to the definition of $\varphi$.  Thus,
@@
    \caption{
-     Lower-left: the density of a trajectory satisfying (\ref{tentmap})
-    that shadows a random trajectory satisfying (\ref{tent_tilted}) for $s=0.1$.
-    Lower-right: the density of a trajectory satisfying (\ref{tentmap})
-    that shadows a random trajectory satisfying (\ref{tent_tilted}) for $s=0.5$.
+     L: the density of a trajectory satisfying Eq. \ref{tentmap}
+    that shadows a random trajectory satisfying Eq. \ref{tent_tilted} for $s=0.1$.
+    R: the density of a trajectory satisfying Eq. \ref{tentmap}
+    that shadows a random trajectory satisfying Eq. \ref{tent_tilted} for $s=0.5$.
     }
@@
 
-Figure \ref{fig:tilted_tent_density} shows how fractal the conjugate map is.
-This fractal conjugate map transforms a uniform distribution in $[0,2]$ into
+Figure \ref{fig:tilted_tent_density} shows how fractal the conjugacy map is.
+This fractal conjugacy map transforms a uniform distribution in $[0,2]$ into
 a fractal distribution similar to the ones plotted in Section \ref{sec:quasiphysical}.

### 2020-10-15 comments.tex
Edits up to proof of conjugacy.
Edits up to proof of shadowing solution being almost surely quasi physical

@@
     then, $\xi_{s,j} =0;$ otherwise, it is one.
-Can $\tilde{h}_s(x)$ ever be $x$ at some point?
-This means that all the bits ($\xi_{s,j}(x)$) must be the same as the binary representation of $x$ itself.
-Also, $\tilde{h}_s$
+    \item This should probably be addressed in 3.1, and relates to what is shadowed by what.
+    Can we add a line that says that the two trajectories that remain close are 
+    $\varphi^j \tilde{h}_s x$ and $\tilde{\varphi}^j_s x$?
 \end{enumerate}

### 2020-10-15 main.tex
Edits up to proof of shadowing solution being almost surely quasi physical

@@
 In fact, we will show that the fractal distribution obtained by mapping a uniform
-distribution under $\tilde{h}_s$ is exactly the family of distributions
+distribution through $\tilde{h}_s$ is exactly the family of distributions
 shown in Figure \ref{fig:tent_quasiphysical}.  We will also show that
-a physical solution of the tilted tent map (\ref{tent_tilted}) is uniformly
+a physical solution of the tilted tent map (Eq. \ref{tent_tilted}) is uniformly
 distributed in $[0,2]$.  Thus, for almost any physical solution of the
@@
 through the conjugate map $\tilde{h}_s$, has a fractal distribution.
-Such shadowing solution is therefore a quasi-physical solution of
+Such a shadowing solution is therefore a quasi-physical solution of
 the tent map.
@@
 any $0\le s<1$.
-According to ergodic theory, we only need to show that the uniform
+According to the ergodic theory and the properties of the SRB measure of 
+one-dimensional chaotic systems (\cite{young}), 
+we only need to show that the uniform
 distribution is stationary under the tilted tent map.
-If $x$ is uniformly distributed in $[0,2]$ with a density of 0.5,
-the left and right sides of the map, from $x=0$ to $x=1+s$ and
-from $x=1+s$ to 2 respectively, both has constant derivatives.
-The resulting density of $\tilde\varphi_s(x)$ is
-therefore uniform in $[0,2]$.
+The function $\tilde{\varphi}_s$ stretches an infinitesimal
+region around each $x$ by the absolute value of the derivative at $x$; 
+hence, a uniform density on an infinitesimal region around $\tilde{\varphi}_s(x)$ 
+is reduced by the same factor. In other words, let $\tilde{\rho}(x)$ 
+be the map of a uniform density on $[0,2]$, through $\tilde{\varphi}_s$. Then, 
+$\tilde{\rho}(\tilde{\varphi}_s (x)) = 0.5/|\tilde \varphi_s'(x)| + 0.5/|\tilde \varphi_s'(2-x)|,$ where the two terms express the conservation of probability 
+from the two pre-images of $\tilde{\varphi}_s$. Substituting the piecewise constant 
+derivative of $\tilde{\varphi}_s$ on the two intervals, $[0,1+s)$ and $[1+s,2]$,
+we have $\tilde{\rho}(\tilde{\varphi}_s (x)) = 0.25(1+s) + 0.25(1-s) = 0.5,$ at 
+all $x.$ Hence, a uniform density of 0.5 is unaltered by mapping through 
+$\tilde\varphi_s(x)$.
 
-We then show that the conjugate map $\tilde{h}_s$ maps a uniform
-distribution to a fractal.  Here we use the closed form
-(\ref{conjugate_tilted}-\ref{conjugate_tilted_helper}).
+We now show that the conjugacy $\tilde{h}_s$ maps a uniform
+distribution to a fractal.  Here we use the closed forms Eqs. \ref{conjugate_tilted}-\ref{conjugate_tilted_helper}.
 If $x$ is a random number drawn uniformly in $[0,2]$,
@@
 Thus, $\xi_{s,0}(x)=0$ with probability $\frac{1+s}2$.
-This means, according to Equation (\ref{conjugate_tilted}),
-$\tilde{h}_s(x)<1$ with probability $\frac{1+s}2$ --
+This means, according to Eq. \ref{conjugate_tilted},
+$\tilde{h}_s(x)<1$ with probability $\frac{1+s}2$:
 $\tilde{h}_s(x)$ is more likely to lie in the left half
@@
 Furthermore, for $j\ge 0$, $\xi_{s,j+1}(x)=\xi_{s,j}(x)$
-with probability $\frac{1+s}2$.  This probability has direct
-implication, again by Equation (\ref{conjugate_tilted}),
-to the binary decimal representation of $\tilde{h}_s(x)$ --
-each binary digit repeat the previous one with probability
-$\frac{1+s}2$.  Now we understand that for a uniformly random
+with probability $\frac{1+s}2$ since each $\tilde{\varphi}^j_s(x)$ is sampled
+from the uniform distribution on $[0,2]$.  This probability has direct
+implication on $\tilde{h}_s(x)$ since, again by Eq. (\ref{conjugate_tilted}),
+$\xi_{s,j}$ is the $j$th bit in the binary representation of $\tilde{h}_s$ --
+each bit repeats the previous one with probability
+$\frac{1+s}2$.  For a uniformly random
 $x$, $\tilde{h}_s(x)$ is exactly the kind of initial condition
@@
 in Section \ref{sec:quasiphysical}, with $p=\frac{1+s}2$.
-
-This analysis implies that the shadowing solution is
-almost surely nonphysical.  Let $x_{s,i},i=0,1,\ldots$ be a solution
+\nisha{This is so well set up and awesome!}
+Since this analysis holds for any $x$ sampled uniformly on $[0,2]$, we can conclude that
+the shadowing solution is almost surely nonphysical.  Let $x_{s,i},i=0,1,\ldots$ be a solution
 to the tilted tent map with an initial condition randomly chosen
@@
     \centering
-    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_15_28_2.67.json.png}
+    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_density_lorenz_zmax_15_28_2.67.json.png}
     \hspace{0.005\textwidth}
@@
     \\
-    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_10_30_2.67.json.png}
+    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_density_lorenz_zmax_10_30_2.67.json.png}
     \hspace{0.005\textwidth}
@@
     \\
-    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_lorenz_shadow_density_lorenz_zmax_10_28_3.33.json.png}
+    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_density_lorenz_zmax_10_28_3.33.json.png}
     \hspace{0.005\textwidth}

### 2020-10-15 code/lorenz_map/plot.py

@@ -1,15 +1,0 @@
-from pylab import *
-from numpy import *
-import sys
-
-density = load(sys.argv[1])
-print(density.shape)
-width = 2100
-assert density.size == width
-x = linspace(29,50,width*2+1)[1:-1:2]
-
-figure(figsize=(16,18))
-fill_between(x, density)
-xticks(fontsize=40)
-yticks(fontsize=40)
-savefig('{}.png'.format(sys.argv[1].rsplit('.', maxsplit=1)[0]))

### 2020-10-15 code/lorenz_map/shadow.py

@@ -98,7 +98,7 @@ if __name__ == '__main__':
     
     zs0 = (zMax + zSep) / 2
     for i in range(99999, -1, -1):
-        fname = 'shadow_baseline_{}_{}.npy'.format(sys.argv[0].rsplit('.', r)[0], i)
+        fname = 'data/shadow_baseline_{}_{}.npy'.format(sys.argv[1].rsplit('.', 1)[0], i)
         if os.path.exists(fname):
             print(i)
             z = load(fname)
@@ -108,4 +108,4 @@ if __name__ == '__main__':
             zs0 = zs[0]
             density += histogram(zs, linspace(29,50,width+1))[0]
     
-    save('lorenz_shadow_density_{}.npy'.format(sys.argv[2]), density)
+    save('lorenz_shadow_density_{}.npy'.format(sys.argv[2]).rsplit('.', 1)[0], density)

### 2020-10-15 code/lorenz_map/shadow_baseline.py

@@ -69,7 +69,7 @@ n = 10000000
 z = (zSep + zMax) / 2
 for i in range(200):
     z = trajectory(z, n, zMax, zSep, pwr, pL, pR, aL, aR, bL, bR)
-    # save('shadow_baseline_{}_{}.npy'.format(sys.argv[1].rsplit('.', 1)[0], i), z)
+    #save('data/shadow_baseline_{}_{}.npy'.format(sys.argv[1].rsplit('.', 1)[0], i), z)
     density += histogram(z, linspace(29,50,width+1))[0]
     z = z[-1]
 

### 2020-10-15 main.tex
Add figure captions for the tent maps.

@@
 the shadowing solution is almost surely nonphysical.  Let $x_{s,i},i=0,1,\ldots$ be a solution
-to the tilted tent map with an initial condition randomly chosen
-in $[0,1]$.  Then with a hundred percent probability
+to the tilted tent map $\tilde{\varphi}_s$ with $x_{s,0}$ chosen 
+randomly in $[0,1+s]$.  Then, with a hundred percent probability,
 it is a physical solution that uniformly visits the domain $[0,2]$.
@@
 is a quasi-physical solution, also with a hundred percent probability.
-It explores $[0,2]$ at a nonuniform frequency with a fractal density function.
+It explores $[0,2]$ at a nonuniform frequency with a fractal probability distribution.
 
@@
 In the previous subsection, we examined a tilting perturbation to the tent map.
-We derived the fractal conjugate map, and showed that a typical shadowing trajectory
+We derived the fractal conjugacy map, and showed that a typical shadowing trajectory
 is exactly the quasi-physical solution we analyzed in Section \ref{sec:quasiphysical}.
-This example contrast with Section \ref{sec:sub:shadow-example}, in which the
+This example contrasts with Section \ref{sec:sub:shadow-example}, in which the
 shadowing solutions of a different perturbation, the scaling perturbation,
@@
     \includegraphics[width=0.48\textwidth]{figure/squashed_tent_map.png}
-    \caption{The ``squashed'' tent map (\ref{eqn:tent_squashed}) for $s=0.01, 0.05, 0,2,$ and 0.5.}
+    \caption{The ``squashed'' tent map (Eq. \ref{eqn:tent_squashed}) for $s=0.01, 0.05, 0,2,$ and 0.5.}
     \label{fig:tent_squashed}
@@
 of the tent map.  The resulting ``squashed'' tent map, illustrated
-in Figure \ref{fig:tent_squashed}, has a closed form of
+in Figure \ref{fig:tent_squashed}, has the closed form 
 \begin{equation} \label{eqn:tent_squashed}
-    \widehat{\varphi^{sq}}_s(x) := \begin{cases}
+\hat{\varphi}^{\rm sq}_s(x) := \begin{cases}
     (2-s)x \quad & x < 1 \\
@@
     \includegraphics[width=0.48\textwidth]{figure/tent_squashed_physical_density_0.2.png}
-    \caption{Caption}
+    \caption{The left column shows the empirical distribution of the shadowing solution for the squashed tent map at $s=0.01$ (top) and at $s=0.2$ (bottom).
+The right column shows the probability distribution of a physical solution at the same two values of $s = 0.01$ (top) and $s = 0.2$ (bottom).}
     \label{fig:tent_squashed_shadow}
@@
     \includegraphics[width=0.48\textwidth]{figure/pinched_tent_map.png}
-    \caption{Caption}
+    \caption{The pinched tent map (Eq. \ref{eqn:tent_pinched}) for $s = 0.01, 0.05, 0.2,$ and 0.5.}
     \label{fig:tent_pinched}
@@
     \includegraphics[width=0.48\textwidth]{figure/tent_pinched_physical_density_0.2.png}
-    \caption{Caption}
+    \caption{The left column shows the empirical distribution of the shadowing solution for the pinched tent map at $s=0.05$ (top) and at $s=0.2$ (bottom).
+The right column shows the probability distribution of a physical solution at the same two values of $s = 0.05$ (top) and $s = 0.2$ (bottom).}
     \label{fig:tent_pinched_shadow}
@@
     \includegraphics[width=0.48\textwidth]{figure/wave_tent_map.png}
-    \caption{Caption}
+    \caption{The wave tent map (Eq. \ref{eqn:tent_wave}) for $s = 0.01, 0.05, 0.2,$ and 0.5.}
     \label{fig:tent_wave}
@@
     \includegraphics[width=0.48\textwidth]{figure/tent_wave_physical_density_0.2.png}
-    \caption{Caption}
+    \caption{The left column shows the empirical distribution of the shadowing solution for the wave tent map (Eq.
+\ref{eqn:tent_wave}) at $s=0.05$ (top) and at $s=0.2$ (bottom).
+The right column shows the probability distribution of a physical solution at the same two values of $s = 0.05$ (top) and $s = 0.2$ (bottom).}
     \label{fig:tent_wave_shadow}

### 2020-10-15 main.tex
Add squashed tent explanation and pinched tent equation

@@
 \end{figure}
-
+In Figure \ref{fig:tent_squashed_shadow}, we show side-by-side the probability distributions generated by observing a long solution (of length 10 billion) 
+of $\hat{\varphi}^{\rm sq}_s$ (left) and its corresponding shadowing solution (right)
+at two different values of $s.$ The left column showing the probability distribution of the shadowing solution is reminiscent of the hairy probability distributions of 
+the quasi-solutions of the tent map in Section \ref{sec:quasiphysical}.
+Note that, unlike the tilted tent maps, the family of squashed tent maps do not have the uniform distribution as their SRB distribution, as indicated on the right column.
+However, the SRB distributions are not fractal.
+That is, the physical solutions of the squashed tent map are distributed according to a regular probability density (right column of Figure \ref{fig:tent_squashed_shadow}).
+But the shadowing solution, which is a true orbit of the original tent map, is neither a physical solution of the original tent map nor of the squashed tent map, as suggested by its fractal-like empirical probability distribution.
+
+Next we consider a second perturbation of the tent map, the \emph{pinched} tent map, which has the following closed form
+\begin{align}
+    \hat{\varphi}^p_s(x) = \begin{cases}
+        \dfrac{4x}{1 + s + \sqrt{(1+s)^2 - 4s x} }, & x < 1 \\
+        \dfrac{4(2-x)}{1 + s + \sqrt{(1+s)^2 - 4s(2- x)} }, & 2 \leq x \leq 1.
+    \end{cases}
+    \label{eqn:tent_pinched}
+\end{align}
+Again, at $s=0$, the original tent map is recovered, and other values of $s,$ the tent map is ``pinched'' by perturbations that are symmetric around $x = 1$, and zero at the end points.
+Figure \ref{fig:tent_pinched} shows the map at different $s$ values.
 \begin{figure}[H]
@@
 \end{figure}
-
+We consider a second perturbation of the tent map with the closed form
+\begin{align}
+    \label{eqn:tent_wave}
+\end{align}
 \begin{figure}[H]

### 2020-10-15 main.tex
Start on the wave tent map

@@
 That is, the physical solutions of the squashed tent map are distributed according to a regular probability density (right column of Figure \ref{fig:tent_squashed_shadow}).
-But the shadowing solution, which is a true orbit of the original tent map, is neither a physical solution of the original tent map nor of the squashed tent map, as suggested by its fractal-like empirical probability distribution.
+But the shadowing solution, which is a true orbit of the original tent map, is distributed neither like a physical solution of the original tent map nor of the squashed tent map, as suggested by its fractal-like empirical probability distribution.
 
@@
 Again, at $s=0$, the original tent map is recovered, and other values of $s,$ the tent map is ``pinched'' by perturbations that are symmetric around $x = 1$, and zero at the end points.
-Figure \ref{fig:tent_pinched} shows the map at different $s$ values.
+Figure \ref{fig:tent_pinched} shows the pinched tent map at different $s$ values.
 \begin{figure}[H]
@@
 \end{figure}
-
+In Figure \ref{fig:tent_pinched_shadow}, we show the corresponding plots of the 
+physical and shadowing distributions for the pinched tent map. From the right 
+column, which shows the probability density of a long physical solution, we 
+can see that the pinching perturbation changes the uniform density of the original tent map to a linearly varying density.
+The more pronounced the perturbation -- the higher the value of $s$ -- the steeper decline in the density from left to the right.
+The shadowing distribution, in this case as well, appears fractal.
+Once again, with probability one -- or, for any randomly chosen initial condition of the solution of the pinched tent map -- any physical solution is shadowed by a quasi-physical solution.
 \begin{figure}[H]
@@
 \end{figure}
-We consider a second perturbation of the tent map with the closed form
+We consider yet another perturbation of the tent map with the closed form
 \begin{align}

### 2020-10-15 main.tex
Add new json files: Qiqi

@@
 \begin{align}
+\hat{\varphi}^w_s(x) = \begin{cases}
+        \dfrac{4x}{1 + s + \sqrt{(1+s)^2 - 4s x} }, & x < 1 \\
+        \dfrac{4(2-x)}{1 - s + \sqrt{(1-s)^2 - 4s(2- x)} }, & 2 \leq x \leq 1.
+    \end{cases}
     \label{eqn:tent_wave}
 \end{align}
+The above map, called the wave tent map, is illustrated in Figure \ref{fig:tent_wave}.
+To construct the pinched tent map (Eq.
+\ref{eqn:tent_pinched}), we perturbed the original tent map in such a way that the perturbations on the two halves of the interval $[0,2]$ are mirror images of each other about the $x=1$ line.
+In the wave tent map, the perturbations on the two halves are a reflection of each other about the
+$x$-axis.
 \begin{figure}[H]
@@
 \end{figure}
-
+The wave perturbation does not noticeably alter the SRB density of the original tent map. As we show in Figure \ref{fig:tent_wave_shadow} (right), a solution of the 
+wave tent map, starting from almost any point on $[0,2]$ -- a physical solution -- has a density that looks almost identical to a uniform density of 0.5.
+On the left column of Figure \ref{fig:tent_wave_shadow}, we observe that the distributions of the shadowing solutions corresponding to the physical solutions on the right, once again appear fractal.
 \begin{figure}[H]
@@
 \end{figure}
+Hence, we can once again conclude that almost every solution, which is physical, 
+has a corresponding shadowing solution, that is non-physical. In this section,
+we have shown that the same conclusion holds for several different perturbations 
+of the tent map. The physical solutions were distributed nonuniformly 
+and differently from one another.
+However, we observed a commonality in the different perturbations: the shadowing solutions corresponding to almost every physical solution had a fractal-like probability distribution that did not resemble the distributions of the physical solutions in any case.
+Thus, we have shown significant evidence, through analytical constructions of perturbed tent maps in this section and
+section \ref{sec:quasiphysical} that shadowing solutions can almost surely be nonphysical.
 

### 2020-10-15 data/plot_lorenz_map_density.py

@@ -1,0 +1,20 @@
+from pylab import *
+from numpy import *
+import glob
+import sys
+
+for fname in sys.argv[1:]:
+    density = load(fname)
+    print(density.shape)
+    width = 2100
+    assert density.size == width
+    x = linspace(29,50,width*2+1)[1:-1:2]
+    density /= density.sum() * (x[1] - x[0])
+    
+    figure(figsize=(16,18))
+    fill_between(x, density)
+    xlim([29, 50])
+    ylim([0, 0.18])
+    xticks(fontsize=40)
+    yticks(fontsize=40)
+    savefig('{}.png'.format(fname.rsplit('.', maxsplit=1)[0]))

### 2020-10-18 main.tex
start lorenz. explain why we need to make an approximation
Edit until caption of Lorenz map

@@
 \subsection{What about the Lorenz equation?}
-
+In order to analyze the behavior of the shadowing solutions of the Lorenz'63 system 
+of equations, we first make a closed form approximation of the Lorenz map. The reason 
+is that except at rare parameter values, the Lorenz system does not have 
+solutions that shadow perturbed orbits for all time, but rather only for a finite time 
+\cite{lorenzPOTP}. Hence, the Lorenz map also does not have infinitely long shadowing solutions. This implies that a numerical solution of the Lorenz map approximates 
+a true solution, as accurately as desired, only for a finite time.
+
+A closed form approximation is constructed by regression performed using long solutions 
+of the Lorenz map (see Figure \ref{fig:lorenz_map}).
+The cusp, around $(39,48)$, is fitted with an exponential function, and the tails are fitted with a sum of a cubic polynomial and a rational function (see supplementary material for the fitting parameters).
 \begin{figure}[H]
@@
     In the left plot, the blue, orange, and green lines represent
-    $\sigma=10,11$, and $12$, respectively, while $\rho$ and $\beta$ stay
-    at 28 and $8/3$.
+    $\sigma=10,11$, and $12$, respectively, while $\rho  = 28$ and $\beta = 8/3.$
     In the center plot, the blue, orange, and green lines represent
-    $\rho=28,29$ and $30$, respectively, while $\sigma$ and $\beta$ stays at 10
-    and $8/3$.
+    $\rho=28,29$ and $30$, respectively, while $\sigma = 10$ and $\beta  = 8/3.$ 
     In the right plot, the blue, orange, and green lines represent
-    $\sigma=8/3,3$ and $10/3$ respectively, while $\sigma$ and $\rho$ stays at 10
-    and $28$.}
+    $\sigma=8/3,3$ and $10/3$ respectively, while $\sigma = 10$ and $\rho = 28.$\nisha{Is the green line overlapping one of the two lines?}}
     \label{fig:lorenz_params}
 \end{figure}
-
+The Lorenz map approximation obtained this way is illustrated in Figure \ref{fig:lorenz_params}, for variations of the parameters around their standard values of 
+$\rho = 28, \sigma = 10, \beta = 8/3.$ For a range of parameter values, it is numerically verified that the approximate Lorenz map qualitatively reproduces the behavior of the true Lorenz map.
+Using the approximate Lorenz map, we calculate next the shadowing solutions.
 \begin{figure}[H]
@@
     \includegraphics[width=0.48\textwidth]{figure/baseline_density_lorenz_zmax_10_28_3.33.json.png}
-    \caption{
+    \caption{The left column shows the empirical distribution of the shadowing solution of the Lorenz map at $\rho = 28$ (top row), $\rho = 29$ (middle row), and $\rho = 30$ (bottom row).
+The right column shows the probability distribution of a physical solution at the corresponding $\rho$ values of each row.
+The other two parameters are held at their standard values of $\sigma = 10, \beta = 8/3.$
     }
@@
 \end{figure}
-----------------------------
 

### 2020-10-19 code/lorenz_map/shadow.py
remove two unnecessary lines in the shadow function

@@ -28,6 +28,7 @@ def fit1(dz, pwr, p):
 
 @numba.jit(nopython=True)
 def fit2(dz, pwr, p, a, b):
+    print(dz)
     f1, f1grad = fit1(dz, pwr, p)
     num, numgrad = polyval(a, dz)
     den, dengrad = polyval(b, dz)
@@ -65,10 +66,8 @@ def shadow(zTraj, zSep0, z0, zMin, zMax, zSep, pwr, pL, pR, aL, aR, bL, bR):
     dzMaxR = zSep - zMin
     for i in range(zTraj.size-1,-1,-1):
         if zTraj[i] > zSep0:
-            dz0 = max(0.002, zTraj[i] - zSep0)
             z = zSep + find2(zMax - z, dzMinR, dzMaxR, pwr, pR, aR, bR)
         else:
-            dz0 = max(0.002, zSep0 - zTraj[i])
             z = zSep - find2(zMax - z, dzMinL, dzMaxL, pwr, pL, aL, bL)
         zShadow[i] = z
         if not isfinite(z):

### 2020-10-19 code/lorenz_map/shadow_baseline.py

@@ -60,12 +60,12 @@ aL = array(params0['aL'])
 aR = array(params0['aR'])
 bL = array(params0['bL'])
 bR = array(params0['bR'])
-
+'''
 width = 2100
 x = linspace(29,50,width*2+1)[1:-1:2]
 density = zeros_like(x)
 
-n = 10000000
+n = 10
 z = (zSep + zMax) / 2
 for i in range(200):
     z = trajectory(z, n, zMax, zSep, pwr, pL, pR, aL, aR, bL, bR)
@@ -74,3 +74,4 @@ for i in range(200):
     z = z[-1]
 
 save('baseline_density_{}.npy'.format(sys.argv[1]), density)
+'''

### 2020-10-19 supp.tex
Start editing supplementary material

\section{Supplementary material}
In this section, we discuss how to numerically compute the shadowing trajectories of the Lorenz map. One of the motivations for approximating the Lorenz map is that having a closed form expression is necessary in the algorithm we discuss for numerical shadowing solution computation. The program used to generate Figure 
\ref{fig:lorenz_map} is in

### 2020-10-19 main.tex
Add grebogi citations

@@
 The existence of arbitrarily long shadowing solutions for sufficiently similar governing equations
-has been proven for invertible hyperbolic maps by Anosov (1967) and Bowen (1970, 1978).
+has been proven for invertible hyperbolic maps by Anosov \cite{anosov} and Bowen \cite{bowen}.
 For non-hyperbolic systems, it is often possible to estimate the length of time for which shadowing
-solutions exist. (Grebogi, Hammel, Yorke and Sauer (1990), Sauer, Grebogi and Yorke (1997)).
+solutions exist \cite{grebogi, hammel}.
 
@@
 % give example of numerics that destroys physicalness
+
 \bibliographystyle{spmpsci}
 \bibliography{refs.bib}
+\include{supp}
 \end{document}

### 2020-10-19 supp.tex
Add grebogi citations

@@
 \section{Supplementary material}
+\subsection{The Lorenz Map}
+\subsection{Approximation of the Lorenz Map}
+
+\subsubsection{Computing shadowing solution of the Lorenz map}
 In this section, we discuss how to numerically compute the shadowing trajectories of the Lorenz map.
@@
 The program used to generate Figure
-\ref{fig:lorenz_map} is in
+\ref{fig:lorenz_params_scaled} is in \verb+lorenz_map/shadow.py+.
+The main logic behind the numerical procedure to compute the shadowing solution can be found in that script, within the function \verb+shadow+, which calls \verb+find2+.

### 2020-10-20 supp.tex
Writing supplementary material.

@@
 The main logic behind the numerical procedure to compute the shadowing solution can be found in that script, within the function \verb+shadow+, which calls \verb+find2+.
+Let $x_n, n = 0,1,2,3\cdots,N$ be a given perturbed solution given to us until time $N$.
+The goal is to find, at every time $n,$ a perturbation $\epsilon_n$ such that $y_n = x_n + \epsilon_n$ is a shadowing solution.
+This means that $y_n$ must satisfy the governing equation: $y_{n+1} = \varphi(y_n),$ where $\varphi$ is the Lorenz map, and, $y_n$ and $x_n$ must stay
+close to each to other for the length of time $N$. We start by initialized $y_N = x_N$ and find $\epsilon_n$ backward in time.

### 2020-10-20 supp.tex
Attempting to understand Qiqi's shadowing procedure.

--- 2020-10-19 supp.tex
@@
 The main logic behind the numerical procedure to compute the shadowing solution can be found in that script, within the function \verb+shadow+, which calls \verb+find2+.
+Let $x_n, n = 0,1,2,3\cdots,N$ be a given perturbed solution given to us until time $N$.
+The goal is to find a sequence $\left\{ y_n\right\}_{n=0}^{N}$ that is a shadowing solution.
+This means that $y_n$ must satisfy the governing equation: $y_{n+1} = \varphi(y_n),$ where $\varphi$ is the Lorenz map, and,
+$y_n$ and $x_n$ must stay close to each to other for the length of time, $N$.
+We find the sequence $y_n$ backward in time.
+Let $z_{\rm sep,s}$ be the $z$-location of the cusp of the perturbed Lorenz map that generated $x_n$; to indicate the original map, we omit the $s$ in the subscript.
+We split the domain into two regions, $[z_{\rm min}, z_{\rm sep}]$ and $(z_{\rm sep}, z_{\rm max}],$ and
+obtain, for $n = N-1, \cdots, 0,$
+\begin{align}
+    y_n = \begin{cases}
+                     z_{\rm sep} + f(y_{n+1}), & x_n > z_{\rm sep, s} \\
+                     z_{\rm sep} - f(y_{n+1}), & x_n \leq z_{\rm sep, s}.
+    \end{cases}
+    \label{eqn:shadowingFormula}
+\end{align}
+Here the function $f$ is implemented in 
+\verb+find2+. Inside \verb+find2+, given a $y$, 
+a recursive procedure is used to 
+find a number $v_k = f(y_{n+1})$ such that if $y_n$ set 
+using Eq. \ref{eqn:shadowingFormula} is (approximately) the shadowing 
+solution. At $n = N-1,$ $y_N$ is initialized with a random guess close to $x_N.$
+Then, the following recursive equation is solved to compute $f(y_{n+1})$.
+\begin{align}
+        v_{k+1} = v_k \;  \exp{\big({y_{n+1} - \varphi(y_{n,k})}\big)} \; \exp{\Big(\dfrac{-1}{\varphi'(y_{n,k}) v_k}\Big)}, \; k = 0,1,\cdots,
+\end{align}
+At $k=0,$ we randomly assign $v_0 = 1,$ and continue the above iteration 
+until $|v_k|$, and $|y_{n+1} - \varphi(y_{n,k})|$ are both less than 
+a tolerance. When the iteration converges, at some $k$,
+$f_i(y_{n+1}) = v_k.$ Notice that by construction, the
+sequence $\left\{ y_n\right\}$ computed this way is a shadowing solution.

### 2020-10-20 main.tex
Writing supplementary material.
Add table of fit parameters values

@@
 For non-hyperbolic systems, it is often possible to estimate the length of time for which shadowing
-solutions exist \cite{grebogi, hammel}.
+solutions exist \cite{grebogi}\cite{hammel}.
 
@@
 In order to analyze the behavior of the shadowing solutions of the Lorenz'63 system 
-of equations, we first make a closed form approximation of the Lorenz map. The reason 
+of equations, we first make a closed form approximation of the Lorenz map (see section \ref{sec:lorenzMapApprox} of the supplementary material). The reason 
 is that except at rare parameter values, the Lorenz system does not have 
@@
 of the Lorenz map (see Figure \ref{fig:lorenz_map}). 
-The cusp, around $(39,48)$, is fitted with an exponential function, and the tails are fitted with a sum of a cubic polynomial and a rational function (see supplementary material for the fitting parameters).
 \begin{figure}[H]

### 2020-10-20 supp.tex
Add table of fit parameters values

@@
 \subsection{The Lorenz Map}
-\subsection{Approximation of the Lorenz Map}
+\subsubsection{Approximation of the Lorenz Map}
+The motivation for approximating the Lorenz map is that a 
+closed form expression for the map is necessary for our numerical shadowing procedure.
+In a small region around the cusp, we approximate the map using an exponential function.
+The tails on both sides are fitted with a sum of a cubic polynomial and a rational function.
+Thus, the approximate Lorenz map has the
+following closed form expression:
+\begin{align}
+    \varphi(x) = 
+\end{align}
+\begin{table}[H]
+    \centering
+    \begin{tabular}{|c|c|}
+    \hline 
+         $z_{\rm sep}$ = 38.55302437476555 &
+         $z_{\rm min}$ =  29.213182255013322 \\ 
+         $z_{\rm max}$ =  47.978140718671284 & 
+         $p$ = 0.28796740575434676   \\
+         $p_{L,1}$ = -0.00024683786275242047 & $p_{L,2} = $ 
+         0.016174566354858824 \\
+         $p_{L,3} = $ 0.40179772568004946 & 
+         $p_{L, 4}  = $ -0.24612651488351725 \\
+         $p_{R,1} = $ -0.00020712463321688308 & 
+         $p_{R,2} = $ 0.017130843276711716 \\
+         $p_{R, 3} = $ 0.3930080703420676 & 
+         $p_{R, 4} = $ -0.23471384266765036 \\
+         $a_{L,1}$ = -0.05405742075580959 & $a_{L,2}$ = -0.05405742075580959 \\
+         $a_{R,1}$ = -0.05351127783496397 & $a_{R,2}$ = 0.22489891059122896 \\
+         $b_{L,1}$ = 0.5609397451213353 & $b_{L,2}$ = -0.3491184293228338 \\  $b_{L,3}$ = 2.419972619058592 & $b_{L,4}$ = 1.0 \\
+         $b_{R,1}$ = 0.6456076059873844 & $b_{R,2}$ = -0.34840383986411055 \\ $b_{R,3}$ = 2.6035438510917692 &  $b_{R,4}$ = 1.0 \\
+         \hline
+    \end{tabular}
+    \caption{The fitting parameters of the Lorenz map at $\sigma = 10, \beta = 8/3$ and $\rho = 28.$}
+    \label{tab:my_label}
+\end{table}
 
-\subsubsection{Computing shadowing solution of the Lorenz map}
-In this section, we discuss how to numerically compute the shadowing trajectories of the Lorenz map.
-One of the motivations for approximating the Lorenz map is that having a closed form expression is necessary in the algorithm we discuss for numerical shadowing solution computation.
-The program used to generate Figure
+\subsubsection{Computing shadowing solutions of the Lorenz map}
+In this section, we discuss how to numerically compute the shadowing trajectories of the Lorenz map.  The program used to generate Figure 
 \ref{fig:lorenz_params_scaled} is in \verb+lorenz_map/shadow.py+.
 The main logic behind the numerical procedure to compute the shadowing solution can be found in that script, within the function \verb+shadow+, which calls \verb+find2+.
-Let $x_n, n = 0,1,2,3\cdots,N$ be a given perturbed solution given to us until time $N$.
+Let $x_n, n = 0,1,2,3,\cdots,N$ be a perturbed solution given to us until time $N$.
 The goal is to find a sequence $\left\{ y_n\right\}_{n=0}^{N}$ that is a shadowing solution.
@@
 Here the function $f$ is implemented in 
-\verb+find2+. Inside \verb+find2+, given a $y$, 
+\verb+find2+. Inside \verb+find2+, given a $y_{n+1}$, 
 a recursive procedure is used to 
-find a number $v_k = f(y_{n+1})$ such that if $y_n$ set 
+find a number $v_k = f(y_{n+1})$ such that $y_n$ set 
 using Eq. \ref{eqn:shadowingFormula} is (approximately) the shadowing 
@@
 a tolerance. When the iteration converges, at some $k$,
-$f_i(y_{n+1}) = v_k.$ Notice that by construction, the
-sequence $\left\{ y_n\right\}$ computed this way is a shadowing solution.
+$f(y_{n+1}) = v_k.$ Notice that by construction, the
+sequence $\left\{ y_n\right\}$ computed this way solves the governing equation.
+To conclude that the sequence approximates the shadowing solution, we only need to check that $x_n$ and $y_n$ are close to each other for all times up to $N.$

### 2020-10-20 main.tex
correct error in plot caption

@@
 $\rho = 28, \sigma = 10, \beta = 8/3.$ For a range of parameter values, it is numerically verified that the approximate Lorenz map qualitatively reproduces the behavior of the true Lorenz map.
-Using the approximate Lorenz map, we calculate next the shadowing solutions.
+Using the approximate Lorenz map, we calculate next the shadowing solutions (see the supplementary material section \ref{sec:lorenzShadowingProcedure} for the numerical procedure).
 \begin{figure}[H]
@@
     \includegraphics[width=0.48\textwidth]{figure/baseline_density_lorenz_zmax_10_28_3.33.json.png}
-    \caption{The left column shows the empirical distribution of the shadowing solution of the Lorenz map at $\rho = 28$ (top row), $\rho = 29$ (middle row), and $\rho = 30$ (bottom row).
-The right column shows the probability distribution of a physical solution at the corresponding $\rho$ values of each row.
-The other two parameters are held at their standard values of $\sigma = 10, \beta = 8/3.$
+    \caption{The left column shows the empirical distribution of the shadowing solution of the Lorenz map at $\rho = 28$ (top row), $\rho = 29$ (middle row), and $\beta = 10/3$ (bottom row).
+The right column shows the probability distribution of the corresponding physical solution.
+In each case, the parameters not mentioned are held at their standard values of $\rho = 28, \sigma = 10, \beta = 8/3.$
     }
@@
 \end{figure}
+The empirical densities computed from long shadowing solutions, of length 10 billion, are shown in the left column of Figure \ref{fig:lorenz_params_scaled},
 

### 2020-10-20 supp.tex
correct error in plot caption

@@
 \begin{align}
-    \varphi(x) = 
+    \varphi(x) = \begin{cases}
+        z_{\rm max} - f_R(x - z_{\rm sep})   &  x > z_{\rm sep} \\
+        z_{\rm max} -  f_L(z_{\rm sep} - x)   &  x \leq z_{\rm sep}
+    \end{cases}
 \end{align}
+where 
+\begin{align}
+    f_R(y) = \Big( (1000 y)^p + 
+        \sum_{n=0}^3 p_{R,n} y^n + \dfrac{a_{L,0} + a_{L,1} y}{\sum_{n=0}^3 b_{L,n} y^n}\Big),
+\end{align}
+and,
+\begin{align}
+    f_L(y) = \Big((1000 y)^p 
+        + \sum_{n=0}^3 p_{L,n} y^n + \dfrac{a_{R,0} + a_{R,1}y}{\sum_{n=0}^3 b_{R,n} y^n}\Big).
+\end{align}
+The location of the cusp is denoted $z_{\rm sep}$ (in which ``sep'' stands 
+for \emph{separation}), and the maximum and minimum values encountered 
+in the Lorenz map iterates are denoted $z_{\rm max}$ and $z_{\rm min}$, respectively.
+The exponent of the cusp is denoted $p$.
+The coefficients of the cubic polynomial modelling the left (right) tail are  denoted $p_{L,n}, n = 0,1,2,3$ ($p_{R,n}, n = 0,1,2,3$).
+The coefficients of the numerator and denominator of the rational function modelling the left (right) tail are denoted $a_{L,0}, a_{L,1}$ ($a_{R,0}, a_{R,1}$) and $b_{L,n}, n=0,1,2,3$ ($b_{R,n}, n=0,1,2,3$), respectively.
+The values of these coefficients are obtained by
+regression on 
 \begin{table}[H]
@@
          $p$ = 0.28796740575434676   \\
-         $p_{L,1}$ = -0.00024683786275242047 & $p_{L,2} = $ 
+         $p_{L,3}$ = -0.00024683786275242047 & $p_{L,2} = $ 
          0.016174566354858824 \\
-         $p_{L,3} = $ 0.40179772568004946 & 
-         $p_{L, 4}  = $ -0.24612651488351725 \\
-         $p_{R,1} = $ -0.00020712463321688308 & 
+         $p_{L,1} = $ 0.40179772568004946 & 
+         $p_{L, 0}  = $ -0.24612651488351725 \\
+         $p_{R,3} = $ -0.00020712463321688308 & 
          $p_{R,2} = $ 0.017130843276711716 \\
-         $p_{R, 3} = $ 0.3930080703420676 & 
-         $p_{R, 4} = $ -0.23471384266765036 \\
-         $a_{L,1}$ = -0.05405742075580959 & $a_{L,2}$ = -0.05405742075580959 \\
-         $a_{R,1}$ = -0.05351127783496397 & $a_{R,2}$ = 0.22489891059122896 \\
-         $b_{L,1}$ = 0.5609397451213353 & $b_{L,2}$ = -0.3491184293228338 \\  $b_{L,3}$ = 2.419972619058592 & $b_{L,4}$ = 1.0 \\
-         $b_{R,1}$ = 0.6456076059873844 & $b_{R,2}$ = -0.34840383986411055 \\ $b_{R,3}$ = 2.6035438510917692 &  $b_{R,4}$ = 1.0 \\
+         $p_{R, 1} = $ 0.3930080703420676 & 
+         $p_{R, 0} = $ -0.23471384266765036 \\
+         $a_{L,1}$ = -0.05405742075580959 & $a_{L,0}$ = -0.05405742075580959 \\
+         $a_{R,1}$ = -0.05351127783496397 & $a_{R,0}$ = 0.22489891059122896 \\
+         $b_{L,3}$ = 0.5609397451213353 & $b_{L,2}$ = -0.3491184293228338 \\  $b_{L,1}$ = 2.419972619058592 & $b_{L,0}$ = 1.0 \\
+         $b_{R,3}$ = 0.6456076059873844 & $b_{R,2}$ = -0.34840383986411055 \\ $b_{R,1}$ = 2.6035438510917692 &  $b_{R,0}$ = 1.0 \\
          \hline

### 2020-10-21 code/tent_sens_stat/map.py
Add new map for last section

@@ -1,0 +1,7 @@
+@numba.jit(nopython=True)
+def tent(x, n, s):
+    if x < 1+s:
+        return 2*x/(1+s)
+    return 2/(1-s)*(2-x)
+
+

### 2020-10-21 main.tex
Updates from Overleaf

@@
 \end{figure}
-The empirical densities computed from long shadowing solutions, of length 10 billion, are shown in the left column of Figure \ref{fig:lorenz_params_scaled},
+The empirical densities computed from long shadowing solutions, of length 10 billion, are shown in the left column of Figure \ref{fig:lorenz_params_scaled} at three different sets of parameter values, which are also different from the standard values of $\rho = 28, \sigma = 10$, and $\beta = 8/3$.
+In the top and middle row, $\rho = 29$ and $\rho = 30$ are used, respectively, while the other two parameters are retained at their standard values.
+In the bottom row, only the parameter $\beta$ is perturbed from the standard values; $\beta = 10/3$ is used.
+On the right column, we plot the densities of physical solutions, the SRB distribution, at the set of parameters corresponding to each row.
+We again notice that, while the physical distributions are smooth, the shadowing solutions have a rough distribution that appears to have hair-like patterns.
+Thus, we draw the same conclusion in the Lorenz map as in the various perturbations of the tent map treated in section \ref{sec:tent}: shadowing solutions do not have the same distribution as a typically observed solution of the governing equation.
+In other words, shadowing solutions may be nonphysical.
 
 \section{Discussions}
-Shown through a rigorous counter-example, we conclude that shadowing can lead
-to nonphysical solutions.  The conclusion has instant implication for
-sensitivity analysis methods based on the concept shadowing.
-If the goal is to compute derivative of ensemble averages, where the ensemble
-is distributed according to the physical measure, i.e., the SRB measure,
-then shadowing-based methods can give wrong results.  This has been observed
-before and appropriately attributed as ``ergodicity breaking error''.
-This paper merely gave a rigorous analytical example of it.
-
-Even more troubling is the implication for any numerical simulation of chaotic
-governing equations.  We now understand that a numerical solution, satisfying
+Through a rigorous counter-example, we show that shadowing can lead
+to nonphysical solutions.  The conclusion has an immediate implication for
+sensitivity analysis methods based on the concept of shadowing \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagne}.
+When the goal is to compute derivatives of ensemble averages, where the ensemble
+is distributed according to the physical measure, i.e., the SRB measure, shadowing-based methods can give wrong results.  This has been observed
+before and appropriately attributed as ``ergodicity breaking error'' \cite{patrick} \cite{angxiu-error}. In this paper, we construct rigorous analytical examples of this error.
+
+Even more troubling is the implication for numerical simulations of chaotic
+governing equations.  Through the examples in this paper, it is clear that a numerical solution, satisfying
 a governing equation perturbed due to numerics, is not expected to shadow a
@@
 the true physics?
-[Nisha: Use tent map as example for point 1. For $10^{-18}$ perturbation (trunction numerically), we get completely different statistics. How can we trust numerical solutions]
+While this question remains open, we conclude with a final example illustrating that not only can shadowing solutions display statistical averages different from physical solutions, but that the difference can be drastic.
+For $10^{-18}$ perturbation (trunction numerically), we get completely different statistics. How can we trust numerical solutions
 

### 2020-10-22 code/tent_sens_stat/map.py
Add accumulate function that collects density.

@@ -1,7 +1,35 @@
+import numba
+from numpy import *
 @numba.jit(nopython=True)
 def tent(x, n, s):
     if x < 1+s:
         return 2*x/(1+s)
     return 2/(1-s)*(2-x)
+
+@numba.jit(nopython=True)
+def accumulate(density, n, m, s):
+    nbins = len(density)
+    dx = 2.0/nbins
+    x = 2*rand()
+    for i in range(n):
+        x = tent(x, m, s)
+        bno = int(x//dx)
+        density[bno] += 1/n
+    return density
+
+nbins = 2**8
+density = zeros(nbins)
+'''
+x = linspace(0, 2, nbins)[1:-1]
+y = zeros_like(x)
+for i, xi in enumerate(x):
+    y[i] = tent(xi, 1, 0)
+
+fig = figure(figsize=(8,5))
+ax = fig.add_subplot(111)
+ax.plot(x,y)
+ax.xaxis.set_tick_params(labelsize=40)
+ax.yaxis.set_tick_params(labelsize=40)
+'''
 
 

### 2020-10-22 code/tent_sens_stat/plot_map.py
plot map and density in separate files.

@@ -1,0 +1,56 @@
+import numba
+from numpy import *
+@numba.jit(nopython=True)
+def tent_basic(x,s):
+    if x < 1:
+        return min(2*x/(1-s), \
+                2 - 2*(1-x)/(1+s))
+    return min(2*(2-x)/(1-s), \
+            2 - 2*(x-1)/(1+s))
+
+@numba.jit(nopython=True)
+def oscillation(x, s, n):
+    return tent_basic(2**n*x - floor(2**n*x),s)/2**n + \
+            2*floor(2**n*x)/2**n
+
+@numba.jit(nopython=True)
+def osc_tent(x, s, n):
+    return min(oscillation(x,s,n), oscillation(2-x,s,n))
+
+@numba.jit(nopython=True)
+def accumulate(nbins, n, m, s):
+    density = zeros(nbins)
+    dx = 2.0/nbins
+    x = 2*rand()
+    for i in range(n):
+        x = osc_tent(x, s, m)
+        bno = int(x//dx)
+        density[bno] += 1/n/dx
+    return density
+'''
+nbins = 2**9
+x = linspace(0, 2, nbins)[1:-1]
+y = zeros_like(x)
+for i, xi in enumerate(x):
+    y[i] = osc_tent(xi, 0.5, 5)
+fig = figure(figsize=(8,5))
+ax = fig.add_subplot(111)
+ax.plot(x, y)
+ax.xaxis.set_tick_params(labelsize=40)
+ax.yaxis.set_tick_params(labelsize=40)
+'''
+
+
+nbins = 2**12
+nrep = 1
+density = zeros(nbins)
+for m in range(nrep):
+    print("repeat {}".format(m))
+    density += accumulate(nbins, 1000000000, 6, 0.5)/nrep
+x = linspace(0, 2, nbins+2)[1:-1]
+fig = figure(figsize=(8,5))
+ax = fig.add_subplot(111)
+ax.fill_between(x, 0, density)
+ax.xaxis.set_tick_params(labelsize=40)
+ax.yaxis.set_tick_params(labelsize=40)
+

### 2020-10-22 code/tent_sens_stat/map.py
Add map figures for oscillating tent map

@@ -1,35 +1,0 @@
-import numba
-from numpy import *
-@numba.jit(nopython=True)
-def tent(x, n, s):
-    if x < 1+s:
-        return 2*x/(1+s)
-    return 2/(1-s)*(2-x)
-
-@numba.jit(nopython=True)
-def accumulate(density, n, m, s):
-    nbins = len(density)
-    dx = 2.0/nbins
-    x = 2*rand()
-    for i in range(n):
-        x = tent(x, m, s)
-        bno = int(x//dx)
-        density[bno] += 1/n
-    return density
-
-nbins = 2**8
-density = zeros(nbins)
-'''
-x = linspace(0, 2, nbins)[1:-1]
-y = zeros_like(x)
-for i, xi in enumerate(x):
-    y[i] = tent(xi, 1, 0)
-
-fig = figure(figsize=(8,5))
-ax = fig.add_subplot(111)
-ax.plot(x,y)
-ax.xaxis.set_tick_params(labelsize=40)
-ax.yaxis.set_tick_params(labelsize=40)
-'''
-
-

### 2020-10-22 code/tent_sens_stat/plot_map.py

@@ -27,30 +27,36 @@ def accumulate(nbins, n, m, s):
         bno = int(x//dx)
         density[bno] += 1/n/dx
     return density
-'''
-nbins = 2**9
-x = linspace(0, 2, nbins)[1:-1]
-y = zeros_like(x)
-for i, xi in enumerate(x):
-    y[i] = osc_tent(xi, 0.5, 5)
-fig = figure(figsize=(8,5))
-ax = fig.add_subplot(111)
-ax.plot(x, y)
-ax.xaxis.set_tick_params(labelsize=40)
-ax.yaxis.set_tick_params(labelsize=40)
-'''
 
+nbins = 2**13
+ns = 5
+nn = 5
+n_arr = [0, 1, 2, 4, 6]
+s_arr = [0, 0.1, 0.2, 0.3, 0.5]
 
-nbins = 2**12
-nrep = 1
-density = zeros(nbins)
-for m in range(nrep):
-    print("repeat {}".format(m))
-    density += accumulate(nbins, 1000000000, 6, 0.5)/nrep
 x = linspace(0, 2, nbins+2)[1:-1]
-fig = figure(figsize=(8,5))
+ys = zeros((ns,nbins))
+yn = zeros((nn,nbins))
+for k,nk in enumerate(n_arr):
+    for i, xi in enumerate(x):
+        yn[k,i] = osc_tent(xi, 0.5, nk)
+fig = figure(figsize=(8,6))
 ax = fig.add_subplot(111)
-ax.fill_between(x, 0, density)
+for k,nk in enumerate(n_arr):
+    ax.plot(x, yn[k],lw=2.5,label="n = {}".format(nk))
 ax.xaxis.set_tick_params(labelsize=40)
 ax.yaxis.set_tick_params(labelsize=40)
+leg = ax.legend(fontsize=30)
+ax.axis("scaled")
 
+for k,sk in enumerate(s_arr):
+    for i, xi in enumerate(x):
+        ys[k,i] = osc_tent(xi, sk, 0)
+fig = figure(figsize=(8,6))
+ax = fig.add_subplot(111)
+for k,sk in enumerate(s_arr):
+    ax.plot(x, ys[k],lw=2.5,label="s = {}".format(sk))
+ax.xaxis.set_tick_params(labelsize=40)
+ax.yaxis.set_tick_params(labelsize=40)
+ax.axis("scaled")
+leg = ax.legend(fontsize=30)

### 2020-10-22 main.tex
Add plucked tent map figures.

@@
     \centering
-    \includegraphics[width=0.35\textwidth,height=2in]{figure/mapn0s0.png}
-    \includegraphics[width=0.35\textwidth,height=2in]{figure/mapn5s0.1.png}
-    \caption{Two chaotic maps (in red) and their associated stationary  distributions (in purple).
-The map on the right is only an \emph{infinitesimal} perturbation of that on the left; for more information see \cite{desmos}.}
+    \includegraphics[width=0.35\textwidth,height=2in]{figure/osc_tent_n.png}
+    \includegraphics[width=0.35\textwidth,height=2in]{figure/osc_tent_s.png}
+    \caption{The plucked tent map shown at different values of $s$ and $n$, that represent the magnitude and frequency of the applied oscillation to the original tent map ($n=0,s=0$).
+On the left, $s$ is held constant at 0.5.
+On the right, $n$ is held constant at 0.}
     \label{fig:map}
@@
 [Nisha: shadowing perturabtions predict tiny effects. Real effect is a lot more dramatic. Desmos plot.]
-% give example of numerics that destroys physicalness
+give example of numerics that destroys physicalness
 

### 2020-10-22 main.tex
Adjust fig size

@@
 before and appropriately attributed as ``ergodicity breaking error'' \cite{patrick} \cite{angxiu-error}. In this paper, we construct rigorous analytical examples of this error.
-
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{figure/osc_tent_n.png}
+    \includegraphics[width=0.45\textwidth]{figure/osc_tent_s.png}
+    \caption{The plucked tent map shown at different values of $s$ and $n$, that represent the magnitude and frequency of the applied oscillation to the original tent map ($n=0,s=0$).
+On the left, $s$ is held constant at 0.5.
+On the right, $n$ is held constant at 0.}
+    \label{fig:map}
+\end{figure}
 Even more troubling is the implication for numerical simulations of chaotic
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.35\textwidth,height=2in]{figure/osc_tent_n.png}
-    \includegraphics[width=0.35\textwidth,height=2in]{figure/osc_tent_s.png}
-    \caption{The plucked tent map shown at different values of $s$ and $n$, that represent the magnitude and frequency of the applied oscillation to the original tent map ($n=0,s=0$).
-On the left, $s$ is held constant at 0.5.
-On the right, $n$ is held constant at 0.}
-    \label{fig:map}
-\end{figure}
 [Nisha: shadowing perturabtions predict tiny effects. Real effect is a lot more dramatic. Desmos plot.]

### 2020-10-22 code/tent_sens_stat/plot_density.py
Add analytical density plots

@@ -27,27 +27,24 @@ def accumulate(nbins, n, m, s):
         bno = int(x//dx)
         density[bno] += 1/n/dx
     return density
-'''
-nbins = 2**9
-x = linspace(0, 2, nbins)[1:-1]
-y = zeros_like(x)
-for i, xi in enumerate(x):
-    y[i] = osc_tent(xi, 0.5, 5)
-fig = figure(figsize=(8,5))
-ax = fig.add_subplot(111)
-ax.plot(x, y)
-ax.xaxis.set_tick_params(labelsize=40)
-ax.yaxis.set_tick_params(labelsize=40)
-'''
 
+@numba.jit(nopython=True)
+def analytical_rho(x,s):
+    if x < 1:
+        return (1-s)/2
+    return (1+s)/2
 
 nbins = 2**12
-nrep = 1
+nrep = 10
 density = zeros(nbins)
-for m in range(nrep):
-    print("repeat {}".format(m))
-    density += accumulate(nbins, 1000000000, 6, 0.5)/nrep
-x = linspace(0, 2, nbins+2)[1:-1]
+s = 0.5
+n = 5
+#for m in range(nrep):
+ #   print("repeat {}".format(m))
+    #density += accumulate(nbins, 1000000000, n, s)/nrep
+x = array(linspace(0, 2, nbins+2)[1:-1])
+for i,xi in enumerate(x):
+    density[i] = analytical_rho(2*2**n*xi - 2*floor(2**n*xi),s)
 fig = figure(figsize=(8,5))
 ax = fig.add_subplot(111)
 ax.fill_between(x, 0, density)

### 2020-10-23 code/tent_sens_stat/plot_density.py
edit map and density to have asymmetric density

@@ -9,13 +9,21 @@ def tent_basic(x,s):
             2 - 2*(x-1)/(1+s))
 
 @numba.jit(nopython=True)
-def oscillation(x, s, n):
-    return tent_basic(2**n*x - floor(2**n*x),s)/2**n + \
+def oscillation(x,s):
+    if x < 0.5:
+        return tent_basic(2*x,s)/2
+    return 2-tent_basic(2-2*x,s)/2 
+
+
+@numba.jit(nopython=True)
+def frequency(x,s,n):
+    return oscillation(2**n*x - floor(2**n*x),s)/2**n + \
             2*floor(2**n*x)/2**n
 
 @numba.jit(nopython=True)
 def osc_tent(x, s, n):
-    return min(oscillation(x,s,n), oscillation(2-x,s,n))
+    return min(frequency(x,s,n), frequency(2-x,s,n))
+
 
 @numba.jit(nopython=True)
 def accumulate(nbins, n, m, s):
@@ -35,16 +43,16 @@ def analytical_rho(x,s):
     return (1+s)/2
 
 nbins = 2**12
-nrep = 10
+nrep = 1
 density = zeros(nbins)
 s = 0.5
-n = 5
-#for m in range(nrep):
- #   print("repeat {}".format(m))
-    #density += accumulate(nbins, 1000000000, n, s)/nrep
+n = 6
+for m in range(nrep):
+    print("repeat {}".format(m))
+    density += accumulate(nbins, 1000000, n, s)/nrep
 x = array(linspace(0, 2, nbins+2)[1:-1])
-for i,xi in enumerate(x):
-    density[i] = analytical_rho(2*2**n*xi - 2*floor(2**n*xi),s)
+#for i,xi in enumerate(x):
+#    density[i] = analytical_rho(2*2**n*xi - 2*floor(2**n*xi),s)
 fig = figure(figsize=(8,5))
 ax = fig.add_subplot(111)
 ax.fill_between(x, 0, density)

### 2020-10-23 code/tent_sens_stat/plot_map.py

@@ -9,13 +9,20 @@ def tent_basic(x,s):
             2 - 2*(x-1)/(1+s))
 
 @numba.jit(nopython=True)
-def oscillation(x, s, n):
-    return tent_basic(2**n*x - floor(2**n*x),s)/2**n + \
+def oscillation(x,s):
+    if x < 0.5:
+        return tent_basic(2*x,s)/2
+    return 2-tent_basic(2-2*x,s)/2 
+
+
+@numba.jit(nopython=True)
+def frequency(x,s,n):
+    return oscillation(2**n*x - floor(2**n*x),s)/2**n + \
             2*floor(2**n*x)/2**n
 
 @numba.jit(nopython=True)
 def osc_tent(x, s, n):
-    return min(oscillation(x,s,n), oscillation(2-x,s,n))
+    return min(frequency(x,s,n), frequency(2-x,s,n))
 
 @numba.jit(nopython=True)
 def accumulate(nbins, n, m, s):
@@ -29,10 +36,10 @@ def accumulate(nbins, n, m, s):
     return density
 
 nbins = 2**13
-ns = 5
-nn = 5
-n_arr = [0, 1, 2, 4, 6]
-s_arr = [0, 0.1, 0.2, 0.3, 0.5]
+ns = 1
+nn = 1
+n_arr = [1]
+s_arr = [0.5]
 
 x = linspace(0, 2, nbins+2)[1:-1]
 ys = zeros((ns,nbins))

### 2020-10-23 code/tent_sens_stat/plot_density.py
Add figures of the new map

@@ -14,7 +14,6 @@ def oscillation(x,s):
         return tent_basic(2*x,s)/2
     return 2-tent_basic(2-2*x,s)/2 
 
-
 @numba.jit(nopython=True)
 def frequency(x,s,n):
     return oscillation(2**n*x - floor(2**n*x),s)/2**n + \

### 2020-10-23 code/tent_sens_stat/plot_map.py

@@ -36,10 +36,10 @@ def accumulate(nbins, n, m, s):
     return density
 
 nbins = 2**13
-ns = 1
-nn = 1
-n_arr = [1]
-s_arr = [0.5]
+ns = 5
+nn = 5
+n_arr = [0, 1, 2, 4, 6]
+s_arr = [0.005, 0.1, 0.2, 0.4, 0.5]
 
 x = linspace(0, 2, nbins+2)[1:-1]
 ys = zeros((ns,nbins))

### 2020-10-23 main.tex
Add text for the discussion section.

@@
 \end{figure}
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{figure/osc_tent_dens_0.5_0.png}
+    \includegraphics[width=0.45\textwidth]{figure/osc_tent_dens_0.5_6.png}
+    \caption{The effect of increasing $n$ on the probability distribution associated to the plucked tent map.
+L: The stationary probability distribution of the plucked tent map at $s=0.5, n = 0$.
+R: The stationary probability distribution of the plucked tent map at $s=0.5, n = 6$.}
+    \label{fig:map}
+\end{figure}
 Even more troubling is the implication for numerical simulations of chaotic
@@
 relation of a numerical solution, such as DNS of turbulent flows, to
-the true physics?
-While this question remains open, we conclude with a final example illustrating that not only can shadowing solutions display statistical averages different from physical solutions, but that the difference can be drastic.
+the true physics? While this question remains open, we conclude with a final example illustrating 
+another problem with assuming the physicality of shadowing solutions: with a small change in a parameter, the stationary probability distribution of a chaotic system \emph{change} drastically, while if shadowing solutions are physical, the change in the statistics must be small when the parameter perturbation is small.
+Once again, a perturbation of a tent map can be used as an example.
+We consider a plucked tent map as shown in Figure only can shadowing solutions display statistical averages different from physical solutions, but that the difference can be drastic. 
 For $10^{-18}$ perturbation (trunction numerically), we get completely different statistics. How can we trust numerical solutions
-
-[Nisha: Eq. 18 and Eq. 19 - shadowing sensitivities are incorrect.]
 

### 2020-10-23 code/tent_sens_stat/plot_density.py
Add analytical density at n = 0

@@ -37,21 +37,25 @@ def accumulate(nbins, n, m, s):
 
 @numba.jit(nopython=True)
 def analytical_rho(x,s):
-    if x < 1:
-        return (1-s)/2
-    return (1+s)/2
+    if x < s:
+        return (1-s)/4
+    elif x < 1:
+        return (1 + s)/4
+    elif x < 1+s:
+        return 1 + s/4
+    return (1+s)/4
 
 nbins = 2**12
-nrep = 1
+nrep = 5
 density = zeros(nbins)
 s = 0.5
-n = 6
-for m in range(nrep):
-    print("repeat {}".format(m))
-    density += accumulate(nbins, 1000000, n, s)/nrep
+n = 0
+#for m in range(nrep):
+#    print("repeat {}".format(m))
+#    density += accumulate(nbins, 1000000000, n, s)/nrep
 x = array(linspace(0, 2, nbins+2)[1:-1])
-#for i,xi in enumerate(x):
-#    density[i] = analytical_rho(2*2**n*xi - 2*floor(2**n*xi),s)
+for i,xi in enumerate(x):
+    density[i] = analytical_rho(xi,s)
 fig = figure(figsize=(8,5))
 ax = fig.add_subplot(111)
 ax.fill_between(x, 0, density)

### 2020-10-23 main.tex

@@
 relation of a numerical solution, such as DNS of turbulent flows, to
-the true physics? While this question remains open, we conclude with a final example illustrating 
-another problem with assuming the physicality of shadowing solutions: with a small change in a parameter, the stationary probability distribution of a chaotic system \emph{change} drastically, while if shadowing solutions are physical, the change in the statistics must be small when the parameter perturbation is small.
+the true physics? For $10^{-18}$ perturbation (due to numerics), 
+we may get completely different statistics from a true, physical solution. 
+How then can we trust numerical solutions that, even if they are shadowed by
+physical solutions, do not share their statistics?
+
+We conclude with a final example illustrating 
+another problem with assuming the physicality of shadowing solutions: with a small change in a parameter, the stationary probability distribution of a chaotic system \emph{can} change drastically, while if shadowing solutions are physical, the change in the statistics must be small when the parameter perturbation is small.
 Once again, a perturbation of a tent map can be used as an example.
 We consider a plucked tent map as shown in Figure only can shadowing solutions display statistical averages different from physical solutions, but that the difference can be drastic.
-For $10^{-18}$ perturbation (trunction numerically), we get completely different statistics. How can we trust numerical solutions
-
-[Nisha: shadowing perturabtions predict tiny effects. Real effect is a lot more dramatic. Desmos plot.]
-give example of numerics that destroys physicalness
 

### 2020-10-23 main.tex
Edit paper for arxiv

@@
 \usepackage{amsmath, amssymb, amsfonts}
-\usepackage{xcolor}
+\usepackage{xcolor,verbatim}
 \definecolor{darkgreen}{rgb}{0.0, 0.5, 0.0}
@@
 \date{June 2020}
-\subtitle{Do you have a subtitle?\\ If so, write it here}
 
@@
 
-\author{Qiqi Wang}
+\author{Qiqi Wang and Nisha Chandramoorthy}
 
@@
 
-\institute{F. Author \at
-              first address \\
-              Tel.: +123-45-678910\\
-              Fax: +123-45-678910\\
-              \email{fauthor@example.com}           %  \\
+\institute{Q. Wang \at 
+        Department of Aeronautics and Astronautics, Center for Computational Science and Engineering, MIT, Cambridge, MA 02139, USA.
+              \email{qiqi@mit.edu}           %  \\
 %             \emph{Present address:} of F. Author  %  if needed
            \and
-           S. Author \at
-              second address
+           N. Chandramoorthy \at Department of Mechanical Engineering, Center for Computational Science and Engineering, MIT, Cambridge, MA 02139, USA. \email{nishac@mit.edu}
 }
 
-\date{Received: date / Accepted: date}
+%\date{Received: date / Accepted: date}
 % The correct dates will be entered by the editor
@@
 \begin{abstract}
-Insert your abstract here. Include keywords, PACS and mathematical
-subject classification numbers as needed.
-\keywords{First keyword \and Second keyword \and More}
+Based on the shadowing property of dynamical systems, we expect that numerical solutions of governing equations lie close to a true solution -- a shadowing solution -- of the governing equation.
+This is used to justify, even when the governing equation is chaotic, that numerical solutions represent a true solution.
+However, shadowing solutions can be \emph{nonphysical}: they may not represent the long-term statistical behavior of an ensemble of true solutions.
+That is, the probability distribution of a shadowing solution can differ, often dramatically, from that of a solution starting from almost every initial condition.
+In this paper, we construct analytical examples of chaotic systems in which shadowing solutions are nonphysical.
+These examples also illustrate that the probability of finding a nonphysical shadowing solution can be 1.
+Through several analytical constructions, we note that the incorrect assumption of the physicality of shadowing raises questions on the trustworthiness of numerical solutions, and shadowing-based sensitivity analysis of chaotic systems.
+\keywords{Numerical simulations \and Shadowing sensitivity analysis \and Chaotic systems}
 % \PACS{PACS code1 \and PACS code2 \and more}
@@
 it is used in computing how long-time-averages in a chaotic system
-respond to small perturbations in the governing equation. In this application, derivative
+respond to small perturbations in the governing equation \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagna}. In this application, derivative
 of the long-time-averages is computed using a solution to the perturbed governing equation
-that shadows a solution to the unperturbed equation.  The Least Squares Shadowing \cite{qiqi-lss} method
+that shadows a solution to the unperturbed equation.  The Least Squares Shadowing \cite{qiqi-lss}\cite{angxiu-lss} method
 uses this concept.
@@
 the shadowing trajectory is a ``physical'' trajectory, a trajectory on which
-the long-time-average of a quantity is equal to the physical \nisha{can we remove the word ``physical'' before ensemble average?} ensemble average.
+the long-time-average of a quantity is equal to the ensemble average.
 Not all solutions satisfying the physical governing equation are considered ``physical''
@@
 What we typically call ``physical'' solutions to a governing equation satisfy the following
-two criteria: \nisha{Is the term ``physical'' to mean the following common in fluid mechanics? If not, can we emphasize that we are defining it here? e.g. }
+two criteria: 
 \begin{enumerate}
     \item Long-time-averaged, ``statistical'' quantities converges as time goes to infinity.
-    Consider $u(t)$ as the solution to a chaotic governing equation, then for \nisha{a continuous/regular observable?}
-    a smooth quantity of interest $J(u(t))$,
+    Consider $u(t)$ as the solution to a chaotic governing equation, then for a regular observable
+    of interest $J(u(t))$,
     \begin{equation} \label{longtimeavg}
@@
 
-\begin{figure}[H] \centering
+\begin{figure} \centering
 \includegraphics[width=0.48\textwidth]{figure/lorenz_trajectory_1000.0.png}
@@
 \end{figure}
-\begin{figure}[H] \centering
+\begin{figure} \centering
 \includegraphics[width=0.48\textwidth]{figure/lorenz_trajectory_10000.0.png}
@@
 \end{figure}
-\begin{figure}[H] \centering
+\begin{figure} \centering
 \includegraphics[width=0.48\textwidth]{figure/lorenz_trajectory_1000000.0.png}
@@
 stationary distribution achieved on long-time evolution of the ensemble that has the absolute continuity property is called the Sinai-Ruelle-Bowen
-(SRB) measure. The absolute continuity property is sufficient to ensure that the SRB measure is the same $\mu.$ That is, the SRB measure is physically observed in the sense that  
-\nisha{what do we mean by ``under sufficient regularity conditions'' here?
-Suggesting a modification to the previous sentence} physical solutions produce long-time-averages which are expectations with respect to the SRB measure.
+(SRB) measure.
+The absolute continuity property is sufficient to ensure that the SRB measure is the same $\mu.$ That is, the SRB measure is physically observed in the sense that physical solutions produce long-time-averages which are expectations with respect to the SRB measure.
 
@@
 
-\begin{figure}[H]\centering
+\begin{figure}\centering
 \includegraphics[width=0.48\textwidth]{figure/lorenz_periodic.png}
@@
 
-\begin{figure}[H]\centering
+\begin{figure}\centering
 \includegraphics[width=0.48\textwidth]{figure/lorenz_zmax_10_28_2.6666666666666665.png}
@@
 line indicates the initial conditions for both the solid and dashed lines in
-Figure \ref{fig:lorenz_periodic}.
-\nisha{In Figure 4, the solid line appears only to have one local maximum, which is shown as the fixed point on the left plot of this figure.
-The two fixed points of the 2-time-iterate of the Lorenz map, shown on the right plot of this figure, both appear to correspond to the local maxima of the dashed curve in Figure 4.}}
+Figure \ref{fig:lorenz_periodic}.}
 \label{fig:lorenz_map}
 \end{figure}
-
+\begin{comment}In Figure 4, the solid line appears only to have one local maximum, which is shown as the fixed point on the left plot of this figure.
+The two fixed points of the 2-time-iterate of the Lorenz map, shown on the right plot of this figure, both appear to correspond to the local maxima of the dashed curve in Figure 4.\end{comment}
 The Lorenz map provides us a tool to find as many periodic solutions as we want.
@@
 of nonphysical solutions, a type that hides in the shadow.
-\nisha{I wish all papers were this intriguing!}
+
 \subsection{Tent map: periodic and physical solutions}
@@
 Figure \ref{fig:tent_map} shows the tent map.
-\begin{figure}[H]\centering
+\begin{figure}\centering
 \includegraphics[width=0.48\textwidth]{figure/tent_map.png}
@@
 
-\begin{figure}[H]\centering
+\begin{figure}\centering
 \includegraphics[width=0.48\textwidth]{figure/tent_quasiphysical_hist_p_0.5001.png}
@@
 appears to be a hallmark of quasi-physical solutions.
-\begin{figure}[H]\centering
+\begin{figure}\centering
     \centering
@@
 
-\begin{figure}[H]   \centering
+\begin{figure}   \centering
     \includegraphics[width=0.48\textwidth]{figure/scaled_tent_map.png}
@@
 we need to analyze two cases: $x<1+s$ and $x\ge 1+s$.
-\begin{figure}[H]
+\begin{figure}
   \centering
@@
 
-\begin{figure}[H]
+\begin{figure}
    \includegraphics[width=0.48\textwidth]{figure/tilted_conjugate_density_0.1.png}
@@
 in Section \ref{sec:quasiphysical}, with $p=\frac{1+s}2$.
-\nisha{This is so well set up and awesome!}
 Since this analysis holds for any $x$ sampled uniformly on $[0,2]$, we can conclude that
@@
 
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
 Figure \ref{fig:tent_pinched} shows the pinched tent map at different $s$ values.
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
 Once again, with probability one -- or, for any randomly chosen initial condition of the solution of the pinched tent map -- any physical solution is shadowed by a quasi-physical solution.
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
 $x$-axis.
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
 On the left column of Figure \ref{fig:tent_wave_shadow}, we observe that the distributions of the shadowing solutions corresponding to the physical solutions on the right, once again appear fractal.
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
 of the Lorenz map (see Figure \ref{fig:lorenz_map}). 
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
     In the right plot, the blue, orange, and green lines represent
-    $\sigma=8/3,3$ and $10/3$ respectively, while $\sigma = 10$ and $\rho = 28.$\nisha{Is the green line overlapping one of the two lines?}}
+    $\sigma=8/3,3$ and $10/3$ respectively, while $\sigma = 10$ and $\rho = 28.$}
     \label{fig:lorenz_params}
 \end{figure}
+\begin{comment}
+Is the green line overlapping one of the two lines?
+\end{comment}
 The Lorenz map approximation obtained this way is illustrated in Figure \ref{fig:lorenz_params}, for variations of the parameters around their standard values of 
 $\rho = 28, \sigma = 10, \beta = 8/3.$ For a range of parameter values, it is numerically verified that the approximate Lorenz map qualitatively reproduces the behavior of the true Lorenz map.
-Using the approximate Lorenz map, we calculate next the shadowing solutions (see the supplementary material section \ref{sec:lorenzShadowingProcedure} for the numerical procedure).
-\begin{figure}[H]
+Using the approximate Lorenz map, we calculate next the shadowing solutions.
+\begin{figure}
     \centering
@@
 to nonphysical solutions.  The conclusion has an immediate implication for
-sensitivity analysis methods based on the concept of shadowing \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagne}.
+sensitivity analysis methods based on the concept of shadowing \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagna}.
 When the goal is to compute derivatives of ensemble averages, where the ensemble
@@
 On the right, $n$ is held constant at 0.}
-    \label{fig:map}
+    \label{fig:osc_map}
 \end{figure}
@@
 
-We conclude with a final example illustrating 
-another problem with assuming the physicality of shadowing solutions: with a small change in a parameter, the stationary probability distribution of a chaotic system \emph{can} change drastically, while if shadowing solutions are physical, the change in the statistics must be small when the parameter perturbation is small.
-Once again, a perturbation of a tent map can be used as an example.
-We consider a plucked tent map as shown in Figure only can shadowing solutions display statistical averages different from physical solutions, but that the difference can be drastic.
+We conclude with an example that illustrates that assuming the 
+physicality of shadowing solutions leads to incorrect results in shadowing-based sensitivity 
+computation methods. We construct yet another perturbed tent map -- the \emph{plucked} 
+tent map -- in which an oscillatory perturbation is introduced.
+In Figure \ref{fig:osc_map}, we show that the magnitude of the oscillatory perturbation is controlled by a parameter $s$, and its frequency is controlled by the parameter $n$; at $s=0,n=0$, we recover the original tent map.
+The plucked tent map illustrates an important feature of some chaotic systems: with a small change in a parameter, the stationary probability distribution of a chaotic system \emph{can} change drastically.
+For the plucked tent map, this feature is illustrated by the stationary probability distributions -- computed empirically over a trajectory of length 10 billion -- at $s=0.5$.
+On the left, $n=0$ and there is already a marked asymmetry developed in the probability distribution compared to the uniform distribution seen at $s=0, n=0$.
+On the right, we show that by increasing $n,$ while keeping the magnitude of the perturbation the same ($s=0.5$), we see a dramatic change in the appearance of the stationary probability distribution: we observe a fractal distribution.
+On the other hand, if perturbed solutions are shadowed by physical solutions for all time, the change in the statistics of the perturbed solutions must be small.
+In other words, the physicality of shadowing predicts that there cannot be a large change in statistics due to small parameter perturbations; the reality is that, as illustrated by the plucked tent map, the effect of small parameter changes on the statistics can be drastic.
+
+In this paper, we have constructed several counterexamples that disprove the common 
+assumption that shadowing solutions are physical solutions, i.e., that their 
+statistical distribution is the same as that typically observed for almost every solution 
+of a governing equation. The existence of long-time shadowing solutions \cite{grebogi} has historically been used to address the issue of whether numerical simulations, which are 
+perturbed solutions, represent the true physics implied by the governing equation. In light of the evidence in this paper, we must reopen this issue. Even when numerical simulations are 
+shadowed by a true solution in that the difference between them is small for a long time, this shadowing solution may not represent the long-term behavior of the physical system.
+The nonphysicality of shadowing solutions also indicates that shadowing-based methods \cite{qiqi-lss} can lead to incorrect values of sensitivities of statistics to parameter changes.
 

### 2020-10-23 supp.tex
Edit paper for arxiv

@@
 \section{Supplementary material}
-\subsection{The Lorenz Map}
-\subsubsection{Approximation of the Lorenz Map}
+%\subsection{The Lorenz Map}
+\subsection{Approximation of the Lorenz Map}
+\label{sec:lorenzMapApprox}
 The motivation for approximating the Lorenz map is that a 
@@
 The values of these coefficients are obtained by
-regression on 
+regression on the map at standard parameters are shown in the table below.
 \begin{table}[H]
@@
 
+\begin{comment}
 \subsubsection{Computing shadowing solutions of the Lorenz map}
@@
 To conclude that the sequence approximates the shadowing solution, we only need to check that $x_n$ and $y_n$ are close to each other for all times up to $N.$
+\end{comment}

### 2020-10-23 main.tex
Qiqi fixes caption

@@
     parameters $\sigma, \rho$, and $\beta$ on the Lorenz map, respectively.
-    In the left plot, the blue, orange, and green lines represent
-    $\sigma=10,11$, and $12$, respectively, while $\rho  = 28$ and $\beta = 8/3.$
-    In the center plot, the blue, orange, and green lines represent
-    $\rho=28,29$ and $30$, respectively, while $\sigma = 10$ and $\beta  = 8/3.$ 
-    In the right plot, the blue, orange, and green lines represent
-    $\sigma=8/3,3$ and $10/3$ respectively, while $\sigma = 10$ and $\rho = 28.$}
+    In the left plot, the blue and orange lines represent
+    $\sigma=10$ and $12$, respectively, while $\rho  = 28$ and $\beta = 8/3.$
+    In the center plot, the blue and orange lines represent
+    $\rho=28$ and $30$, respectively, while $\sigma = 10$ and $\beta  = 8/3.$ 
+    In the right plot, the blue and orange lines represent
+    $\sigma=8/3$ and $10/3$ respectively, while $\sigma = 10$ and $\rho = 28.$}
     \label{fig:lorenz_params}

### 2020-10-26 main.tex
Add all figures in main directory. modify text to arxiv version.

@@
 \usepackage{amsmath, amssymb, amsfonts}
+\usepackage{mathtools,hyperref}
 \usepackage{xcolor,verbatim}
@@
 \newcommand{\nisha}[1]{{\color{darkgreen} #1}}
+\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
 \begin{document}
@@
 
-\author{Qiqi Wang and Nisha Chandramoorthy}
+\author{Nisha Chandramoorthy and Qiqi Wang}
 
@@
 would trip it into turbulence. Unstable steady-state solutions are not the only
-non-physical solution. Many chaotic dynamical systems have infinitely many periodic
+nonphysical solution. Many chaotic dynamical systems have infinitely many periodic
 solutions that are, similar to their steady counterpart, unstable.
@@
 \begin{figure} \centering
-\includegraphics[width=0.48\textwidth]{figure/lorenz_trajectory_1000.0.png}
-\hspace{0.02\textwidth}
-\includegraphics[width=0.48\textwidth]{figure/lorenz_ensemble_10.png}
+\includegraphics[width=0.48\textwidth]{lorenz_trajectory_1000.0.png}
+\hspace{0.02\textwidth}
+\includegraphics[width=0.48\textwidth]{lorenz_ensemble_10.png}
 \caption{
@@
 \begin{figure} \centering
-\includegraphics[width=0.48\textwidth]{figure/lorenz_trajectory_10000.0.png}
-\hspace{0.02\textwidth}
-\includegraphics[width=0.48\textwidth]{figure/lorenz_ensemble_15.png}
+\includegraphics[width=0.48\textwidth]{lorenz_trajectory_10000.0.png}
+\hspace{0.02\textwidth}
+\includegraphics[width=0.48\textwidth]{lorenz_ensemble_15.png}
 \caption{
@@
 \begin{figure} \centering
-\includegraphics[width=0.48\textwidth]{figure/lorenz_trajectory_1000000.0.png}
-\hspace{0.02\textwidth}
-\includegraphics[width=0.48\textwidth]{figure/lorenz_ensemble_50.png}
+\includegraphics[width=0.48\textwidth]{lorenz_trajectory_1000000.0.png}
+\hspace{0.02\textwidth}
+\includegraphics[width=0.48\textwidth]{lorenz_ensemble_50.png}
 \caption{
@@
 We first introduce a type of obviously nonphysical solutions in the next
-subsection, before discussing a less obvious type in Section \ref{sec:quasiphysical}.
+subsection, before discussing a less obvious type in section \ref{sec:quasiphysical}.
 
@@
 \begin{figure}\centering
-\includegraphics[width=0.48\textwidth]{figure/lorenz_periodic.png}
+\includegraphics[width=0.48\textwidth]{lorenz_periodic.png}
 \caption{Periodic solutions of the Lorenz equation, overlaid on top of
@@
 \begin{figure}\centering
-\includegraphics[width=0.48\textwidth]{figure/lorenz_zmax_10_28_2.6666666666666665.png}
-\hspace{0.02\textwidth}
-\includegraphics[width=0.48\textwidth]{figure/lorenz_zmax2_10_28_2.6666666666666665.png}
-\caption{Left: the Lorenz map.  The x-axis is the $n$-th local maximum of $u_3(t)$ over a
+\includegraphics[width=0.48\textwidth]{lorenz_zmax_10_28_2.6666666666666665.png}
+\hspace{0.02\textwidth}
+\includegraphics[width=0.48\textwidth]{lorenz_zmax2_10_28_2.6666666666666665.png}
+\caption{L: the Lorenz map.  The x-axis is the $n$-th local maximum of $u_3(t)$ over a
 long solution; the y-axis is the $(n+1)$-th local maximum of $u_3(t)$.  The intersection of this curve
-with the diagonal dashed line indicates the initial condition for the solid line
+		with the dashed line ($y=x$) indicates the initial condition for the solid line
 in Figure \ref{fig:lorenz_periodic}.
-Right: the Lorenz map iterated twice.  The x-axis is the $n$-th local maximum
-of $u_3(t)$; the y-axis is the $(n+2)$-th maximum.  The intersection with the diagonal dashed
-line indicates the initial conditions for both the solid and dashed lines in
+R: the Lorenz map iterated twice.  The x-axis is the $n$-th local maximum
+of $u_3(t)$; the y-axis is the $(n+2)$-th maximum.  The intersections with the diagonal dashed
+line indicate the initial conditions for both the dotted and dashed lines in
 Figure \ref{fig:lorenz_periodic}.}
@@
 to the Lorenz equation, the solution will visit only a discrete set of points.
-It is thus a periodic, non-physical solution.  What we learned about
+It is thus a periodic, nonphysical solution.  What we learned about
 the Lorenz equation could have all been learned from the Lorenz map.
@@
 with a closed form, one that qualitatively resembles the Lorenz map.  This map
-will help us, in Section \ref{sec:quasiphysical}, to study a more insidious type
+will help us, in section \ref{sec:quasiphysical}, to study a more insidious type
 of nonphysical solutions, a type that hides in the shadow.
@@
 \begin{figure}\centering
-\includegraphics[width=0.48\textwidth]{figure/tent_map.png}
+\includegraphics[width=0.48\textwidth]{tent_map.png}
 \caption{The tent map $\varphi$.}
@@
 \begin{figure}\centering
-\includegraphics[width=0.48\textwidth]{figure/tent_quasiphysical_hist_p_0.5001.png}
-\hspace{0.02\textwidth}
-\includegraphics[width=0.48\textwidth]{figure/tent_quasiphysical_hist_p_0.51.png}
+\includegraphics[width=0.48\textwidth]{tent_quasiphysical_hist_p_0.5001.png}
+\hspace{0.02\textwidth}
+\includegraphics[width=0.48\textwidth]{tent_quasiphysical_hist_p_0.51.png}
 \\ \vspace{0.05\textwidth}
-\includegraphics[width=0.48\textwidth]{figure/tent_quasiphysical_hist_p_0.55.png}
-\hspace{0.02\textwidth}
-\includegraphics[width=0.48\textwidth]{figure/tent_quasiphysical_hist_p_0.9.png}
+\includegraphics[width=0.48\textwidth]{tent_quasiphysical_hist_p_0.55.png}
+\hspace{0.02\textwidth}
+\includegraphics[width=0.48\textwidth]{tent_quasiphysical_hist_p_0.9.png}
 \caption{Empirical density function of long solutions (a billion steps)
@@
 pattern emerges from the distribution.  A fractal distribution
-appears to be a hallmark of quasi-physical solutions.
+appears to be a signature of quasi-physical solutions.
 \begin{figure}\centering
     \centering
-    \includegraphics[width=0.48\textwidth]{figure/tent_quasiphysical_alternaive1_hist_p_0.6.png}
-    \hspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tent_quasiphysical_alternaive2_hist_p_0.6.png}
+    \includegraphics[width=0.48\textwidth]{tent_quasiphysical_alternaive1_hist_p_0.6.png}
+    \hspace{0.02\textwidth}
+    \includegraphics[width=0.48\textwidth]{tent_quasiphysical_alternaive2_hist_p_0.6.png}
     \caption{Empirical distribution function of other quasi-physical solutions.
@@
 extension, the Lorenz equation.  It is then natural to conjecture
-that such aperiodic, non-physical, quasi-physical solutions
+that such aperiodic, nonphysical, quasi-physical solutions
 exist in general for chaotic dynamical systems.
@@
 \begin{figure}   \centering
-    \includegraphics[width=0.48\textwidth]{figure/scaled_tent_map.png}
-    \hspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tent_ivp.png}
+    \includegraphics[width=0.48\textwidth]{scaled_tent_map.png}
+    \hspace{0.02\textwidth}
+    \includegraphics[width=0.48\textwidth]{tent_ivp.png}
     \caption{L: The scaled tent map $\hat{\varphi}_s$ at different values of $s$ between 0 and 1. R: sensitivity to small perturbation in the governing
@@
 \end{equation}
-or equivalently, $\varphi\circ h_s \equiv h_s\circ\hat\varphi_s$.
+or equivalently, $\varphi\circ \hat{h}_s \equiv \hat{h}_s\circ\hat\varphi_s$.
 Such conjugacy maps can generally help us construct shadowing solutions.
 For every solution satisfying
-$\hat{\varphi}_s$, $h_s$ maps it to a shadowing solution satisfying $\varphi$.
+$\hat{\varphi}_s$, $h_s$ maps it to a shadowing solution satisfying $\varphi$ because iterating Eq.
+\ref{conjugate_def}, we get $\varphi^n \circ \hat{h}_s (x) = \hat{h}_s \circ \hat{\varphi}_s^n(x)$, for $n = 1,2,3\cdots, $ where $f^n$ stands for the function composition of $f$ $n$-times.
 
@@
 distributed in $[0, 2(1+s)]$ -- we can repeat our analysis
-in Section \ref{sec:tent} but represent our initial condition as
+in section \ref{sec:tent} but represent our initial condition as
 $x_0= (1+s) \sum_{j=0}^\infty x_{0,j} / 2^j$.  Its shadowing solution,
@@
   \centering
-    \includegraphics[width=0.48\textwidth]{figure/tilted_tent_map.png}
-    \hspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tilted_conjugate_map.png}
+    \includegraphics[width=0.48\textwidth]{tilted_tent_map.png}
+    \hspace{0.02\textwidth}
+    \includegraphics[width=0.48\textwidth]{tilted_conjugate_map.png}
     \\ \vspace{0.02\textwidth}
@@
 \begin{figure}
-   \includegraphics[width=0.48\textwidth]{figure/tilted_conjugate_density_0.1.png}
-    \hspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tilted_conjugate_density_0.5.png}
+   \includegraphics[width=0.48\textwidth]{tilted_conjugate_density_0.1.png}
+    \hspace{0.02\textwidth}
+    \includegraphics[width=0.48\textwidth]{tilted_conjugate_density_0.5.png}
    \caption{
@@
 This fractal conjugacy map transforms a uniform distribution in $[0,2]$ into
-a fractal distribution similar to the ones plotted in Section \ref{sec:quasiphysical}.
+a fractal distribution similar to the ones plotted in section \ref{sec:quasiphysical}.
 In fact, we will show that the fractal distribution obtained by mapping a uniform
@@
 we used to construct the quasi-physical solution
-in Section \ref{sec:quasiphysical}, with $p=\frac{1+s}2$.
+in section \ref{sec:quasiphysical}, with $p=\frac{1+s}2$.
 Since this analysis holds for any $x$ sampled uniformly on $[0,2]$, we can conclude that
@@
 \subsection{Are general shadowing solutions physical?}
-
+\label{sec:tentPerts}
 In the previous subsection, we examined a tilting perturbation to the tent map.
 We derived the fractal conjugacy map, and showed that a typical shadowing trajectory
-is exactly the quasi-physical solution we analyzed in Section \ref{sec:quasiphysical}.
-This example contrasts with Section \ref{sec:sub:shadow-example}, in which the
+is exactly the quasi-physical solution we analyzed in section \ref{sec:quasiphysical}.
+This example contrasts with section \ref{sec:sub:shadow-example}, in which the
 shadowing solutions of a different perturbation, the scaling perturbation,
@@
     \centering
-    \includegraphics[width=0.48\textwidth]{figure/squashed_tent_map.png}
+    \includegraphics[width=0.48\textwidth]{squashed_tent_map.png}
     \caption{The ``squashed'' tent map (Eq. \ref{eqn:tent_squashed}) for $s=0.01, 0.05, 0,2,$ and 0.5.}
@@
     \centering
-    \includegraphics[width=0.48\textwidth]{figure/tent_squashed_shadow_density_0.01.png}
-    \hspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tent_squashed_physical_density_0.01.png}
+    \includegraphics[width=0.48\textwidth]{tent_squashed_shadow_density_0.01.png}
+    \hspace{0.02\textwidth}
+    \includegraphics[width=0.48\textwidth]{tent_squashed_physical_density_0.01.png}
     \\ \vspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tent_squashed_shadow_density_0.2.png}
-    \hspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tent_squashed_physical_density_0.2.png}
-    \caption{The left column shows the empirical distribution of the shadowing solution for the squashed tent map at $s=0.01$ (top) and at $s=0.2$ (bottom).
-The right column shows the probability distribution of a physical solution at the same two values of $s = 0.01$ (top) and $s = 0.2$ (bottom).}
+    \includegraphics[width=0.48\textwidth]{tent_squashed_shadow_density_0.2.png}
+    \hspace{0.02\textwidth}
+    \includegraphics[width=0.48\textwidth]{tent_squashed_physical_density_0.2.png}
+    \caption{The left column shows the empirical probability distribution of the shadowing solution for the squashed tent map at $s=0.01$ (top) and at $s=0.2$ (bottom).
+The right column shows the empirical probability distribution of a physical solution at the same two values of $s = 0.01$ (top) and $s = 0.2$ (bottom).}
     \label{fig:tent_squashed_shadow}
@@
 at two different values of $s.$ The left column showing the probability distribution of the shadowing solution is reminiscent of the hairy probability distributions of 
-the quasi-solutions of the tent map in Section \ref{sec:quasiphysical}.
+the quasi-solutions of the tent map in section \ref{sec:quasiphysical}.
+The reader is referred to the Supplementary Material section \ref{sec:computingShadowing} for a note on the computation of the shadowing solution.
 Note that, unlike the tilted tent maps, the family of squashed tent maps do not have the uniform distribution as their SRB distribution, as indicated on the right column.
@@
 That is, the physical solutions of the squashed tent map are distributed according to a regular probability density (right column of Figure \ref{fig:tent_squashed_shadow}).
-But the shadowing solution, which is a true orbit of the original tent map, is distributed neither like a physical solution of the original tent map nor of the squashed tent map, as suggested by its fractal-like empirical probability distribution.
+But the shadowing solution, which is a true orbit of the original tent map, is distributed neither like a physical solution of the original tent map nor of the squashed tent map, as suggested by its fractal-like probability distribution.
 
@@
     \centering
-    \includegraphics[width=0.48\textwidth]{figure/pinched_tent_map.png}
+    \includegraphics[width=0.48\textwidth]{pinched_tent_map.png}
     \caption{The pinched tent map (Eq. \ref{eqn:tent_pinched}) for $s = 0.01, 0.05, 0.2,$ and 0.5.}
@@
     \centering
-    \includegraphics[width=0.48\textwidth]{figure/tent_pinched_shadow_density_0.05.png}
-    \hspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tent_pinched_physical_density_0.05.png}
+    \includegraphics[width=0.48\textwidth]{tent_pinched_shadow_density_0.05.png}
+    \hspace{0.02\textwidth}
+    \includegraphics[width=0.48\textwidth]{tent_pinched_physical_density_0.05.png}
     \\ \vspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tent_pinched_shadow_density_0.2.png}
-    \hspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tent_pinched_physical_density_0.2.png}
+    \includegraphics[width=0.48\textwidth]{tent_pinched_shadow_density_0.2.png}
+    \hspace{0.02\textwidth}
+    \includegraphics[width=0.48\textwidth]{tent_pinched_physical_density_0.2.png}
     \caption{The left column shows the empirical distribution of the shadowing solution for the pinched tent map at $s=0.05$ (top) and at $s=0.2$ (bottom).
@@
     \centering
-    \includegraphics[width=0.48\textwidth]{figure/wave_tent_map.png}
+    \includegraphics[width=0.48\textwidth]{wave_tent_map.png}
     \caption{The wave tent map (Eq. \ref{eqn:tent_wave}) for $s = 0.01, 0.05, 0.2,$ and 0.5.}
@@
     \centering
-    \includegraphics[width=0.48\textwidth]{figure/tent_wave_shadow_density_0.05.png}
-    \hspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tent_wave_physical_density_0.05.png}
+    \includegraphics[width=0.48\textwidth]{tent_wave_shadow_density_0.05.png}
+    \hspace{0.02\textwidth}
+    \includegraphics[width=0.48\textwidth]{tent_wave_physical_density_0.05.png}
     \\ \vspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tent_wave_shadow_density_0.2.png}
-    \hspace{0.02\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/tent_wave_physical_density_0.2.png}
+    \includegraphics[width=0.48\textwidth]{tent_wave_shadow_density_0.2.png}
+    \hspace{0.02\textwidth}
+    \includegraphics[width=0.48\textwidth]{tent_wave_physical_density_0.2.png}
     \caption{The left column shows the empirical distribution of the shadowing solution for the wave tent map (Eq.
@@
 Hence, we can once again conclude that almost every solution, which is physical, 
-has a corresponding shadowing solution, that is non-physical. In this section,
+has a corresponding shadowing solution, that is nonphysical. In this section,
 we have shown that the same conclusion holds for several different perturbations 
@@
     \centering
-    \includegraphics[width=0.32\textwidth]{figure/lorenz_zmax_sigma.png}
+    \includegraphics[width=0.32\textwidth]{lorenz_zmax_sigma.png}
     \hspace{0.005\textwidth}
-    \includegraphics[width=0.32\textwidth]{figure/lorenz_zmax_rho.png}
+    \includegraphics[width=0.32\textwidth]{lorenz_zmax_rho.png}
     \hspace{0.005\textwidth}
-    \includegraphics[width=0.32\textwidth]{figure/lorenz_zmax_beta.png}
-    \caption{The left, center, and right plots show the effect of the
+    \includegraphics[width=0.32\textwidth]{lorenz_zmax_beta.png}
+    \caption{
+The left, center, and right plots show the effect of the
     parameters $\sigma, \rho$, and $\beta$ on the Lorenz map, respectively.
@@
     In the right plot, the blue and orange lines represent
-    $\sigma=8/3$ and $10/3$ respectively, while $\sigma = 10$ and $\rho = 28.$}
+    $\sigma=8/3$ and $10/3$ respectively, while $\sigma = 10$ and 
+		$\rho = 28.$}
     \label{fig:lorenz_params}
@@
     \centering
-    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_density_lorenz_zmax_15_28_2.67.json.png}
+    \includegraphics[width=0.48\textwidth]{lorenz_shadow_density_lorenz_zmax_15_28_2.67.json.png}
     \hspace{0.005\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/baseline_density_lorenz_zmax_15_28_2.67.json.png}
+    \includegraphics[width=0.48\textwidth]{baseline_density_lorenz_zmax_15_28_2.67.json.png}
     \\
-    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_density_lorenz_zmax_10_30_2.67.json.png}
+    \includegraphics[width=0.48\textwidth]{lorenz_shadow_density_lorenz_zmax_10_30_2.67.json.png}
     \hspace{0.005\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/baseline_density_lorenz_zmax_10_30_2.67.json.png}
+    \includegraphics[width=0.48\textwidth]{baseline_density_lorenz_zmax_10_30_2.67.json.png}
     \\
-    \includegraphics[width=0.48\textwidth]{figure/lorenz_shadow_density_lorenz_zmax_10_28_3.33.json.png}
+    \includegraphics[width=0.48\textwidth]{lorenz_shadow_density_lorenz_zmax_10_28_3.33.json.png}
     \hspace{0.005\textwidth}
-    \includegraphics[width=0.48\textwidth]{figure/baseline_density_lorenz_zmax_10_28_3.33.json.png}
+    \includegraphics[width=0.48\textwidth]{baseline_density_lorenz_zmax_10_28_3.33.json.png}
     \caption{The left column shows the empirical distribution of the shadowing solution of the Lorenz map at $\rho = 28$ (top row), $\rho = 29$ (middle row), and $\beta = 10/3$ (bottom row).
@@
 We again notice that, while the physical distributions are smooth, the shadowing solutions have a rough distribution that appears to have hair-like patterns.
-Thus, we draw the same conclusion in the Lorenz map as in the various perturbations of the tent map treated in section \ref{sec:tent}: shadowing solutions do not have the same distribution as a typically observed solution of the governing equation.
+Thus, we draw the same conclusion in the Lorenz map as in the various perturbations of the tent map treated in section \ref{sec:tentPerts}: shadowing solutions do not have the same distribution as a typically observed solution of the governing equation.
 In other words, shadowing solutions may be nonphysical.
 
-\section{Discussions}
+\section{Discussion and outlook}
+
 Through a rigorous counter-example, we show that shadowing can lead
-to nonphysical solutions.  The conclusion has an immediate implication for
-sensitivity analysis methods based on the concept of shadowing \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagna}.
+to nonphysical solutions. This has a troubling implication for numerical simulations of chaotic
+governing equations.  Through the examples in this paper, it is clear that a numerical solution, satisfying
+a governing equation perturbed due to numerics, is not 
+expected to be shadowed by a
+physical solution of the real governing physics.  What is, if any, the
+relation of a numerical solution, such as DNS of turbulent flows, to
+the true physics? For an ${\cal O}(10^{-18})$ 
+perturbation (due to numerical error), 
+we may get completely different statistics from a true, physical solution.
+How then can we trust numerical solutions when they are not guaranteed to share the long-term statistical behavior of the governing equation?
+We conclude with an example that illustrates a counterintuitive feature of some chaotic systems: a small perturbation to the 
+governing equation can significantly change the statistics/long-term 
+behavior of its physical solutions. We construct yet another 
+perturbed tent map -- the \emph{plucked} 
+tent map -- in which an oscillatory perturbation is introduced (see Supplementary Material section \ref{sec:pluckedTentMap} for the map equation).
+In Figure \ref{fig:pluckedTentMap}, we show that the magnitude of the oscillatory perturbation is controlled by a parameter $s$, and its frequency is controlled by the parameter $n$; at $s=0,n=0$, we recover the original tent map.
+The stationary probability distributions -- computed empirically over a trajectory of length 10 billion -- at $s=0.5$, are shown in Figure \ref{fig:pluckedTentMapDistribution}.
+On the left, $n=0$ and there is already a marked asymmetry developed in the probability distribution compared to the uniform distribution seen at $s=0, n=0$.
+On the right, we show that by increasing $n,$ while keeping the magnitude of the perturbation the same ($s=0.5$), we see a dramatic change in the appearance of the stationary probability distribution: we observe a fractal distribution.
+
+In view of the plucked tent map, we can question further the
+validity of numerical solutions. Can numerical play 
+the role of the plucking perturbation to the 
+governing equation? That is, can numerical solutions represent 
+slight perturbations to the governing equations that nevertheless 
+exhibit drastically ensemble statistics? Uniformly hyperbolic 
+dynamical systems exhibit \emph{linear response} 
+(\cite{ruelle}\cite{baladi}) by which small parameter perturbations 
+lead to small changes in statistics, which can be expanded as Taylor series around the reference parameter value.
+But, uniform hyperbolicity is a mathematical idealization, and although some physical systems have been observed to behave as if they were uniformly hyperbolic \cite{gallavotti}, a violation of uniform hyperbolicity is more likely
+\cite{wormell}\cite{luca}. The plucked tent map provides a pathological example of a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).
+
+Both the nonphysicality of shadowing solutions, and the existence of pathological statistical response, undermine the validity of using shadowing for sensitivity analysis of statistics \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagna}.
 When the goal is to compute derivatives of ensemble averages, where the ensemble
-is distributed according to the physical measure, i.e., the SRB measure, shadowing-based methods can give wrong results.  This has been observed
-before and appropriately attributed as ``ergodicity breaking error'' \cite{patrick} \cite{angxiu-error}. In this paper, we construct rigorous analytical examples of this error.
+is distributed according to the physical measure, i.e., the SRB measure, shadowing-based methods can give wrong results.
+This is because shadowing-based methods compute the sensitivities of ensemble statistics along shadowing solutions, but these may not be physical solutions that reproduce ensemble statistics.
+The error in shadowing sensitivities has been observed
+before and appropriately attributed as ``ergodicity breaking error'' \cite{patrick} \cite{angxiu-error}. 
 \begin{figure}
     \centering
-    \includegraphics[width=0.45\textwidth]{figure/osc_tent_n.png}
-    \includegraphics[width=0.45\textwidth]{figure/osc_tent_s.png}
+    \includegraphics[width=0.45\textwidth]{osc_tent_n.png}
+    \includegraphics[width=0.45\textwidth]{osc_tent_s.png}
     \caption{The plucked tent map shown at different values of $s$ and $n$, that represent the magnitude and frequency of the applied oscillation to the original tent map ($n=0,s=0$).
@@
 On the right, $n$ is held constant at 0.}
-    \label{fig:osc_map}
+    \label{fig:pluckedTentMap}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{figure/osc_tent_dens_0.5_0.png}
-    \includegraphics[width=0.45\textwidth]{figure/osc_tent_dens_0.5_6.png}
+    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.5_0.png}
+    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.5_6.png}
     \caption{The effect of increasing $n$ on the probability distribution associated to the plucked tent map.
@@
 R: The stationary probability distribution of the plucked tent map at $s=0.5, n = 6$.}
-    \label{fig:map}
-\end{figure}
-Even more troubling is the implication for numerical simulations of chaotic
-governing equations.  Through the examples in this paper, it is clear that a numerical solution, satisfying
-a governing equation perturbed due to numerics, is not expected to shadow a
-physical solution of the real governing physics.  What is, if any, the
-relation of a numerical solution, such as DNS of turbulent flows, to
-the true physics? For $10^{-18}$ perturbation (due to numerics), 
-we may get completely different statistics from a true, physical solution. 
-How then can we trust numerical solutions that, even if they are shadowed by
-physical solutions, do not share their statistics?
-
-We conclude with an example that illustrates that assuming the 
-physicality of shadowing solutions leads to incorrect results in shadowing-based sensitivity 
-computation methods. We construct yet another perturbed tent map -- the \emph{plucked} 
-tent map -- in which an oscillatory perturbation is introduced.
-In Figure \ref{fig:osc_map}, we show that the magnitude of the oscillatory perturbation is controlled by a parameter $s$, and its frequency is controlled by the parameter $n$; at $s=0,n=0$, we recover the original tent map.
-The plucked tent map illustrates an important feature of some chaotic systems: with a small change in a parameter, the stationary probability distribution of a chaotic system \emph{can} change drastically.
-For the plucked tent map, this feature is illustrated by the stationary probability distributions -- computed empirically over a trajectory of length 10 billion -- at $s=0.5$.
-On the left, $n=0$ and there is already a marked asymmetry developed in the probability distribution compared to the uniform distribution seen at $s=0, n=0$.
-On the right, we show that by increasing $n,$ while keeping the magnitude of the perturbation the same ($s=0.5$), we see a dramatic change in the appearance of the stationary probability distribution: we observe a fractal distribution.
-On the other hand, if perturbed solutions are shadowed by physical solutions for all time, the change in the statistics of the perturbed solutions must be small.
+    \label{fig:pluckedTentMapDistribution}
+\end{figure}
+Violations of smooth response, like in the plucked tent map, has an immediate implication for shadowing sensitivities as well.
+In particular, if perturbed solutions are shadowed by physical solutions for all time, the change in the statistics of the perturbed solutions must be small.
 In other words, the physicality of shadowing predicts that there cannot be a large change in statistics due to small parameter perturbations; the reality is that, as illustrated by the plucked tent map, the effect of small parameter changes on the statistics can be drastic.
 
-In this paper, we have constructed several counterexamples that disprove the common 
-assumption that shadowing solutions are physical solutions, i.e., that their 
+In this paper, we have constructed several counterexamples that 
+dispel the notion that shadowing solutions are physical solutions, i.e., that their 
 statistical distribution is the same as that typically observed for almost every solution 
@@
 perturbed solutions, represent the true physics implied by the governing equation. In light of the evidence in this paper, we must reopen this issue. Even when numerical simulations are 
-shadowed by a true solution in that the difference between them is small for a long time, this shadowing solution may not represent the long-term behavior of the physical system.
+shadowed by a true solution in that the difference between them is small for a long time, this shadowing solution may not represent the long-term or ensemble behavior of the physical system.
 The nonphysicality of shadowing solutions also indicates that shadowing-based methods \cite{qiqi-lss} can lead to incorrect values of sensitivities of statistics to parameter changes.
@@
 \bibliography{refs.bib}
-\include{supp}
+\section{Supplementary material}
+%\subsection{The Lorenz Map}
+\subsection{Approximation of the Lorenz Map}
+\label{sec:lorenzMapApprox}
+The motivation for approximating the Lorenz map is that a 
+closed form expression for the map is necessary for our numerical shadowing procedure.
+In a small region around the cusp, we approximate the map using an exponential function.
+The tails on both sides are fitted with a sum of a cubic polynomial and a rational function.
+Thus, the approximate Lorenz map has the
+following closed form expression:
+\begin{align}
+    \varphi(x) = \begin{cases}
+        z_{\rm max} - f_R(x - z_{\rm sep})   &  x > z_{\rm sep} \\
+        z_{\rm max} -  f_L(z_{\rm sep} - x)   &  x \leq z_{\rm sep}
+    \end{cases}
+\end{align}
+where 
+\begin{align}
+    f_R(y) = \Big( (1000 y)^p + 
+        \sum_{n=0}^3 p_{R,n} y^n + \dfrac{a_{R,0} + a_{R,1} y}{\sum_{n=0}^3 b_{R,n} y^n}\Big),
+\end{align}
+and,
+\begin{align}
+    f_L(y) = \Big((1000 y)^p 
+        + \sum_{n=0}^3 p_{L,n} y^n + \dfrac{a_{L,0} + a_{L,1}y}{\sum_{n=0}^3 b_{L,n} y^n}\Big).
+\end{align}
+\begin{figure}
+		\includegraphics[width=0.45\textwidth]{lorenz_map_fit_lorenz_zmax_10_28_2.6666666666666665.npy.png}
+		\includegraphics[width=0.45\textwidth]{lorenz_map_fit_error_lorenz_zmax_10_28_2.6666666666666665.npy.png}
+		\caption{L: the blue and orange lines are the Lorenz map at the standard parameters, and the dotted black lines indicate the approximate Lorenz map (Table \ref{tab:lorenzFit}).
+R: the regression error is shown as a function of $z$}
+	\label{fig:lorenzMapComp}
+\end{figure}
+The location of the cusp is denoted $z_{\rm sep}$ (in which ``sep'' stands 
+for \emph{separation}), and the maximum and minimum values encountered 
+in the Lorenz map iterates are denoted $z_{\rm max}$ and $z_{\rm min}$, respectively.
+The exponent of the cusp is denoted $p$.
+The coefficients of the cubic polynomial modelling the left (right) tail are  denoted $p_{L,n}, n = 0,1,2,3$ ($p_{R,n}, n = 0,1,2,3$).
+The coefficients of the numerator and denominator of the rational function modelling the left (right) tail are denoted $a_{L,0}, a_{L,1}$ ($a_{R,0}, a_{R,1}$) and $b_{L,n}, n=0,1,2,3$ ($b_{R,n}, n=0,1,2,3$), respectively.
+The values of these coefficients are obtained by
+regression on the map at standard parameters are shown in the table below. As shown in Figure \ref{fig:lorenzMapComp}, the fit obtained matches the Lorenz map closely.
+\begin{table}[H]
+    \centering
+    \begin{tabular}{|c|c|}
+    \hline 
+         $z_{\rm sep}$ = 38.55302437476555 &
+         $z_{\rm min}$ =  29.213182255013322 \\ 
+         $z_{\rm max}$ =  47.978140718671284 & 
+         $p$ = 0.28796740575434676   \\
+         $p_{L,3}$ = -0.00024683786275242047 & $p_{L,2} = $ 
+         0.016174566354858824 \\
+         $p_{L,1} = $ 0.40179772568004946 & 
+         $p_{L, 0}  = $ -0.24612651488351725 \\
+         $p_{R,3} = $ -0.00020712463321688308 & 
+         $p_{R,2} = $ 0.017130843276711716 \\
+         $p_{R, 1} = $ 0.3930080703420676 & 
+         $p_{R, 0} = $ -0.23471384266765036 \\
+         $a_{L,1}$ = -0.05405742075580959 & $a_{L,0}$ = -0.05405742075580959 \\
+         $a_{R,1}$ = -0.05351127783496397 & $a_{R,0}$ = 0.22489891059122896 \\
+         $b_{L,3}$ = 0.5609397451213353 & $b_{L,2}$ = -0.3491184293228338 \\  $b_{L,1}$ = 2.419972619058592 & $b_{L,0}$ = 1.0 \\
+         $b_{R,3}$ = 0.6456076059873844 & $b_{R,2}$ = -0.34840383986411055 \\ $b_{R,1}$ = 2.6035438510917692 &  $b_{R,0}$ = 1.0 \\
+         \hline
+    \end{tabular}
+    \caption{The fitting parameters of the Lorenz map at $\sigma = 10, \beta = 8/3$ and $\rho = 28.$}
+    \label{tab:lorenzFit}
+\end{table}
+
+\subsection{Computing shadowing solutions}
+\label{sec:computingShadowing}
+		In general, to numerically compute the shadowing solutions, one could use existing methods such as 
+		the least squares shadowing method \cite{qiqi-lss}.
+However, since these maps are all one-dimensional chaotic systems that have in common non-invertibility with two inverse \emph{branches}, we devise a simpler approach.
+The map $\varphi_s$ in this section refers to any of the perturbations of the tent map (section \ref{sec:tentPerts}) or the Lorenz map.
+Suppose we are given a perturbed solution $x_n, n =0,1,\cdots,N$ that we must compute a shadow of.
+The shadowing solution, $y_n, n=0,1,2,\cdots,N$ 
+must a) satisfy the governing equation: $y_{n+1} = \varphi_s(y_n)$
+and b) lie close to the given perturbed solution at all times up to 
+$N$, i.e., $|x_n - y_n| < \epsilon$, for some $\epsilon > 0$, for all $n \leq N$.
+Suppose we set $y_0 = x_0$, the difference $y_1 - x_1$ will amplify on further iterations under $\varphi_s$ for any perturbed solution.
+On the other hand, backward iteration of the map is contracting, and thus we set $y_N = x_N$ and proceed backward in time to
+construct a solution of $\varphi_s$, noting that any difference that emerges at a given time $n$ will be made smaller starting at $n-1$.
+Each point has two pre-images under $\varphi_s$, and thus there are two possible choices for $y_{N-1}$, each lying in one of two fixed sub-intervals separated by the cusp of $\varphi_s$.
+Due to the contraction of errors backward in time, as long as we choose a pre-image in the same subinterval as the perturbed solution, we are guaranteed to approximate a shadowing solution.
+Marching backward by choosing at each step, the pre-image $y_n$ in the same subinterval as $x_n$, $y_n$ approximates the shadowing solution better as $n$ decreases.
+Thus, the procedure to find a shadowing solution simply reduces to solving for a backward trajectory (specifically one among the possible $2^N$), starting at a given final condition $x_N$.
+
+		Hence, it is clear that all we need is the inverse of 
+		$\varphi_s$, which is propagated backward by choosing the same \emph{branch} of the inverse as $x_n$ at time $n$.
+This logic is implemented for each map of section \ref{sec:tentPerts} in the function \verb+shadow+ that can be found in the files named for each map (for example, the shadowing solution of the pinched tent map can be found by executing the \verb+shadow+ function of \verb+tent_shadow/tent_shadow_pinched.py+).
+These functions use the analytical inverses of the maps, which are easy to derive for the tent map perturbations of section \ref{sec:tentPerts}.
+For the Lorenz map, we use Newton's method to solve for the inverse, and this is implemented in the file \verb+lorenz_map/shadow.py+.
+Note that we need a closed form expression of the map, for the Newton's method, and this is indeed the reason why we approximate the map, as described in section \ref{sec:lorenzMapApprox}.
+
+\subsection{The \emph{plucked} tent map}
+\label{sec:pluckedTentMap}
+We provide a recursive definition of the plucked tent map, which 
+is illustrated at different values of $n$ and $s$ in Figure \ref{fig:pluckedTentMap}. First we define a function $f_s(x)$, which 
+creates a bend in the tent map that increases with $s$, around 
+$((1-s)/2,1)$:
+\begin{align}
+		f_s(x) = 		{\rm min}\Big(
+				\dfrac{2x}{1-s}, 2 - \dfrac{2(1-x)}{1+s}\Big),
+				\;\;\; x < 1 \\
+\end{align}
+Then, we introduce oscillations using repititions of
+the above map within the unit interval,
+\begin{align}
+		o_s(x) = \begin{cases}
+				f_s(2x)/2, & \; x < 0.5 \\
+				2 - f_s(2-2x)/2, & \; x \ge 0.5.
+		\end{cases}
+\end{align}
+We can modulate
+the frequency or repititions -- proportional to $n$ -- 
+of the oscillations through
+\begin{align}
+		\label{eqn:frequency}
+		\lambda_{s,n}(x) = \dfrac{o_s(2^n x - \floor*{2^n x})}{2^n} 
+		+ 2 \frac{\floor*{2^n x}}{2^n},
+\end{align}
+where $\floor*{x}$ is the greatest integer less than or equal to $x$.
+Finally, the plucked tent map is defined as the above function 
+in the unit interval, and as its reflection about $x=1$,
+in the interval $[1,2)$.
+\begin{align}
+		\label{eqn:plucked}
+		\varphi_{s,n}(x) = {\rm min}\left(
+		\lambda_{s,n}(x), \lambda_{s,n}(2-x)\right), \; 0< x < 2. 
+\end{align}
+Recall that our motivation is to construction a slight perturbation of the tent map whose stationary probability distribution is not just 
+non-uniform (e.g.
+like the other tent map perturbations of section \ref{sec:tentPerts}) but in which the nonuniformity can be controlled, and made skewed as desired.
+Thus, we choose to construct a base perturbation of the tent map, ${\rm min}(o_s(x), o_s(2-x))$ --
+which is also obtained by setting $n=0$ in $\varphi_{s,n}$ --  
+in which an asymmetry is produced in the probability distribution (left hand side of Figure \ref{fig:pluckedTentMapDistribution}).
+Upon repeating, with appropriate scaling, the oscillation $o_s$, 
+the frequency of which is controlled by $n$, we obtain $\varphi_{s,n}$, 
+as indicated by Eq.
+\ref{eqn:frequency}-\ref{eqn:plucked}.
+Thus, the asymmetric probability distribution, through this process of scaled repitition, acquires the fractal-like structure seen on increasing $n$, as shown on the right hand side of Figure \ref{fig:pluckedTentMapDistribution}.
+
+Consider $s=0.5$. 
+One way to see the asymmetry in the probability distribution 
+about $x=1$, at $n=0$, is to construct a Markov chain with nodes 00, 01, 10, and 11, that indicate the subintervals $(0,0.5)$, $[0.5,1)$, 
+$[1,1.5)$ and $[1.5,2)$, respectively. From Eq. \ref{eqn:plucked}, we can see that the application of $\varphi_{0.5,0}$ yields the following transition matrix:
+\[
+\begin{bmatrix}
+		1/4 & 3/4 & 0 & 0 \\
+		0 & 0 & 3/4 & 1/4 \\
+		0 & 0 & 3/4 & 1/4 \\
+		1/4 & 3/4 & 0 & 0
+\end{bmatrix}
+.\] 
+The stationary probability distribution of this Markov chain is $[1/8, 3/8, 9/8, 3/8]^T$, consistent with the left hand side of Figure \ref{fig:pluckedTentMapDistribution}.
+Clearly, the left subinterval, $[0,1)$ is three times less likely to be visited by a long trajectory of $\varphi_{0.5,0}$, compared to $[1,2)$.
+In fact, the probability of visiting $[0,1)$ remains at 0.25 for all values of $n$, but for $n >> 0$, fractal distributions are produced within smaller subsets of the two subintervals.
 \end{document}

### 2020-10-26 main.tex
Updates from Overleaf

--- 2020-10-23 main.tex
@@
 The question therefore arises as to whether or not numerical solutions to chaotic systems, such
-as turbulent flows, are related to the physics at all.  The existence of shadowing solutions
+as turbulent flows, represent the true physics.  The existence of shadowing solutions
 is used to argue for the usefulness of such numerical solutions.
@@
 \end{figure}
-\begin{comment}In Figure 4, the solid line appears only to have one local maximum, which is shown as the fixed point on the left plot of this figure.
-The two fixed points of the 2-time-iterate of the Lorenz map, shown on the right plot of this figure, both appear to correspond to the local maxima of the dashed curve in Figure 4.\end{comment}
+\nisha{In Figure 4, the solid line appears only to have one local maximum, which is shown as the fixed point on the left plot of this figure.
+The two fixed points of the 2-time-iterate of the Lorenz map, shown on the right plot of this figure, both appear to correspond to the local maxima of the dashed curve in Figure 4.}
 The Lorenz map provides us a tool to find as many periodic solutions as we want.

### 2020-10-26 code/tent_sens_stat/plot_map.py

@@ -38,32 +38,21 @@ def accumulate(nbins, n, m, s):
 nbins = 2**13
 ns = 5
 nn = 5
-n_arr = [0, 1, 2, 4, 6]
-s_arr = [0.005, 0.1, 0.2, 0.4, 0.5]
+n_arr = [0, 2, 4, 6]
+s_arr = [0.2, 0.5]
 
 x = linspace(0, 2, nbins+2)[1:-1]
 ys = zeros((ns,nbins))
 yn = zeros((nn,nbins))
-for k,nk in enumerate(n_arr):
-    for i, xi in enumerate(x):
-        yn[k,i] = osc_tent(xi, 0.5, nk)
-fig = figure(figsize=(8,6))
-ax = fig.add_subplot(111)
-for k,nk in enumerate(n_arr):
-    ax.plot(x, yn[k],lw=2.5,label="n = {}".format(nk))
-ax.xaxis.set_tick_params(labelsize=40)
-ax.yaxis.set_tick_params(labelsize=40)
-leg = ax.legend(fontsize=30)
-ax.axis("scaled")
-
-for k,sk in enumerate(s_arr):
-    for i, xi in enumerate(x):
-        ys[k,i] = osc_tent(xi, sk, 0)
-fig = figure(figsize=(8,6))
-ax = fig.add_subplot(111)
-for k,sk in enumerate(s_arr):
-    ax.plot(x, ys[k],lw=2.5,label="s = {}".format(sk))
-ax.xaxis.set_tick_params(labelsize=40)
-ax.yaxis.set_tick_params(labelsize=40)
-ax.axis("scaled")
-leg = ax.legend(fontsize=30)
+for kn,nk in enumerate(n_arr):
+    for k,sk in enumerate(s_arr):
+        for i, xi in enumerate(x):
+            yn[k,i] = osc_tent(xi, sk, nk)
+    fig = figure(figsize=(8,6))
+    ax = fig.add_subplot(111)
+    for k,sk in enumerate(s_arr):
+        ax.plot(x, yn[k],lw=2.5,label="s = {}".format(sk))
+    ax.xaxis.set_tick_params(labelsize=40)
+    ax.yaxis.set_tick_params(labelsize=40)
+    leg = ax.legend(fontsize=30)
+    ax.axis("scaled")

### 2020-10-26 main.tex
Add readme and modified tex file as on arxiv

--- 2020-10-26 main.tex
@@
 %\subsection{The Lorenz Map}
+The supplementary material for this paper including the code, data and figures can be found on Github \cite{supp}.
+The code can be found under the \verb+code+ subdirectory inside of which section-wise code is separated into further subdirectories.
+The data used to plot the figures can be found under \verb+data+.
+The files referred to in this section can be found in the appropriate subdirectory under \verb+code+.
 \subsection{Approximation of the Lorenz Map}

### 2020-10-26 main.tex
Fix type errors up to page 8

@@
 
-\institute{Q. Wang \at 
-        Department of Aeronautics and Astronautics, Center for Computational Science and Engineering, MIT, Cambridge, MA 02139, USA.
-              \email{qiqi@mit.edu}           %  \\
+\institute{           %  \\
 %             \emph{Present address:} of F. Author  %  if needed
-           \and
-           N. Chandramoorthy \at Department of Mechanical Engineering, Center for Computational Science and Engineering, MIT, Cambridge, MA 02139, USA. \email{nishac@mit.edu}
+           Center for Computational Science and Engineering, MIT, Cambridge, MA 02139, USA. \email{nishac@mit.edu 
+           }
 }
@@
 
-In this section, we first describe mathematically what distinguishes a physical
-and a nonphysical solution.  We also explain theoretically why it is unlikely
+In this section, we first describe what distinguishes, mathematically, a physical
+and a nonphysical solution.  We also explain why it is theoretically unlikely
 to observe a nonphysical solution in an experiment.  We then give a few
@@
 \begin{enumerate}
-    \item Long-time-averaged, ``statistical'' quantities converges as time goes to infinity.
+    \item Time-averaged quantities converge in the limit of infinite averaging time.
     Consider $u(t)$ as the solution to a chaotic governing equation, then for a regular observable
@@
 \caption{
-Left: An ensemble of initial conditions distributed uniformly in the box $u_1\in[0,1], u_2\in[0,1], u_3\in[28,39]$ 
-after 10 time units of evolution, shown in in the $u_1-u_3$ plane.
-Right: distribution of a trajectory of $1000$ time units in length.
+L: An ensemble of initial conditions distributed uniformly in the box $u_1\in[0,1], u_2\in[0,1], u_3\in[28,39]$ 
+after 10 time units of evolution, shown on the $u_1-u_3$ plane.
+R: distribution of a trajectory of $1000$ time units in length.
 In both plots, the color represent the number of samples in a
@@
 \caption{
-Left: distribution of the same ensemble as in Figure \ref{fig:lorenz_ergodicity1}
+L: distribution of the same ensemble as in Figure \ref{fig:lorenz_ergodicity1}
 after another 5 time units of evolution.
-Right: distribution of the same trajectory as in Figure \ref{fig:lorenz_ergodicity1}
+R: distribution of the same trajectory as in Figure \ref{fig:lorenz_ergodicity1}
 evolved for $10,000$ time units.
@@
 \caption{
-Left: distribution of the same ensemble as in Figures \ref{fig:lorenz_ergodicity1}
+L: distribution of the same ensemble as in Figures \ref{fig:lorenz_ergodicity1}
 and \ref{fig:lorenz_ergodicity2} after a total of 50 time units of evolution.
-Right: distribution of the same trajectory as in Figures \ref{fig:lorenz_ergodicity1}
+R: distribution of the same trajectory as in Figures \ref{fig:lorenz_ergodicity1}
 and \ref{fig:lorenz_ergodicity2} evolved for $100,000$ time units.
@@
 the solution has visited a large portion of the $u_1-u_3$ plane with a varying, but well-defined, frequency.  We can use a probability distribution, $\mu$, to quantify
-how frequently a physical solution visits portions the phase space.
+how frequently a physical solution visits portions of the phase space.
 Specifically, for a subset $A$ of the phase space, $\mu(A)$ measures the
@@
 larger portion of the $u_1-u_3$ plane.
-After a long time, the ensemble settles into a set known as an \emph{attractor} that does not change with time, and contains the \emph{unstable manifold}, which appears as filaments forming the attractor.
-The distribution of the ensemble on the attractor becomes identical to the distribution of a single, very long physical
+After a long time, the ensemble settles into a time-invariant \emph{attractor} that contains the \emph{unstable manifold}, which appears as filaments forming the attractor.
+The probability distribution of the ensemble on the attractor becomes identical to the distribution of a single, very long physical
 solution, which is also contained in the attractor (Figure \ref{fig:lorenz_ergodicity3}).
@@
 the solid, dashed, and dotted lines, respectively.
-We observe from the table that although the mean of $J_1$ over the periodic
+We observe from the table that the mean of $J_1$ over the periodic
 solutions are different but comparable to the mean over physical solutions.  The mean of $J_2$ over the periodic solutions, on the other hand,
@@
 \end{figure}
-\nisha{In Figure 4, the solid line appears only to have one local maximum, which is shown as the fixed point on the left plot of this figure.
-The two fixed points of the 2-time-iterate of the Lorenz map, shown on the right plot of this figure, both appear to correspond to the local maxima of the dashed curve in Figure 4.}
 The Lorenz map provides us a tool to find as many periodic solutions as we want.

### 2020-10-26 main.tex
Edited most sections. Last pass before submitting.

@@
 for which the next maximum is almost the same value.  We then look up the values of
-$u_1(t)$ and $u_2(t)$ when $u_3(t)$ achieved this maximum.  This gives us an initial
+$u_1(t)$ and $u_2(t)$ when $u_3(t)$ achieves this maximum.  This gives us an initial
 condition starting from which the solution is nearly periodic.   We can similarly
-intersect the second iterate of the Lorenz map (right plot of Figure \ref{fig:lorenz_map},)
-and the third iterate, etc, with a diagonal line to find increasingly
+intersect the second iterate of the Lorenz map (right plot of Figure \ref{fig:lorenz_map}),
+and the third iterate, etc, with a diagonal line, to find increasingly
 complex periodic solutions.
@@
 can lead to a large change in the y-axis. This sensitivity grows exponentially
-for the $n$-th iterate of this map. A solution to the Lorenz map can be obtained
+for iterates of this map. A solution to the Lorenz map can be obtained
 by extracting the consecutive local maxima of a solution to the Lorenz equation.
@@
 will help us, in section \ref{sec:quasiphysical}, to study a more insidious type
-of nonphysical solutions, a type that hides in the shadow.
+of nonphysical solutions that hides in shadowing solutions.
 
@@
 The tent map has been widely studied \cite{tent1}\cite{tent2}.  It is qualitatively similar
-to the Lorenz map and has a simple, analytic form:
+to the Lorenz map and has a simple analytical form:
 \begin{equation} \label{tentmap}
@@
 with ergodic theory because all rational numbers in $[0,2]$ comprise a subset of Lebesgue measure zero. That is, we would have zero likelihood of getting a
-rational number if sampled the Lebesgue measure (the uniform distribution) on $[0,2]$.  Instead, with a 100\% probability, one would get an irrational number
+rational number if we sampled the Lebesgue measure (the uniform distribution) on $[0,2]$.  Instead, with a 100\% probability, one would get an irrational number
 that, when iterated under the tent map, leads to a physical trajectory that distributes uniformly on the interval $[0,2]$.
@@
 \end{equation}
-would satisfy
+satisfies
 \begin{equation} \label{tent_xor}
@@
 and each bit is still independent of the other bits.
-Thus, every $x_i$ in this physical solution is uniformly distributed in $[0,2]$.
 As the map iterates starting from almost any $x_0$, a physical solution explores the entire interval $[0,2]$ uniformly.
@@
 \label{sec:quasiphysical}
-The simplicity of the tent map, as well as its binary form (\ref{tent_xor}),
+The simplicity of the tent map, as well as its binary form (Eq. \ref{tent_xor}),
 enables us to study nonphysical solutions that are not periodic.
@@
 \begin{equation}
-    x_{i+1,0} = x_{i,0} \veebar x_{i,1}
+    x_{i+1,0} = x_{i,0} \veebar x_{i,1}.
 \end{equation}
@@
 
-If $p = 1$, the bits of $x_0$ are either all zeros or all ones.
+If $p = 1$, the bits of $x_0$ are either all zeros or all ones, which is the same as all zeros in mod 2 arithmetic.
 So, all further iterates when $x_0$ is 0, are 0, and this is a trivial nonphysical solution of type 1.
-When $x_0$ consists of all ones, the solution eventually becomes 0, and hence is also a nonphysical solution that is eventually periodic (of period 1).
-Even when this solution does not attain zero in a large number of iterations, it does not
-distribute uniformly on $[0,2]$ like a physical solution 
-because all $x_i, i\geq 1$ are of the form $1/2^k$ for some $k \in \mathbb{N}$; other numbers are not visited. Now when $p > \frac12$ but strictly less than 1, 
+Now when $p > \frac12$ but strictly less than 1,
 the solutions we just constructed are both nonphysical and aperiodic.
@@
 \includegraphics[width=0.48\textwidth]{tent_quasiphysical_hist_p_0.9.png}
-\caption{Empirical density function of long solutions (a billion steps)
+\caption{Empirical probability distribution of long solutions (a billion steps)
 starting from four points whose binary digits have probability $p$ of
@@
 top-right: $p=0.51$; bottom-left: $p=0.55$; bottom-right: $p=0.9$.
-Note that the empirical densities at larger $p'$s are well above 1.}
+}
 \label{fig:tent_quasiphysical}
@@
 When $p=0.9$, the fractal pattern is so intensified that most
-of the solution seem to concentrate in a collection of tiny intervals.
+of the solution seems to concentrate in a collection of tiny intervals.
 These plots expose the diversity of quasi-physical solutions.
@@
 more likely to be 1 than 0 only if it follows two consecutive 0's.
-Such $x_0$ would lead to a nonphysical solution.  Its statistical
+Such an $x_0$ would lead to a nonphysical solution.  Its statistical
 distribution would differ from any of the plots in Figure
@@
 On the left, $n=0$ and there is already a marked asymmetry developed in the probability distribution compared to the uniform distribution seen at $s=0, n=0$.
-On the right, we show that by increasing $n,$ while keeping the magnitude of the perturbation the same ($s=0.5$), we see a dramatic change in the appearance of the stationary probability distribution: we observe a fractal distribution.
+On the right, we show that by increasing $n,$ while keeping the magnitude of the perturbation the same ($s=0.5$), we see a dramatic change in the appearance of the stationary probability distribution: we observe an apparently fractal distribution reminiscent of that of the quasi-physical solutions in
+section \ref{sec:tentPerts}.
 
@@
 slight perturbations to the governing equations that nevertheless 
-exhibit drastically ensemble statistics? Uniformly hyperbolic 
+exhibit drastically different statistics from that of the ensemble? Uniformly hyperbolic 
 dynamical systems exhibit \emph{linear response} 
@@
 
-Both the nonphysicality of shadowing solutions, and the existence of pathological statistical response, undermine the validity of using shadowing for sensitivity analysis of statistics \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagna}.
+Both the nonphysicality of shadowing solutions, and the existence of extremely non-smooth statistical response, undermine the validity of using shadowing for sensitivity analysis of statistics \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagna}.
 When the goal is to compute derivatives of ensemble averages, where the ensemble
@@
 \end{figure}
-Violations of smooth response, like in the plucked tent map, has an immediate implication for shadowing sensitivities as well.
+Violations of smooth response, like in the plucked tent map, have an immediate implication for shadowing sensitivities as well.
 In particular, if perturbed solutions are shadowed by physical solutions for all time, the change in the statistics of the perturbed solutions must be small.
@@
 \end{align}
-Then, we introduce oscillations using repititions of
+Then, we introduce oscillations using repetitions of
 the above map within the unit interval,
@@
 We can modulate
-the frequency or repititions -- proportional to $n$ -- 
+the frequency or repetitions -- proportional to $n$ -- 
 of the oscillations through
@@
 \ref{eqn:frequency}-\ref{eqn:plucked}.
-Thus, the asymmetric probability distribution, through this process of scaled repitition, acquires the fractal-like structure seen on increasing $n$, as shown on the right hand side of Figure \ref{fig:pluckedTentMapDistribution}.
+Thus, the asymmetric probability distribution, through this process of scaled repetition, acquires the fractal-like structure seen on increasing $n$, as shown on the right hand side of Figure \ref{fig:pluckedTentMapDistribution}.
 

### 2020-10-27 main.tex
plot_map update by qiqi
Add Sauer reference.

@@
 But, uniform hyperbolicity is a mathematical idealization, and although some physical systems have been observed to behave as if they were uniformly hyperbolic \cite{gallavotti}, a violation of uniform hyperbolicity is more likely
-\cite{wormell}\cite{luca}. The plucked tent map provides a pathological example of a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).
+\cite{wormell}\cite{luca}. Sauer \cite{sauer} has shown 
+examples in which numerical error due to computation in double-precision floating point arithmetic cause significant change in the stationary probability distribution on the attractor.
+The plucked tent map adds to this list of examples by exhibiting a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).
 
@@
 In fact, the probability of visiting $[0,1)$ remains at 0.25 for all values of $n$, but for $n >> 0$, fractal distributions are produced within smaller subsets of the two subintervals.
+The scripts that generate the plucked tent map and its
+stationary probability distributions can be found under \verb+tent_sens_stat+. 
 \end{document}

### 2020-10-27 main.tex
Arxiv version 2

@@
            Center for Computational Science and Engineering, MIT, Cambridge, MA 02139, USA. \email{nishac@mit.edu 
-           }
+		   }  \email{qiqi@mit.edu}
 }
@@
 \cite{wormell}\cite{luca}. Sauer \cite{sauer} has shown 
-examples in which numerical error due to computation in double-precision floating point arithmetic cause significant change in the stationary probability distribution on the attractor.
+examples in which numerical error due to computation in double-precision floating point arithmetic causes significant change in the stationary probability distribution on the attractor.
 The plucked tent map adds to this list of examples by exhibiting a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).

### 2020-10-27 main.tex
move 0.5 to extras and plot analytical density at 0.1

@@
     \centering
-    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.5_0.png}
+    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_0.png}
     \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.5_6.png}
     \caption{The effect of increasing $n$ on the probability distribution associated to the plucked tent map.
-L: The stationary probability distribution of the plucked tent map at $s=0.5, n = 0$.
-R: The stationary probability distribution of the plucked tent map at $s=0.5, n = 6$.}
+L: The stationary probability distribution of the plucked tent map at $s=0.1, n = 0$.
+R: The stationary probability distribution of the plucked tent map at $s=0.1, n = 6$.}
     \label{fig:pluckedTentMapDistribution}

### 2020-10-27 code/tent_sens_stat/plot_density.py
Put map in a different file so density and plot map functions can access it

@@ -1,29 +1,6 @@
 import numba
 from numpy import *
-@numba.jit(nopython=True)
-def tent_basic(x,s):
-    if x < 1:
-        return min(2*x/(1-s), \
-                2 - 2*(1-x)/(1+s))
-    return min(2*(2-x)/(1-s), \
-            2 - 2*(x-1)/(1+s))
-
-@numba.jit(nopython=True)
-def oscillation(x,s):
-    if x < 0.5:
-        return tent_basic(2*x,s)/2
-    return 2-tent_basic(2-2*x,s)/2 
-
-@numba.jit(nopython=True)
-def frequency(x,s,n):
-    return oscillation(2**n*x - floor(2**n*x),s)/2**n + \
-            2*floor(2**n*x)/2**n
-
-@numba.jit(nopython=True)
-def osc_tent(x, s, n):
-    return min(frequency(x,s,n), frequency(2-x,s,n))
-
-
+from plucked_map import *
 @numba.jit(nopython=True)
 def accumulate(nbins, n, m, s):
     density = zeros(nbins)
@@ -37,25 +14,25 @@ def accumulate(nbins, n, m, s):
 
 @numba.jit(nopython=True)
 def analytical_rho(x,s):
-    if x < s:
-        return (1-s)/4
+    if x < 0.5:
+        return (1-s)**2/2
     elif x < 1:
-        return (1 + s)/4
-    elif x < 1+s:
-        return 1 + s/4
-    return (1+s)/4
+        return (1-s*s)/2
+    elif x < 1.5:
+        return (1+s)**2/2
+    return (1-s*s)/2
 
 nbins = 2**12
 nrep = 5
 density = zeros(nbins)
-s = 0.5
+s = 0.1
 n = 0
-#for m in range(nrep):
-#    print("repeat {}".format(m))
-#    density += accumulate(nbins, 1000000000, n, s)/nrep
+for m in range(nrep):
+    print("repeat {}".format(m))
+    density += accumulate(nbins, 1000000000, n, s)/nrep
 x = array(linspace(0, 2, nbins+2)[1:-1])
-for i,xi in enumerate(x):
-    density[i] = analytical_rho(xi,s)
+#for i,xi in enumerate(x):
+#    density[i] = analytical_rho(xi,s)
 fig = figure(figsize=(8,5))
 ax = fig.add_subplot(111)
 ax.fill_between(x, 0, density)

### 2020-10-27 code/tent_sens_stat/plot_map.py

@@ -1,40 +1,6 @@
 import numba
 from numpy import *
-@numba.jit(nopython=True)
-def tent_basic(x,s):
-    if x < 1:
-        return min(2*x/(1-s), \
-                2 - 2*(1-x)/(1+s))
-    return min(2*(2-x)/(1-s), \
-            2 - 2*(x-1)/(1+s))
-
-@numba.jit(nopython=True)
-def oscillation(x,s):
-    if x < 0.5:
-        return tent_basic(2*x,s)/2
-    return 2-tent_basic(2-2*x,s)/2 
-
-
-@numba.jit(nopython=True)
-def frequency(x,s,n):
-    return oscillation(2**n*x - floor(2**n*x),s)/2**n + \
-            2*floor(2**n*x)/2**n
-
-@numba.jit(nopython=True)
-def osc_tent(x, s, n):
-    return min(frequency(x,s,n), frequency(2-x,s,n))
-
-@numba.jit(nopython=True)
-def accumulate(nbins, n, m, s):
-    density = zeros(nbins)
-    dx = 2.0/nbins
-    x = 2*rand()
-    for i in range(n):
-        x = osc_tent(x, s, m)
-        bno = int(x//dx)
-        density[bno] += 1/n/dx
-    return density
-
+from plucked_map import *
 nbins = 2**13
 ns = 5
 nn = 5

### 2020-10-27 code/tent_sens_stat/plucked_map.py

@@ -1,0 +1,26 @@
+from numpy import *
+import numba
+@numba.jit(nopython=True)
+def tent_basic(x,s):
+    if x < 1:
+        return min(2*x/(1-s), \
+                2 - 2*(1-x)/(1+s))
+    return min(2*(2-x)/(1-s), \
+            2 - 2*(x-1)/(1+s))
+
+@numba.jit(nopython=True)
+def oscillation(x,s):
+    if x < 0.5:
+        return tent_basic(2*x,s)/2
+    return 2-tent_basic(2-2*x,s)/2
+
+
+@numba.jit(nopython=True)
+def frequency(x,s,n):
+    return oscillation(2**n*x - floor(2**n*x),s)/2**n + \
+            2*floor(2**n*x)/2**n
+
+@numba.jit(nopython=True)
+def osc_tent(x, s, n):
+    return min(frequency(x,s,n), frequency(2-x,s,n))
+

### 2020-10-27 main.tex
Updates from Overleaf

@@
 \ref{eqn:frequency}-\ref{eqn:plucked}.
-Thus, the asymmetric probability distribution, through this process of scaled repetition, acquires the fractal-like structure seen on increasing $n$, as shown on the right hand side of Figure \ref{fig:pluckedTentMapDistribution}.
-
-Consider $s=0.5$. 
+Thus, the asymmetric probability distribution, through this process of scaled repetition, acquires an apparent fractal-like structure seen on increasing $n$, as shown on the right hand side of Figure \ref{fig:pluckedTentMapDistribution}.
+
 One way to see the asymmetry in the probability distribution 
 about $x=1$, at $n=0$, is to construct a Markov chain with nodes 00, 01, 10, and 11, that indicate the subintervals $(0,0.5)$, $[0.5,1)$, 
-$[1,1.5)$ and $[1.5,2)$, respectively. From Eq. \ref{eqn:plucked}, we can see that the application of $\varphi_{0.5,0}$ yields the following transition matrix:
+$[1,1.5)$ and $[1.5,2)$, respectively. From Eq. \ref{eqn:plucked}, we can see that the application of $\varphi_{s,0}$ yields the following transition matrix:
 \[
 \begin{bmatrix}
-		1/4 & 3/4 & 0 & 0 \\
-		0 & 0 & 3/4 & 1/4 \\
-		0 & 0 & 3/4 & 1/4 \\
-		1/4 & 3/4 & 0 & 0
+		(1-s)/2\quad & (1+s)/2 \quad & 0 \quad & 0 \\
+		0 \quad & 0 \quad & (1+s)/2 \quad & (1-s)/2 \\
+		0 \quad & 0  \quad & (1+s)/2 \quad & (1-s)/2 \\
+		(1-s)/2 \quad & (1+s)/2 \quad & 0 \quad & 0
 \end{bmatrix}
 .\] 
-The stationary probability distribution of this Markov chain is $[1/8, 3/8, 9/8, 3/8]^T$, consistent with the left hand side of Figure \ref{fig:pluckedTentMapDistribution}.
-Clearly, the left subinterval, $[0,1)$ is three times less likely to be visited by a long trajectory of $\varphi_{0.5,0}$, compared to $[1,2)$.
+The stationary probability distribution of this Markov chain, which 
+is the left eigenvector of the transition matrix corresponding to eigenvalue 1, is $[(1-s)^2/2, \; (1-s^2)/2, \; (1+s)^2/2,\; (1-s^2)/2]^T$.
+One can verify this is consistent with the left hand side of Figure \ref{fig:pluckedTentMapDistribution} for $\varphi_{0.1,0}$.
+Clearly, the left subinterval, $[0,1),$ is three times less likely to be visited by a long trajectory of $\varphi_{0.5,0}$, compared to $[1,2)$.
 In fact, the probability of visiting $[0,1)$ remains at 0.25 for all values of $n$, but for $n >> 0$, fractal distributions are produced within smaller subsets of the two subintervals.

### 2020-10-27 main.tex
Edit supp text and conclusion text to emphasize small perturbation, after changing the figure to s=0.1

@@
 In Figure \ref{fig:pluckedTentMap}, we show that the magnitude of the oscillatory perturbation is controlled by a parameter $s$, and its frequency is controlled by the parameter $n$; at $s=0,n=0$, we recover the original tent map.
-The stationary probability distributions -- computed empirically over a trajectory of length 10 billion -- at $s=0.5$, are shown in Figure \ref{fig:pluckedTentMapDistribution}.
+The stationary probability distributions -- computed empirically over a trajectory of length 10 billion -- at $s=0.1$, are shown in Figure \ref{fig:pluckedTentMapDistribution}.
 On the left, $n=0$ and there is already a marked asymmetry developed in the probability distribution compared to the uniform distribution seen at $s=0, n=0$.
-On the right, we show that by increasing $n,$ while keeping the magnitude of the perturbation the same ($s=0.5$), we see a dramatic change in the appearance of the stationary probability distribution: we observe an apparently fractal distribution reminiscent of that of the quasi-physical solutions in
+On the right, we show that by increasing $n,$ despite the fact that the magnitude of the perturbation remains small, we see a dramatic change in the appearance of the stationary probability distribution: we observe an apparent fractal distribution reminiscent of that of the quasi-physical solutions in
 section \ref{sec:tentPerts}.
+By construction of the map (Supplementary Material section \ref{sec:pluckedTentMap}), the nonuniformities of the probability distribution at $n=0$ are transferred to smaller and smaller scales,
+as $n$ is increased.
+Although this effect is more pronounced as $s$, and hence the magnitude of the perturbation, is increased, it must be emphasized that remarkable changes in the physical distribution can be effected, by applying even a small perturbation.
 
@@
     \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_0.png}
-    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.5_6.png}
+    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_6.png}
     \caption{The effect of increasing $n$ on the probability distribution associated to the plucked tent map.
@@
 One can verify this is consistent with the left hand side of Figure \ref{fig:pluckedTentMapDistribution} for $\varphi_{0.1,0}$.
-Clearly, the left subinterval, $[0,1),$ is three times less likely to be visited by a long trajectory of $\varphi_{0.5,0}$, compared to $[1,2)$.
-In fact, the probability of visiting $[0,1)$ remains at 0.25 for all values of $n$, but for $n >> 0$, fractal distributions are produced within smaller subsets of the two subintervals.
+Clearly, the time spent by an infinitely long physical solution in the second quarter interval, $[0.5,1]$ is 1.2 times the time spent in the first quarter, $[0,0.5]$; the time spent in the third quarter $[1,1.5]$ is 1.5 times that spent in the first quarter.
+This nonuniformity is more pronounced with increasing $s$.
+For instance, at $s=0.5$, the left subinterval, [0,1),is three times less likely to be visited by a long trajectory compared to $[1,2)$.
+
+But, as mentioned in the main text, even a small perturbation -- very small $s$ -- is sufficient to trigger a remarkable variation in the statistics.
+This is because this particular construction transfers
+the nonuniformities in the probability distribution at larger scales to smaller scales, with increasing $n.$ The nonuniformity at the largest 
+scale of half intervals is retained at higher values of $n.$ That is,
+the probability of visiting the interval $[0,1]$ is the same: $(1-s)$, at all $n$.
+However, with increasing $n,$ the nonuniformity emerges within smaller subintervals, due to the construction of the map that relies on repeating the behavior at larger scales at smaller scales (Eq.
+\ref{eqn:frequency}).
+This repetitive construction is responsible for the apparent fractal structure of the probability distribution (Figure \ref{fig:pluckedTentMapDistribution}, right).
 The scripts that generate the plucked tent map and its

### 2020-10-27 main.tex
emphasize smallness of perts

@@
 tent map -- in which an oscillatory perturbation is introduced (see Supplementary Material section \ref{sec:pluckedTentMap} for the map equation).
-In Figure \ref{fig:pluckedTentMap}, we show that the magnitude of the oscillatory perturbation is controlled by a parameter $s$, and its frequency is controlled by the parameter $n$; at $s=0,n=0$, we recover the original tent map.
+In Figure \ref{fig:pluckedTentMap}, we show that the magnitude of the oscillatory perturbation is controlled by parameters $s$ and $n$, and its frequency is controlled by the parameter $n$; at $s=0,n=0$, we recover the original tent map.
 The stationary probability distributions -- computed empirically over a trajectory of length 10 billion -- at $s=0.1$, are shown in Figure \ref{fig:pluckedTentMapDistribution}.
 On the left, $n=0$ and there is already a marked asymmetry developed in the probability distribution compared to the uniform distribution seen at $s=0, n=0$.
-On the right, we show that by increasing $n,$ despite the fact that the magnitude of the perturbation remains small, we see a dramatic change in the appearance of the stationary probability distribution: we observe an apparent fractal distribution reminiscent of that of the quasi-physical solutions in
+On the right, we show that by increasing $n,$ despite the fact that the magnitude of the perturbation becomes smaller with $n$, we see a dramatic change in the appearance of the stationary probability distribution: we observe an apparent fractal distribution reminiscent of that of the quasi-physical solutions in
 section \ref{sec:tentPerts}.
@@
 as $n$ is increased.
-Although this effect is more pronounced as $s$, and hence the magnitude of the perturbation, is increased, it must be emphasized that remarkable changes in the physical distribution can be effected, by applying even a small perturbation.
+It is worth emphasizing that remarkable changes in the physical distribution are effected (as shown in the right hand side of Figure \ref{fig:pluckedTentMapDistribution}) with a tiny perturbation -- the perturbed map at $n=10$, and the original tent map appear indiscernible on the left hand side of Figure \ref{fig:pluckedTentMap}.
 

### 2020-10-27 main.tex
Add new figures of pert tent map

@@
     \centering
-    \includegraphics[width=0.45\textwidth]{osc_tent_n.png}
-    \includegraphics[width=0.45\textwidth]{osc_tent_s.png}
-    \caption{The plucked tent map shown at different values of $s$ and $n$, that represent the magnitude and frequency of the applied oscillation to the original tent map ($n=0,s=0$).
-On the left, $s$ is held constant at 0.5.
-On the right, $n$ is held constant at 0.}
+    \includegraphics[width=0.45\textwidth]{osc_tent_0.png}
+    \includegraphics[width=0.45\textwidth]{osc_tent_3.png}
+    \includegraphics[width=0.45\textwidth]{osc_tent_6.png}
+     \includegraphics[width=0.45\textwidth]{osc_tent_10.png}
+    \caption{The plucked tent map shown at different values of $s$ and $n = 0$ (top-left), $n=3$ (top-right), $n=6$ (bottom-left), and $n=10$ (bottom-right).
+The original tent map is at $s=0$ on each plot.}
     \label{fig:pluckedTentMap}
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_0.png}
-    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_6.png}
+%    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.png}
+ %   \includegraphics[width=0.45\textwidth]{osc_tent_dens_3.png}
+  %  \includegraphics[width=0.45\textwidth]{osc_tent_dens_6.png}
+   % \includegraphics[width=0.45\textwidth]{osc_tent_dens_10.png}
     \caption{The effect of increasing $n$ on the probability distribution associated to the plucked tent map.

### 2020-10-27 main.tex
Change text to indicate change in figure

@@
 tent map -- in which an oscillatory perturbation is introduced (see Supplementary Material section \ref{sec:pluckedTentMap} for the map equation).
-In Figure \ref{fig:pluckedTentMap}, we show that the magnitude of the oscillatory perturbation is controlled by parameters $s$ and $n$, and its frequency is controlled by the parameter $n$; at $s=0,n=0$, we recover the original tent map.
+In Figure \ref{fig:pluckedTentMap}, we show that the magnitude of the oscillatory perturbation is controlled by parameters $s$ and $n$, and its frequency is controlled by the parameter $n$; at $s=0$, we recover the original tent map at all $n$.
 The stationary probability distributions -- computed empirically over a trajectory of length 10 billion -- at $s=0.1$, are shown in Figure \ref{fig:pluckedTentMapDistribution}.
-On the left, $n=0$ and there is already a marked asymmetry developed in the probability distribution compared to the uniform distribution seen at $s=0, n=0$.
-On the right, we show that by increasing $n,$ despite the fact that the magnitude of the perturbation becomes smaller with $n$, we see a dramatic change in the appearance of the stationary probability distribution: we observe an apparent fractal distribution reminiscent of that of the quasi-physical solutions in
+On the top-left, $n=0$, and there is already a marked asymmetry developed in the probability distribution compared to the uniform distribution seen at $s=0, n=0$.
+The figure shows that by increasing $n,$ despite the fact that the magnitude of the perturbation becomes smaller with $n$, we see a dramatic change in the appearance of the stationary probability distribution: we observe an apparent fractal distribution reminiscent of that of the quasi-physical solutions in
 section \ref{sec:tentPerts}.
@@
 as $n$ is increased.
-It is worth emphasizing that remarkable changes in the physical distribution are effected (as shown in the right hand side of Figure \ref{fig:pluckedTentMapDistribution}) with a tiny perturbation -- the perturbed map at $n=10$, and the original tent map appear indiscernible on the left hand side of Figure \ref{fig:pluckedTentMap}.
+It is worth emphasizing that these remarkable changes in the physical distribution are effected (as shown in the right hand side of Figure \ref{fig:pluckedTentMapDistribution}) with a tiny perturbation -- the perturbed map at $n=10$, and the original tent map appear indiscernible on the bottom-right of Figure \ref{fig:pluckedTentMap}.
 
@@
     \centering
-%    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.png}
- %   \includegraphics[width=0.45\textwidth]{osc_tent_dens_3.png}
-  %  \includegraphics[width=0.45\textwidth]{osc_tent_dens_6.png}
-   % \includegraphics[width=0.45\textwidth]{osc_tent_dens_10.png}
+   \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_0.png}
+    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_3.png}
+    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_6.png}
+    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_10.png}
     \caption{The effect of increasing $n$ on the probability distribution associated to the plucked tent map.
-L: The stationary probability distribution of the plucked tent map at $s=0.1, n = 0$.
-R: The stationary probability distribution of the plucked tent map at $s=0.1, n = 6$.}
+The stationary probability distribution of the plucked tent map at $s=0.1$ and $n = 0$ (top-left), $n=3$ (top-right), $n = 6$ (bottom-left) and $n=10$ (bottom-right).}
     \label{fig:pluckedTentMapDistribution}
@@
 which is also obtained by setting $n=0$ in $\varphi_{s,n}$ --  
-in which an asymmetry is produced in the probability distribution (left hand side of Figure \ref{fig:pluckedTentMapDistribution}).
+in which an asymmetry is produced in the probability distribution (top-left of Figure \ref{fig:pluckedTentMapDistribution}).
 Upon repeating, with appropriate scaling, the oscillation $o_s$, 
@@
 \ref{eqn:frequency}-\ref{eqn:plucked}.
-Thus, the asymmetric probability distribution, through this process of scaled repetition, acquires an apparent fractal-like structure seen on increasing $n$, as shown on the right hand side of Figure \ref{fig:pluckedTentMapDistribution}.
+Thus, the asymmetric probability distribution, through this process of scaled repetition, acquires an apparent fractal-like structure seen on increasing $n$, as shown on the bottom row of Figure \ref{fig:pluckedTentMapDistribution}.
 
@@
 is the left eigenvector of the transition matrix corresponding to eigenvalue 1, is $[(1-s)^2/2, \; (1-s^2)/2, \; (1+s)^2/2,\; (1-s^2)/2]^T$.
-One can verify this is consistent with the left hand side of Figure \ref{fig:pluckedTentMapDistribution} for $\varphi_{0.1,0}$.
+One can verify this is consistent with the top-left of Figure \ref{fig:pluckedTentMapDistribution} for $\varphi_{0.1,0}$.
 Clearly, the time spent by an infinitely long physical solution in the second quarter interval, $[0.5,1]$ is 1.2 times the time spent in the first quarter, $[0,0.5]$; the time spent in the third quarter $[1,1.5]$ is 1.5 times that spent in the first quarter.
@@
 \ref{eqn:frequency}).
-This repetitive construction is responsible for the apparent fractal structure of the probability distribution (Figure \ref{fig:pluckedTentMapDistribution}, right).
+This repetitive construction is responsible for the apparent fractal structure of the probability distribution (Figure \ref{fig:pluckedTentMapDistribution}, bottom-right).
 The scripts that generate the plucked tent map and its

### 2020-10-27 letter/letter.tex
Add letter to the editor

\documentclass{letter}
\title{Letter to the Editor}
\begin{document}
{\hspace{1.6in} 
\begin{large}{Letter to the Editor}\end{large}
\\
\\
Dear Editor,

We are happy to submit our manuscript titled 
``On the probability of finding a nonphysical solution through shadowing ''
for consideration for publication in Applied Mathematics and Mechanics. Thank you very much for giving us the 
opportunity to have our work reviewed!

The purpose of this work is to raise an important question on the reliability of numerical simulations of chaos that resonates across all scientific disciplines. We ask whether perturbed solutions of a 
governing equation reproduce its physical behavior. This is a question distinct from whether perturbed solutions of a chaotic system lie close to a true solution. This latter question has historically been answered in the affirmative, at least in hyperbolic dynamical systems, by shadowing. In this paper, we show that while
perturbed solutions such as a numerical simulation, 
may have their corresponding true solutions that shadow them, these shadowing solutions can be \emph{nonphysical}. We define the concept of nonphysical solution as a true solution to a governing equation along which long-time averages do not converge to expectations with respect to the stationary probability distribution on the attractor. Equivalently, the infinite-time averages along a nonphysical solution do not converge to the same value as that along almost every solution of the system.

We argue, through analytical constructions, that shadowing solutions are typically nonphysical solutions. We show that the probability of finding a nonphysical shadowing solution can be 1. This has the implication that perturbed solutions such as numerical solutions may be close to nonphysical solutions, and hence produce incorrect long-term or statistical behavior. In addition, sensitivity analysis methods based on the concept of shadowing also give incorrect results for the response of statistics to small perturbations. We construct an example to show that 
there are chaotic systems in which small parameter perturbations can in fact have a dramatic change in the statistics. This extreme non-smooth response further strengthens our conclusion that the assumption of physicality of shadowing solutions is easily violated.

This paper raises a fundamental and profound issue with numerical simulation and sensitivity analysis of chaotic governing equations. It therefore has immediate implications for the entire computational science and engineering community, which must work together to rethink chaotic simulations. 
Thank you very much for your time and consideration,\\
Nisha Chandramoorthy and Qiqi Wang.
\end{document}

### 2020-10-27 main.tex
Updates from Overleaf

@@
 This implies that the solution may not eventually converge to 
-any fixed point, nor show any repetitive patterns.  
- As the map iterates, these patterns
+any fixed point, nor follow any strict period.  
+ As the map iterates, the nonrepetitive bit patterns
 shift towards more significant digits, and the solution visits an infinite set of
@@
 composition of $\tilde{\varphi}_s.$ That is, if $\tilde{x}_{j+1} = \tilde{\varphi}_s(\tilde{x}_j), j = 0,1,\cdots,$ is a solution, $\tilde{\varphi}_s^i(\tilde{x}_j) = \tilde{x}_{j+i},\quad i,j = 0,1,\cdots$.
+
 To see why $\tilde{h}_s$ constructed in Eq. \ref{conjugate_tilted} satisfies the definition of a conjugacy map --
@@
 
-Figure \ref{fig:tilted_tent_density} shows how fractal the conjugacy map is.
+Figure \ref{fig:tilted_tent_conjugate} shows how fractal the conjugacy map is.
 This fractal conjugacy map transforms a uniform distribution in $[0,2]$ into
-a fractal distribution similar to the ones plotted in section \ref{sec:quasiphysical}.
+a fractal distribution, shown in Figure \ref{fig:tilted_tent_density},
+similar to the ones plotted in section \ref{sec:quasiphysical}.
 In fact, we will show that the fractal distribution obtained by mapping a uniform
@@
 be the map of a uniform density on $[0,2]$, through $\tilde{\varphi}_s$. Then, 
-$\tilde{\rho}(\tilde{\varphi}_s (x)) = 0.5/|\tilde \varphi_s'(x)| + 0.5/|\tilde \varphi_s'(2-x)|,$ where the two terms express the conservation of probability 
+$\tilde{\rho}(\tilde{\varphi}_s (x)) = 0.5/|\tilde \varphi_s'(x)| + 0.5/|\tilde \varphi_s'(x')|,$ where
+$x$ and $x'$ lie on two sides of $1+s$ and
+$\tilde{\varphi}_s(x)=\tilde{\varphi}_s(x')$; the two terms express the conservation of probability 
 from the two pre-images of $\tilde{\varphi}_s$. Substituting the piecewise constant 

### 2020-10-28 letter/letter.tex
change the last figure and text.

@@
 \begin{document}
-{\hspace{1.6in} 
+\hspace{1.6in} 
 \begin{large}{Letter to the Editor}\end{large}
@@
 We define the concept of nonphysical solution as a true solution to a governing equation along which long-time averages do not converge to expectations with respect to the stationary probability distribution on the attractor.
-Equivalently, the infinite-time averages along a nonphysical solution do not converge to the same value as that along almost every solution of the system.
+Equivalently, the infinite-time averages along a nonphysical solution do not converge to the same value as that along almost every solution.
 
@@
 We construct an example to show that
-there are chaotic systems in which small parameter perturbations can in fact have a dramatic change in the statistics.
-This extreme non-smooth response further strengthens our conclusion that the assumption of physicality of shadowing solutions is easily violated.
+there are chaotic systems in which small parameter perturbations can, in fact, cause a dramatic change in the statistics.
+This extreme non-smooth response of a chaotic attractor further strengthens our conclusion that the assumption of physicality of shadowing solutions is easily violated.
 
@@
 It therefore has immediate implications for the entire computational science and engineering community, which must work together to rethink chaotic simulations.
-Thank you very much for your time and consideration,\\
+Thank you very much for your time and consideration,
+\\
+\\
 Nisha Chandramoorthy and Qiqi Wang.

### 2020-10-28 main.tex
change the last figure and text.

@@
 In Figure \ref{fig:pluckedTentMap}, we show that the magnitude of the oscillatory perturbation is controlled by parameters $s$ and $n$, and its frequency is controlled by the parameter $n$; at $s=0$, we recover the original tent map at all $n$.
-The stationary probability distributions -- computed empirically over a trajectory of length 10 billion -- at $s=0.1$, are shown in Figure \ref{fig:pluckedTentMapDistribution}.
-On the top-left, $n=0$, and there is already a marked asymmetry developed in the probability distribution compared to the uniform distribution seen at $s=0, n=0$.
+The stationary probability distributions -- computed empirically over a trajectory of length 10 billion -- at $s=0.1$, are shown on the right column of Figure \ref{fig:pluckedTentMap}.
+On the top row, $n=0$, and there is already a marked asymmetry developed in the probability distribution compared to the uniform distribution seen at $s=0, n=0$.
 The figure shows that by increasing $n,$ despite the fact that the magnitude of the perturbation becomes smaller with $n$, we see a dramatic change in the appearance of the stationary probability distribution: we observe an apparent fractal distribution reminiscent of that of the quasi-physical solutions in
@@
 as $n$ is increased.
-It is worth emphasizing that these remarkable changes in the physical distribution are effected (as shown in the right hand side of Figure \ref{fig:pluckedTentMapDistribution}) with a tiny perturbation -- the perturbed map at $n=10$, and the original tent map appear indiscernible on the bottom-right of Figure \ref{fig:pluckedTentMap}.
+It is worth emphasizing that these remarkable changes in the physical distribution (shown on the bottom-right of Figure \ref{fig:pluckedTentMap}) are effected with a tiny perturbation -- the perturbed map at $n=6$, and the original tent map appear indiscernible on the bottom-left of Figure \ref{fig:pluckedTentMap}.
 
@@
     \includegraphics[width=0.45\textwidth]{osc_tent_0.png}
+     \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_0.png}
     \includegraphics[width=0.45\textwidth]{osc_tent_3.png}
+     \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_3.png}
     \includegraphics[width=0.45\textwidth]{osc_tent_6.png}
-     \includegraphics[width=0.45\textwidth]{osc_tent_10.png}
-    \caption{The plucked tent map shown at different values of $s$ and $n = 0$ (top-left), $n=3$ (top-right), $n=6$ (bottom-left), and $n=10$ (bottom-right).
-The original tent map is at $s=0$ on each plot.}
+     \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_6.png}
+    \caption{
+    The effect of increasing $n$ on the probability distribution associated to the plucked tent map. The left column shows the plucked tent map at 
+    different values of $s$ and $n = 0$ (top), $n=3$ (middle), $n=6$ (bottom). The original tent map is at $s=0$ on each plot.
+    The right column shows the stationary, physical probability distribution of the plucked tent map at $s=0.1$ and $n = 0$ (top), $n=3$ (middle), $n = 6$ (bottom).}
     \label{fig:pluckedTentMap}
 \end{figure}
-\begin{figure}
-    \centering
-   \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_0.png}
-    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_3.png}
-    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_6.png}
-    \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_10.png}
-    \caption{The effect of increasing $n$ on the probability distribution associated to the plucked tent map.
-The stationary probability distribution of the plucked tent map at $s=0.1$ and $n = 0$ (top-left), $n=3$ (top-right), $n = 6$ (bottom-left) and $n=10$ (bottom-right).}
-    \label{fig:pluckedTentMapDistribution}
-\end{figure}
+
 Violations of smooth response, like in the plucked tent map, have an immediate implication for shadowing sensitivities as well.
@@
 which is also obtained by setting $n=0$ in $\varphi_{s,n}$ --  
-in which an asymmetry is produced in the probability distribution (top-left of Figure \ref{fig:pluckedTentMapDistribution}).
+in which an asymmetry is produced in the probability distribution (top-right of Figure \ref{fig:pluckedTentMap}).
 Upon repeating, with appropriate scaling, the oscillation $o_s$, 
@@
 \ref{eqn:frequency}-\ref{eqn:plucked}.
-Thus, the asymmetric probability distribution, through this process of scaled repetition, acquires an apparent fractal-like structure seen on increasing $n$, as shown on the bottom row of Figure \ref{fig:pluckedTentMapDistribution}.
+Thus, the asymmetric probability distribution, through this process of scaled repetition, acquires an apparent fractal-like structure seen on increasing $n$, as shown on the bottom-right of Figure \ref{fig:pluckedTentMap}.
 
@@
 is the left eigenvector of the transition matrix corresponding to eigenvalue 1, is $[(1-s)^2/2, \; (1-s^2)/2, \; (1+s)^2/2,\; (1-s^2)/2]^T$.
-One can verify this is consistent with the top-left of Figure \ref{fig:pluckedTentMapDistribution} for $\varphi_{0.1,0}$.
+One can verify this is consistent with the top-right of Figure \ref{fig:pluckedTentMap} for $\varphi_{0.1,0}$.
 Clearly, the time spent by an infinitely long physical solution in the second quarter interval, $[0.5,1]$ is 1.2 times the time spent in the first quarter, $[0,0.5]$; the time spent in the third quarter $[1,1.5]$ is 1.5 times that spent in the first quarter.
@@
 \ref{eqn:frequency}).
-This repetitive construction is responsible for the apparent fractal structure of the probability distribution (Figure \ref{fig:pluckedTentMapDistribution}, bottom-right).
+This repetitive construction is responsible for the apparent fractal structure of the probability distribution (Figure \ref{fig:pluckedTentMap}, bottom-right).
 The scripts that generate the plucked tent map and its

### 2020-10-29 main.tex
Change final figure

@@
 In Figure \ref{fig:pluckedTentMap}, we show that the magnitude of the oscillatory perturbation is controlled by parameters $s$ and $n$, and its frequency is controlled by the parameter $n$; at $s=0$, we recover the original tent map at all $n$.
-The stationary probability distributions -- computed empirically over a trajectory of length 10 billion -- at $s=0.1$, are shown on the right column of Figure \ref{fig:pluckedTentMap}.
-On the top row, $n=0$, and there is already a marked asymmetry developed in the probability distribution compared to the uniform distribution seen at $s=0, n=0$.
+The physical probability distributions -- computed empirically over a trajectory of length 10 billion -- at $s=0.1$, are shown on the right column of Figure \ref{fig:pluckedTentMap}.
+On the top row, $n=0$, and there is already a marked asymmetry developed in the physical distribution compared to the uniform distribution, which is the physical distribution seen at $s=0, n=0$.
 The figure shows that by increasing $n,$ despite the fact that the magnitude of the perturbation becomes smaller with $n$, we see a dramatic change in the appearance of the stationary probability distribution: we observe an apparent fractal distribution reminiscent of that of the quasi-physical solutions in
@@
 as $n$ is increased.
-It is worth emphasizing that these remarkable changes in the physical distribution (shown on the bottom-right of Figure \ref{fig:pluckedTentMap}) are effected with a tiny perturbation -- the perturbed map at $n=6$, and the original tent map appear indiscernible on the bottom-left of Figure \ref{fig:pluckedTentMap}.
+It is worth emphasizing that these remarkable changes in the physical distribution (shown on the bottom-right of Figure \ref{fig:pluckedTentMap}) are effected with a tiny perturbation -- the perturbed map at $n=6$ and the original tent map appear indiscernible on the bottom-left of Figure \ref{fig:pluckedTentMap}.
 
 In view of the plucked tent map, we can question further the
-validity of numerical solutions. Can numerical play 
+validity of numerical solutions. Can numerical solutions play 
 the role of the plucking perturbation to the 
 governing equation? That is, can numerical solutions represent 
-slight perturbations to the governing equations that nevertheless 
-exhibit drastically different statistics from that of the ensemble? Uniformly hyperbolic 
+slight perturbations to the governing equations for which the physical 
+distribution is drastically different from the physical distribution 
+of the original governing equation? Sauer \cite{sauer} has shown 
+examples in which numerical error due to computation in double-precision floating point arithmetic causes significant change in the stationary probability distribution on the attractor.
+The plucked tent map adds to this list of examples by demonstrating a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).
+Uniformly hyperbolic
 dynamical systems exhibit \emph{linear response} 
@@
 But, uniform hyperbolicity is a mathematical idealization, and although some physical systems have been observed to behave as if they were uniformly hyperbolic \cite{gallavotti}, a violation of uniform hyperbolicity is more likely
-\cite{wormell}\cite{luca}. Sauer \cite{sauer} has shown 
-examples in which numerical error due to computation in double-precision floating point arithmetic causes significant change in the stationary probability distribution on the attractor.
-The plucked tent map adds to this list of examples by exhibiting a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).
+\cite{wormell}\cite{luca}.
 
@@
 The supplementary material for this paper including the code, data and figures can be found on Github \cite{supp}.
-The code can be found under the \verb+code+ subdirectory inside of which section-wise code is separated into further subdirectories.
+The code can be found under the \verb+code+ subdirectory inside which section-wise code is separated into further subdirectories.
 The data used to plot the figures can be found under \verb+data+.
@@
 The motivation for approximating the Lorenz map is that a 
-closed form expression for the map is necessary for our numerical shadowing procedure.
+closed form expression for the map is necessary for our numerical shadowing procedure (section \ref{sec:computingShadowing}).
 In a small region around the cusp, we approximate the map using an exponential function.
@@
 The coefficients of the numerator and denominator of the rational function modelling the left (right) tail are denoted $a_{L,0}, a_{L,1}$ ($a_{R,0}, a_{R,1}$) and $b_{L,n}, n=0,1,2,3$ ($b_{R,n}, n=0,1,2,3$), respectively.
-The values of these coefficients are obtained by
-regression on the map at standard parameters are shown in the table below. As shown in Figure \ref{fig:lorenzMapComp}, the fit obtained matches the Lorenz map closely.
+The values of these coefficients, which are obtained by
+regression, are shown for the map at standard parameters, in the table below. As shown in Figure \ref{fig:lorenzMapComp}, the fit obtained matches the Lorenz map closely.
 \begin{table}[H]
@@
     \end{tabular}
-    \caption{The fitting parameters of the Lorenz map at $\sigma = 10, \beta = 8/3$ and $\rho = 28.$}
+    \caption{The fitting parameters of the Lorenz map at $\sigma = 10, \beta = 8/3$, and $\rho = 28.$}
     \label{tab:lorenzFit}
@@
 		the least squares shadowing method \cite{qiqi-lss}.
-However, since these maps are all one-dimensional chaotic systems that have in common non-invertibility with two inverse \emph{branches}, we devise a simpler approach.
+However, since the  maps we consider are all one-dimensional chaotic systems that have in common non-invertibility with two inverse \emph{branches}, we devise a simpler approach.
 The map $\varphi_s$ in this section refers to any of the perturbations of the tent map (section \ref{sec:tentPerts}) or the Lorenz map.
@@
 Suppose we set $y_0 = x_0$, the difference $y_1 - x_1$ will amplify on further iterations under $\varphi_s$ for any perturbed solution.
-On the other hand, backward iteration of the map is contracting, and thus we set $y_N = x_N$ and proceed backward in time to
+On the other hand, backward iteration of the map is contracting, and thus we set $y_N = x_N$.
+Then, we proceed backward in time to
 construct a solution of $\varphi_s$, noting that any difference that emerges at a given time $n$ will be made smaller starting at $n-1$.
@@
 Marching backward by choosing at each step, the pre-image $y_n$ in the same subinterval as $x_n$, $y_n$ approximates the shadowing solution better as $n$ decreases.
-Thus, the procedure to find a shadowing solution simply reduces to solving for a backward trajectory (specifically one among the possible $2^N$), starting at a given final condition $x_N$.
+Thus, the procedure to find a shadowing solution simply reduces to solving for a backward trajectory (specifically one among the possible $2^N$), starting at a given final condition,  $x_N$.
 
@@
 				\dfrac{2x}{1-s}, 2 - \dfrac{2(1-x)}{1+s}\Big),
-				\;\;\; x < 1 \\
+				\;\;\; x < 1. 
 \end{align}
@@
 We can modulate
-the frequency or repetitions -- proportional to $n$ -- 
+the frequency of repetitions -- proportional to $n$ -- 
 of the oscillations through
@@
 non-uniform (e.g.
-like the other tent map perturbations of section \ref{sec:tentPerts}) but in which the nonuniformity can be controlled, and made skewed as desired.
+like the other tent map perturbations of section \ref{sec:tentPerts}), but in which the nonuniformity can be controlled, and made skewed as desired.
 Thus, we choose to construct a base perturbation of the tent map, ${\rm min}(o_s(x), o_s(2-x))$ --
@@
 One can verify this is consistent with the top-right of Figure \ref{fig:pluckedTentMap} for $\varphi_{0.1,0}$.
-Clearly, the time spent by an infinitely long physical solution in the second quarter interval, $[0.5,1]$ is 1.2 times the time spent in the first quarter, $[0,0.5]$; the time spent in the third quarter $[1,1.5]$ is 1.5 times that spent in the first quarter.
+Clearly, the time spent by an infinitely long physical solution in the second quarter interval, $[0.5,1]$ is about 1.2 times the time spent in the first quarter, $[0,0.5]$; the time spent in the third quarter $[1,1.5]$ is about 1.5 times that spent in the first quarter.
 This nonuniformity is more pronounced with increasing $s$.
@@
 The scripts that generate the plucked tent map and its
-stationary probability distributions can be found under \verb+tent_sens_stat+. 
+physical probability distributions can be found under \verb+tent_sens_stat+. 
 \end{document}

### 2020-10-29 main.tex
Add AMM reference

@@
 of the original governing equation? Sauer \cite{sauer} has shown 
-examples in which numerical error due to computation in double-precision floating point arithmetic causes significant change in the stationary probability distribution on the attractor.
+examples in which numerical error due to computation in double-precision floating point arithmetic causes significant change in the stationary probability distribution on the attractor; other works \cite{cns} further analyze the unreliability due to double-precision arithmetic of statistical measures such as power spectra and correlation functions.
 The plucked tent map adds to this list of examples by demonstrating a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{osc_tent_0.png}
-     \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_0.png}
-    \includegraphics[width=0.45\textwidth]{osc_tent_3.png}
-     \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_3.png}
-    \includegraphics[width=0.45\textwidth]{osc_tent_6.png}
-     \includegraphics[width=0.45\textwidth]{osc_tent_dens_0.1_6.png}
+    \includegraphics[width=0.3\textwidth,height=1.35in]{osc_tent_0.png}
+     \includegraphics[width=0.6\textwidth,height=1.35in]{osc_tent_dens_0.1_0.png}
+    \includegraphics[width=0.3\textwidth,height=1.35in]{osc_tent_3.png}
+     \includegraphics[width=0.6\textwidth,height=1.35in]{osc_tent_dens_0.1_3.png}
+    \includegraphics[width=0.3\textwidth,height=1.35in]{osc_tent_6.png}
+     \includegraphics[width=0.6\textwidth,height=1.35in]{osc_tent_dens_0.1_6.png}
     \caption{
-    The effect of increasing $n$ on the probability distribution associated to the plucked tent map. The left column shows the plucked tent map at 
+    The effect of increasing $n$ on the physical probability distribution associated to the plucked tent map. The left column shows the plucked tent map at 
     different values of $s$ and $n = 0$ (top), $n=3$ (middle), $n=6$ (bottom). The original tent map is at $s=0$ on each plot.
@@
 shadowed by a true solution in that the difference between them is small for a long time, this shadowing solution may not represent the long-term or ensemble behavior of the physical system.
-The nonphysicality of shadowing solutions also indicates that shadowing-based methods \cite{qiqi-lss} can lead to incorrect values of sensitivities of statistics to parameter changes.
+The nonphysicality of shadowing solutions also indicates that shadowing-based methods can lead to incorrect values of sensitivities of statistics to parameter changes.
 

### 2020-10-30 main.tex
Correct the paramters of the lorenz plot

@@
 In order to analyze the behavior of the shadowing solutions of the Lorenz'63 system 
-of equations, we first make a closed form approximation of the Lorenz map (see section \ref{sec:lorenzMapApprox} of the supplementary material). The reason 
-is that except at rare parameter values, the Lorenz system does not have 
+of equations, we first make a closed form approximation of the Lorenz map (see section \ref{sec:lorenzMapApprox} of the supplementary material).
+Except at rare parameter values, the Lorenz system does not have
 solutions that shadow perturbed orbits for all time, but rather only for a finite time 
@@
 A closed form approximation is constructed by regression performed using long solutions 
-of the Lorenz map (see Figure \ref{fig:lorenz_map}). 
+of the Lorenz map (Figure \ref{fig:lorenz_map}). 
 \begin{figure}
@@
 The Lorenz map approximation obtained this way is illustrated in Figure \ref{fig:lorenz_params}, for variations of the parameters around their standard values of 
-$\rho = 28, \sigma = 10, \beta = 8/3.$ For a range of parameter values, it is numerically verified that the approximate Lorenz map qualitatively reproduces the behavior of the true Lorenz map.
-Using the approximate Lorenz map, we calculate next the shadowing solutions.
+$\rho = 28, \sigma = 10,$ and $\beta = 8/3.$ Using the approximate Lorenz map, we calculate next the shadowing solutions. 
 \begin{figure}
@@
     \includegraphics[width=0.48\textwidth]{baseline_density_lorenz_zmax_10_28_3.33.json.png}
-    \caption{The left column shows the empirical distribution of the shadowing solution of the Lorenz map at $\rho = 28$ (top row), $\rho = 29$ (middle row), and $\beta = 10/3$ (bottom row).
-The right column shows the probability distribution of the corresponding physical solution.
-In each case, the parameters not mentioned are held at their standard values of $\rho = 28, \sigma = 10, \beta = 8/3.$
+    \caption{The left column shows the empirical distribution of the shadowing solution, and the right column shows the physical distribution of the Lorenz map at the following sets of parameters: top row: $\sigma = 15, \rho = 28, \beta = 8/3$, $\rho = 29$, middle row: $\sigma = 10, \rho = 30, \beta = 10/3$, and bottom row: $\sigma = 10, \rho = 28, \beta = 10/3$.
     }
@@
 \end{figure}
-The empirical densities computed from long shadowing solutions, of length 10 billion, are shown in the left column of Figure \ref{fig:lorenz_params_scaled} at three different sets of parameter values, which are also different from the standard values of $\rho = 28, \sigma = 10$, and $\beta = 8/3$.
-In the top and middle row, $\rho = 29$ and $\rho = 30$ are used, respectively, while the other two parameters are retained at their standard values.
+The empirical distribution computed from long shadowing solutions, of length 10 billion, are shown in the left column of Figure \ref{fig:lorenz_params_scaled} at three different sets of parameter values, which are also different from the standard values of $\rho = 28, \sigma = 10$, and $\beta = 8/3$.
+In each row, one of the parameters is  $\rho = 29$ and $\rho = 30$ are used, respectivel, while the other two parameters are retained at their standard values.
 In the bottom row, only the parameter $\beta$ is perturbed from the standard values; $\beta = 10/3$ is used.
-On the right column, we plot the densities of physical solutions, the SRB distribution, at the set of parameters corresponding to each row.
-We again notice that, while the physical distributions are smooth, the shadowing solutions have a rough distribution that appears to have hair-like patterns.
+On the right column, we plot the distribution of physical solutions, the SRB distribution, at the set of parameters corresponding to each row.
+We again notice that, while the physical distributions appear smooth, the shadowing solutions clearly exhibit roughness.
 Thus, we draw the same conclusion in the Lorenz map as in the various perturbations of the tent map treated in section \ref{sec:tentPerts}: shadowing solutions do not have the same distribution as a typically observed solution of the governing equation.
-In other words, shadowing solutions may be nonphysical.
+In other words, shadowing solutions are nonphysical.
 

### 2020-10-30 main.tex
Final edits

@@
 to the Lorenz equation and other, more complex, governing
-equations which typically produce chaotic solutions.  Neverthless, the similarity 
+equations which typically produce chaotic solutions.  Nevertheless, the similarity 
 between the tent map and the Lorenz map \cite{tent2}, shown in Figure \ref{fig:lorenz_map}, suggests that
@@
 shadowing solutions?
-
 %%
@@
     (2-s)x \quad & x < 1 \\
-    (2-s)(2 - x) \quad & x \ge 1
+    (2-s)(2 - x) \quad & 2 >  x \ge 1.
     \end{cases}
@@
 the quasi-solutions of the tent map in section \ref{sec:quasiphysical}.
-The reader is referred to the Supplementary Material section \ref{sec:computingShadowing} for a note on the computation of the shadowing solution.
-Note that, unlike the tilted tent maps, the family of squashed tent maps do not have the uniform distribution as their SRB distribution, as indicated on the right column.
-However, the SRB distributions are not fractal.
-That is, the physical solutions of the squashed tent map are distributed according to a regular probability density (right column of Figure \ref{fig:tent_squashed_shadow}).
-But the shadowing solution, which is a true orbit of the original tent map, is distributed neither like a physical solution of the original tent map nor of the squashed tent map, as suggested by its fractal-like probability distribution.
+The reader is referred to the Supplementary Material section \ref{sec:computingShadowing} for the computational method used in this paper, for the shadowing solutions.
+Note that, unlike the tilted tent maps, the family of squashed tent maps do not have the uniform distribution as their SRB distribution, but have a different regular probability distribution as indicated on the right column of Figure \ref{fig:tent_squashed_shadow}.
+On the other hand, the shadowing solution is distributed neither like a physical solution of the original tent map nor of the squashed tent map, as suggested by its fractal-like probability distribution on the left column of Figure \ref{fig:tent_squashed_shadow}.
 
@@
 \end{align}
-Again, at $s=0$, the original tent map is recovered, and other values of $s,$ the tent map is ``pinched'' by perturbations that are symmetric around $x = 1$, and zero at the end points.
+Again, at $s=0$, the original tent map is recovered, and at other values of $s,$ the tent map is ``pinched'' by perturbations that are symmetric around $x = 1$, and zero at the end points.
 Figure \ref{fig:tent_pinched} shows the pinched tent map at different $s$ values.
@@
     \includegraphics[width=0.48\textwidth]{pinched_tent_map.png}
-    \caption{The pinched tent map (Eq. \ref{eqn:tent_pinched}) for $s = 0.01, 0.05, 0.2,$ and 0.5.}
+    \caption{The pinched tent map (Eq. \ref{eqn:tent_pinched}) for $s = 0.01$ (blue), $s=0.05$ (orange), $s=0.2,$ (green) and $s=0.5$(red).}
     \label{fig:tent_pinched}
@@
 can see that the pinching perturbation changes the uniform density of the original tent map to a linearly varying density.
-The more pronounced the perturbation -- the higher the value of $s$ -- the steeper decline in the density from left to the right.
+The more pronounced the perturbation -- the higher the value of $s$ -- the steeper the slope.
 The shadowing distribution, in this case as well, appears fractal.
-Once again, with probability one -- or, for any randomly chosen initial condition of the solution of the pinched tent map -- any physical solution is shadowed by a quasi-physical solution.
+Once again, with probability one -- or, for any randomly chosen initial condition of the solution of the pinched tent map -- a physical solution is shadowed by a quasi-physical solution.
 \begin{figure}
@@
 Hence, we can once again conclude that almost every solution, which is physical, 
-has a corresponding shadowing solution, that is nonphysical. In this section,
+has a corresponding shadowing solution that is nonphysical. In this section,
 we have shown that the same conclusion holds for several different perturbations 
@@
 The empirical distribution computed from long shadowing solutions, of length 10 billion, are shown in the left column of Figure \ref{fig:lorenz_params_scaled} at three different sets of parameter values, which are also different from the standard values of $\rho = 28, \sigma = 10$, and $\beta = 8/3$.
-In each row, one of the parameters is  $\rho = 29$ and $\rho = 30$ are used, respectivel, while the other two parameters are retained at their standard values.
-In the bottom row, only the parameter $\beta$ is perturbed from the standard values; $\beta = 10/3$ is used.
+In each row, one of the parameters is perturbed from the standard values.
 On the right column, we plot the distribution of physical solutions, the SRB distribution, at the set of parameters corresponding to each row.

### 2020-11-06 main.tex
Change a few sentences too similar to scholarpedia.

@@
 %             \emph{Present address:} of F. Author  %  if needed
-           Center for Computational Science and Engineering, MIT, Cambridge, MA 02139, USA. \email{nishac@mit.edu 
-		   }  \email{qiqi@mit.edu}
+           Center for Computational Science and Engineering, MIT, Cambridge, MA 02139, USA.
+Corresponding author email: \email{nishac@mit.edu}  \email{qiqi@mit.edu}.
+Preprint available at \cite{preprint}
 }
@@
 
-The existence of arbitrarily long shadowing solutions for sufficiently similar governing equations
-has been proven for invertible hyperbolic maps by Anosov \cite{anosov} and Bowen \cite{bowen}.
-For non-hyperbolic systems, it is often possible to estimate the length of time for which shadowing
-solutions exist \cite{grebogi}\cite{hammel}.
+When the slightly different governing equations are uniformly hyperbolic, it has been long know that infinitely 
+long shadowing solutions exist \cite{anosov}\cite{bowen}.
+Numerical experiments in non-hyperbolic systems have shown that shadowing
+solutions may still exist \cite{grebogi}\cite{hammel} for finite amounts of time.
 
@@
 The nonphysicality of shadowing solutions also indicates that shadowing-based methods can lead to incorrect values of sensitivities of statistics to parameter changes.
+
+\vspace{1in}
+
+\textbf{Funding:} This work was supported by Air Force Office of Scientific Research Grant No. FA8650-19-C-2207.
+\textbf{Conflict of Interest:} The authors declare that they have no conflict of interest.
 

### 2020-11-06 main.tex
Create and add new bst file for this journal
Updates from Overleaf

@@
 useful, despite the butterfly effect. Consider a numerical solution to a deterministic chaotic system with an initial condition $x$.
-Because of numerical and modeling errors, there will typically be a small difference between the
-true governing physics and the equations solved on a computer.  As the system is evolved forward
-in time, this difference will be amplified exponentially due to the chaotic nature of the system.
-The question therefore arises as to whether or not numerical solutions to chaotic systems, such
-as turbulent flows, represent the true physics.  The existence of shadowing solutions
-is used to argue for the usefulness of such numerical solutions.
+Because of numerical and modeling errors, we expect there will be a small difference between the
+true governing physics and the equations solved on a computer.
+With time evolution, by nature of chaos, any such small difference grows exponentially.
+Justifiably, this calls into question the fidelity of numerical solutions to chaotic systems, such as turbulent flows, to the true physics.
+The existence of shadowing solutions
+is used to argue for the veracity of numerical solutions.
 When certain conditions for shadowing theorems are met, the numerical solution would be an

### 2020-11-06 main.tex
Minor edits everywhere.

--- 2020-11-06 main.tex
@@
 
-\title{On the probability of finding a nonphysical solution through shadowing}
+\title{On the probability of finding a nonphysical solution through shadowing\footnote{Preprint available at \cite{preprint}}}
 \date{June 2020}
@@
 
-\author{Nisha Chandramoorthy and Qiqi Wang}
+\author{Nisha Chandramoorthy \and Qiqi Wang}
 
 %\authorrunning{Short form of author list} % if too long for running head
-
-\institute{           %  \\
+\institute{N. Chandramoorthy, Corresponding Author \at
+              Mechanical Engineering and Center for Computational Science and Engineering \\
+              77 Massachusetts Avenue, Cambridge MA
+               \email{nishac@mit.edu}           %  \\
 %             \emph{Present address:} of F. Author  %  if needed
-           Center for Computational Science and Engineering, MIT, Cambridge, MA 02139, USA.
-Corresponding author email: \email{nishac@mit.edu}  \email{qiqi@mit.edu}.
-Preprint available at \cite{preprint}
+           \and
+           Q. Wang \at
+              Department of Aeronautics and Astronautics and Center for Computational Science and Engineering \\
+              77 Massachusetts Avenue, Cambridge MA \\
+              \email{qiqi@mit.edu}
 }
-
 %\date{Received: date / Accepted: date}
@@
 
-When the slightly different governing equations are uniformly hyperbolic, it has been long know that infinitely 
+When the slightly different governing equations are uniformly hyperbolic, it has been long known that infinitely 
 long shadowing solutions exist \cite{anosov}\cite{bowen}.
 Numerical experiments in non-hyperbolic systems have shown that shadowing
-solutions may still exist \cite{grebogi}\cite{hammel} for finite amounts of time.
+solutions may still exist \cite{grebogi}\cite{hammel} for a 
+finite amount of time.
 
@@
 It is argued, based on shadowing, that such numerical simulations of chaotic dynamical systems can be
-useful, despite the butterfly effect. Consider a numerical solution to a deterministic chaotic system with an initial condition $x$.
-Because of numerical and modeling errors, there will typically be a small difference between the
-true governing physics and the equations solved on a computer.  As the system is evolved forward
-in time, this difference will be amplified exponentially due to the chaotic nature of the system.
-The question therefore arises as to whether or not numerical solutions to chaotic systems, such
-as turbulent flows, represent the true physics.  The existence of shadowing solutions
+useful, despite the \emph{butterfly effect} that causes infinitesimal errors in the state to grow exponentially in time.
+Because of numerical and modeling errors, there is invariably a small difference between the true governing physics and the equations solved on a computer.
+By nature of chaos, this difference increases exponentially, as the system is evolved forward in time.
+Thus, the fidelity of numerical solutions of chaotic systems, such
+as turbulent flows, to the true physics is called into question.  The existence of shadowing solutions
 is used to argue for the usefulness of such numerical solutions.
@@
 it is used in computing how long-time-averages in a chaotic system
-respond to small perturbations in the governing equation \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagna}. In this application, derivative
+respond to small perturbations in the governing equation \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagna}. In this application, the derivative
 of the long-time-averages is computed using a solution to the perturbed governing equation
@@
 There is an implicit assumption in both of these applications.  The assumption is that
-the shadowing trajectory is a ``physical'' trajectory, a trajectory on which
+the shadowing trajectory is a \emph{physical} trajectory, a trajectory on which
 the long-time-average of a quantity is equal to the ensemble average.
-Not all solutions satisfying the physical governing equation are considered ``physical''
-in this sense.
+Not all solutions satisfying the physical governing equation are considered physical in this sense.
 In high-Reynolds number fluid flows, for example, a steady-state, laminar flow solution may satisfy the Navier-Stokes equation.
@@
 It is also possible to effect significant change in the statistics of the true governing dynamics by introducing a minor parameter perturbation.
-Thus, although two trajectories, at slightly different parameters, may shadow each other, the long-time-averages of observables calculated using them, may not be close.
+Thus, although trajectories of governing equations at slightly different parameters may shadow each other, the long-time-averages that are typically observed at one parameter may not be close to that of the other.
 As we will see
@@
 
-Intuitively, we call a solution to a governing equation ``physical'' if
+Intuitively, we call a solution to a governing equation \emph{physical} if
 it represents what one would observe in a physical experiment.  In particular, the statistics measured from a physical solution match the statistics observed
@@
 solution, despite satisfying the governing equation, does not produce the
-turbulent statistics one would observe in a high-Reynolds number experiment. Such solutions are thus called nonphysical.
+turbulent statistics one would observe in a high-Reynolds number experiment. Such solutions are thus called \emph{nonphysical}.
 
@@
 \subsection{Physical solutions}
-What we typically call ``physical'' solutions to a governing equation satisfy the following
+What we have intuitively defined as physical solutions to a governing equation satisfy the following
 two criteria: 
@@
 R: distribution of a trajectory of $1000$ time units in length.
-In both plots, the color represent the number of samples in a
+In both plots, the color represents the number of samples in a
 $2048\times2048$ uniform grid.  The trajectory is sampled every 0.001 time units.
@@
 \begin{align}
-    J_1(u) = u_3\;,\quad J_2(u) = e^{-\dfrac{u_3^2}2}.
+		J_1(u) = u_3\;{\rm and}\;\quad J_2(u) = e^{-\dfrac{u_3^2}2}.
 \end{align}
@@
 We observe from the table that the mean of $J_1$ over the periodic
-solutions are different but comparable to the mean over physical solutions.  The mean of $J_2$ over the periodic solutions, on the other hand,
+solutions is different but comparable to the mean over physical solutions.  The mean of $J_2$ over the periodic solutions, on the other hand,
 is orders of magnitude different from that of the physical solutions.
@@
 This type is more difficult to find and study than the periodic solutions.
-To make it easier, we switch our example to a one-dimensional, discrete time dynamical
+To make it easier, we switch our example to a one-dimensional, 
+discrete-time dynamical
 systems like the Lorenz map.  Because the Lorenz map lacks a closed form,
@@
 \label{sec:tent}
-The tent map has been widely studied \cite{tent1}\cite{tent2}.  It is qualitatively similar
-to the Lorenz map and has a simple analytical form:
+The tent map is qualitatively similar
+to the Lorenz map,
 \begin{equation} \label{tentmap}
@@
     4 - 2x \quad & 2 \ge x \ge 1
-    \end{cases}
-\end{equation}
-Figure \ref{fig:tent_map} shows the tent map.
+    \end{cases}.
+\end{equation}
+Due to its simple analytical form, the properties of the tent map have been studied extensively \cite{tent1}\cite{tent2}. Figure \ref{fig:tent_map} shows the tent map.
 \begin{figure}\centering
@@
 \begin{equation} \label{tent_xor}
-    x_{i+1,j} = x_{i,0} \veebar x_{i,j+1}
+    x_{i+1,j} = x_{i,0} \veebar x_{i,j+1},
 \end{equation}
@@
 This fractal pattern further amplifies when $p=0.55$.  Meanwhile,
-the density on the left side, $[0,1]$, becomes obviously higher than
-the density on the right side, $[1,2]$.  This is consistent with our
+the density on the left sub-interval, $[0,1]$, becomes obviously higher than
+the density on the right sub-interval, $[1,2]$.  This is consistent with our
 theoretical analysis at the beginning of this subsection.
@@
 (\ref{tent_tilted}) is uniformly distributed in $[0,2]$, for
-any $0\le s<1$.
-According to the ergodic theory and the properties of the SRB measure of 
-one-dimensional chaotic systems (\cite{young}), 
-we only need to show that the uniform
+any $0\le s<1$. For this, we only need to show that the uniform
 distribution is stationary under the tilted tent map.
@@
 $x$, $\tilde{h}_s(x)$ is exactly the kind of initial condition
-we used to construct the quasi-physical solution
-in section \ref{sec:quasiphysical}, with $p=\frac{1+s}2$.
+we used to construct the quasi-physical solution in section \ref{sec:quasiphysical}, with $p=\frac{1+s}2$.
 Since this analysis holds for any $x$ sampled uniformly on $[0,2]$, we can conclude that
@@
 at two different values of $s.$ The left column showing the probability distribution of the shadowing solution is reminiscent of the hairy probability distributions of 
-the quasi-solutions of the tent map in section \ref{sec:quasiphysical}.
+the quasi-solutions of the tent map (section \ref{sec:quasiphysical}).
 The reader is referred to the Supplementary Material section \ref{sec:computingShadowing} for the computational method used in this paper, for the shadowing solutions.
@@
 wave tent map, starting from almost any point on $[0,2]$ -- a physical solution -- has a density that looks almost identical to a uniform density of 0.5.
-On the left column of Figure \ref{fig:tent_wave_shadow}, we observe that the distributions of the shadowing solutions corresponding to the physical solutions on the right, once again appear fractal.
+On the left column of Figure \ref{fig:tent_wave_shadow}, we observe that the distributions of the shadowing solutions corresponding to the physical solutions on the right once again appear fractal.
 \begin{figure}
@@
 
-Through a rigorous counter-example, we show that shadowing can lead
+Through rigorous counterexamples, we show that shadowing can lead
 to nonphysical solutions. This has a troubling implication for numerical simulations of chaotic
@@
 
-In view of the plucked tent map, we can question further the
+In view of the plucked tent map, we can further question the
 validity of numerical solutions. Can numerical solutions play 
@@
 
-\bibliographystyle{spmpsci}
+\bibliographystyle{amm}
 \bibliography{refs.bib}

### 2020-11-18 main.tex
Updates from Overleaf

--- 2020-11-06 main.tex
@@
 
-\author{Nisha Chandramoorthy and Qiqi Wang}
+\author{Nisha Chandramoorthy \and Qiqi Wang}
 
@@
 %             \emph{Present address:} of F. Author  %  if needed
-           Center for Computational Science and Engineering, MIT, Cambridge, MA 02139, USA.
-Corresponding author email: \email{nishac@mit.edu}  \email{qiqi@mit.edu}.
-Preprint available at \cite{preprint}
-}
+           Center for Computational Science and Engineering, MIT, Cambridge, MA 02139, USA. Corresponding author email: \email{nishac@mit.edu}  \email{qiqi@mit.edu}. }
 
@@
 \begin{abstract}
+In this work, we show that a small parameter perturbation can lead to a dramatic change in the statistics of a chaotic system.
+This theoretical observation implies that chaotic numerical simulations are not guaranteed to emulate the true  able to c ability of chaotic numerical simulations to capture the true long-term behavior of underlying governing equation.
+the failure of numerical simulations to capture the long-term statistical behavior of the governing equations.
+whether  the fidelity of chaotic numerical simulations to the underlying physics.
 Based on the shadowing property of dynamical systems, we expect that numerical solutions of governing equations lie close to a true solution -- a shadowing solution -- of the governing equation.

### 2020-11-18 main.tex

--- 2020-11-06 main.tex
@@
 is used to argue for the usefulness of such numerical solutions.
+
 When certain conditions for shadowing theorems are met, the numerical solution would be an

### 2020-11-20 main.tex
Add intro, change abstract

--- 2020-11-06 main.tex
@@
 \begin{abstract}
-Based on the shadowing property of dynamical systems, we expect that numerical solutions of governing equations lie close to a true solution -- a shadowing solution -- of the governing equation.
-This is used to justify, even when the governing equation is chaotic, that numerical solutions represent a true solution.
-However, shadowing solutions can be \emph{nonphysical}: they may not represent the long-term statistical behavior of an ensemble of true solutions.
-That is, the probability distribution of a shadowing solution can differ, often dramatically, from that of a solution starting from almost every initial condition.
-In this paper, we construct analytical examples of chaotic systems in which shadowing solutions are nonphysical.
-These examples also illustrate that the probability of finding a nonphysical shadowing solution can be 1.
-Through several analytical constructions, we note that the incorrect assumption of the physicality of shadowing raises questions on the trustworthiness of numerical solutions, and shadowing-based sensitivity analysis of chaotic systems.
-\keywords{Numerical simulations \and Shadowing sensitivity analysis \and Chaotic systems}
+This paper proves that shadowing solutions can be almost surely nonphysical.
+This finding invalidates the argument that small perturbations in a chaotic system can only have a small impact on its statistical behavior.
+This theoretical finding has implications for many applications
+in which chaotic mechanics play an important role.
+It suggests, for example, that we can control the climate through subtle perturbations.
+It also suggests that numerical simulations of chaotic dynamics, such as turbulent flows, may fail to predict the true long-term or statistical behavior.
+\keywords{Climate \and Chaotic systems \and Numerical simulations \and Turbulence \and Shadowing sensitivity analysis}
 % \PACS{PACS code1 \and PACS code2 \and more}
@@
 \maketitle
-\section{Shadowing and some applications}
-
+\section{Introduction}
+In this paper, we show that a large mismatch is possible in the statistical or long-time averages measured in a pair of chaotic solutions to slightly different models.
+We construct numerical and analytical examples that illustrate this feature of chaotic dynamics.
+The analysis of the long-term behavior of {\em shadowing} solutions in the majority of the chaotic models constructed in this paper reveals that shadowing solutions are almost always {\em nonphysical}: they do not represent the behavior of a physically observed solution starting almost everywhere.
+This also leads us to the conclusion that shadowing-based sensitivities \cite{qiqi-lss} of long-time averages to parameter changes can be incorrect.
+Thus, the main conjecture of this paper urges us to reconsider the use of chaotic numerical simulations, and statistical sensitivity analysis on these simulations.
+These implications are particularly important for climate studies, where chaotic models are widely used to make long-term predictions.
+In particular, three implications of the evidence in this paper may be stated, the third of which harnesses the statistical inconsistency between slightly different models:
+\begin{itemize}
+    \item we cannot trust the long-time averages generated by numerical simulations of chaotic processes, even when we eliminate model uncertainties stemming from imperfect modeling of the underlying physics, and statistical noise due to finite-time averaging;
+    \item we cannot trust the sensitivities of long-time averages or statistical averages to small parameter changes computed by using shadowing-based methods for sensitivity computation;
+    \item it is possible for two models separated by tiny changes in parameter values to have drastically different statistical behavior.
+This means that the Lorenzian butterfly can significantly alter the climate (statistical/long-term average) of Texas, and not just its weather (short-term events).
+\end{itemize}
+The paper is organized as follows. In section \ref{sec:shadowingIntro}, we introduce the dynamical systems concept of shadowing and its applications in the context of chaotic numerical simulations.
+
+\section{Shadowing and applications}
+\label{sec:shadowingIntro}
+Our main analysis tool in advancing our conjecture is {\em shadowing}.
+Shadowing and other dynamical systems concepts used in this paper are introduced in detail, where they appear, so as to make our presentation accessible to the large audience of computational scientists across different disciplines.
 Shadowing refers to the relationship between a pair of solutions to slightly different governing equations.
@@
 \section{What are physical and nonphysical solutions?}
-
+\label{sec:nonphysical}
 Intuitively, we call a solution to a governing equation \emph{physical} if
@@
 \section{Are shadowing solutions physical?}
-
+\label{sec:shadowingnonphysical}
 \subsection{Example of a shadowing solution for the tent map}
@@
 \section{Discussion and outlook}
-
+\label{sec:discussion}
 Through rigorous counterexamples, we show that shadowing can lead
@@
 %\subsection{The Lorenz Map}
-The supplementary material for this paper including the code, data and figures can be found on Github \cite{supp}.
+The supplementary material for this paper including the code and data to generate the figures can be found on Github \cite{supp}.
 The code can be found under the \verb+code+ subdirectory inside which section-wise code is separated into further subdirectories.

### 2020-11-21 main.tex
Changes to abstract and introduction

@@
 We construct numerical and analytical examples that illustrate this feature of chaotic dynamics.
-The analysis of the long-term behavior of {\em shadowing} solutions in the majority of the chaotic models constructed in this paper reveals that shadowing solutions are almost always {\em nonphysical}: they do not represent the behavior of a physically observed solution starting almost everywhere.
+The analysis of the long-term behavior of {\em shadowing} solutions in the majority of the chaotic models constructed in this paper reveals that shadowing solutions are almost surely {\em nonphysical}: they do not represent the behavior of a physically observed solution starting almost everywhere.
 This also leads us to the conclusion that shadowing-based sensitivities \cite{qiqi-lss} of long-time averages to parameter changes can be incorrect.
@@
 \end{itemize}
-The paper is organized as follows. In section \ref{sec:shadowingIntro}, we introduce the dynamical systems concept of shadowing and its applications in the context of chaotic numerical simulations.
+The paper is organized as follows.
+In section \ref{sec:shadowingIntro}, we introduce the dynamical systems concept of shadowing and its applications in the context of chaotic numerical simulations.
+In section \ref{sec:nonphysical}, we define the notion of physicality and give examples to illustrate the different ways in which to obtain nonphysical solutions.
+Section \ref{sec:shadowingnonphysical} is dedicated to several examples of perturbed tent maps that have nonphysical shadowing solutions.
+In the discussion in section \ref{sec:discussion}, we illustrate with an example that a small parameter perturbation can change the statistics significantly.
+The supplementary material, section \ref{sec:supp}, contains the numerical procedure to generate the probability distributions of the shadowing solutions.
 
@@
 \section{Supplementary material}
+\label{sec:supp}
 %\subsection{The Lorenz Map}

### 2020-12-08 main.tex
Correct type error
change to general comput phys

@@
 This theoretical finding has implications for many applications
-in which chaotic mechanics play an important role.
+in which chaotic mechanics plays an important role.
 It suggests, for example, that we can control the climate through subtle perturbations.
-It also suggests that numerical simulations of chaotic dynamics, such as turbulent flows, may fail to predict the true long-term or statistical behavior.
+It also suggests that numerical simulations of chaotic dynamics, such as turbulent flows and global atmosphere and ocean circulation, may fail to predict the true long-term or statistical behavior.
 \keywords{Climate \and Chaotic systems \and Numerical simulations \and Turbulence \and Shadowing sensitivity analysis}

### 2020-12-08 main_els.tex
Add new elsarticle class.

--- 2020-12-08 main.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% This is a general template file for the LaTeX package SVJour3
-% for Springer journals.          Springer Heidelberg 2010/09/16
-%
-% Copy it to a new file with a new name and use it as the basis
-% for your article. Delete % signs as needed.
-%
-% This template includes a few options for different layouts and
-% content for various journals. Please consult a previous issue of
-% your journal as needed.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% First comes an example EPS file -- just ignore it and
-% proceed on the \documentclass line
-% your LaTeX will extract the file if required
-%\begin{filecontents*}{example.eps}
-%!PS-Adobe-3.0 EPSF-3.0
-%%BoundingBox: 19 19 221 221
-%%CreationDate: Mon Sep 29 1997
-%%Creator: programmed by hand (JK)
-%%EndComments
-%gsave
-%newpath
-%  20 20 moveto
-%  20 220 lineto
-%  220 220 lineto
-%  220 20 lineto
-%closepath
-%2 setlinewidth
-%gsave
-%  .4 setgray fill
-%grestore
-%stroke
-%grestore
-%\end{filecontents*}
-%
-\RequirePackage{fix-cm}
-%
-%\documentclass{svjour3}                     % onecolumn (standard format)
-%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
-\documentclass[smallextended]{svjour3}       % onecolumn (second format)
-%\documentclass[twocolumn]{svjour3}          % twocolumn
-%
-\smartqed  % flush right qed marks, e.g. at end of proof
-%
-% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
-%
-% insert here the call for the packages your document requires
-%\usepackage{latexsym}
-% etc.
-%
-% please place your own definitions here and don't use \def but
-% \newcommand{}{}
-%
-% Insert the name of "your journal" with
-% \journalname{myjournal}
+%% 
+%% Copyright 2007-2020 Elsevier Ltd
+%% 
+%% This file is part of the 'Elsarticle Bundle'.
+%% ---------------------------------------------
+%% 
+%% It may be distributed under the conditions of the LaTeX Project Public
+%% License, either version 1.2 of this license or (at your option) any
+%% later version.  The latest version of this license is in
+%%    http://www.latex-project.org/lppl.txt
+%% and version 1.2 or later is part of all distributions of LaTeX
+%% version 1999/12/01 or later.
+%% 
+%% The list of all files belonging to the 'Elsarticle Bundle' is
+%% given in the file `manifest.txt'.
+%%
+
+%% Template article for Elsevier's document class `elsarticle'
+%% with numbered style bibliographic references
+%% SP 2008/03/01
+%%
+%% 
+%%
+%% $Id: elsarticle-template-num.tex 190 2020-11-23 11:12:32Z rishi $
+%%
+%%
+\documentclass[preprint,12pt]{elsarticle}
+
+%% Use the option review to obtain double line spacing
+%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
+
+%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
+%% for a journal layout:
+%% \documentclass[final,1p,times]{elsarticle}
+%% \documentclass[final,1p,times,twocolumn]{elsarticle}
+%% \documentclass[final,3p,times]{elsarticle}
+%% \documentclass[final,3p,times,twocolumn]{elsarticle}
+%% \documentclass[final,5p,times]{elsarticle}
+%% \documentclass[final,5p,times,twocolumn]{elsarticle}
+
+%% For including figures, graphicx.sty has been loaded in
+%% elsarticle.cls. If you prefer to use the old commands
+%% please give \usepackage{epsfig}
+
+%% The amssymb package provides various useful mathematical symbols
 \usepackage[utf8]{inputenc}
@@
 \usepackage{amsmath, amssymb, amsfonts}
-\usepackage{mathtools,hyperref}
+\usepackage{mathtools}
 \usepackage{xcolor,verbatim}
@@
 \DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
+
+%% The amsthm package provides extended theorem environments
+%% \usepackage{amsthm}
+
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers.
+%% \usepackage{lineno}
+
+\journal{Journal of Computational Physics}
+
 \begin{document}
 
-\title{On the probability of finding a nonphysical solution through shadowing\footnote{Preprint available at \cite{preprint}}}
-\date{June 2020}
-
-%\titlerunning{Short form of title}        % if too long for running head
-
-\author{Nisha Chandramoorthy \and Qiqi Wang}
-
-%\authorrunning{Short form of author list} % if too long for running head
-\institute{N. Chandramoorthy, Corresponding Author \at
-              Mechanical Engineering and Center for Computational Science and Engineering \\
-              77 Massachusetts Avenue, Cambridge MA
-               \email{nishac@mit.edu}           %  \\
-%             \emph{Present address:} of F. Author  %  if needed
-           \and
-           Q. Wang \at
-              Department of Aeronautics and Astronautics and Center for Computational Science and Engineering \\
-              77 Massachusetts Avenue, Cambridge MA \\
-              \email{qiqi@mit.edu}
-}
-%\date{Received: date / Accepted: date}
-% The correct dates will be entered by the editor
-
-\maketitle
-
+\begin{frontmatter}
+
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for theassociated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for theassociated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for theassociated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \affiliation{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+%% \fntext[label3]{}
+
+		\title{On the probability of finding a nonphysical solution through shadowing\tnoteref{1}}
+		\tnotetext[1]{Preprint available at \cite{preprint}}
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{}
+%% \affiliation[label1]{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+%%
+%% \affiliation[label2]{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+
+\author[N,ccse]{Nisha Chandramoorthy\corref{cor}}
+\cortext[cor]{Corresponding author}
+\ead{nishac@mit.edu}
+\affiliation[ccse]{organization={Center for Computational Science and Engineering, Massachusetts Institute of Technology},%Department and Organization
+            addressline={77 Massachusetts Avenue}, 
+            city={Cambridge},
+            postcode={02139}, 
+            state={MA},
+            country={USA}}
+\author[Q,ccse]{Qiqi Wang}
+\ead{qiqi@mit.edu}
+\affiliation[N]{Department of Mechanical Engineering}
+\affiliation[Q]{Department of Aeronautics and Astronautics}
 \begin{abstract}
@@
 It also suggests that numerical simulations of chaotic dynamics, such as turbulent flows and global atmosphere and ocean circulation, may fail to predict the true long-term or statistical behavior.
-\keywords{Climate \and Chaotic systems \and Numerical simulations \and Turbulence \and Shadowing sensitivity analysis}
 % \PACS{PACS code1 \and PACS code2 \and more}
@@
 
-\maketitle
+%%Research highlights
+\begin{highlights}
+\item Research highlight 1
+\item Research highlight 2
+\end{highlights}
+
+\begin{keyword}
+Chaotic dynamics \sep numerical simulations \sep scientific computing 
+\sep shadowing sensitivity analysis %% keywords here, in the form: keyword \sep keyword
+
+%% PACS codes here, in the form: \PACS code \sep code
+
+%% MSC codes here, in the form: \MSC code \sep code
+%% or \MSC[2008] code \sep code (2000 is the default)
+
+\end{keyword}
+
+\end{frontmatter}
+
+%% \linenumbers
+
 \section{Introduction}
@@
 	\item it is possible for two models separated by tiny changes in parameter values to have drastically different statistical behavior.
-This means that the Lorenzian butterfly can significantly alter the climate (statistical/long-term average) of Texas, and not just its weather (short-term events).
+This means that the Lorenzian butterfly (a small perturbation) can significantly alter the climate of Texas (statistical/long-term average), and not just its weather (short-term events).
 \end{itemize}
@@
 
-\bibliographystyle{amm}
+\bibliographystyle{elsarticle-num} 
 \bibliography{refs.bib}
@@
 physical probability distributions can be found under \verb+tent_sens_stat+.
+
 \end{document}
+\endinput
+%%

### 2020-12-08 letter/letter.tex
Remove amm related files

@@
 ``On the probability of finding a nonphysical solution through shadowing ''
-for consideration for publication in Applied Mathematics and Mechanics. Thank you very much for giving us the 
+for consideration for publication in the Journal of Computational Physics. Thank you very much for giving us the 
 opportunity to have our work reviewed!

### 2020-12-08 code/tent_quasiphysical/hist0.py

@@ -3,7 +3,6 @@ from pylab import *
 from numpy import *
 
 width = 2**13
-
 @numba.jit(nopython=True)
 def transform(x, p):
     addition = (random.rand() > p) ^ (x % 2)
@@ -22,7 +21,7 @@ def accumulate(n, p):
     return xhist
 
 n = 10000000000
-
+'''
 figure(figsize=(16,18))
 density = array(accumulate(n, 0.5001), float) / n * width / 2
 x = linspace(0,2,width*2+1)[1:-1:2]
@@ -55,9 +54,8 @@ for tick in gca().xaxis.get_major_ticks():
 for tick in gca().yaxis.get_major_ticks():
     tick.label.set_fontsize(40)
 savefig('tent_quasiphysical_hist_p_0.55.png')
-
+'''
 figure(figsize=(16,18))
-density = array(accumulate(n, 0.9), float) / n * width / 2
 x = linspace(0,2,width*2+1)[1:-1:2]
 fill_between(x, density)
 ylim([0,1])
@@ -66,3 +64,4 @@ for tick in gca().xaxis.get_major_ticks():
 for tick in gca().yaxis.get_major_ticks():
     tick.label.set_fontsize(40)
 savefig('tent_quasiphysical_hist_p_0.9.png')
+

### 2020-12-08 code/tent_sens_stat/plot_density.py

@@ -26,7 +26,7 @@ nbins = 2**12
 nrep = 5
 density = zeros(nbins)
 s = 0.1
-n = 0
+n = 3
 for m in range(nrep):
     print("repeat {}".format(m))
     density += accumulate(nbins, 1000000000, n, s)/nrep

### 2020-12-08 code/tent_sens_stat/plot_map.py

@@ -4,8 +4,8 @@ from plucked_map import *
 nbins = 2**13
 ns = 5
 nn = 5
-n_arr = [0, 2, 4, 6]
-s_arr = [0.2, 0.5]
+n_arr = [0, 3, 6]
+s_arr = [0., 0.1,0.2, 0.5]
 
 x = linspace(0, 2, nbins+2)[1:-1]
 ys = zeros((ns,nbins))
@@ -14,7 +14,7 @@ for kn,nk in enumerate(n_arr):
     for k,sk in enumerate(s_arr):
         for i, xi in enumerate(x):
             yn[k,i] = osc_tent(xi, sk, nk)
-    fig = figure(figsize=(8,6))
+    fig = figure(figsize=(10,8))
     ax = fig.add_subplot(111)
     for k,sk in enumerate(s_arr):
         ax.plot(x, yn[k],lw=2.5,label="s = {}".format(sk))

### 2020-12-08 code/tent_shadow/tent_shadow_pinched.py

@@ -22,7 +22,7 @@ def shadow(n, s, width):
         count[int(floor(y * width / 2))] += 1
         x = x / 2 + random.randint(0,2)
     return count
-
+'''
 n = 10000000000
 for s in [0.01, 0.05, 0.2, 0.9]:
     width = 2**13
@@ -51,3 +51,4 @@ for tick in gca().xaxis.get_major_ticks():
 for tick in gca().yaxis.get_major_ticks():
     tick.label.set_fontsize(30)
 savefig('pinched_tent_map.png')
+'''

### 2020-12-12 main_els.tex
Edit intro again

@@
 \section{Introduction}
-In this paper, we show that a large mismatch is possible in the statistical or long-time averages measured in a pair of chaotic solutions to slightly different models.
+In this paper, we show that a large mismatch is possible in the ensemble averages of two slightly different chaotic systems.
 We construct numerical and analytical examples that illustrate this feature of chaotic dynamics.

### 2020-12-12 main_els.tex
Updates from Overleaf

@@
 
-		\title{On the probability of finding a nonphysical solution through shadowing\tnoteref{1}}
+		\title{On the probability of finding nonphysical solutions through shadowing\tnoteref{1}}
 		\tnotetext[1]{Preprint available at \cite{preprint}}
@@
 \section{Introduction}
-In this paper, we show that a large mismatch is possible in the ensemble averages of two slightly different chaotic systems.
-We construct numerical and analytical examples that illustrate this feature of chaotic dynamics.
+Owing to the {\em shadowing} property of dynamical systems, it is generally believed that a numerical simulation of chaotic governing equation, which are small perturbations of the true solution, are {\em shadowed} or closely followed by a true solution of the physics.
+In this paper, we show through rigorous analytical and numerical examples that these shadowing orbits are not representative solutions of the true governing equation in the sense that they are statistically inaccurate.
+In other words, shadowing orbits
+In this paper, we show that a large mismatch is possible in the statistical or long-time averages measured in a pair of chaotic solutions to slightly different models.
+that  We construct numerical and analytical examples that illustrate this feature of chaotic dynamics.
 The analysis of the long-term behavior of {\em shadowing} solutions in the majority of the chaotic models constructed in this paper reveals that shadowing solutions are almost surely {\em nonphysical}: they do not represent the behavior of a physically observed solution starting almost everywhere.

### 2020-12-12 main_els.tex

--- 2020-12-12 main_els.tex
@@
 In this paper, we show that a large mismatch is possible in the ensemble averages of two slightly different chaotic systems.
-We construct numerical and analytical examples that illustrate this feature of chaotic dynamics.
+We illustrate this feature of chaotic dynamics by studying the long-time behavior of {\em shadowing} solutions in analytically or numerically constructed examples.{\em Shadowing} solutions are true orbits that closely follow a slightly perturbed solution, say due to a small parameter perturbation in the original governing equation, for a long, and sometimes infinitely long, time.
+In this paper, we show through rigorous analytical and numerical examples that these shadowing orbits can be statistically mismatched.
+In other words, the probability distribution of shadowing orbits is different, often drastically so, from that of almost every solution.
+
 The analysis of the long-term behavior of {\em shadowing} solutions in the majority of the chaotic models constructed in this paper reveals that shadowing solutions are almost surely {\em nonphysical}: they do not represent the behavior of a physically observed solution starting almost everywhere.
@@
 In particular, three implications of the evidence in this paper may be stated, the third of which harnesses the statistical inconsistency between slightly different models:
+
+Owing to the {\em shadowing} property of dynamical systems, it is generally believed that a numerical simulation of chaotic governing equation, which are small perturbations of the true solution, are {\em shadowed} or closely followed by a true solution of the physics.
+
+In this paper, we show that a large mismatch is possible in the statistical or long-time averages measured in a pair of chaotic solutions to slightly different models.
+that  We construct numerical and analytical examples that illustrate this feature of chaotic dynamics.
+The analysis of the long-term behavior of {\em shadowing} solutions in the majority of the chaotic models constructed in this paper reveals that shadowing solutions are almost surely {\em nonphysical}: they do not represent the behavior of a physically observed solution starting almost everywhere.
+This also leads us to the conclusion that shadowing-based sensitivities \cite{qiqi-lss} of long-time averages to parameter changes can be incorrect.
+Thus, the main conjecture of this paper urges us to reconsider the use of chaotic numerical simulations, and statistical sensitivity analysis on these simulations.
+These implications are particularly important for climate studies, where chaotic models are widely used to make long-term predictions.
+In particular, three implications of the evidence in this paper may be stated, the third of which harnesses the statistical inconsistency between slightly different models:
+>>>>>>> origin/overleaf-2020-12-12-1530
 \begin{itemize}

### 2020-12-12 main_els.tex
First para of intro complete.

@@
 \section{Introduction}
-In this paper, we show that a large mismatch is possible in the ensemble averages of two slightly different chaotic systems.
-We illustrate this feature of chaotic dynamics by studying the long-time behavior of {\em shadowing} solutions in analytically or numerically constructed examples.{\em Shadowing} solutions are true orbits that closely follow a slightly perturbed solution, say due to a small parameter perturbation in the original governing equation, for a long, and sometimes infinitely long, time.
-In this paper, we show through rigorous analytical and numerical examples that these shadowing orbits can be statistically mismatched.
-In other words, the probability distribution of shadowing orbits is different, often drastically so, from that of almost every solution.
+In this paper, we show that a large mismatch is possible in the long-time or statistical behavior of two slightly different chaotic systems.
+We illustrate this feature of chaotic dynamics by studying long-time averages along {\em shadowing} solutions in analytically or numerically constructed
+examples.
+Shadowing solutions are true orbits of a dynamical system that closely follow slightly perturbed orbits, or {\em pseudo-orbits},  for a long, and sometimes infinitely long, time.
+In this paper, we show through rigorous analytical and numerical examples that shadowing orbits may not represent the long-time behavior of almost every orbit.
+In other words, the probability distribution of shadowing orbits is different, often drastically so, from that of almost every solution of the unperturbed system.
+The pseudo-orbits we study are solutions to slightly different governing equations that are generated by small parameter perturbations of the original system.
+Because we show that shadowing solutions following these pseudo-orbits almost surely have different statistics when compared to an ensemble of unperturbed orbits, we prove that small parameter perturbations can produce a large statistical mismatch between the perturbed and unperturbed systems.
 

### 2020-12-12 main_els.tex
Shadowing second para

@@
 In this paper, we show that a large mismatch is possible in the long-time or statistical behavior of two slightly different chaotic systems.
-We illustrate this feature of chaotic dynamics by studying long-time averages along {\em shadowing} solutions in analytically or numerically constructed
-examples.
+We illustrate this feature of chaotic dynamics by studying long-time averages along {\em shadowing} solutions in analytically or numerically constructed examples.
 Shadowing solutions are true orbits of a dynamical system that closely follow slightly perturbed orbits, or {\em pseudo-orbits},  for a long, and sometimes infinitely long, time.
@@
 
-The analysis of the long-term behavior of {\em shadowing} solutions in the majority of the chaotic models constructed in this paper reveals that shadowing solutions are almost surely {\em nonphysical}: they do not represent the behavior of a physically observed solution starting almost everywhere.
+We refer to the statistically mismatched shadowing solutions as {\em nonphysical} solutions since they do not represent the behavior of a physically observed solution starting almost everywhere.
+We identify different ways in which nonphysical solutions occur, adding to the known examples such as periodic orbits embedded within a chaotic attractor.
+The shadowing lemma \cite{anosov}\cite{bowen} establishes the existence of shadowing orbits corresponding to small perturbations but does not preclude the possibility of nonphysical shadowing orbits (\cite{katok} Theorem 18.1.2).
+
+Owing to the {\em shadowing} property of dynamical systems, it is generally believed that a numerical simulation of chaotic governing equation, which are small perturbations of the true solution, are {\em shadowed} or closely followed by a true solution of the physics.
+
 This also leads us to the conclusion that shadowing-based sensitivities \cite{qiqi-lss} of long-time averages to parameter changes can be incorrect.
@@
 In particular, three implications of the evidence in this paper may be stated, the third of which harnesses the statistical inconsistency between slightly different models:
-
-Owing to the {\em shadowing} property of dynamical systems, it is generally believed that a numerical simulation of chaotic governing equation, which are small perturbations of the true solution, are {\em shadowed} or closely followed by a true solution of the physics.
 

### 2020-12-12 main_els.tex
Intro complete.

@@
 We identify different ways in which nonphysical solutions occur, adding to the known examples such as periodic orbits embedded within a chaotic attractor.
-The shadowing lemma \cite{anosov}\cite{bowen} establishes the existence of shadowing orbits corresponding to small perturbations but does not preclude the possibility of nonphysical shadowing orbits (\cite{katok} Theorem 18.1.2).
-
-Owing to the {\em shadowing} property of dynamical systems, it is generally believed that a numerical simulation of chaotic governing equation, which are small perturbations of the true solution, are {\em shadowed} or closely followed by a true solution of the physics.
-
-This also leads us to the conclusion that shadowing-based sensitivities \cite{qiqi-lss} of long-time averages to parameter changes can be incorrect.
-Thus, the main conjecture of this paper urges us to reconsider the use of chaotic numerical simulations, and statistical sensitivity analysis on these simulations.
+The shadowing lemma \cite{anosov}\cite{bowen} establishes the existence of shadowing orbits corresponding to small perturbations but does not preclude the nonphysicality of these shadowing orbits (\cite{katok} Theorem 18.1.2).
+One contribution of the present work is the construction of rigorous examples that show that shadowing orbits can be nonphysical, including relating them to previously identified forms of nonphysical orbits, and computing the probability of finding nonphysical shadowing orbits.
+Another contribution is the interpretation of this finding in the context of chaotic numerical simulations and sensitivity analysis on such simulations.
+
+Owing to the shadowing lemma, it is generally believed that numerical simulations of chaotic governing equations, which are small perturbations of true solutions, are shadowed by true solutions of the governing physics.
+However, if these shadowing solutions are nonphysical, numerical simulations do not produce the correct long-time averages.
+Furthermore, the computation of parametric derivatives or {\em linear response} of
+long-time averages by using the idea of shadowing \cite{qiqi-lss}, leads to incorrect values for those derivatives.
+Thus, the evidence of nonphysicality of shadowing presented in this paper urges us to reconsider the use of chaotic numerical simulations, and statistical sensitivity analysis on these simulations, which are critical for cross-disciplinary applications.
 These implications are particularly important for climate studies, where chaotic models are widely used to make long-term predictions.
-In particular, three implications of the evidence in this paper may be stated, the third of which harnesses the statistical inconsistency between slightly different models:
-
-In this paper, we show that a large mismatch is possible in the statistical or long-time averages measured in a pair of chaotic solutions to slightly different models.
-that  We construct numerical and analytical examples that illustrate this feature of chaotic dynamics.
-The analysis of the long-term behavior of {\em shadowing} solutions in the majority of the chaotic models constructed in this paper reveals that shadowing solutions are almost surely {\em nonphysical}: they do not represent the behavior of a physically observed solution starting almost everywhere.
-This also leads us to the conclusion that shadowing-based sensitivities \cite{qiqi-lss} of long-time averages to parameter changes can be incorrect.
-Thus, the main conjecture of this paper urges us to reconsider the use of chaotic numerical simulations, and statistical sensitivity analysis on these simulations.
-These implications are particularly important for climate studies, where chaotic models are widely used to make long-term predictions.
-In particular, three implications of the evidence in this paper may be stated, the third of which harnesses the statistical inconsistency between slightly different models:
->>>>>>> origin/overleaf-2020-12-12-1530
+In particular, three broad implications of the evidence in this paper may be stated, the third of which harnesses the statistical inconsistency between slightly different models:
 \begin{itemize}
-    \item we cannot trust the long-time averages generated by numerical simulations of chaotic processes, even when we eliminate model uncertainties stemming from imperfect modeling of the underlying physics, and statistical noise due to finite-time averaging;
-    \item we cannot trust the sensitivities of long-time averages or statistical averages to small parameter changes computed by using shadowing-based methods for sensitivity computation;
+    \item we cannot trust the long-time averages generated by numerical simulations of chaotic processes, even when we account for model uncertainties and statistical noise due to finite-time averaging;
+    \item we cannot trust the sensitivities of long-time averages or statistical averages to small parameter changes computed by using shadowing-based methods;
 	\item it is possible for two models separated by tiny changes in parameter values to have drastically different statistical behavior.
-This means that the Lorenzian butterfly (a small perturbation) can significantly alter the climate of Texas (statistical/long-term average), and not just its weather (short-term events).
+This means that the Lorenzian butterfly (a small perturbation) can significantly alter the climate of Texas (statistical/long-term average), and not just its weather (short-term averages).
 \end{itemize}

### 2020-12-12 main_els.tex
Write highlights

@@
 %%Research highlights
+This paper highlights through rigorous examples that
 \begin{highlights}
-\item Research highlight 1
-\item Research highlight 2
+\item numerical solutions of chaotic dynamical systems are not guaranteed to produce the correct statistics or long-term behavior;
+\item it is theoretically possible to significant alter the statistics of a chaotic system, e.g. the earth's climate, using tiny parameter perturbations.
 \end{highlights}
@@
 \end{itemize}
+Our subject -- shadowing -- and other dynamical systems concepts used in this paper are introduced in detail, where they appear, so as to make our presentation accessible to computational scientists across disciplines.
 The paper is organized as follows.
-In section \ref{sec:shadowingIntro}, we introduce the dynamical systems concept of shadowing and its applications in the context of chaotic numerical simulations.
+In section \ref{sec:shadowingIntro}, we introduce the concept of shadowing and its applications in the context of chaotic numerical simulations.
 In section \ref{sec:nonphysical}, we define the notion of physicality and give examples to illustrate the different ways in which to obtain nonphysical solutions.
 Section \ref{sec:shadowingnonphysical} is dedicated to several examples of perturbed tent maps that have nonphysical shadowing solutions.
-In the discussion in section \ref{sec:discussion}, we illustrate with an example that a small parameter perturbation can change the statistics significantly.
-The supplementary material, section \ref{sec:supp}, contains the numerical procedure to generate the probability distributions of the shadowing solutions.
+In the discussion in section \ref{sec:discussion}, we illustrate with an example that a small parameter perturbation can change statistics significantly.
+The supplementary material, section \ref{sec:supp}, contains the numerical procedure used to generate shadowing solutions.
 
@@
 \label{sec:shadowingIntro}
-Our main analysis tool in advancing our conjecture is {\em shadowing}.
-Shadowing and other dynamical systems concepts used in this paper are introduced in detail, where they appear, so as to make our presentation accessible to the large audience of computational scientists across different disciplines.
 Shadowing refers to the relationship between a pair of solutions to slightly different governing equations.

### 2020-12-12 main_els.tex
Made edits throughout.

@@
 This theoretical finding has implications for many applications
-in which chaotic mechanics plays an important role.
+in which chaotic dynamics plays an important role.
 It suggests, for example, that we can control the climate through subtle perturbations.
@@
 %%Research highlights
-This paper highlights through rigorous examples that
+
 \begin{highlights}
-\item numerical solutions of chaotic dynamical systems are not guaranteed to produce the correct statistics or long-term behavior;
-\item it is theoretically possible to significant alter the statistics of a chaotic system, e.g. the earth's climate, using tiny parameter perturbations.
+\item This paper highlights through rigorous examples that numerical solutions of chaotic dynamical systems are not guaranteed to produce the correct statistics or long-term behavior;
+\item We show through examples that it is theoretically possible to significantly alter the statistics of a chaotic system, e.g. the earth's climate, by minor parameter perturbations.
 \end{highlights}
@@
 The existence of shadowing trajectories underlies many applications.
-Numerical simulations of turbulent flows have been widely used to study its statistical behavior.
+Numerical simulations of turbulent flows have been extensively used to study their statistical behavior.
 It is argued, based on shadowing, that such numerical simulations of chaotic dynamical systems can be
@@
 Shadowing is also used in sensitivity analysis of chaotic dynamical systems.  In particular,
-it is used in computing how long-time-averages in a chaotic system
+it is used in computing how long-time averages in a chaotic system
 respond to small perturbations in the governing equation \cite{qiqi-lss}\cite{angxiu-lss}\cite{lasagna}. In this application, the derivative
-of the long-time-averages is computed using a solution to the perturbed governing equation
-that shadows a solution to the unperturbed equation.  The Least Squares Shadowing \cite{qiqi-lss}\cite{angxiu-lss} method
-uses this concept.
-
-There is an implicit assumption in both of these applications.  The assumption is that
-the shadowing trajectory is a \emph{physical} trajectory, a trajectory on which
-the long-time-average of a quantity is equal to the ensemble average.
-Not all solutions satisfying the physical governing equation are considered physical in this sense.
+of the long-time averages is computed using a solution of the perturbed governing equation
+that shadows a solution of the unperturbed equation.  The Least Squares Shadowing \cite{qiqi-lss}\cite{angxiu-lss} method
+uses this idea.
+
+There is an implicit assumption in both of these applications: chaotic numerical simulations and shadowing-based sensitivity analysis.
+The assumption is that a shadowing trajectory is a \emph{physical} trajectory, a trajectory on which
+the long-time average of a quantity is equal to the ensemble average.
+Not all solutions satisfying the governing equation are considered physical in this sense.
 In high-Reynolds number fluid flows, for example, a steady-state, laminar flow solution may satisfy the Navier-Stokes equation.
@@
 These trajectories can have a probability distribution that is remarkably different from that of a typically observed solution of the governing equation.
-It is also possible to effect significant change in the statistics of the true governing dynamics by introducing a minor parameter perturbation.
-Thus, although trajectories of governing equations at slightly different parameters may shadow each other, the long-time-averages that are typically observed at one parameter may not be close to that of the other.
 As we will see
@@
 (SRB) measure.
-The absolute continuity property is sufficient to ensure that the SRB measure is the same $\mu.$ That is, the SRB measure is physically observed in the sense that physical solutions produce long-time-averages which are expectations with respect to the SRB measure.
+The absolute continuity property is sufficient to ensure that the SRB measure is the same $\mu.$ That is, the SRB measure is physically observed in the sense that physical solutions produce long-time averages which are expectations with respect to the SRB measure.
 
@@
 How then can we trust numerical solutions when they are not guaranteed to share the long-term statistical behavior of the governing equation?
+
 We conclude with an example that illustrates a counterintuitive feature of some chaotic systems: a small perturbation to the 
@@
 lead to small changes in statistics, which can be expanded as Taylor series around the reference parameter value.
-But, uniform hyperbolicity is a mathematical idealization, and although some physical systems have been observed to behave as if they were uniformly hyperbolic \cite{gallavotti}, a violation of uniform hyperbolicity is more likely
-\cite{wormell}\cite{luca}.
+But, uniform hyperbolicity is a mathematical idealization, and although some physical systems have been observed to behave as if they were uniformly hyperbolic \cite{gallavotti}, a violation of uniform hyperbolicity is also prevalent in physical systems
+\cite{wormell}\cite{lucarini}.
 
@@
 shadowed by a true solution in that the difference between them is small for a long time, this shadowing solution may not represent the long-term or ensemble behavior of the physical system.
-The nonphysicality of shadowing solutions also indicates that shadowing-based methods can lead to incorrect values of sensitivities of statistics to parameter changes.
-
+Because of nonphysicality of shadowing solutions, shadowing-based methods can lead to incorrect values of sensitivities of statistics to parameter changes.
+Finally, nonphysicality of shadowing also hints at a distinctive feature of some chaotic systems: physical solutions of two slightly different models may have remarkably different statistics.
+This suggests the possibility that we may be able to construct minor parameter perturbations that favorably alter climate statistics.
 \vspace{1in}
 
+{\em Acknowledgments:} The authors are grateful to Prof. Tim Sauer and Prof. John Williams for comments on the paper.
+\\
 \textbf{Funding:} This work was supported by Air Force Office of Scientific Research Grant No. FA8650-19-C-2207.

### 2020-12-13 letter/letter.tex
Add significance.

@@
 there are chaotic systems in which small parameter perturbations can, in fact, cause a dramatic change in the statistics.
-This extreme non-smooth response of a chaotic attractor further strengthens our conclusion that the assumption of physicality of shadowing solutions is easily violated.
+This extreme non-smooth response is an important finding because it suggests the theoretical possibility that we can favorably control our climate with subtle perturbations.
 
 This paper raises a fundamental and profound issue with numerical simulation and sensitivity analysis of chaotic governing equations.
-It therefore has immediate implications for the entire computational science and engineering community, which must work together to rethink chaotic simulations.
+It therefore has immediate implications for the entire computational science and engineering community.
 Thank you very much for your time and consideration,

### 2020-12-13 letter/significance.tex
Add significance.

\documentclass{letter}
\title{Letter to the Editor}
\begin{document}
\hspace{1.6in} 
\begin{large}{Significance and novelty}\end{large}
\\
\\

The major contribution of this work is to show, through rigorous examples, that shadowing solutions can be nonphysical, and to expose the implications of this finding. That is, we show that long-time averages computed along shadowing orbits are often drastically different from ensemble averages. This finding dispels the notion that small parameter perturbations in
a chaotic system result in small differences in its statistics or long-time averages. Hence, the trustworthiness of numerical simulations of chaos, which are small perturbations of the true governing equation, in representing physically observed long-term behavior, is called into question. This finding is especially important in climate studies and turbulent flows in engineering, where detailed chaotic simulations are used for statistical predictions. It also hints at the theoretical possibility of harnessing this distinctive feature of chaotic dynamics: it implies that we can favorably alter climate statistics significantly by applying subtle perturbations. Finally, the nonphysicality of shadowing also reveals that shadowing-based sensitivity computation of parametric derivatives of statistics can lead to incorrect results.

Nisha Chandramoorthy and Qiqi Wang.
\end{document}

### 2020-12-13 highlights.tex
Add highlights and edit intro. Final JCP version.

\documentclass[12pt]{article}
\title{Research highlights}
\begin{document}
\maketitle
\begin{itemize}
\item This paper highlights through rigorous examples that numerical solutions of chaotic dynamical systems are not guaranteed to produce the correct statistics or long-term behavior;
\item We show through examples that it is theoretically possible to significantly alter the statistics of a chaotic system by minor parameter perturbations.
\item The nonphysicality of shadowing orbits also implies that methods to compute parametric derivatives of statistics in chaotic systems, based on shadowing, may produce incorrect results.
\end{itemize}
\end{document}

### 2020-12-13 main_els.tex
Add highlights and edit intro. Final JCP version.

@@
 %%Research highlights
-
-\begin{highlights}
-\item This paper highlights through rigorous examples that numerical solutions of chaotic dynamical systems are not guaranteed to produce the correct statistics or long-term behavior;
-\item We show through examples that it is theoretically possible to significantly alter the statistics of a chaotic system, e.g. the earth's climate, by minor parameter perturbations.
-\end{highlights}
-
 \begin{keyword}
@@
 Furthermore, the computation of parametric derivatives or {\em linear response} of
-long-time averages by using the idea of shadowing \cite{qiqi-lss}, leads to incorrect values for those derivatives.
+long-time averages by using the idea of shadowing \cite{qiqi-lss}, leads to incorrect values of those derivatives.
 Thus, the evidence of nonphysicality of shadowing presented in this paper urges us to reconsider the use of chaotic numerical simulations, and statistical sensitivity analysis on these simulations, which are critical for cross-disciplinary applications.
@@
     \item we cannot trust the sensitivities of long-time averages or statistical averages to small parameter changes computed by using shadowing-based methods;
-	\item it is possible for two models separated by tiny changes in parameter values to have drastically different statistical behavior.
+	\item it is possible for two models separated by minor changes in parameter values to have drastically different statistical behavior.
 This means that the Lorenzian butterfly (a small perturbation) can significantly alter the climate of Texas (statistical/long-term average), and not just its weather (short-term averages).
@@
 
-{\em Acknowledgments:} The authors are grateful to Prof. Tim Sauer and Prof. John Williams for comments on the paper.
-\\
 \textbf{Funding:} This work was supported by Air Force Office of Scientific Research Grant No. FA8650-19-C-2207.

### 2021-04-20 letter/responseToReviewers.tex
Add template for response to reviewers.

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1.1in]{geometry}
\usepackage{graphicx}
\usepackage[numbers]{natbib}
\usepackage{verbatim}
\usepackage[textsize=tiny]{todonotes}
\usepackage{amsmath,amssymb}
\usepackage{fancyhdr}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue]{hyperref}
\pagestyle{fancy}
\lhead{Response to the Reviewers}
\rhead{30th October 2020}
\lfoot{\href{mailto:nishac@mit.edu}{Nisha Chandramoorthy} and \href{mailto:qiqi@mit.edu}{Qiqi Wang}}
\cfoot{}
\rfoot{\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

\definecolor{blue(pigment)}{rgb}{0.2, 0.2, 0.6}
\definecolor{burgundy}{rgb}{0.5, 0.0, 0.13}
\newcommand{\highlight}[1]{{\color{burgundy}\textbf{#1}}}

\title{}
\author{}
\date{30th October 2020}
\setlength{\marginparwidth}{2.1cm}
\begin{document}

\begin{center}
		\Large{\textbf{Manuscript NODY-D-20-01636R1: Response to Reviewers}}
\end{center}
\medskip
We would first like to thank both the reviewers for investing their time to read our work carefully and provide us with thoughtful feedback. We have substantially revised our work to incorporate all of both the reviewers' suggestions. In the following letter, we answer each of their comments and hope that both reviewers would find that the 
important issues they raised have been satisfactorily addressed by us.
Quotes from the Reviewers' comments are highlighted in 
\highlight{red} in order to distinguish them from our responses to the comments, which immediately follow the comments. 
\section{Response to Reviewer 1}
We are very grateful to you for asking thought-provoking
questions on extending our approach to higher-dimensional unstable manifolds 
and on the applications of the CLV self-derivative computations.

1. \highlight{Although equation (13) can be utilized to compute 
the self-derivative for all unstable CLV vectors, this has not
been demonstrated. How does the proposed method work for 
higher dimensional systems with at $d_u > 3$ dimensional unstable subspace?}

Indeed, as you point out, our examples involve computing only the first CLV self-derivative. However, our derivation of Eq. 13 is also valid for other unstable CLV self-derivatives; convergence to the self-derivative is guaranteed for the $i$th CLV self-derivative when $2\lambda_i > \lambda_1$. This criterion, which is obtained in Appendix C, suggests that the convergence of the differential CLV method is always guaranteed for first CLV self-derivative. We have rewritten section 3.4 to emphasize this point. Due to this unconditional convergence guarantee, we have focused, in this paper, on the unstable CLV self-derivative computation in the case of one-dimensional unstable manifolds.

Moreover, a convincing validation through the 
super-contracting Solenoid map was possible because there is an alternative analytical route to the computation of the first CLV's self-derivative; this is in turn because its norm is also the curvature of the attractor manifold. In the case of higher-dimensional attractors, the local unstable manifolds are not one-parameter curves, but rather, embedded images of $d_u$ dimensional Euclidean spaces. Thus, the local unstable manifolds can be endowed with a Riemannian metric, and several different notions of curvature that exist for higher-dimensional manifolds, including the Riemannian curvature tensor, can theoretically be defined for local unstable manifolds. However, to the best of our knowledge, it is not yet well-understood how the CLV self-derivatives are connected to these curvatures in higher dimensions.

We believe the differential CLV method provides a starting point for this more complicated geometric picture in higher dimensions, exploration of which we defer to a future work. With these considerations, we have added text 
to the introduction, section 3 and conclusion that clarifies our focus on one-dimensional unstable manifolds.

2. \highlight{The application of these derivatives to compute sensitivity of ergodic chaotic systems
is potentially quite significant. 
However, the authors do not present any demonstration of 
this important application and for the connection of derivations the readers
are referred to an unrefereed publication (Ref. [9]). Obviously, from application perspective, the high-dimensionality is again the real issue in computing sensitivities
in chaotic systems, which is raised in item (1) above.}

This is indeed a valid concern. However, the connection between the differential expansion equation (Eq. 17 in the revised manuscript) and linear response is fundamental, and independent of the S3 algorithm (Ref. 9, which is Ref. 12 in the revised manuscript). To clarify this point, we have rewritten section 5, which now provides a full explanation of this fundamental connection, directly from Ruelle's formula, without relying on the reference. The purpose of our paper is to show this link itself rather than full-fledged applications of linear response. In particular, the link we expose is between the fundamental quantity, the density gradient (denoted $g$), and the change in the expansion of tangent space-volumes, in the unstable manifold, which is captured by the scalar fields $\alpha_{x,i}.$

Because of this link, even when one uses a different basis -- for instance, forward/backward Lyapunov vectors -- as opposed to the CLV basis to represent unstable tangent vectors, the second-order tangent equations in the form of Eq. 17 (differential expansion equation) or Eq. 18 (differential CLV method) in the revised manuscript, still prove useful for linear response. Further, a different treatment of the unstable contribution to Ruelle's response, such as the Linear Response Algorithm (Ref. 26), also uses a second-order tangent equation of the form we present in this paper.

To the best of our knowledge, this algorithm and our S3 algorithm, which are both under active development, are the only 
practical ways for exact computation of Ruelle's formula in chaotic systems, and both use the recursive second-order tangent equations.

Another important reason for our exposition of this link (between $g$ and $\alpha_{\cdot,i}$) is that it could trigger applications beyond linear response, where the change in the SRB distribution on unstable manifolds is directly or indirectly needed, e.g., in variational data assimilation, Bayesian inference and inverse problems in chaotic systems.

We now describe this theoretical link in greater detail, independently of the computational aspects. Then, we present the major ideas in the S3 algorithm in order to present a complete picture of linear response computation, which is beyond the scope of this paper, and is part of our work under review (Ref. 12).

Focusing on the one-dimensional unstable manifold case, the solution to the differential expansion equation, $d/dt(1/z_{\cdot, 1})$, appears in the recursive computation of the logarithmic {\em density gradient} of the SRB measure on the unstable manifold, $(1/\rho) d(\rho\circ\mathcal{C}_{x,1})/dt.$ While Ref. 9 (Ref. 12 in the revised manuscript) provides a different derivation of this recursive computation (Eq. 33 in revised manuscript), it can also be derived from the fact that the SRB density is preserved in the following sense by $\varphi$ (alternatively, it is a simple eigenfunction of the transfer operator for eigenvalue 1):
\begin{align*}
		\rho(\varphi(x)) = \dfrac{\rho(x)}{z_{x,1}}. 
\end{align*}
Taking the logarithm of the above measure-preservation statement, and then differentiating with respect to the unstable coordinate $t$, we obtain Eq. 33 of the revised manuscript. 
Since the main component of linear response (Eq. 32, for which we have included the complete derivation in the Appendix section D) is the time-correlation between the density gradient and the objective function, the connection with the density gradient reinforces our observation that the differential expansion factor, $\alpha_{x,1}$, is a fundamental quantity for linear response.

The S3 algorithm, which is the primary contribution of the first author's ongoing PhD thesis work, has another crucial component besides the density gradient computation: a certain tangent space-splitting of a general perturbation. We did not consider a general differentiable perturbation field but rather a differentiable perturbation tangent to the unstable manifold, in the present paper (section 5). The splitting of a general perturbation, which makes a contribution to the overall sensitivity due to its stable components, and the computation of this stable contribution, are beyond the scope of this paper. This is because the CLV derivatives and the differential expansion factors, which are the subject of this paper, are directly linked to the sensitivity to unstable perturbations. For completion, here we briefly describe the main ideas of S3, including the stable contribution computation.

Given the setting in section 5 of the revised manuscript, we now consider a general perturbation $V.$ In the case of one-dimensional unstable manifolds, we find a vector field $a$, where $V = a V^1 + \chi$, such that the sensitivity to the component $\chi$, $\langle d(J \circ \varphi^n) \cdot \chi, \mu\rangle$, known as the {\em stable contribution}, is efficiently computable. The stable contribution can be computed using the following stable tangent equation:  
\begin{align}
		v_{n+1} = (d\varphi)_n\; v_n + \chi_{n+1},
\end{align}
where we have used the shorthand notation as in the manuscript. The solution to the stable tangent equation gives the stable contribution as an ergodic average,
\begin{align}
	\langle d(J\circ\varphi^n) \cdot \chi,\mu \rangle = 
		\dfrac{1}{N}\sum_{n=0}^{N-1} (dJ)_n \cdot v_n.
\end{align}
For the remaining unstable contribution -- $\langle d(J\circ\varphi^n) \cdot (aq) \rangle$ -- we use the method (regularization through integration by parts) described in section 5 of our revised manuscript. However, the scalar derivative $b := d(a\circ \mathcal{C}_{x,1})/dt$, which is required for the unstable contribution (in the second term of Eq. 32) is no longer given by the expression in the revised manuscript (in the paragraph below the definitions of $b$ and $g$). This is because $V$ now consists of the component $\chi$ in addition to $a\: V^1$. Thus, to determine $b$, we must also know the derivative of $\chi$ with respect to the unstable coordinate $t$.

A recursive algorithm for the derivative of $\chi$ in the unstable direction, including the proof of existence of this derivative, is part of our new work on linear response computation and beyond the scope of the present paper. The S3 algorithm thus comprises of the following ingredients:
\begin{itemize}
	\item In order to obtain the stable contribution, the stable tangent equation is solved so that $v_n \cdot q_n = 0$. The scalar field $a_n$ is obtained from this orthogonality.
		\item We obtain $g_n$ along a trajectory by solving Eq. 33. 
		\item We obtain $b_n$ along a trajectory by solving a second-order tangent equation for $d\chi_n/dt$.
\end{itemize}
Putting the stable and unstable contributions together, we obtain the following ergodic average for the linear response:
\begin{align}
	\langle J, \partial_s\mu_s\rangle = \lim_{N\to\infty}
		\dfrac{1}{N} \sum_{n=0}^{N-1}(dJ)_n\cdot v_n \; +\; \sum_{k=0}^\infty \lim_{N\to\infty}
		\dfrac{1}{N} \sum_{n=0}^{N-1}  J_{n+k} \; (a_n\; g_n + b_n )   .
\end{align}
In section 5, we have presented the complete picture for the computation of the unstable contribution, by considering an unstable perturbation. Thank you very much for encouraging us to improve this section!

\section*{Response to Reviewer 2}
We are very grateful to you for recommending edits that have substantially improved our manuscript.

\highlight{Rather than first mentioning CLVs, I suggest starting with sensitivity analysis (and predictions under
uncertainty), then explaining that computing CLVs (and their self-derivatives) are useful for these.}

We have fully restructured the introduction and rewritten a majority of it. In the revised manuscript's introduction, sensitivity analysis of statistics is first described along with applications. After this, we introduce the differential CLV method and discuss that it is needed for the computation of sensitivities of statistics or linear response. Thank you very much for this suggestion!

\highlight{Section 2, Line 6: I recommend not making $p_0 = p$ to make it clear it is the initial point along a trajectory. This
would allow you to use p for some generic point, not the initial point on a trajectory.}

Thank you for this suggestion!
In order to conform better with the standard notation in the literature, we have denoted phase points with the letter $x$ rather than $p$ in the revised manuscript. Since everywhere in the paper $x_0$ is a $\mu$-typical point, interchangeably using $x$ and $x_0$ is intentional. We mention prominently that the initial condition $x_0$, or simply $x$, is a generic phase point chosen $\mu$-a.e. on the attractor, at the start of the algorithm (section 3) and in the background section (section 2), where the $\mu$-typicality of the initial condition plays a role, e.g., in the definition of LEs. To achieve a $\mu$-typical point $x$, we evolve the dynamics for a sufficiently long spin-up time, upon starting from a Lebesgue-a.e. initial condition.

\highlight{Section 3.1, Equation 10: I think one or two intermediate equations to explain the coordinate changes mentioned
in the previous paragraph would be helpful here for improving understanding of this equation.}

We have added more explanation of the coordinate changes in the first paragraph of section 3.1 and the first two paragraphs of section 3.2.
The definition of the CLV self-derivative in Eq. 10 follows from the limit-based definition in Eq. 8, which states that the self-derivative is with respect to the $t$ coordinate function along the curve $\mathcal{C}_{x,i}$. These $d$ curves (at different indices $1\leq i\leq d$) represent our local coordinate system that is completely determined (up to a linear transformation) by the following constraints: i) $\mathcal{C}_{x,i}(0) = x$ and ii) $\mathcal{C}_{x,i}'(t) = V^i(\mathcal{C}_{x,i}(t)).$ In order to present more intuition about the coordinate system, we describe i) that its existence can also be explained due to existence and uniqueness theorems of ODEs (section 2.5) ii) properties of the connecting maps $f_{x,i}$ that can be proved from Pesin theory of unstable/stable manifolds (e.g. Ch. 6 of Ref. 19; section 2 of Ref. 22) in section 3.2.  
Thank you for suggesting an improvement in our explanation!

\highlight{Figure 1: Use a dotted or dashed line for one of the curves to explicitly show that they overlap.
Figure 2: See suggestion for figure 1}

Thank you very much for this tip! We have revised both these figures by using different markers for the analytical and numerically computed values such that their overlap is now visible.

\highlight{Figure 4: It's difficult to see some of the points, you might want to plot this with slightly larger points, especially
for the red points on the right.}

We have increased the size of the points and they are more visible now. Thank you very much for your feedback!

\highlight{Figure 5: These curves look orange and gray to me. I don't see any blue.}

Thank you very much for noticing! We have corrected the reference to ``blue'' to ``gray''.

\highlight{Section 4.4, line 16: The bottom left subfigure shows $[0,0,1,0]^T$, not $[0,1,0,0]^T$.}

Thanks for catching this error! We have made the correction.

\highlight{Section 4.4: What impact to perturbations to $s_3$ have? You've described the others but omitted $s_3$.}

Turning on $s_3$ produces compressions and rarefactions in the unstable manifold. This is because the perturbations we introduce are non-uniform in the unstable direction and are aligned with the unstable direction. We have now added the effects of all 4 parameters, while we had only explained those of $s_0$ and $s_2$ before, which served as examples for a small and significant curvature in the unstable manifold, respectively. Thank you for your recommendation!

\highlight{Section 4.5, figure 12: It's not easy to see the variation in color, please consider another colormap for this plot. If
you use a colormap with many colors, you might not need figure 13}

We have modified the colormap to a more visibly diverging colormap, which brings out the range of $\|W^1\|$ more clearly. Thank you for improving this figure! The curvature can be 8 orders of magnitude higher in the sharp (possibly non-differentiable) turns when compared to the straight portions, and a few orders of magnitude higher than the visibly curved portions of the attractor. In order to convey this clearly, we have retained Figure 13.

\highlight{Section 4.5, figure 13: You should consider plotting $V^1$ in gray or black behind the colored portions of the plot to
reveal the entire unstable manifold.}

Thank you for this excellent suggestion! We have added a thin black line that shows the entire unstable manifold. Since the colormap for the curvature contains a range of clearly distinguished light colors, which are also shown on thicker lines, the subfigures are easily legible and interpretable now. Thank you again!

\highlight{Section 5, line 10: Since you used "f" earlier, I would strongly suggest using another variable to indicate a scalar
observable here or change the earlier definition.}

Thank you very much for spotting this oversight! We have changed the scalar observable to $J$ in the linear response section (section 5).

\highlight{Section 5, equation 31: You should consider putting an abbreviated derivation in an appendix at least. Also, what
is $q_n$?}

We have rewritten this section to improve our explanation of the computation of linear response, and particularly where the differential expansion equation (Eq. 17 in the revised paper) appears. The equation in question is now Eq. 32 in the revised manuscript, which can be derived by integration by parts of Ruelle's original formula (Eq. 31). In order to obtain the regularized expression, we use the following sequence of steps: i) using the disintegration theorem on the SRB measure, ii) integration by parts on local unstable manifolds, and iii) expressing the resulting ensemble average as an ergodic average. These steps are briefly outlined in a new Appendix section (section D).
Thank you very much for suggesting this addition, which has led to a more complete presentation of linear response.

The type errors you point out were automatically fixed when we rewrote the section. Thank you for spotting them!
\end{document}

### 2021-04-20 letter/responseToReviewers.tex
Comments of reviewer 1

@@
 
-\definecolor{blue(pigment)}{rgb}{0.2, 0.2, 0.6}
+\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
 \definecolor{burgundy}{rgb}{0.5, 0.0, 0.13}
-\newcommand{\highlight}[1]{{\color{burgundy}\textbf{#1}}}
+\newcommand{\reviewerOne}[1]{{\color{burgundy}\textbf{#1}}}
+\newcommand{\reviewerTwo}[1]{{\color{azure}\textbf{#1}}}
 
@@
 \begin{center}
-		\Large{\textbf{Manuscript NODY-D-20-01636R1: Response to Reviewers}}
+		\Large{\textbf{Manuscript JCOMP-D-20-01958: Response to Reviewers}}
 \end{center}
@@
 important issues they raised have been satisfactorily addressed by us.
-Quotes from the Reviewers' comments are highlighted in 
-\highlight{red} in order to distinguish them from our responses to the comments, which immediately follow the comments. 
+Quotes from the Reviewer 1's comments are highlighted in \reviewerOne{red}, and those from Reviewer 2's are in \reviewerTwo{blue} in order to distinguish them from our responses to the comments, which immediately follow the comments.
+Consistently, in our revised manuscript, we highlight the changes inspired by Reviewer 1's and Reviewer 2's comments in \reviewerOne{red} and \reviewerTwo{blue} respectively.
 \section{Response to Reviewer 1}
@@
 
-1. \highlight{Although equation (13) can be utilized to compute 
-the self-derivative for all unstable CLV vectors, this has not
-been demonstrated. How does the proposed method work for 
-higher dimensional systems with at $d_u > 3$ dimensional unstable subspace?}
+1. \reviewerOne{There are some published articles that support some conclusions of this manuscript...All of these should give the authors more information about the topic related to this manuscript}
 
-Indeed, as you point out, our examples involve computing only the first CLV self-derivative.
-However, our derivation of Eq.
-13 is also valid for other unstable CLV self-derivatives; convergence to the self-derivative is guaranteed for the $i$th CLV self-derivative when $2\lambda_i > \lambda_1$.
-This criterion, which is obtained in Appendix C, suggests that the convergence of the differential CLV method is always guaranteed for first CLV self-derivative.
-We have rewritten section 3.4 to emphasize this point.
-Due to this unconditional convergence guarantee, we have focused, in this paper, on the unstable CLV self-derivative computation in the case of one-dimensional unstable manifolds.
+We thank the reviewer for pointing out multiple references that support the conclusions of the present paper.
+In our revised manuscript, we have added them to our discussion section (the edits are highlighted in \reviewerOne{red}), where we had originally referred to Clean Numerical Simulation (CNS).
 
-Moreover, a convincing validation through the 
-super-contracting Solenoid map was possible because there is an alternative analytical route to the computation of the first CLV's self-derivative; this is in turn because its norm is also the curvature of the attractor manifold.
-In the case of higher-dimensional attractors, the local unstable manifolds are not one-parameter curves, but rather, embedded images of $d_u$ dimensional Euclidean spaces.
-Thus, the local unstable manifolds can be endowed with a Riemannian metric, and several different notions of curvature that exist for higher-dimensional manifolds, including the Riemannian curvature tensor, can theoretically be defined for local unstable manifolds.
-However, to the best of our knowledge, it is not yet well-understood how the CLV self-derivatives are connected to these curvatures in higher dimensions.
+2.
+\reviewerOne{There is another thing that might be mentioned.
+Periodic solutions of some chaotic systems, such as three-body system, have physical meanings.
+Although three-body systems are chaotic in essence, periodic orbits of three-body system indeed exist.}
 
-We believe the differential CLV method provides a starting point for this more complicated geometric picture in higher dimensions, exploration of which we defer to a future work.
-With these considerations, we have added text
-to the introduction, section 3 and conclusion that clarifies our focus on one-dimensional unstable manifolds.
-
-2. \highlight{The application of these derivatives to compute sensitivity of ergodic chaotic systems
-is potentially quite significant. 
-However, the authors do not present any demonstration of 
-this important application and for the connection of derivations the readers
-are referred to an unrefereed publication (Ref. [9]). Obviously, from application perspective, the high-dimensionality is again the real issue in computing sensitivities
-in chaotic systems, which is raised in item (1) above.}
-
-This is indeed a valid concern.
-However, the connection between the differential expansion equation (Eq.
-17 in the revised manuscript) and linear response is fundamental, and independent of the S3 algorithm (Ref.
-9, which is Ref.
-12 in the revised manuscript).
-To clarify this point, we have rewritten section 5, which now provides a full explanation of this fundamental connection, directly from Ruelle's formula, without relying on the reference.
-The purpose of our paper is to show this link itself rather than full-fledged applications of linear response.
-In particular, the link we expose is between the fundamental quantity, the density gradient (denoted $g$), and the change in the expansion of tangent space-volumes, in the unstable manifold, which is captured by the scalar fields $\alpha_{x,i}.$
-
-Because of this link, even when one uses a different basis -- for instance, forward/backward Lyapunov vectors -- as opposed to the CLV basis to represent unstable tangent vectors, the second-order tangent equations in the form of Eq.
-17 (differential expansion equation) or Eq.
-18 (differential CLV method) in the revised manuscript, still prove useful for linear response.
-Further, a different treatment of the unstable contribution to Ruelle's response, such as the Linear Response Algorithm (Ref.
-26), also uses a second-order tangent equation of the form we present in this paper.
-
-To the best of our knowledge, this algorithm and our S3 algorithm, which are both under active development, are the only 
-practical ways for exact computation of Ruelle's formula in chaotic systems, and both use the recursive second-order tangent equations.
-
-Another important reason for our exposition of this link (between $g$ and $\alpha_{\cdot,i}$) is that it could trigger applications beyond linear response, where the change in the SRB distribution on unstable manifolds is directly or indirectly needed, e.g., in variational data assimilation, Bayesian inference and inverse problems in chaotic systems.
-
-We now describe this theoretical link in greater detail, independently of the computational aspects.
-Then, we present the major ideas in the S3 algorithm in order to present a complete picture of linear response computation, which is beyond the scope of this paper, and is part of our work under review (Ref.
-12).
-
-Focusing on the one-dimensional unstable manifold case, the solution to the differential expansion equation, $d/dt(1/z_{\cdot, 1})$, appears in the recursive computation of the logarithmic {\em density gradient} of the SRB measure on the unstable manifold, $(1/\rho) d(\rho\circ\mathcal{C}_{x,1})/dt.$ While Ref.
-9 (Ref.
-12 in the revised manuscript) provides a different derivation of this recursive computation (Eq.
-33 in revised manuscript), it can also be derived from the fact that the SRB density is preserved in the following sense by $\varphi$ (alternatively, it is a simple eigenfunction of the transfer operator for eigenvalue 1):
-\begin{align*}
-		\rho(\varphi(x)) = \dfrac{\rho(x)}{z_{x,1}}. 
-\end{align*}
-Taking the logarithm of the above measure-preservation statement, and then differentiating with respect to the unstable coordinate $t$, we obtain Eq. 33 of the revised manuscript. 
-Since the main component of linear response (Eq.
-32, for which we have included the complete derivation in the Appendix section D) is the time-correlation between the density gradient and the objective function, the connection with the density gradient reinforces our observation that the differential expansion factor, $\alpha_{x,1}$, is a fundamental quantity for linear response.
-
-The S3 algorithm, which is the primary contribution of the first author's ongoing PhD thesis work, has another crucial component besides the density gradient computation: a certain tangent space-splitting of a general perturbation.
-We did not consider a general differentiable perturbation field but rather a differentiable perturbation tangent to the unstable manifold, in the present paper (section 5).
-The splitting of a general perturbation, which makes a contribution to the overall sensitivity due to its stable components, and the computation of this stable contribution, are beyond the scope of this paper.
-This is because the CLV derivatives and the differential expansion factors, which are the subject of this paper, are directly linked to the sensitivity to unstable perturbations.
-For completion, here we briefly describe the main ideas of S3, including the stable contribution computation.
-
-Given the setting in section 5 of the revised manuscript, we now consider a general perturbation $V.$ In the case of one-dimensional unstable manifolds, we find a vector field $a$, where $V = a V^1 + \chi$, such that the sensitivity to the component $\chi$, $\langle d(J \circ \varphi^n) \cdot \chi, \mu\rangle$, known as the {\em stable contribution}, is efficiently computable.
-The stable contribution can be computed using the following stable tangent equation:
-\begin{align}
-		v_{n+1} = (d\varphi)_n\; v_n + \chi_{n+1},
-\end{align}
-where we have used the shorthand notation as in the manuscript. The solution to the stable tangent equation gives the stable contribution as an ergodic average,
-\begin{align}
-	\langle d(J\circ\varphi^n) \cdot \chi,\mu \rangle = 
-		\dfrac{1}{N}\sum_{n=0}^{N-1} (dJ)_n \cdot v_n.
-\end{align}
-For the remaining unstable contribution -- $\langle d(J\circ\varphi^n) \cdot (aq) \rangle$ -- we use the method (regularization through integration by parts) described in section 5 of our revised manuscript.
-However, the scalar derivative $b := d(a\circ \mathcal{C}_{x,1})/dt$, which is required for the unstable contribution (in the second term of Eq.
-32) is no longer given by the expression in the revised manuscript (in the paragraph below the definitions of $b$ and $g$).
-This is because $V$ now consists of the component $\chi$ in addition to $a\: V^1$.
-Thus, to determine $b$, we must also know the derivative of $\chi$ with respect to the unstable coordinate $t$.
-
-A recursive algorithm for the derivative of $\chi$ in the unstable direction, including the proof of existence of this derivative, is part of our new work on linear response computation and beyond the scope of the present paper.
-The S3 algorithm thus comprises of the following ingredients:
-\begin{itemize}
-	\item In order to obtain the stable contribution, the stable tangent equation is solved so that $v_n \cdot q_n = 0$. The scalar field $a_n$ is obtained from this orthogonality.
-		\item We obtain $g_n$ along a trajectory by solving Eq. 33. 
-		\item We obtain $b_n$ along a trajectory by solving a second-order tangent equation for $d\chi_n/dt$.
-\end{itemize}
-Putting the stable and unstable contributions together, we obtain the following ergodic average for the linear response:
-\begin{align}
-	\langle J, \partial_s\mu_s\rangle = \lim_{N\to\infty}
-		\dfrac{1}{N} \sum_{n=0}^{N-1}(dJ)_n\cdot v_n \; +\; \sum_{k=0}^\infty \lim_{N\to\infty}
-		\dfrac{1}{N} \sum_{n=0}^{N-1}  J_{n+k} \; (a_n\; g_n + b_n )   .
-\end{align}
-In section 5, we have presented the complete picture for the computation of the unstable contribution, by considering an unstable perturbation.
-Thank you very much for encouraging us to improve this section!
+Thank you for pointing this out. In section \ref{sec:nonphysical}, we have included the idea that periodic solutions may be atypical but not necessarily nonphysical.
 
@@
 
-\highlight{Rather than first mentioning CLVs, I suggest starting with sensitivity analysis (and predictions under
+\reviewerOne{Rather than first mentioning CLVs, I suggest starting with sensitivity analysis (and predictions under
 uncertainty), then explaining that computing CLVs (and their self-derivatives) are useful for these.}
@@
 
-\highlight{Section 2, Line 6: I recommend not making $p_0 = p$ to make it clear it is the initial point along a trajectory. This
+\reviewerOne{Section 2, Line 6: I recommend not making $p_0 = p$ to make it clear it is the initial point along a trajectory. This
 would allow you to use p for some generic point, not the initial point on a trajectory.}
@@
 
-\highlight{Section 3.1, Equation 10: I think one or two intermediate equations to explain the coordinate changes mentioned
+\reviewerOne{Section 3.1, Equation 10: I think one or two intermediate equations to explain the coordinate changes mentioned
 in the previous paragraph would be helpful here for improving understanding of this equation.}
@@
 
-\highlight{Figure 1: Use a dotted or dashed line for one of the curves to explicitly show that they overlap.
+\reviewerOne{Figure 1: Use a dotted or dashed line for one of the curves to explicitly show that they overlap.
 Figure 2: See suggestion for figure 1}
@@
 
-\highlight{Figure 4: It's difficult to see some of the points, you might want to plot this with slightly larger points, especially
+\reviewerOne{Figure 4: It's difficult to see some of the points, you might want to plot this with slightly larger points, especially
 for the red points on the right.}
@@
 
-\highlight{Figure 5: These curves look orange and gray to me. I don't see any blue.}
+\reviewerOne{Figure 5: These curves look orange and gray to me. I don't see any blue.}
 
@@
 
-\highlight{Section 4.4, line 16: The bottom left subfigure shows $[0,0,1,0]^T$, not $[0,1,0,0]^T$.}
+\reviewerOne{Section 4.4, line 16: The bottom left subfigure shows $[0,0,1,0]^T$, not $[0,1,0,0]^T$.}
 
@@
 
-\highlight{Section 4.4: What impact to perturbations to $s_3$ have? You've described the others but omitted $s_3$.}
+\reviewerOne{Section 4.4: What impact to perturbations to $s_3$ have? You've described the others but omitted $s_3$.}
 
@@
 
-\highlight{Section 4.5, figure 12: It's not easy to see the variation in color, please consider another colormap for this plot. If
+\reviewerOne{Section 4.5, figure 12: It's not easy to see the variation in color, please consider another colormap for this plot. If
 you use a colormap with many colors, you might not need figure 13}
@@
 
-\highlight{Section 4.5, figure 13: You should consider plotting $V^1$ in gray or black behind the colored portions of the plot to
+\reviewerOne{Section 4.5, figure 13: You should consider plotting $V^1$ in gray or black behind the colored portions of the plot to
 reveal the entire unstable manifold.}
@@
 
-\highlight{Section 5, line 10: Since you used "f" earlier, I would strongly suggest using another variable to indicate a scalar
+\reviewerOne{Section 5, line 10: Since you used "f" earlier, I would strongly suggest using another variable to indicate a scalar
 observable here or change the earlier definition.}
@@
 
-\highlight{Section 5, equation 31: You should consider putting an abbreviated derivation in an appendix at least. Also, what
+\reviewerOne{Section 5, equation 31: You should consider putting an abbreviated derivation in an appendix at least. Also, what
 is $q_n$?}

### 2021-04-21 letter/responseToReviewers.tex
Add reviewer 2's comments

@@
 
-\reviewerOne{Rather than first mentioning CLVs, I suggest starting with sensitivity analysis (and predictions under
-uncertainty), then explaining that computing CLVs (and their self-derivatives) are useful for these.}
+\reviewerTwo{My main technical objection is the claim that the findings here imply that all numerical solutions of chaotic systems might be "unphysical".
+...In other words, why must a numerical solution closely follow a specific realization of the exact chaotic system?
+}
 
@@
 
-\reviewerOne{Section 2, Line 6: I recommend not making $p_0 = p$ to make it clear it is the initial point along a trajectory. This
-would allow you to use p for some generic point, not the initial point on a trajectory.}
-
-Thank you for this suggestion!
-In order to conform better with the standard notation in the literature, we have denoted phase points with the letter $x$ rather than $p$ in the revised manuscript.
-Since everywhere in the paper $x_0$ is a $\mu$-typical point, interchangeably using $x$ and $x_0$ is intentional.
-We mention prominently that the initial condition $x_0$, or simply $x$, is a generic phase point chosen $\mu$-a.e.
-on the attractor, at the start of the algorithm (section 3) and in the background section (section 2), where the $\mu$-typicality of the initial condition plays a role, e.g., in the definition of LEs.
-To achieve a $\mu$-typical point $x$, we evolve the dynamics for a sufficiently long spin-up time, upon starting from a Lebesgue-a.e.
-initial condition.
-
-\reviewerOne{Section 3.1, Equation 10: I think one or two intermediate equations to explain the coordinate changes mentioned
-in the previous paragraph would be helpful here for improving understanding of this equation.}
-
-We have added more explanation of the coordinate changes in the first paragraph of section 3.1 and the first two paragraphs of section 3.2.
-The definition of the CLV self-derivative in Eq.
-10 follows from the limit-based definition in Eq.
-8, which states that the self-derivative is with respect to the $t$ coordinate function along the curve $\mathcal{C}_{x,i}$.
-These $d$ curves (at different indices $1\leq i\leq d$) represent our local coordinate system that is completely determined (up to a linear transformation) by the following constraints: i) $\mathcal{C}_{x,i}(0) = x$ and ii) $\mathcal{C}_{x,i}'(t) = V^i(\mathcal{C}_{x,i}(t)).$ In order to present more intuition about the coordinate system, we describe i) that its existence can also be explained due to existence and uniqueness theorems of ODEs (section 2.5) ii) properties of the connecting maps $f_{x,i}$ that can be proved from Pesin theory of unstable/stable manifolds (e.g.
-Ch.
-6 of Ref.
-19; section 2 of Ref.
-22) in section 3.2.
-Thank you for suggesting an improvement in our explanation!
-
-\reviewerOne{Figure 1: Use a dotted or dashed line for one of the curves to explicitly show that they overlap.
-Figure 2: See suggestion for figure 1}
-
-Thank you very much for this tip! We have revised both these figures by using different markers for the analytical and numerically computed values such that their overlap is now visible.
-
-\reviewerOne{Figure 4: It's difficult to see some of the points, you might want to plot this with slightly larger points, especially
-for the red points on the right.}
-
-We have increased the size of the points and they are more visible now. Thank you very much for your feedback!
-
-\reviewerOne{Figure 5: These curves look orange and gray to me. I don't see any blue.}
-
-Thank you very much for noticing! We have corrected the reference to ``blue'' to ``gray''.
-
-\reviewerOne{Section 4.4, line 16: The bottom left subfigure shows $[0,0,1,0]^T$, not $[0,1,0,0]^T$.}
-
-Thanks for catching this error! We have made the correction.
-
-\reviewerOne{Section 4.4: What impact to perturbations to $s_3$ have? You've described the others but omitted $s_3$.}
-
-Turning on $s_3$ produces compressions and rarefactions in the unstable manifold.
-This is because the perturbations we introduce are non-uniform in the unstable direction and are aligned with the unstable direction.
-We have now added the effects of all 4 parameters, while we had only explained those of $s_0$ and $s_2$ before, which served as examples for a small and significant curvature in the unstable manifold, respectively.
-Thank you for your recommendation!
-
-\reviewerOne{Section 4.5, figure 12: It's not easy to see the variation in color, please consider another colormap for this plot. If
-you use a colormap with many colors, you might not need figure 13}
-
-We have modified the colormap to a more visibly diverging colormap, which brings out the range of $\|W^1\|$ more clearly.
-Thank you for improving this figure!
-The curvature can be 8 orders of magnitude higher in the sharp (possibly non-differentiable) turns when compared to the straight portions, and a few orders of magnitude higher than the visibly curved portions of the attractor.
-In order to convey this clearly, we have retained Figure 13.
-
-\reviewerOne{Section 4.5, figure 13: You should consider plotting $V^1$ in gray or black behind the colored portions of the plot to
-reveal the entire unstable manifold.}
-
-Thank you for this excellent suggestion!
-We have added a thin black line that shows the entire unstable manifold.
-Since the colormap for the curvature contains a range of clearly distinguished light colors, which are also shown on thicker lines, the subfigures are easily legible and interpretable now.
-Thank you again!
-
-\reviewerOne{Section 5, line 10: Since you used "f" earlier, I would strongly suggest using another variable to indicate a scalar
-observable here or change the earlier definition.}
-
-Thank you very much for spotting this oversight! We have changed the scalar observable to $J$ in the linear response section (section 5).
-
-\reviewerOne{Section 5, equation 31: You should consider putting an abbreviated derivation in an appendix at least. Also, what
-is $q_n$?}
-
-We have rewritten this section to improve our explanation of the computation of linear response, and particularly where the differential expansion equation (Eq.
-17 in the revised paper) appears.
-The equation in question is now Eq.
-32 in the revised manuscript, which can be derived by integration by parts of Ruelle's original formula (Eq.
-31).
-In order to obtain the regularized expression, we use the following sequence of steps: i) using the disintegration theorem on the SRB measure, ii) integration by parts on local unstable manifolds, and iii) expressing the resulting ensemble average as an ergodic average.
-These steps are briefly outlined in a new Appendix section (section D).
-Thank you very much for suggesting this addition, which has led to a more complete presentation of linear response.
-
-The type errors you point out were automatically fixed when we rewrote the section. Thank you for spotting them!
 \end{document}

### 2021-04-21 letter/responseToReviewers.tex
Make the shadowing argument for numerical simulation

@@
 
-We have fully restructured the introduction and rewritten a majority of it.
-In the revised manuscript's introduction, sensitivity analysis of statistics is first described along with applications.
-After this, we introduce the differential CLV method and discuss that it is needed for the computation of sensitivities of statistics or linear response.
-Thank you very much for this suggestion!
+Thank you for your question!
+Traditionally, the justification for using numerical solutions to model chaotic governing physics is shadowing.
+While we do not contend that there may be alternative perspectives on numerical solutions, we question this traditional justification, based on the atypicality of shadowing trajectories that we show in the paper.
+To elaborate on this traditional argument for shadowing justifying the correctness of numerical solution, let us consider a chaotic ODE of the form
+\begin{align}
+    \dfrac{d x}{dt} = f(x, s),
+\end{align}
+where $x \in \mathbb{R}^d$ is a $d$-dimensional state vector, $s$ is a set of
+parameters in $\mathbb{R}^p$, and $f:\mathbb{R}^d \times \mathbb{R}^p \to \mathbb{R}^d$ is the generating flow. A numerical solution of the system may be represented as $\hat{x}(t)$ such that 
+\begin{align}
+    \dfrac{d\hat{x}}{dt} = f(\hat{x}, s) + \delta f(\hat{x},s), 
+\end{align}
+where $\delta f:\mathbb{R}^d \times \mathbb{R}^p \to \mathbb{R}^d$ is the model discrepancy introduced due to numerical discretization.
+A corollary of the shadowing lemma is that for every $\epsilon > 0,$ there exists  $\sup_{x} \|\delta f(x,s)\|$ is small,
 

### 2021-04-22 letter/responseToReviewers.tex
Finish reviewer 2's comments.

--- 2021-04-21 letter/responseToReviewers.tex
@@
 
-Thank you for pointing this out. In section \ref{sec:nonphysical}, we have included the idea that periodic solutions may be atypical but not necessarily nonphysical.
+Thank you for pointing this out. In section , we have included the idea that periodic solutions may be atypical but not necessarily nonphysical.
 
@@
 
-We have fully restructured the introduction and rewritten a majority of it.
-In the revised manuscript's introduction, sensitivity analysis of statistics is first described along with applications.
-After this, we introduce the differential CLV method and discuss that it is needed for the computation of sensitivities of statistics or linear response.
-Thank you very much for this suggestion!
+Thank you for your question!
+The traditional tenet used to justify using numerical solutions to model chaotic governing physics is shadowing.
+While we do not contend that there may be alternative perspectives on numerical solutions, we question this traditional tenet, based on the atypicality of shadowing trajectories that we show in the paper.
+To elaborate on this point, let us consider a chaotic ODE of the form
+\begin{align}
+\label{eqn:true}
+    \dfrac{d x}{dt} = f(x, s),
+\end{align}
+where $x \in \mathbb{R}^d$ is a $d$-dimensional state vector, $s$ is a set of
+parameters in $\mathbb{R}^p$, and $f:\mathbb{R}^d \times \mathbb{R}^p \to \mathbb{R}^d$ is the generating vector field.
+As you point out, the traditional view is that the numerical solution of the system is considered to be an exact solution of a slightly perturbed model.
+According to this view, a numerical solution to Eq.
+\ref{eqn:true} can be represented as $\hat{x}(t)$ such that
+\begin{align}
+\label{eqn:numerical}
+    \dfrac{d\hat{x}}{dt} = f(\hat{x}, s) + \delta f(\hat{x},s), 
+\end{align}
+where $\delta f:\mathbb{R}^d \times \mathbb{R}^p \to \mathbb{R}^d$ is the model discrepancy introduced due to numerical error.
+One may consider a different model for the numerical solution such as the flow of a vector field $f(\hat{x}, s + \delta s)$ with perturbations of the parameter.
+However, for small perturbations $\delta s$, performing a Taylor expansion of $f$ around the unperturbed $s$ leads to the model in Eq.
+\ref{eqn:numerical}, up to first order in $\|\delta s\|$.
+Thus, Eq.
+\ref{eqn:numerical} is a rather general model for a numerical solution of Eq.
+\ref{eqn:true}.
+
+For an $\epsilon >0$, an $\epsilon$-orbit or a {\em pseudo-orbit} of the original system (Eq.
+\ref{eqn:true}) is any solution $\hat{x}(t)$ of Eq.
+\ref{eqn:numerical} wherein $\|\delta f(\hat{x}(t), s)\| \leq \epsilon$ for all $t$.
+Any reasonably accurate numerical solution can be considered to be an $\epsilon$-orbit of the ground truth dynamics in Eq.
+\ref{eqn:true}.
+
+The shadowing lemma says that for every $\delta > 0,$ there exists $\epsilon > 0$ such that for every $\epsilon$-orbit $\hat{x}(t)$ that satisfies Eq.
+\ref{eqn:numerical}, there exists a true orbit $x(t)$ satisfying Eq.
+\ref{eqn:true} that $\delta$-shadows it.
+This means that there exists some scalar function $\gamma(t)$ with $|\gamma'(t) - 1| < \delta$ such that $\| \hat{x}(\gamma(t)) - x(t) \| < \delta$ for all $t.$
+
+Now, since $x(t)$ and $\hat{x}(t)$ are $\delta$-close to each other, their long-term (empirical) probability distributions are also similar.
+That is, for any regular scalar function $J,$ the time-averages $\langle J\rangle_x^T := (1/T)\int_0^T J(x(t))\: dt$ and
+$\langle J\rangle_{\hat{x}}^T := (1/T) \int_0^T J(\hat{x}(t)) \: dt$ are expected to be close to each other for large $T.$
+
+However, we show through examples in the paper that $x(t)$, a shadowing orbit, which is a true solution of Eq.
+\ref{eqn:true}, need not be a typical or a {\em physical} orbit.
+That is, the infinitely-long time average along $x(t)$, $\lim_{T\to\infty} \langle J\rangle_x^T$, need not be equal to the average of $J$ over an ensemble of solutions to Eq.
+\ref{eqn:true}.
+In fact, we show that the discrepancy can be large, when compared to the true value of the ensemble average.
+Note that the ensemble average over all possible solutions of Eq.
+\ref{eqn:true} is observed along any randomly chosen orbit, i.e., if we chose an initial condition $x(0)$ by sampling a uniform distribution in the attractor basin of the dynamics in Eq.
+\ref{eqn:true}, then $\lim_{T\to\infty}\langle J\rangle_x^T$ is equal to the ensemble average of $J.$ However, when $x(0)$ is a special initial condition that is chosen so that $x(t)$ $\delta$-shadows a pseudo-orbit $\hat{x}(t),$ it is not guaranteed to reproduce ensemble statistics.
+
+Since the statistics or long-term averages of pseudo-orbits and shadowing orbits are close, we expect that the pseudoorbits (the numerical solution $\hat{x}(t)$) is also statistically inconsistent.
+
+As you rightly pointed out, in some scenarios, the pseudo-orbits may in fact inhabit the same or almost the same attractor.
+In other words, the attractor, or the set that all orbits of Eq.
+\ref{eqn:true} converge to asymptotically, may be identical or very close to that of Eq.
+\ref{eqn:numerical}.
+However, the probability distribution of an ensemble of orbits on the attractor, which is proportional to the amount of time almost every orbit spends in each portion of the attractor, are not guaranteed to be the same for the two models.
+This is the main argument in the paper.
+
+We hope that this clarifies our position: we refute the belief that, due to shadowing, numerical solutions, which are considered to be pseudo-orbits that are close to true orbits, are also capable of modeling the true long-term behavior of a chaotic system.
+We support our claim by proving that shadowing orbits, despite being true orbits, are not representative of almost every orbit of the system: they can have atypical statistical behavior.
+
+Thank you again for raising this question! We have added more explanation (in \reviewerTwo{blue}) to clarify this point in section 5.
 

### 2021-04-22 letter/responseToReviewers.tex
atypical nonphysical

@@
 
-Thank you for pointing this out. In section , we have included the idea that periodic solutions may be atypical but not necessarily nonphysical.
+Thank you for pointing this out.
+In our paper, we use ``nonphysicality'' to mean ``atypicality'' in the following sense.
+In section 3.2, we have included this idea that periodic solutions may be atypical but not necessarily nonphysical.
 

### 2021-04-22 letter/responseToReviewers.tex
Add reviewer 2 last comment
Add refs for UPOs

@@
 We would first like to thank both the reviewers for investing their time to read our work carefully and provide us with thoughtful feedback.
-We have substantially revised our work to incorporate all of both the reviewers' suggestions.
+We have revised our work to incorporate all of both the reviewers' suggestions.
 In the following letter, we answer each of their comments and hope that both reviewers would find that the
@@
 \section{Response to Reviewer 1}
-We are very grateful to you for asking thought-provoking
-questions on extending our approach to higher-dimensional unstable manifolds 
-and on the applications of the CLV self-derivative computations.
+We are very grateful to you for making our literature review more complete and for strengthening our results.
 
@@
 We thank the reviewer for pointing out multiple references that support the conclusions of the present paper.
-In our revised manuscript, we have added them to our discussion section (the edits are highlighted in \reviewerOne{red}), where we had originally referred to Clean Numerical Simulation (CNS).
+In our revised manuscript, we have added them to our discussion section 5 (the edits are highlighted in \reviewerOne{red}), where we had originally referred to Clean Numerical Simulation (CNS).
 
@@
 Thank you for pointing this out.
-In our paper, we use ``nonphysicality'' to mean ``atypicality'' in the following sense.
+In our paper, we use ``nonphysicality'' to mean ``atypicality'' in the following sense, which, as you point out, may not be physically meaningless.
+Whenever an initial condition is sampled from a set of full Lebesgue measure in an open set, known as the attractor basin, that contains the chaotic attractor, the ergodic average along the orbit starting from that initial condition is equal to the ensemble average.
+This latter ensemble average or phase-space average is with respect to an SRB-type probability measure, which is called a {\em physical} measure \cite{young} since it is ``observed'' through ergodic averages along almost every orbit starting in the attractor basin.
+Thus, we refer to as ``nonphysical'' atypical orbits that do not produce the correct statistics associated with the SRB measure.
+
 In section 3.2, we have included this idea that periodic solutions may be atypical but not necessarily nonphysical.
@@
 Thank you again for raising this question! We have added more explanation (in \reviewerTwo{blue}) to clarify this point in section 5.
+\vspace{0.2in}
+
+\reviewerTwo{The only very minor exception is that the figure captions use "L" and "R" for "left" and "right", which took me some time to understand. I would prefer these words to be written out.
+}
+
+We have abandoned ``L''s and ``R''s and written out ``left'' and ``right'' in their place respectively. Thank you for eliminating this source of confusion!
 

### 2021-04-22 main_els.tex
Add refs for UPOs

@@
 \newcommand{\nisha}[1]{{\color{darkgreen} #1}}
+\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
+\definecolor{burgundy}{rgb}{0.5, 0.0, 0.13}
+\newcommand{\reviewerOne}[1]{{\color{burgundy}\textbf{#1}}}
+\newcommand{\reviewerTwo}[1]{{\color{azure}\textbf{#1}}}
 \DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
@@
 \caption{
-L: An ensemble of initial conditions distributed uniformly in the box $u_1\in[0,1], u_2\in[0,1], u_3\in[28,39]$ 
+Left: An ensemble of initial conditions distributed uniformly in the box $u_1\in[0,1], u_2\in[0,1], u_3\in[28,39]$ 
 after 10 time units of evolution, shown on the $u_1-u_3$ plane.
-R: distribution of a trajectory of $1000$ time units in length.
+Right: distribution of a trajectory of $1000$ time units in length.
 In both plots, the color represents the number of samples in a
@@
 \caption{
-L: distribution of the same ensemble as in Figure \ref{fig:lorenz_ergodicity1}
+Left: distribution of the same ensemble as in Figure \ref{fig:lorenz_ergodicity1}
 after another 5 time units of evolution.
-R: distribution of the same trajectory as in Figure \ref{fig:lorenz_ergodicity1}
+Right: distribution of the same trajectory as in Figure \ref{fig:lorenz_ergodicity1}
 evolved for $10,000$ time units.
@@
 \caption{
-L: distribution of the same ensemble as in Figures \ref{fig:lorenz_ergodicity1}
+Left: distribution of the same ensemble as in Figures \ref{fig:lorenz_ergodicity1}
 and \ref{fig:lorenz_ergodicity2} after a total of 50 time units of evolution.
-R: distribution of the same trajectory as in Figures \ref{fig:lorenz_ergodicity1}
+Right: distribution of the same trajectory as in Figures \ref{fig:lorenz_ergodicity1}
 and \ref{fig:lorenz_ergodicity2} evolved for $100,000$ time units.
@@
 of the phase space as a physical solution does.
+\reviewerOne{Although we refer to them as nonphysical, periodic solutions, may not, in a literal sense be physically meaningless.
+Rather, by nonphysicality, we mean atypicality, in this paper.
+Atypical solutions, such as periodic solutions, which do not produce the same statistics as a chaotic solution, are almost never physically observed.
+This is because, an initial condition randomly generated from the attractor basin, according to any absolutely continuous probability distribution, produces a chaotic orbit, with probability 1.
+The long-term statistics along such an orbit are the same as phase-space averages according to the SRB measure, i.e., such orbits are physical.
+In other words, when randomly sampling initial conditions, finding a periodic solution is a zero probability event.
+This lack of physical observability of periodic solutions is what has lent them a classification under nonphysical solutions, although periodic solutions may have some problem-specific, meaningful physical interpretations.}
+
+\reviewerOne{
+At the same time, the set of all periodic solutions is a dense subset of the attractor. That is,}
+
 Figure \ref{fig:lorenz_periodic} shows
@@
 \includegraphics[width=0.48\textwidth]{lorenz_zmax2_10_28_2.6666666666666665.png}
-\caption{L: the Lorenz map.  The x-axis is the $n$-th local maximum of $u_3(t)$ over a
+\caption{Left: the Lorenz map.  The x-axis is the $n$-th local maximum of $u_3(t)$ over a
 long solution; the y-axis is the $(n+1)$-th local maximum of $u_3(t)$.  The intersection of this curve
@@
 in Figure \ref{fig:lorenz_periodic}.
-R: the Lorenz map iterated twice.  The x-axis is the $n$-th local maximum
+Right: the Lorenz map iterated twice.  The x-axis is the $n$-th local maximum
 of $u_3(t)$; the y-axis is the $(n+2)$-th maximum.  The intersections with the diagonal dashed
@@
     \includegraphics[width=0.48\textwidth]{tent_ivp.png}
-    \caption{L: The scaled tent map $\hat{\varphi}_s$ at different values of $s$ between 0 and 1. R: sensitivity to small perturbation in the governing
+    \caption{Left: The scaled tent map $\hat{\varphi}_s$ at different values of $s$ between 0 and 1. Right: sensitivity to small perturbation in the governing
     equation.
@@
     \\ \vspace{0.02\textwidth}
-     \caption{L: the tilted tent map Eq. \ref{tent_tilted} for $s=0, 0.1, 0.2, 0.3, 0.4,$ and 0.5. R: the conjugacy $\tilde{h}_s$ between the tent map (Eq. \ref{tentmap})
+     \caption{Left: the tilted tent map Eq. \ref{tent_tilted} for $s=0, 0.1, 0.2, 0.3, 0.4,$ and 0.5. Right: the conjugacy $\tilde{h}_s$ between the tent map (Eq. \ref{tentmap})
     and the tilted tent map (Eq. \ref{tent_tilted}), evaluated using
@@
    \caption{
-     L: the density of a trajectory satisfying Eq. \ref{tentmap}
+     Left: the density of a trajectory satisfying Eq. \ref{tentmap}
     that shadows a random trajectory satisfying Eq. \ref{tent_tilted} for $s=0.1$.
-    R: the density of a trajectory satisfying Eq. \ref{tentmap}
+    Right: the density of a trajectory satisfying Eq. \ref{tentmap}
     that shadows a random trajectory satisfying Eq. \ref{tent_tilted} for $s=0.5$.
@@
 		\includegraphics[width=0.45\textwidth]{lorenz_map_fit_error_lorenz_zmax_10_28_2.6666666666666665.npy.png}
-		\caption{L: the blue and orange lines are the Lorenz map at the standard parameters, and the dotted black lines indicate the approximate Lorenz map (Table \ref{tab:lorenzFit}).
-R: the regression error is shown as a function of $z$}
+		\caption{Left: the blue and orange lines are the Lorenz map at the standard parameters, and the dotted black lines indicate the approximate Lorenz map (Table \ref{tab:lorenzFit}).
+Right: the regression error is shown as a function of $z$}
 	\label{fig:lorenzMapComp}

### 2021-04-23 letter/responseToReviewers.tex
Add all of Liao's refs.

@@
 
-Thank you for pointing this out.
+Thank you for pointing this important point out.
 In our paper, we use ``nonphysicality'' to mean ``atypicality'' in the following sense, which, as you point out, may not be physically meaningless.
 Whenever an initial condition is sampled from a set of full Lebesgue measure in an open set, known as the attractor basin, that contains the chaotic attractor, the ergodic average along the orbit starting from that initial condition is equal to the ensemble average.
-This latter ensemble average or phase-space average is with respect to an SRB-type probability measure, which is called a {\em physical} measure \cite{young} since it is ``observed'' through ergodic averages along almost every orbit starting in the attractor basin.
+This latter ensemble average or phase-space average is with respect to an SRB-type probability measure, which is called a {\em physical} measure (Ref.
+10) since it is ``observed'' through ergodic averages along almost every orbit starting in the attractor basin.
 Thus, we refer to as ``nonphysical'' atypical orbits that do not produce the correct statistics associated with the SRB measure.
 
-In section 3.2, we have included this idea that periodic solutions may be atypical but not necessarily nonphysical.
+Nevertheless, the collection of unstable periodic orbits is intimately connected with the statistical properties of the strange attractor.
+Ensemble averages can be realized as weighted ergodic averages along long unstable periodic orbits (e.g.
+see Property 3.10 of \cite{llave}).
+Dynamical determinants, which are calculated on the set of periodic orbits, are connected with spectral resonances or the slowest correlation decay rates \cite{baladi-book}.
+Apart from these theoretical connections, the set of unstable periodic orbits is also recognized as a physically interpretable ``skeleton'' of an attractor \cite{cvitanovic1}.
+Besides providing alternative methods to calculate dynamical properties, the tangencies on unstable periodic orbits are thought to play an important role in attractor crises: a large statistical response to small system parameter perturbations \cite{berry}\cite{grebogi1}.
+
+In section 3.2, we have expanded on these ideas without going into mathematical details of unstable periodic orbits, which would be beyond the scope of this paper.
+Thank you very much once again for this suggestion!
 
@@
 We have abandoned ``L''s and ``R''s and written out ``left'' and ``right'' in their place respectively. Thank you for eliminating this source of confusion! 
-
+\bibliographystyle{plain}
+\bibliography{refs.bib}
 \end{document}

### 2021-04-23 main_els.tex
Add all of Liao's refs.

@@
 of the original governing equation? Sauer \cite{sauer} has shown 
-examples in which numerical error due to computation in double-precision floating point arithmetic causes significant change in the stationary probability distribution on the attractor; other works \cite{cns} further analyze the unreliability due to double-precision arithmetic of statistical measures such as power spectra and correlation functions.
-The plucked tent map adds to this list of examples by demonstrating a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).
+examples in which numerical error due to computation in double-precision floating point arithmetic causes significant change in the stationary probability distribution on the attractor.
+\reviewerOne{Some authors \cite{cns2}\cite{cns3} resort to Clean Numerical Simulations to tackle scenarios that demand highly accurate long-term prediction of chaotic orbits.
+When using double-precision arithmetic however, these works  \cite{cns}\cite{cns1}\cite{cns2}\cite{cns3} show that large errors result in statistical measures such as power spectra and spatio-temporal correlation functions.} The plucked tent map adds to this list of examples by demonstrating a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).
+
 Uniformly hyperbolic 

### 2021-04-23 letter/summaryOfRevisions.tex
Add summary of revisions.

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1.1in]{geometry}
\usepackage{graphicx}
\usepackage[numbers]{natbib}
\usepackage{verbatim}
\usepackage[textsize=tiny]{todonotes}
\usepackage{amsmath,amssymb}
\usepackage{fancyhdr}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue]{hyperref}
\pagestyle{fancy}
\lhead{Response to the Reviewers}
\rhead{30th October 2020}
\lfoot{\href{mailto:nishac@mit.edu}{Nisha Chandramoorthy} and \href{mailto:qiqi@mit.edu}{Qiqi Wang}}
\cfoot{}
\rfoot{\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
\definecolor{burgundy}{rgb}{0.5, 0.0, 0.13}
\newcommand{\reviewerOne}[1]{{\color{burgundy}\textbf{#1}}}
\newcommand{\reviewerTwo}[1]{{\color{azure}\textbf{#1}}}

\title{}
\author{}
\date{23rd April 2021}
\setlength{\marginparwidth}{2.1cm}
\begin{document}

\begin{center}
		\Large{\textbf{Manuscript JCOMP-D-20-01958: Response to Reviewers}}
\end{center}
\medskip
We would first like to thank both the reviewers for investing their time to read our work carefully and provide us with thoughtful feedback. We have revised our work to incorporate all of both the reviewers' suggestions. Here is a summary of the changes we have made:
\begin{itemize}
    \item We have added citations to works pointed out to us by Reviewer 1 that strengthen the conclusions of this paper, especially on the implications of our results for the statistical inaccuracy of numerical solutions to chaotic systems. A brief explanation of the cited works is added to section 5, and is highlighted in \reviewerOne{red}.
    \item We have added a couple of paragraphs in section 3.2 (in \reviewerOne{blue}) to explain the sense in which we refer to periodic solutions as nonphysical. This is in response to a comment by Reviewer 1 on periodic solutions having physical interpretations.
    \item 
\end{itemize}
\end{document}

### 2021-04-23 main_els.tex
Add summary of revisions.

@@
 examples in which numerical error due to computation in double-precision floating point arithmetic causes significant change in the stationary probability distribution on the attractor.
-\reviewerOne{Some authors \cite{cns2}\cite{cns3} resort to Clean Numerical Simulations to tackle scenarios that demand highly accurate long-term prediction of chaotic orbits.
-When using double-precision arithmetic however, these works  \cite{cns}\cite{cns1}\cite{cns2}\cite{cns3} show that large errors result in statistical measures such as power spectra and spatio-temporal correlation functions.} The plucked tent map adds to this list of examples by demonstrating a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).
+\reviewerOne{Some authors} \cite{cns2}\cite{cns3} \reviewerOne{resort to Clean Numerical Simulations to tackle scenarios that demand highly accurate long-term prediction of chaotic orbits.
+When using double-precision arithmetic however, these works}\cite{cns}\cite{cns1}\cite{cns2}\cite{cns3}\reviewerOne{show that large errors result in statistical measures such as power spectra and spatio-temporal correlation functions.} The plucked tent map adds to this list of examples by demonstrating a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).
+
+\reviewerTwo{Here, we raise questions on the statistical consistency on numerical simulations of chaotic systems because they are traditionally interpreted as being shadowed by a true solution.
+}\reviewerTwo{Since shadowing solutions can be nonphysical, so can numerical solutions that lie close to them at all times, and hence have a long-term statistical distribution close to their shadowing counterparts.} \reviewerTwo{This argument holds even when numerical solutions are interpreted as perturbed solutions to the true dynamics that produce the same or a nearby attractor.}\reviewerTwo{The probability distribution on the attractor, measured by the amount of time a solution spends in each location on the attractor, can be drastically different for a numerical solution than a typical solution of the dynamics.}
 

### 2021-04-23 letter/responseToReviewers.tex
Add summary of revisions.
Add few more refs and complete writing on periodic orbits

@@
 \lhead{Response to the Reviewers}
-\rhead{30th October 2020}
+\rhead{23rd April 2021}
 \lfoot{\href{mailto:nishac@mit.edu}{Nisha Chandramoorthy} and \href{mailto:qiqi@mit.edu}{Qiqi Wang}}
@@
 
-Nevertheless, the collection of unstable periodic orbits is intimately connected with the statistical properties of the strange attractor.
+Nevertheless, as you point out, in some problems, such as the three-body problem, families of periodic orbits have been observed and studied in their own right \cite{henon3body}.
+Furthermore, the collection of unstable periodic orbits is intimately connected with the statistical properties of the strange attractor.
 Ensemble averages can be realized as weighted ergodic averages along long unstable periodic orbits (e.g.

### 2021-04-23 letter/summaryOfRevisions.tex
Add few more refs and complete writing on periodic orbits

@@
 \begin{center}
-		\Large{\textbf{Manuscript JCOMP-D-20-01958: Response to Reviewers}}
+		\Large{\textbf{Manuscript JCOMP-D-20-01958: Summary of changes}}
 \end{center}
@@
 This is in response to a comment by Reviewer 1 on periodic solutions having physical interpretations.
-    \item 
+    \item We have provided a detailed answer to Reviewer 2's question in the response to Reviewers on the interpretation of numerical solutions to chaotic systems.
+We have included some of this clarification also in section 5 of the revised manuscript in \reviewerTwo{red}.
+    \item We have changed the ``L''s and ``R''s in the figure captions to ``left''s and ``right''s respectively.
 \end{itemize}

### 2021-04-23 main_els.tex
Add few more refs and complete writing on periodic orbits

@@
 \reviewerOne{
-At the same time, the set of all periodic solutions is a dense subset of the attractor. That is,}
-
+At the same time, families of periodic solutions have been well-studied for some problems}\cite{henon3body}.\reviewerOne{The set of all unstable periodic orbits have been recognized as the ``skeleton'' of the chaotic attractor \cite{cvitanovic1} forming a dense subset of the attractor and providing an alternative characterization of the SRB distribution \cite{llave}.
+This implies that ergodic/ensemble averages with respect to the SRB distribution are also weighted averages along unstable periodic orbits.} \reviewerOne{The significance of the set of unstable periodic orbits extends even further as it can be used to compute dynamical invariants such as entropy and Lyapunov dimension \cite{grebogi1}\cite{cvitanovic1}.
+A quantity of great mathematical interest is the dynamical determinant (see \cite{baladi-book} for a recent review), which is computed as a weighted sum over all periodic orbits.
+The zeros of the dynamical determinant have a one-to-one correspondence with the slowest decay rates of correlations between observables \cite{ruelleDeterminant}.
+This connection has been exploited both in mathematical as well as physical studies of the distribution of energy levels and resonances in quantum scattering \cite{semyonZworksi}\cite{berry}\cite{cvitanovic2005chaos}.
+Furthermore, theoretical and numerical studies have attributed cataclysmic changes in the shape and position of the attractor in response to small parameter changes, known as attractor crises \cite{grebogi1}, to the emergence of tangencies between the stable and unstable manifolds along periodic orbits.}
 Figure \ref{fig:lorenz_periodic} shows

### 2021-04-23 main_els.tex
Add refs for attractor crises

--- 2021-04-23 main_els.tex
@@
 \reviewerOne{
-At the same time, the set of all periodic solutions is a dense subset of the attractor. That is,}
+At the same time, families of periodic solutions have been well-studied for some problems}\cite{henon3body}.\reviewerOne{The set of all unstable periodic orbits have been recognized as the ``skeleton'' of the chaotic attractor \cite{cvitanovic1} forming a dense subset of the attractor and providing an alternative characterization of the SRB distribution \cite{llave}.
+This implies that ergodic/ensemble averages with respect to the SRB distribution are also weighted averages along the set of periodic orbits.} \reviewerOne{The significance of the set of unstable periodic orbits extends even further as it can be used to compute dynamical invariants such as entropy and Lyapunov dimension \cite{grebogi1}\cite{cvitanovic1}.
+A quantity of great mathematical interest is the dynamical determinant (see \cite{baladi-book} for a recent review), which is computed as a weighted sum over all periodic orbits.
+The zeros of the dynamical determinant have a one-to-one correspondence with the slowest decay rates of correlations between observables \cite{baladi-book}.
+This connection has been exploited both in mathematical as well as physical studies of the distribution of energy levels and resonances in quantum scattering \cite{semyonZworski}\cite{berry}\cite{cvitanovic2005chaos}.
+Furthermore, theoretical and numerical studies have attributed cataclysmic changes in the shape and position of the attractor in response to small parameter changes, known as attractor crises \cite{grebogi1}, to the emergence of tangencies between the stable and unstable manifolds along periodic orbits.}
 

### 2021-04-24 letter/responseToReviewers.tex
Current changes

@@
 Thank you for pointing this important point out.
-In our paper, we use ``nonphysicality'' to mean ``atypicality'' in the following sense, which, as you point out, may not be physically meaningless.
+In our paper, we use ``nonphysicality'' to mean ``atypicality'' in the following sense, which, as you point out, may not be the same as physically meaningless.
 Whenever an initial condition is sampled from a set of full Lebesgue measure in an open set, known as the attractor basin, that contains the chaotic attractor, the ergodic average along the orbit starting from that initial condition is equal to the ensemble average.
@@
 
-The shadowing lemma says that for every $\delta > 0,$ there exists $\epsilon > 0$ such that for every $\epsilon$-orbit $\hat{x}(t)$ that satisfies Eq.
-\ref{eqn:numerical}, there exists a true orbit $x(t)$ satisfying Eq.
+This also holds even when a numerical solution may not be modelled by Eq.
+\ref{eqn:numerical} because $\epsilon$-orbits could be defined more generally without recourse to the particular form of Eq.
+\ref{eqn:numerical}.
+Any curve $\hat{x}(t)$ such that $\|\hat{x}'(t) - x'(\hat{x}(t))\| < \epsilon$ for all $t$ is an $\epsilon$-orbit of Eq.
+\ref{eqn:true}.
+
+The shadowing lemma says that for every $\delta > 0,$ there exists $\epsilon > 0$ such that for every $\epsilon$-orbit $\hat{x}(t)$, there exists a true orbit $x(t)$ satisfying Eq.
 \ref{eqn:true} that $\delta$-shadows it.
@@
 That is, for any regular scalar function $J,$ the time-averages $\langle J\rangle_x^T := (1/T)\int_0^T J(x(t))\: dt$ and
-$\langle J\rangle_{\hat{x}}^T := (1/T) \int_0^T J(\hat{x}(t)) \: dt$ are expected to be close to each other for large $T.$
+$\langle J\rangle_{\hat{x}}^T := (1/T) \int_0^T J(\hat{x}(t)) \: dt$ are close to each other for large $T.$
 
@@
 \ref{eqn:true}.
-In fact, we show that the discrepancy can be large, when compared to the true value of the ensemble average.
+In fact, we show in section 5 with the plucked tent map that the discrepancy can be large, when compared to the true value of the ensemble average.
 Note that the ensemble average over all possible solutions of Eq.
@@
 
-Since the statistics or long-term averages of pseudo-orbits and shadowing orbits are close, we expect that the pseudoorbits (the numerical solution $\hat{x}(t)$) is also statistically inconsistent.
+Since the statistics or long-term averages of pseudo-orbits and shadowing orbits are close, we expect that the pseudoorbit (the numerical solution $\hat{x}(t)$) is also not guaranteed to be statistically consistent.
 

### 2021-04-24 main_els.tex
Current changes

@@
 This connection has been exploited both in mathematical as well as physical studies of the distribution of energy levels and resonances in quantum scattering \cite{semyonZworski}\cite{berry}\cite{cvitanovic2005chaos}.
-Furthermore, theoretical and numerical studies have attributed cataclysmic changes in the shape and position of the attractor in response to small parameter changes, known as attractor crises \cite{grebogi1}, to the emergence of tangencies between the stable and unstable manifolds along periodic orbits.}
+Furthermore, theoretical and numerical studies of unstable periodic orbits have been crucial to better understanding transitions to chaos \cite{kreilos} and cataclysmic changes in the shape and position of the attractor in response to small parameter changes, known as attractor crises \cite{grebogi1}\cite{lucariniCrisis}, which are sometimes due to the emergence of tangencies between the stable and unstable manifolds along periodic orbits.}
 

### 2021-04-24 letter/responseToReviewers.tex
Edit last part of response

@@
 
-Thank you again for raising this question! We have added more explanation (in \reviewerTwo{blue}) to clarify this point in section 5.
+Thank you again for raising this question!
+As you point out, we have only proved that numerical solutions are not guaranteed to produce the true long-term/statistical behavior, and not that numerical solutions are always statistically inconsistent (after all, some shadowing trajectories are physical -- e.g.
+section 4.1).
+However, several references strengthen the observation of statistical inconsistency of numerics.
+The new references sent by Reviewer 1 that introduce clean numerical simulation, which are now cited in the revised manuscript, and the previously cited references (Sauer et al, which takes an analytical approach) in section 5 all observe that numerical solutions produce statistical/long-term inconsistencies.
+We have added more explanation (in \reviewerTwo{blue}) to clarify this point in section 5.
 \vspace{0.2in}

### 2021-05-06 highlights.tex
highlights modify and change supp to appendix

@@
 \begin{itemize}
-\item This paper highlights through rigorous examples that numerical solutions of chaotic dynamical systems are not guaranteed to produce the correct statistics or long-term behavior;
-\item We show through examples that it is theoretically possible to significantly alter the statistics of a chaotic system by minor parameter perturbations.
-\item The nonphysicality of shadowing orbits also implies that methods to compute parametric derivatives of statistics in chaotic systems, based on shadowing, may produce incorrect results.
+\item Shadowing orbits may not be typical orbits or physical solutions
+\item This atypicality implies shadowing-based sensitivity can be biased.
+\item Numerical solutions of chaotic dynamical systems can be statistically inconsistent;
+\item Minor parameter changes can significantly alter the statistics of a chaotic system.
+
 \end{itemize}

### 2021-05-06 highlights.tex
Changes made

@@
 \item Shadowing orbits may not be typical orbits or physical solutions
-\item This atypicality implies shadowing-based sensitivity can be biased.
-\item Numerical solutions of chaotic dynamical systems can be statistically inconsistent;
-\item Minor parameter changes can significantly alter the statistics of a chaotic system.
+\item This atypicality implies shadowing-based sensitivity can be biased
+\item Numerical solutions of chaotic dynamical systems can be statistically inconsistent
+\item Minor parameter changes can significantly alter the statistics of a chaotic system
 

### 2021-05-06 main_els.tex
highlights modify and change supp to appendix
Changes made

@@
 In the discussion in section \ref{sec:discussion}, we illustrate with an example that a small parameter perturbation can change statistics significantly.
-The supplementary material, section \ref{sec:supp}, contains the numerical procedure used to generate shadowing solutions.
+Appendix section \ref{sec:supp} contains the numerical procedure used to generate shadowing solutions.
 
@@
 the quasi-solutions of the tent map (section \ref{sec:quasiphysical}).
-The reader is referred to the Supplementary Material section \ref{sec:computingShadowing} for the computational method used in this paper, for the shadowing solutions.
+The reader is referred to the Appendix section \ref{sec:computingShadowing} for the computational method used in this paper, for the shadowing solutions.
 Note that, unlike the tilted tent maps, the family of squashed tent maps do not have the uniform distribution as their SRB distribution, but have a different regular probability distribution as indicated on the right column of Figure \ref{fig:tent_squashed_shadow}.
@@
 In order to analyze the behavior of the shadowing solutions of the Lorenz'63 system 
-of equations, we first make a closed form approximation of the Lorenz map (see section \ref{sec:lorenzMapApprox} of the supplementary material).
-Except at rare parameter values, the Lorenz system does not have
+of equations, we first make a closed form approximation of the Lorenz map (see section \ref{sec:lorenzMapApprox} of the Appendix). Except at rare parameter values, the Lorenz system does not have 
 solutions that shadow perturbed orbits for all time, but rather only for a finite time 
@@
 perturbed tent map -- the \emph{plucked} 
-tent map -- in which an oscillatory perturbation is introduced (see Supplementary Material section \ref{sec:pluckedTentMap} for the map equation).
+tent map -- in which an oscillatory perturbation is introduced (see Appendix section \ref{sec:pluckedTentMap} for the map equation).
 In Figure \ref{fig:pluckedTentMap}, we show that the magnitude of the oscillatory perturbation is controlled by parameters $s$ and $n$, and its frequency is controlled by the parameter $n$; at $s=0$, we recover the original tent map at all $n$.
@@
 section \ref{sec:tentPerts}.
-By construction of the map (Supplementary Material section \ref{sec:pluckedTentMap}), the nonuniformities of the probability distribution at $n=0$ are transferred to smaller and smaller scales,
+By construction of the map (Appendix section \ref{sec:pluckedTentMap}), the nonuniformities of the probability distribution at $n=0$ are transferred to smaller and smaller scales,
 as $n$ is increased.
@@
 \reviewerOne{Some authors} \cite{cns2}\cite{cns3} \reviewerOne{resort to Clean Numerical Simulations to tackle scenarios that demand highly accurate long-term prediction of chaotic orbits.
-When using double-precision arithmetic however, these works}\cite{cns}\cite{cns1}\cite{cns2}\cite{cns3}\reviewerOne{show that large errors result in statistical measures such as power spectra and spatio-temporal correlation functions.} The plucked tent map adds to this list of examples by demonstrating a mechanism to produce an extreme non-smooth response (Supplementary Material section \ref{sec:pluckedTentMap}).
+When using double-precision arithmetic however, these works}\cite{cns}\cite{cns1}\cite{cns2}\cite{cns3}\reviewerOne{show that large errors result in statistical measures such as power spectra and spatio-temporal correlation functions.} The plucked tent map adds to this list of examples by demonstrating a mechanism to produce an extreme non-smooth response (Appendix section \ref{sec:pluckedTentMap}).
 
@@
 \bibliography{refs.bib}
-\section{Supplementary material}
+
+\section{Appendix}
 \label{sec:supp}
 %\subsection{The Lorenz Map}
-The supplementary material for this paper including the code and data to generate the figures can be found on Github \cite{supp}.
+The code and data to generate the figures can be found on Github \cite{supp}.
 The code can be found under the \verb+code+ subdirectory inside which section-wise code is separated into further subdirectories.
